### 2019-04-29 Analysis/Simulation/Formula_sensitivity.Rmd
Add code to evaluate formula sensitivy to variables

---
title: "Formula_sensitivity"
author: "Lucas Avezum"
date: "April 29, 2019"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())
getwd()
knitr::opts_chunk$set(fig.width=6, fig.height=4)

## Load the libraries we will be using
library(tidyverse)
library(scales)
library(plyr)
library(gridExtra)
```

## Introduction

The growing literature on IRB risk-weights focus on the drivers and consequences of average risk-weights dispersion *across* banks. I intend to show that bank-level risk-weight dispersion, i.e., the standard deviation of risk-weight density *within* a bank, also plays a role. The study has two parts. First, I plan to investigate whether bank-level risk-weight dispersion is associated with better capitalization, and higher return on equity. In the second part, I want to uncover which variables explain the variation of bank-level risk-weight dispersion across banks, i.e., why some banks have a risk-weight distribution wider than others.

## Effects

```{r parameters and data, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 0.2
M   <- 2.5
PD  <- seq(0.0001, 0.2, by=0.00001) 
PD_1 <- rlnorm(length(PD),-4,0.1)
PD_1[PD_1<0] <- 0.0001
PD_2 <- rlnorm(length(PD),-4,0.5)
PD_2[PD_2<0] <- 0.0001
```

```{r formulas, echo=FALSE, include=FALSE}
mapping <- function(PD,LGD,M){
  R  <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
  b  <- (0.11852-0.05478*log(PD))^2
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(0.999))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
  RW <- K*12.5
  data <- data.frame(RW,PD)
  return(data)
}

```

### Why does risk-weight dispersion matters for capitalization and ROE?

According to the IRB framework, banks can use their own estimates of probability of default (PD).^[This is valid for banks approved both under the Foundation-IRB and Advanced-IRB. Banks under the latter approach may also use their own estimates of loss-given-default (LGD) and exposure-at-defualt (EAD).] Risk-weights are then mapped from these estimated PDs according to a pre-determined formula. The figure 1 shows how estimated PDs map into regulatory risk-weights.^[The figure shows the map for loans in the corporate sector, to large firms, assuming a 20% of loss given defaults and loan maturity of 2.5 years. The choice of assumptions does not affect my reasoning.]

```{r, echo=FALSE}
data.1 <- mapping(PD,LGD,M)

p <- ggplot(data = subset(data.1,subset = PD <1),
            mapping = aes(x = PD, y=RW)) 
p + geom_line(alpha = 1) + 
  labs(x = "PD", 
         y = "RW",title = "Figure 1:") +
  theme_bw()
```

The concativity of the risk-weight function means that, for a given average PD, an increase in the standard deviation of PD reduces the average risk-weight. Conversely, for a given average RW, greater RW dispersion implies higher average PD. Figure 2 illustrate the last point using simulated data.^[PDs are drawn from log-normal distributions and RW computed using the function shown in figure 1.]

```{r,echo=FALSE}
data.1 <- mapping(PD_1,LGD,M)
data.1 <- mutate(data.1,sample=factor("Sample 1"))
data.2 <- mapping(PD_2,LGD,M)
data.2 <- mutate(data.2,sample=factor("Sample 2"))
data <- rbind(data.1, data.2)
mu <- ddply(data, "sample", summarise, rw.mean=mean(RW),pd.mean=mean(PD))
d.mu <- ts(mu)
d.mu <- d.mu/stats::lag(d.mu,-1) - 1

p1 <- ggplot(data = subset(data,subset = PD<0.04),
            mapping = aes(x = PD, fill = sample, color = sample)) + 
  geom_density(alpha = 0.3) +
  geom_vline(data=mu,aes(xintercept = pd.mean,color=sample), linetype="dashed",size=1)+
  annotate(geom = "text", x = 0.021, y = 200,
                            label = "12.5% higher",
                            hjust = 0)+
  labs(x = "PD",
         title = "Figure 2:") +
  theme_classic() + theme(legend.position = "none")

p2 <- ggplot(data = subset(data,subset = RW<0.7 & RW >0.3),
            mapping = aes(x = RW, fill = sample, color = sample)) + 
  geom_density(alpha = 0.3) + 
  geom_vline(data=mu,aes(xintercept = rw.mean,color=sample), linetype="dashed",size=1)+
  annotate(geom = "text", x = 0.53, y = 25,
                            label = "0.3% higher",
                            hjust = 0)+
  labs(x = "RW") +
 theme_classic() + theme(legend.position = "none")

grid.arrange(p1, p2, nrow = 1)

```

Hence, by widening their RW density, banks can either reduce their risk-weighted asset and maintain their risk-taking level and/or increase risk-taking without affecting their capital ratios.

**Hypothesis 1** *Banks with wider risk-weight distribution are better capitalized*

**Hypothesis 2** *Banks with wider risk-weight distribution have higher return on equity*

# Empirical strategy

The main analysis relies on the following regressions:\newline

$\text{CAR}_{b,t}=\alpha_0+\alpha_1\widehat{sd(\text{RW})}_{b,t}+\alpha_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline

$\text{ROE}_{b,t}=\beta_0+\beta_1\widehat{sd(\text{RW})}_{b,t}+\beta_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline

where the independent variables, $\text{CAR}_{b,t}$ and $\text{ROE}_{b,t}$, are the capital ratio and return on equity of bank $b$ at year $t$, respectively. The variable of interest, $\widehat{sd(\text{RW})}_{b,t}$, captures exogenous changes in risk-weight dispersion and it is obtained as the fitted value for the following first stage regression: \newline

$sd(\text{RW})_{b,t}=\kappa_0+\kappa_1\text{IRB}_{b,t}+\kappa_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline

$sd(\text{RW})_{b,t}$ is the standard deviation of risk-weight of bank $b$ at time $t$. For the years before IRB approval $sd(\text{RW})_{b,t}$ is close to zero.^[Risk-weights under the standardized approach are flat within an asset class.] After approval we replace this variable by the standard deviation of the IRB risk-weight density. The instrument, $IRB_{b,t}$, is a dummy variable that takes 1 for the periods after bank $b$ is allowed to calculate risk-weights using IRB models.

## Causes and policy 
After stabilishing the importance of risk-weight dispersion, I focus on the policy implications of this finding.

$\text{sd(RW)}_{b,c,t}=\gamma_0+\gamma_1\text{Regulation}_{c,t}+\gamma_2\text{Bank}_{b,t}+\varepsilon_{b,c,t}$,  \newline

$\text{Regulation}_{c,t}$ is a vector of variables that may include regulatory standards, for instance, from the Barth dataset on bank regulation. $\text{Bank}_{b,t}$ is a vector bank characteristics that may affect the dispersion of risk-weight, such as, NPL ratio, size and diversification. Comparing the economic and statistical significance between $\gamma_1$ and $\gamma_2$ shows to what extent risk-weight dispersion is the unintended result of regulation or specific to banks' strategies and possibilities.\newline

## Data

For this project, I will have to merge the Bankscope dataset to risk-weight information from Pillar-III reports, which we are already collecting.

### 2019-05-28 Analysis/Simulation/Formula_sensitivity.Rmd
Add few temp files to analyze data on road

@@
 library(gridExtra)
+library(DescTools)
+library(lfe)
+library(stargazer)
 ```
-
-## Introduction
-
-The growing literature on IRB risk-weights focus on the drivers and consequences of average risk-weights dispersion *across* banks.
-I intend to show that bank-level risk-weight dispersion, i.e., the standard deviation of risk-weight density *within* a bank, also plays a role.
-The study has two parts.
-First, I plan to investigate whether bank-level risk-weight dispersion is associated with better capitalization, and higher return on equity.
-In the second part, I want to uncover which variables explain the variation of bank-level risk-weight dispersion across banks, i.e., why some banks have a risk-weight distribution wider than others.
-
-## Effects
 
@@
 M   <- 2.5
-PD  <- seq(0.0001, 0.2, by=0.00001) 
-PD_1 <- rlnorm(length(PD),-4,0.1)
+grid  <- seq(0.0001, 1, by=0.00001) 
+grid_2  <- seq(0, 10, by=0.01) 
+PD_1 <- rlnorm(length(grid),-4,0.1)
 PD_1[PD_1<0] <- 0.0001
-PD_2 <- rlnorm(length(PD),-4,0.5)
+PD_1_avg <- mean(PD_1)
+PD_1_sd <- sd(PD_1)
+PD_2 <- rlnorm(length(grid),-4,0.5)
 PD_2[PD_2<0] <- 0.0001
+PD_2[PD_2>0.9999] <- 0.9999
+PD_2_avg <- mean(PD_2)
+PD_2_sd <- sd(PD_2)
 ```
@@
 ```{r formulas, echo=FALSE, include=FALSE}
-mapping <- function(PD,LGD,M){
+mapping_wholesale <- function(PD,LGD,M,alpha,portfolio){
   R  <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
   b  <- (0.11852-0.05478*log(PD))^2
-  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(0.999))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
+  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
   RW <- K*12.5
-  data <- data.frame(RW,PD)
+  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
   return(data)
@@
 
+mapping_retail <- function(PD,LGD,R,alpha,portfolio){
+  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
+  RW <- K*12.5
+  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
+  return(data)
+}
+
+mapping_other_retail <- function(PD,LGD,alpha,portfolio){
+  R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
+  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
+  RW <- K*12.5
+  data <- data.frame(K,RW,PD,LGD,R,portfolio)
+  return(data)
+}
 ```
 
-### Why does risk-weight dispersion matters for capitalization and ROE?
+```{r wholesale vs retail, echo=FALSE}
+data <- mapping_wholesale(grid,0.45,2.5,0.999,factor("Corporate"))
+data <- rbind(data,mapping_retail(grid,0.45,0.15,0.999,factor("Mortgage")))
 
-According to the IRB framework, banks can use their own estimates of probability of default (PD).^[This is valid for banks approved both under the Foundation-IRB and Advanced-IRB.
-Banks under the latter approach may also use their own estimates of loss-given-default (LGD) and exposure-at-defualt (EAD).] Risk-weights are then mapped from these estimated PDs according to a pre-determined formula.
-The figure 1 shows how estimated PDs map into regulatory risk-weights.^[The figure shows the map for loans in the corporate sector, to large firms, assuming a 20% of loss given defaults and loan maturity of 2.5 years.
-The choice of assumptions does not affect my reasoning.]
-
-```{r, echo=FALSE}
-data.1 <- mapping(PD,LGD,M)
-
-p <- ggplot(data = subset(data.1,subset = PD <1),
-            mapping = aes(x = PD, y=RW)) 
+p <- ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K, colour = portfolio)) 
 p + geom_line(alpha = 1) + 
   labs(x = "PD", 
-         y = "RW",title = "Figure 1:") +
+         y = "Capital requirement per unit of exposure",title = "Figure 1: Wholesale vs Retail") +
   theme_bw()
@@
 
-The concativity of the risk-weight function means that, for a given average PD, an increase in the standard deviation of PD reduces the average risk-weight.
-Conversely, for a given average RW, greater RW dispersion implies higher average PD.
-Figure 2 illustrate the last point using simulated data.^[PDs are drawn from log-normal distributions and RW computed using the function shown in figure 1.]
+```{r LGD sensitivity, echo=FALSE}
+data <- mapping_wholesale(grid,0.4,2.5,0.999,factor("LGD = 40%"))
+data <- rbind(data,mapping_wholesale(grid,0.45,2.5,0.999,factor("LGD = 45%")))
+data <- rbind(data,mapping_wholesale(grid,0.5,2.5,0.999,factor("LGD = 50%")))
 
-```{r,echo=FALSE}
-data.1 <- mapping(PD_1,LGD,M)
-data.1 <- mutate(data.1,sample=factor("Sample 1"))
-data.2 <- mapping(PD_2,LGD,M)
-data.2 <- mutate(data.2,sample=factor("Sample 2"))
-data <- rbind(data.1, data.2)
-mu <- ddply(data, "sample", summarise, rw.mean=mean(RW),pd.mean=mean(PD))
-d.mu <- ts(mu)
-d.mu <- d.mu/stats::lag(d.mu,-1) - 1
+p <- ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K, colour = portfolio)) 
+p + geom_line(alpha = 1) + 
+  labs(x = "PD", 
+         y = "Capital requirement per unit of exposure",title = "Figure 2: PD mapping") +
+  theme_bw()
+```
+```{r M sensitivity, echo=FALSE}
+data <- mapping_wholesale(grid,0.1,0,0.999,factor("M = 0 year"))
+data <- rbind(data,mapping_wholesale(grid,0.1,2.5,0.999,factor("M = 2.5 years")))
+data <- rbind(data,mapping_wholesale(grid,0.1,5,0.999,factor("M = 5 years")))
 
-p1 <- ggplot(data = subset(data,subset = PD<0.04),
-            mapping = aes(x = PD, fill = sample, color = sample)) + 
-  geom_density(alpha = 0.3) +
-  geom_vline(data=mu,aes(xintercept = pd.mean,color=sample), linetype="dashed",size=1)+
-  annotate(geom = "text", x = 0.021, y = 200,
-                            label = "12.5% higher",
-                            hjust = 0)+
+p <- ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K, colour = portfolio)) 
+p + geom_line(alpha = 1) + 
   labs(x = "PD", 
-         title = "Figure 2:") +
-  theme_classic() + theme(legend.position = "none")
+         y = "Capital requirement per unit of exposure",title = "Figure 3: PD mapping") +
+  theme_bw()
+```
 
-p2 <- ggplot(data = subset(data,subset = RW<0.7 & RW >0.3),
-            mapping = aes(x = RW, fill = sample, color = sample)) + 
-  geom_density(alpha = 0.3) + 
-  geom_vline(data=mu,aes(xintercept = rw.mean,color=sample), linetype="dashed",size=1)+
-  annotate(geom = "text", x = 0.53, y = 25,
-                            label = "0.3% higher",
-                            hjust = 0)+
-  labs(x = "RW") +
- theme_classic() + theme(legend.position = "none")
+```{r LGD_mapping, echo=FALSE}
+data <- mapping_wholesale(0.01,grid,2.5,0.999,factor("PD = 1%"))
+data <- rbind(data,mapping_wholesale(0.05,grid,2.5,0.999,factor("PD = 5%")))
+data <- rbind(data,mapping_wholesale(0.1,grid,2.5,0.999,factor("PD = 10%")))
 
-grid.arrange(p1, p2, nrow = 1)
+p <- ggplot(data = subset(data,subset = PD <1),
+            mapping = aes(x = LGD, y=K, colour = portfolio)) 
+p + geom_line(alpha = 1) + 
+  labs(x = "LGD", 
+         y = "Capital requirement per unit of exposure",title = "Figure 4: LGD mapping") +
+  theme_bw()
+```
+
+```{r M_mapping, echo=FALSE}
+data <- mapping_wholesale(0.01,0.45,grid_2,0.999,factor("PD = 1%"))
+data <- rbind(data,mapping_wholesale(0.05,0.45,grid_2,0.999,factor("PD = 5%")))
+data <- rbind(data,mapping_wholesale(0.1,0.45,grid_2,0.999,factor("PD = 10%")))
+
+p <- ggplot(data = data,
+            mapping = aes(x = M, y=K, colour = portfolio)) 
+p + geom_line(alpha = 1) + 
+  labs(x = "Maturity", 
+         y = "Capital requirement per unit of exposure",title = "Figure 5: Maturity mapping") +
+  theme_bw()
+```
+
+```{r savings simulation, echo=FALSE,include=FALSE}
+
+K_bar <- 99991*mapping_wholesale(mean(PD_1),0.45,2.5,0.999,factor("Average PD"))["K"]
+K_i <- mapping_wholesale(PD_1,0.45,2.5,0.999,factor("PD distribution"))
+new <- summarise(K_i,EAD = Freq(K))
+K_total <- K_i  %>%
+  group_by(portfolio) %>% 
+  summarise_all(sum)
+
+print(K_bar-K_total["K"])
+print(K_total["K"]/K_bar-1)
+
+K_bar <- 99991*mapping_wholesale(mean(PD_2),0.45,2.5,0.999,factor("Average PD"))["K"]
+K_i <- mapping_wholesale(PD_2,0.45,2.5,0.999,factor("PD distribution"))
+K_total <- K_i  %>%
+  group_by(portfolio) %>% 
+  summarise_all(sum)
+
+print(K_bar-K_total["K"])
+print(K_total["K"]/K_bar-1)
+
+K_bar <- 99991*mapping_wholesale(mean(PD_2),0.55,2.5,0.999,factor("Average PD"))["K"]
+K_i <- mapping_wholesale(PD_2,0.55,2.5,0.999,factor("PD distribution"))
+K_total <- K_i  %>%
+  group_by(portfolio) %>% 
+  summarise_all(sum)
+
+print(K_bar-K_total["K"])
+print(K_total["K"]/K_bar-1)
+
+K_bar <- 99991*mapping_wholesale(mean(PD_2),0.45,0,0.999,factor("Average PD"))["K"]
+K_i <- mapping_wholesale(PD_2,0.45,0,0.999,factor("PD distribution"))
+K_total <- K_i  %>%
+  group_by(portfolio) %>% 
+  summarise_all(sum)
+
+print(K_bar-K_total["K"])
+print(K_total["K"]/K_bar-1)
 
@@
 
-Hence, by widening their RW density, banks can either reduce their risk-weighted asset and maintain their risk-taking level and/or increase risk-taking without affecting their capital ratios.
+```{r PD LGD correlation, echo=FALSE}
+load("../../Data/Temp/PD_LGD_correlation.Rda")
 
-**Hypothesis 1** *Banks with wider risk-weight distribution are better capitalized*
+reg.hdfe1 <- felm(LGD~ PD | Portfolio+Bank|0|0,
+                  data = subset(data,subset = PD<1))
+robust.hdfe1 <- summary(reg.hdfe1,robust=TRUE)$coefficients[,2]
 
-**Hypothesis 2** *Banks with wider risk-weight distribution have higher return on equity*
+reg.hdfe2 <- felm(LGD~ PD | Bank|0|0,
+                  data = subset(data,subset = PD<1 & Portfolio == "Real estate"))
+robust.hdfe2 <- summary(reg.hdfe2,robust=TRUE)$coefficients[,2]
 
-# Empirical strategy
+reg.hdfe3 <- felm(LGD~ PD | Bank|0|0,
+                  data = subset(data,subset = PD<1 & Portfolio == "Qualifying revolving" ))
+robust.hdfe3 <- summary(reg.hdfe3,robust=TRUE)$coefficients[,2]
 
-The main analysis relies on the following regressions:\newline
+reg.hdfe4 <- felm(LGD~ PD |  Bank|0|0,
+                  data = subset(data,subset = PD<1 & Portfolio == "Other retail"))
+robust.hdfe4 <- summary(reg.hdfe4,robust=TRUE)$coefficients[,2]
 
-$\text{CAR}_{b,t}=\alpha_0+\alpha_1\widehat{sd(\text{RW})}_{b,t}+\alpha_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline
+reg.hdfe5 <- felm(LGD~ PD |  Bank|0|0,
+                  data = subset(data,subset = PD<1 & Portfolio == "Corporate"))
+robust.hdfe5 <- summary(reg.hdfe5,robust=TRUE)$coefficients[,2]
 
-$\text{ROE}_{b,t}=\beta_0+\beta_1\widehat{sd(\text{RW})}_{b,t}+\beta_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline
+reg.hdfe6 <- felm(LGD~ PD |  Bank|0|0,
+                  data = subset(data,subset = PD<1 & Portfolio == "Banks"))
+robust.hdfe6 <- summary(reg.hdfe6,robust=TRUE)$coefficients[,2]
 
-where the independent variables, $\text{CAR}_{b,t}$ and $\text{ROE}_{b,t}$, are the capital ratio and return on equity of bank $b$ at year $t$, respectively.
-The variable of interest, $\widehat{sd(\text{RW})}_{b,t}$, captures exogenous changes in risk-weight dispersion and it is obtained as the fitted value for the following first stage regression: \newline
+reg.hdfe7 <- felm(LGD~ PD |  Bank|0|0,
+                  data = subset(data,subset = PD<1 & Portfolio == "Sovereign"))
+robust.hdfe7 <- summary(reg.hdfe7,robust=TRUE)$coefficients[,2]
+```
+```{r PD LGD table, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
-$sd(\text{RW})_{b,t}=\kappa_0+\kappa_1\text{IRB}_{b,t}+\kappa_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline
+stargazer(reg.hdfe1,reg.hdfe2,reg.hdfe3,reg.hdfe4,reg.hdfe5,reg.hdfe6,reg.hdfe7,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small", 
+          label = "",
+          style = "qje", 
+          column.labels   = c("All", "Real estate","Credit cards", "Other retail","Corporates","Banks","Sovereign"),
+          dep.var.labels = c("LGD"), column.sep.width = "2pt",
+          add.lines = list(c("Portfolio FE", "Yes", "-","-","-","-","-","-"),
+                           c("Bank FE","Yes", "Yes","Yes","Yes","Yes","Yes","Yes")))
 
-$sd(\text{RW})_{b,t}$ is the standard deviation of risk-weight of bank $b$ at time $t$.
-For the years before IRB approval $sd(\text{RW})_{b,t}$ is close to zero.^[Risk-weights under the standardized approach are flat within an asset class.] After approval we replace this variable by the standard deviation of the IRB risk-weight density.
-The instrument, $IRB_{b,t}$, is a dummy variable that takes 1 for the periods after bank $b$ is allowed to calculate risk-weights using IRB models.
-
-## Causes and policy 
-After stabilishing the importance of risk-weight dispersion, I focus on the policy implications of this finding.
-
-$\text{sd(RW)}_{b,c,t}=\gamma_0+\gamma_1\text{Regulation}_{c,t}+\gamma_2\text{Bank}_{b,t}+\varepsilon_{b,c,t}$,  \newline
-
-$\text{Regulation}_{c,t}$ is a vector of variables that may include regulatory standards, for instance, from the Barth dataset on bank regulation.
-$\text{Bank}_{b,t}$ is a vector bank characteristics that may affect the dispersion of risk-weight, such as, NPL ratio, size and diversification.
-Comparing the economic and statistical significance between $\gamma_1$ and $\gamma_2$ shows to what extent risk-weight dispersion is the unintended result of regulation or specific to banks' strategies and possibilities.\newline
-
-## Data
-
-For this project, I will have to merge the Bankscope dataset to risk-weight information from Pillar-III reports, which we are already collecting.
+```

### 2019-05-28 Analysis/Simulation/report.Rmd
Add few temp files to analyze data on road

---
title: "JMP report"
author: "Lucas Avezum"
date: "May 14, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)                                                             ## excel files and,
library(readxl)                                                            ## load the libraries.
library(data.table)
library(foreign)
library(lubridate)
library(lfe)
library(stargazer)
library(ineq)
library(DescTools)
library(gglorenz)

```
```{r, include=FALSE}
load("../../Data/Temp/Raw_data_2.Rda")
lc.wholesale <- data %>% filter(Portfolio_lv1 == "Wholesale") 
lc.wholesale <- Lc(lc.wholesale$PD,lc.wholesale$EAD*lc.wholesale$LGD)

lc.retail <- data %>% filter(Portfolio_lv1 == "Retail") 
lc.retail <- Lc(lc.retail$PD,lc.retail$EAD*lc.retail$LGD)
```

- Average Capital savings using $\hat{P}$ and LGD original distribution: 32.8%

- Average Gini coefficient $PD$ weighted by $LGD$: 0.67

- Correlation between Gini coefficient and capital savings: 0.79

- $\hat{P}=\frac{\sum_i^nEAD_iLGD_iPD_i}{\sum_i^nEAD_iLGD_i}$ \newline

```{r lorenz curve, echo=FALSE,}
plot(lc.wholesale,
     col="black",
     type="l",      # !is not working
     lty=1,
     lwd=3,
     main="Lorenz Curve by Portfolio"     
)

lines(lc.retail, lty=2, lwd=3)

legend("topleft",
       c("Wholesale", "Retail"),
       lty=c(1,2,3),
       lwd=3)
```

### 2019-05-28 Data/Data_Import.R

@@ -1,0 +1,362 @@
+##-----------------------------------------------------------------------------##
+## Project: Bank Regulation and Capital Structure                              ##
+## Author: Lucas Avezum, Tilburg University                                    ##
+## Date: 10/10/2018                                                            ##
+## Description: this file opens, merges and saves in .Rda format the following ## 
+## datasets: Dealscan, EBA-transparency exercises                              ##
+##-----------------------------------------------------------------------------##
+  
+##=============================================================================##
+## Code setup                                                                  ##
+##=============================================================================##
+
+rm(list = ls())                                                                ## Clean enviroment,
+library(tidyverse)                                                             ## excel files and,
+library(readxl)                                                            ## load the libraries.
+library(data.table)
+library(foreign)
+library(lubridate)
+library(lfe)
+library(stargazer)
+library(ineq)
+library(DescTools)
+library(gglorenz)
+
+##============================================================================##
+## Dealscan                                                                   ##
+##============================================================================##
+
+
+data.path  <- "../../../Data/Financial/Exposure_data_collection/Data/Second_list"
+data.names <-  list.files(path = data.path,pattern="\\.xlsx",full.names = TRUE)
+data.list  <- data.names %>%
+  lapply(read_xlsx,sheet = 1,  skip=1)   
+ 
+
+data <-  rbindlist(data.list, use.names=TRUE,fill = TRUE) %>%
[…]

### 2019-07-08 Data/Clean.R
Create codes for importing, merging, and cleaning datasets

@@ -1,0 +1,117 @@
+##-----------------------------------------------------------------------------##
+## Project: PD distribution and capital savings                                ##
+## Author: Lucas Avezum, Tilburg University                                    ##
+## Description: this file cleans the raw dataset in .R format                  ##
+##-----------------------------------------------------------------------------##
+
+##=============================================================================##
+## Code setup                                                                  ##
+##=============================================================================##
+
+rm(list = ls())  
+library(tidyverse)                ## Data manipulation, pipe operator                                                    
+source("Auxiliar/Function.R")     ## Capital requirements functions
+
+##============================================================================##
+## Pillar-III reports                                                         ##
+##============================================================================##
+
+# Open dirt data frame
+load("Data/Temp/Pillar3Data.Rda")
+
+# Create and adjust variables 
+pillar3.data <- pillar3.data %>%
+  rename(Portfolio_1 = `Portfolio - level 1`,
+         Portfolio_2 = `Portfolio - level 2`) %>%
+  mutate(month       = as.numeric(sapply(Date,str_sub,start= 4,end=5)),
+         year        = as.numeric(sapply(Date,str_sub,start= -4)),
+         year        = ifelse(is.na(month),year,ifelse(month<7,year-1,year)),
+         PD          = as.numeric(PD),
+         RW          = as.numeric(RW),
+         EAD         = as.numeric(EAD),
+         bvdid       = as.factor(bvdid),
+         Method      = as.factor(Method),
+         Method2     = as.factor(ifelse(Method %in% c("IRB"), "IRB",sapply(Method,str_sub,start= 2,end=4))),
+         Portfolio_1 = as.factor(Portfolio_1),
[…]

### 2019-07-08 Data/Import.R

@@ -1,0 +1,91 @@
+##-----------------------------------------------------------------------------##
+## Project: PD distribution and capital savings                                ##
+## Author: Lucas Avezum, Tilburg University                                    ##
+## Description: this file opens, merges and saves in .Rda format               ##
+##-----------------------------------------------------------------------------##
+  
+##=============================================================================##
+## Code setup                                                                  ##
+##=============================================================================##
+
+rm(list = ls()) 
+library(tidyverse)    ## Data manipulation, pipe operator                                                             
+library(readxl)       ## Command to open xlsx files                                                     
+library(data.table)   ## Command to bind lists
+library(reshape2)     ## Command to transform data from wide to long
+library(stringr)
+
+# Load auxiliar data frame: list of parent banks
+guo.list <- read_xlsx("../../../Data/Financial/Exposure_data_collection/Auxiliar/List.xlsx",sheet = 1) %>%
+  select(name,bvdid)
+
+##============================================================================##
+## Pillar-III reports                                                         ##
+##============================================================================##
+
+# Import collected datasets from xlsx files and merge all in data frame data.pillar3
+data.path  <- "Data/Raw/Pillar3"
+data.names <-  list.files(path = data.path,pattern="\\.xlsx",full.names = TRUE)
+data.list  <- data.names %>%
+  lapply(read_xlsx,sheet = 1,  skip=1)   
+pillar3.data <-  rbindlist(data.list, use.names=TRUE,fill = TRUE) %>%
+  select(1:25)
+
+# Add short names to data frame
+pillar3.data <- inner_join(pillar3.data, guo.list)
+
+# Save data frame
+save(pillar3.data,file=paste0("Data/Temp/Pillar3Data.Rda"))
+
+##============================================================================##
+## Orbis Bank Focus (BankScope)                                               ##
+##============================================================================##
+
+# Import Orbis dataset from txt file 
+data.path  <- "Data/Raw/BankScope"
[…]

### 2019-07-08 Analysis/Draft/Draft.Rmd
Create draft code and auxiliar files

---
title: "Distribution of portfolio credit risk and capital savings"
author:
- affiliation: Tilburg University; l.avezum@uvt.nl
  name: Lucas Avezum
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
  html_document:
    df_print: paged
  word_document: default
bibliography: null
citecolor: null
anonymous: no
fontsize: 12pt
geometry: margin= 1in
keywords: Key1, Key2, Key3
linkcolor: blue
biblio-style: apsr
subtitle: ''
thanks: Very preliminary draft
abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
  a concave mapping from probabilities of default to capital requiremets. As a consequence,
  banks can reduce requirements or increase risk-taking by widening their portfolio
  credit risk distribution.'
toc: no
urlcolor: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = TRUE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

rm(list = ls())

library(tidyverse)           ## Data manipulation, pipe operator
library(splitstackshape)     ## Command to expand data
library(data.table)          ## Command to bind lists
library(ggridges)            ## Density ridges plot
library(stargazer)           ## Latex tables
library(lfe)                 ## High-dimensional fixed effects
library(gridExtra)           ## Plot side-by-side
library(DescTools)           ## Command to winsorize variables

p_w <- c(0.0, 1)
```

```{r load data, include=FALSE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")

# Expand Pillar-III data frame to plot densities and lorenz curves
expanded.list <- list()
for(i in c("EAD","EL")) {
  expanded.list[[i]] <- pillar3.data %>%
  filter(Method2 %in% c("IRB")) %>%
  select(name,Country,Method,Portfolio_1,EAD,LGD,PD,year)%>%
  mutate(EL  = round(EAD*LGD/1000),
         EAD = round(EAD/1000))%>%
    na.omit(cols=c(i))%>%
    expandRows(i) %>%
    mutate(freq_type = as.factor(i))
}
pillar3.expanded <-  rbindlist(expanded.list, use.names=TRUE,fill = TRUE)

# Aggregate Pillar-III data by bank and merge to bankscope
bank.data <- pillar3.data %>% 
  group_by(name,bvdid,Country,year) %>%
  mutate(weight=EAD*LGD)%>%
  filter(weight>0,
         PD<1&PD>0,
         Portfolio_1 %in% c("Wholesale","Retail"), 
         Method2 %in% c("IRB"))%>%
  dplyr::summarize(gini             = DescTools::Gini(PD,weight),
                   PD_sd            = sd(PD)*100,
                   PD_mean          = mean(PD)*100,
                   K_calculated     = sum(K_calculated),
                   K_original       = sum(K_original),
                   K_counterfactual = sum(K_counterfactual)) %>%
  ungroup()%>% 
  left_join(bank.data,by=c("bvdid","year"))%>%
group_by(bvdid) %>% 
  mutate(savings_log  = log(K_counterfactual-K_calculated),
         Total.assets = Total.assets/10^6,
         Total.Capital= Total.Capital/10^6,
         Tier.1.Capital=Tier.1.Capital/10^6,
         asset_log    = log(Total.assets),
         capital_log  = log(Total.Capital),
         tier.1_log   = log(Tier.1.Capital),
         savings_rate = (K_counterfactual-K_calculated)/K_calculated,
         savings_alt  = (K_counterfactual-K_original)/K_original,
         error        = (K_calculated - K_original)/K_original,
         asset_d      = Winsorize(log(Total.assets)-lag(log(Total.assets),order_by=year), 
                                    probs = p_w ,na.rm = TRUE),
         capital_d    = Winsorize(log(Total.Capital)-lag(log(Total.Capital),order_by=year), 
                                    probs = p_w ,na.rm = TRUE),
         PD_mean_d    = Winsorize(PD_mean-lag(PD_mean,order_by=year), 
                                    probs = p_w ,na.rm = TRUE),
         savings_d    = Winsorize(savings_log-lag(savings_log,order_by=year), 
                                    probs = p_w ,na.rm = TRUE),
         car_d  = Winsorize(Total.Capital.Ratio/100-lag(Total.Capital.Ratio/100,order_by=year), 
                                    probs = p_w ,na.rm = TRUE),
         roe_d  = Winsorize(ROE.using.P.L.before.tax..../100-lag(ROE.using.P.L.before.tax..../100,order_by=year),
                                    probs = p_w ,na.rm = TRUE)) %>%
  ungroup()

# Aggregate Pillar-III data by Portfolio
portfolio.data <- pillar3.data %>% 
  group_by(Country,Bank,year,Portfolio_1,Portfolio_2) %>%
  mutate(weight=EAD*LGD)%>%
  filter(weight>0, Portfolio_1 %in% c("Wholesale","Retail"))%>%
  dplyr::summarize(gini             = DescTools::Gini(PD,weight),
                   K_calculated     = sum(K_calculated),
                   K_counterfactual = sum(K_counterfactual)) %>%
  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)

```
# Introduction

# Hypotheses and Data
In this section I describe the hypotheses of the paper by explaining the Basel II regulatory formulae under the IRB approach to calculate RWA for credit risk.

## Capital savings measure

In this section I describe the chosen method to measure differences in risk distributions:

\begin{equation}
\text{Capital savings}_{i}= f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i),
\end{equation}

where $i$ is one asset of the portfolio, $f()$ is the regulatory formula and,

\begin{equation}
\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
\end{equation}

## Data

Figure 1 shows that the log of _Capital savings_ varies considerably across banks. While most canadian banks seem to have not explore this possilibity to save capital, the european banks in our sample appear to have more than halved their capital requirement due to the distribution mechanism. Importantly, this comparison does not means that european banks are riskier than their canadian counterparties but only that their credit risk distribution is more disperse. The figure also shows that there capital savings have changed in time. However, the figure is not appropriated to see how this variable evolved in time.

```{r}
fig1_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
```

```{r 1, fig.height = 4, fig.width = 5,results="hide",fig.cap= fig1_cap,fig.align = 'center'}
bank.data %>% 
 ggplot(mapping = aes(x = reorder(name, savings_rate,FUN = median, na.rm=TRUE),
                          y = savings_rate,
                          color = Country,
                          fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital savings") +

  theme_bw()

```

Figure 2 shows that both _Capital savings_ and total capital ratio have increased slighly on average in time. After a spike in 2010, the average return on equity appears to have remained constant in time.

```{r}
fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE. The solid line is the average across banks for each year. The first and third quartiles are shown in the gray area.  "
```

```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth"}
mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
  dat <- data.frame(y = mean(x, na.rm = TRUE),
                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
  return(dat)
}

plot1 <- bank.data %>% 
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
 geom_smooth(aes( y=savings_log, color = savings_log),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Capital savings') +
  theme_bw()

plot2 <- bank.data %>% mutate(Tier.1.Ratio= Tier.1.Ratio/100) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=Tier.1.Ratio, color = Tier.1.Ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Total capital ratio') +
  theme_bw()

plot3 <- bank.data %>% mutate(ROE.using.P.L.before.tax....= ROE.using.P.L.before.tax..../100) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=ROE.using.P.L.before.tax...., color = ROE.using.P.L.before.tax....),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'ROE') +
  theme_bw()

grid.arrange(plot1, plot2, plot3, ncol=3)
```

Figure 3 shows the linear relationship between our variable of interest, _Capital savings_ and three potential outcome variables: core tier 1 capital ratio, return on equity and, a measure of riskness, Z-score.

```{r}
fig3_cap = "PD distribution by bank. The figure shows smoothed PD distributions for each bank in our data. Banks are ranked by median PD and colored by host-country"
```

```{r, fig.height = 10, fig.width = 7, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
pillar3.expanded %>% 
  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
                       scale = 2 , from = 0, to = 0.05) +
  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    x = 'Probability of default') +
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())
```

Finally, we supplement the pillar-3 report data with annual information on banks' balance sheets obtained from the Bureau van Dijk's BankFocus database. Table 1 show summary statistics of the final dataset. The sample includes 25 banks. We restrict the analysis to banks for which we are able to collect information from Pillar-III reports.

The current geographical composition of the sample is: 1 norwegian, 1 italian, 1 spanish, 2 british, 4 australian, 6 canadian, and 10 american banks. The target geographical composition is: 1 norwegian, 1 austrian, 2 finish, 3 belgium, 3 danish, 4 italian, 4 spanish banks, 4 dutch, 4 australian, 4 swedish, 5 french, 5 british, 6 canadian, 7 german, and 10 american banks. This sums up to 63 IRB banks. The selection criteria is asset size.

```{r}
fig1_cap = "The figure plots the linear relationship between the capital savings measure and,in descending order, the tier 1 capital ratio, the return on equity using PL before taxes, and the Z-score. In each graph, a point is the average across time for each bank."
```

```{r relation, fig.height = 3, fig.width = 6, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
plot1 <- bank.data %>% group_by(name)%>%
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
 ggplot(mapping = aes(x = savings_log,
                      y = Tier.1.Ratio)) +
geom_point( alpha = 0.5,size = 1) +
  labs(x="Capital savings",y="CT1 capital ratio") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()

plot2 <- bank.data %>% group_by(name)%>%
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
 ggplot(mapping = aes(x = savings_log,
                      y = ROE.using.P.L.before.tax....)) +
geom_point( alpha = 0.5,size = 1) +
  labs(x="Capital savings",y="ROE") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()

plot3 <- bank.data %>% group_by(name)%>%
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
 ggplot(mapping = aes(x = savings_log,
                      y = `Z-score`)) +
geom_point( alpha = 0.5,size = 1) +
  labs(x="Capital savings",y="Z-score") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()

grid.arrange(plot1, plot2,plot3, ncol=3)

```
Banks that only implement the standardized approach (SA) for credit risk most of the sample period will also be included in the dataset. The current geographical composition of SA banks is: 1 portuguese, 1 dutch, 1 belgium, 1 finish, 1 hungarian, 2 maltese, 2 british, 2 greek, 2 polish, 3 french, 3 slovenian, 4 cypriot, 5 german, 5 italian, 8 spanish, 37 american. This sums up to 78 SA banks. The selection criteria is the same, asset size.

```{r summary, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
bank.data %>% 
  subset(select = c("savings_log","PD_mean","gini","Tier.1.Ratio","ROE.using.P.L.before.tax....","Z-score","asset_log","tier.1_log")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("log Savings","Portfolio mean PD","Gini coefficient","CT1 ratio","ROE","Z-score", "log Total asset","log CT1 capital"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{11cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

# Empirical strategy
In this section I briefly describe the empirical strategy of the paper.

## EBA capital exercise and the effect on capital savings

To find out if banks adjust their portfolio risk distribution in response to capital requirement shock, I intend to explore the 2011 EBA capital exercise. This was a strong and unexpected test to the solvency of 63 european banks that resulted in a recommendation for these banks to build up their capital buffert to reach a 9.0% core tier 1 capital ratio by June 2012. Identification of the effect of higher capital requirements on banks'portfolio risk distribution comes from comparing the change in the _Capital savings_ measure (or other measure of the distribution such as the Gini coefficient) around the exercise between banks that participated in the exercise and banks that did not. In particular, I (will) estimate the following specification::

\begin{equation}
\Delta\text{Capital savings}_{i}= \beta\text{CEB}_{i} + \varepsilon_{i},
\end{equation}

where $\text{CEB}_{i}$ is an indicator function that equals one if bank $i$ was included in the exercise and $\beta$ is the treatment effect of higher requirements on portfolio risk distribution.

##  Basel II and the effect of capital savings

Once we know that banks adjust their portolio risk distribution to achieve regulatory capital requirements, an natural extension is to quantify the effect of these changes on capital ratio and evaluate if these changes impact other variables of interest, for instance, return on equity and risk-taking. Endogeneity is an important issue to consider. For instance, worse capitalized banks have a stronger incentive to adjust their portfolios compared to their better capitalized peers resulting in downward biased estimates. Conversely, better-capitalized banks may have more leeway to achieve a portfolio distribution that maximizes return by increasing risk-taking. In this case the estimated effects of capital savings on return on equity and risk-taking will have a positive bias.

To solve the endogeneity concerns and estimate the causal effect of portfolio risk distribution, I (will) explore the time differences in introduction of Basel II across countries. Only after Basel II that the IRB models were allowed to be used and, consequently, the distribution of PD matter for the calculation of RWA. In particular, I (will) estimate the following specification:

\begin{equation}
Y_{it}= \beta\widehat{\text{Capital savings}}_{it}+\varepsilon_{i},
\end{equation}

where $Y_{it} \in \{\text{CT1 ratio},\text{ROE},\text{Z-score}\}$ and $\widehat{\text{Capital savings}}_{it}$ is the fitted values of the first stage equation:

\begin{equation}
\text{Capital savings}_{it}= \alpha\text{Basel II}_{it}+\epsilon_{i},
\end{equation}

where $\text{Basel II}_{it}$ is an indicator function that equals one if bank $i$ in year $t$ is hosted in country that has adopted Basel II and has its own IRB model approved.

# (Preliminary) Results

```{r, echo=FALSE}

reg.hdfe1 <- felm(Tier.1.Ratio  ~ savings_log  | 0  |0|0,
                  data = bank.data)

reg.hdfe1.1 <- felm(Tier.1.Ratio  ~ savings_log + asset_log + tier.1_log +PD_mean | 0  |0|0,
                  data = bank.data)

reg.hdfe1.2 <- felm(Tier.1.Ratio  ~ savings_log + asset_log + tier.1_log +PD_mean | year +name  |0|0,
                  data = bank.data)

reg.hdfe2 <- felm(ROE.using.P.L.before.tax....~ savings_log  | 0 |0|0,
                  data = bank.data)
reg.hdfe2.1 <- felm(ROE.using.P.L.before.tax....~ savings_log + asset_log + tier.1_log +PD_mean | 0 |0|0,
                  data = bank.data)
reg.hdfe2.2 <- felm(ROE.using.P.L.before.tax....~ savings_log + asset_log + tier.1_log +PD_mean | year +name |0|0,
                  data = bank.data)

reg.hdfe3 <- felm(`Z-score`~ savings_log  | 0 |0|0,
                  data = bank.data)
reg.hdfe3.1 <- felm(`Z-score`~ savings_log + asset_log + tier.1_log +PD_mean | 0 |0|0,
                  data = bank.data)
reg.hdfe3.2 <- felm(`Z-score`~ savings_log + asset_log + tier.1_log +PD_mean | year+Country |0|0,
                  data = bank.data)

```

```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}

stargazer(reg.hdfe1,reg.hdfe1.1,reg.hdfe1.2,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Savings and Capital Requirement Ratio",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and risk-weighted capital ratio. The dependent variable in each regression is the Core Tier 1 ratio. In column (2) and (3) I control for asset size, tier one capital size, and mean PD. In regressions (3) I also include bank and year fixed effects.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          dep.var.labels = c("CT1 ratio"), column.sep.width = "2pt",
          covariate.labels = c("log Savings", "log Total asset","log CT1 capital","Portfolio mean PD"),
          add.lines = list(c("Year FE", "No", "No","Yes"),
                           c("Bank FE","No", "No","Yes")))

```

```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}

stargazer(reg.hdfe2,reg.hdfe2.1,reg.hdfe2.2,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small", 
          title = "Capital Savings and Return on Equity",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and return on equity. The dependent variable in each regression is the return on equity using PL before taxes. In column (2) and (3) I control for asset size, tier one capital size, and mean PD. In regressions (3) I also include bank and year fixed effects.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          dep.var.labels = c( "ROE"), column.sep.width = "2pt",
          covariate.labels = c("log Savings", "log Total asset","log CT1 capital","Portfolio mean PD"),
          add.lines = list(c("Year FE", "No", "No","Yes"),
                           c("Bank FE","No", "No","Yes")))

```

```{r, fig.height = 10, fig.width = 7, results="asis",  out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}

stargazer(reg.hdfe3,reg.hdfe3.1,reg.hdfe3.2,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small", 
          title = "Capital Savings and Risk",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and overall risk. The dependent variable in each regression is the Z-score measure. In column (2) and (3) I control for asset size, tier one capital size, and mean PD. In regressions (3) I also include country and year fixed effects.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          dep.var.labels = c("Z-score"), column.sep.width = "2pt",
          covariate.labels = c("log Savings", "log Total asset","log CT1 capital","Portfolio mean PD"),
          add.lines = list(c("Year FE", "No", "No","Yes"),
                           c("Country FE","No", "No","Yes")))

```  
## Robustness

## Conclusion

## References

### 2019-08-03 Analysis/Descriptive/Descriptive_analysis.Rmd
Add code for extended abstract

---
title: "Descripte analysis"
author: "Lucas Avezum"
date: "June 17, 2019"
output:
  html_document: 
    toc: true
  pdf_document:default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)                                                           
library(data.table)
library(lubridate)
library(lfe)
library(stargazer)
library(ineq)
library(DescTools)
library(gglorenz)
library(ggplot2)
library(psych)
library(reporttools)
library(ggridges)
library(splitstackshape)
library(viridis)
library(Hmisc)
library(scales)

```
```{r load data, include=FALSE}
rm(list = ls())   
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankScope.Rda")
data.pillar3 <- data.pillar3 %>%
  mutate(Method2 = as.factor(ifelse(Method %in% c("IRB"), "IRB",
                                                     sapply(Method,str_sub,start= 2,end=4)))) 
bankscope <- bankscope %>%
  mutate(sample = as.factor(ifelse(bvdid %in% data.pillar3$bvdid,"Yes","No")),
         mean.ROA= ave(ROA.using.P.L.before.tax....,bvdid,FUN = function(x) mean(x, na.rm = TRUE)),
         sd.ROA= ave(ROA.using.P.L.before.tax....,bvdid,FUN = function(x) sd(x, na.rm = TRUE)),
         mean.CAR = ave(Equity...Total.assets....,bvdid,FUN = function(x) mean(x, na.rm = TRUE)))%>%
mutate('Z-score'=(mean.ROA+mean.CAR)/sd.ROA)
```

```{r expand data, include=FALSE}

expanded.pillar3 <- data.pillar3 %>%
  filter(Method2 %in% c("IRB")) %>%
  select(Bank,Country,Method,`Portfolio - level 1`,EAD,LGD,PD,year)%>%
  mutate(EL=round(EAD*LGD/1000),EAD=round(EAD/1000),PD=PD*100)

expanded.list <- list()
for(i in c("EAD","EL")) {
expanded.list[[i]] <- expanded.pillar3  %>%
  na.omit(cols=c(i))%>%
  expandRows(i) %>%
  mutate(freq_type = as.factor(i))
}
expanded.pillar3 <-  rbindlist(expanded.list, use.names=TRUE,fill = TRUE)

```

## Data

In this report I briefly describe the data collected so far and provide some descriptive analysis of two measures of capital savings. \newline

Table 1 reports summary statistics for a few variables from Bankscope. The level division means if the bank's Pillar-III reports information has been collected. The p-values from Kruskal test reject the hypothesis that the two groups are draw from the same distribution. The banks collected tend to have higher leverage (both risk-weighted or not), be safer (higher Z-score and lower NPL to gross loans ratio), and higher return on equity.

```{r sample vs bankscope, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
options(xtable.comment = FALSE)
vars <-bankscope%>%
  select(ROA = ROA.using.P.L.before.tax....,
         ROE =ROE.using.P.L.before.tax....,
         `NPL/gross loans` = Non.Perf..Loans...Gross.Loans....,
         `Total capital ratio` = Total.Capital.Ratio....,
         `Equity/total assets` = Equity...Total.assets....,
         `Z-score`) %>%
  mutate_all(Winsorize,na.rm = TRUE, probs = c(0.025, 0.975))

tableContinuous(vars = vars, group = bankscope$sample, longtable = FALSE,prec = 1, 
                cap = "Summary statistics", lab = "tab: descr stat", print.pval = "kruskal",
                stats = c("n","mean", "s", "min","median", "max"))

```
```{r  PD distribution by bank,fig.height = 7, fig.width = 7,eval=FALSE, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(Bank, PD,FUN = median, na.rm=TRUE))) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,
                       scale = 2 , from = 0, to = 2.5) +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    title = 'PD distribution by bank',
    subtitle = 'Data: Pillar-III reports'
  ) +
  theme_ridges(font_size = 13, grid = TRUE) + theme(axis.title.y = element_blank())
```

Figure 1 plots the PD distribution collected by bank.

```{r  PD distribution by bank and country,fig.height = 7, fig.width = 8, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(Bank, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
                       scale = 2 , from = 0, to = 2.5) +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    title = 'Figure 1: PD distribution by bank',
    subtitle = 'Ordered by median PD'
  ) +
  theme_ridges(font_size = 13, grid = TRUE) + theme(axis.title.y = element_blank())
```

## Basel formulas and capital savings

$K=\sum_{i}EAD_i\times LGD_i\times\bigg[\Phi\bigg(\frac{\Phi^{-1}(PD_i)+\sqrt{\rho_i}\Phi^{-1}(0.999)}{\sqrt{1-\rho_i}}\bigg)-PD_i\bigg]$ \newline

For wholesale portfolio:  \newline
$K_{\text{wholesale}}=K\times \frac{1+(M-2.5)b(PD_i))}{1-1.5b(PD_i)}$ \newline
with\newline
$\rho_i = 0.12\times\bigg(\frac{1-e^{-50\times PD_i}}{1-e^{-50}}+0.24\times\bigg(1-\frac{1-e^{-50\times PD_i}}{1-e^{-50}}\bigg)\bigg)$\newline

and $b(PD_i)=(0.11852-0.05478\times log(PD_i))^2$\newline

For retail portfolios the formula is equal to $K$ with different formulas for the correlation coefficient:  \newline

Real estate: $\rho_{i}=0.15$\newline

Qualifying revolving: $\rho_{i}=0.04$ \newline

Other retail: $0.03\times\bigg(\frac{1-e^{-35\times PD_i}}{1-e^{-35}}+0.16\times\bigg(1-\frac{1-e^{-35\times PD_i}}{1-e^{-35}}\bigg)\bigg)$\newline

Using these formulas I computed two measures of capital requirement: $K_{distribution}$ using the distribution of $PD_i$ (as it is required by the Basel accords), and $K_{average}$ using $\widetilde{PD}=\sum_iEAD_iLGD_iPD_i/\sum_iEAD_iLGD_i$. The percentage change $K_{distribution}/K_{average}-1$ is called capital savings and is a measure of the amount of capital saved due to the concavity in the formula for $K$.\newline

```{r mapping formulas, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

mapping_wholesale <- function(PD,LGD,M,alpha,portfolio){
  R  <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
  b  <- (0.11852-0.05478*log(PD))^2
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
  return(data)
}

mapping_retail <- function(PD,LGD,R,alpha,portfolio){
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,R,portfolio)
  return(data)
}

mapping_other_retail <- function(PD,LGD,alpha,portfolio){
  R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,R,portfolio)
  return(data)
}
```

```{r savings calculation, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

data.pillar3   <- data.pillar3 %>%
  filter(Method2 == "IRB",PD<1)%>%
  mutate(K_original=RW*EAD*0.08,
         PD_w=PD*EAD*LGD,
         EAD_w=LGD*EAD) %>%
  mutate(EAD_sum=ave(EAD_w,Bank,year,`Portfolio - level 1`,`Portfolio - level 2`,FUN = function(x) sum(x,na.rm = TRUE)),
         PD_w_sum=ave(PD_w,Bank,year,`Portfolio - level 1`,`Portfolio - level 2`,FUN = function(x) sum(x,na.rm = TRUE)))%>%
  mutate(PD_w_avg = PD_w_sum/EAD_sum)

for (i in 1:nrow(data.pillar3)){
  if (data.pillar3[i,9]!="Wholesale"){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_wholesale(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],2.5,0.999,"")["K"]

  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_wholesale(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],2.5,0.999,"")["K"]
}

for (i in 1:nrow(data.pillar3)){
  if (is.na(data.pillar3[i,10])){
    next
  }
  if (!data.pillar3[i,10]%in%c("Real estate")){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],0.15,0.999,"")["K"]

  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],0.15,0.999,"")["K"]

}

for (i in 1:nrow(data.pillar3)){
  if (is.na(data.pillar3[i,10])){
    next
  }
  if (!data.pillar3[i,10]%in%c("Qualifying revolving")){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],0.04,0.999,"")["K"]

  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],0.04,0.999,"")["K"]

}
for (i in 1:nrow(data.pillar3)){
  if (is.na(data.pillar3[i,10])){
    next
  }
  if (data.pillar3[i,10]!="Other retail"){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_other_retail(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],0.999,"")["K"]

  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_other_retail(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],0.999,"")["K"]

}

data.reduced <- data.pillar3 %>% group_by(Bank,Country,year) %>%
  summarise_if(is.numeric,funs(sum),na.rm=TRUE) %>%
  mutate(error = (K_calculated - K_original)/K_original,
         savings = (K_calculated-K_counterfactual)/K_counterfactual) %>%
  ungroup()

```

Figure 2 plots the calculated capital savings by country. Figure 3 plots the calculated capital savings by country. Figure 4 plots the evolution of calculated capital savings in time. \newline

```{r savings by bank, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
data.reduced %>%
 ggplot(
            mapping = aes(x = reorder(Bank, savings,FUN = median, na.rm=TRUE),
                          y = savings,
                          color = Country,
                          fill = Country)) +
geom_boxplot(alpha = 0.5) +
    labs(x=NULL) +
    coord_flip() +
  labs(
       y =  "Capital savings",
         title = "Figure 2: Capital savings by bank") +

  theme_bw()

```

```{r savings by country, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
data.reduced %>%
 ggplot(
            mapping = aes(x = reorder(Country, savings,FUN = median, na.rm=TRUE),
                          y = savings)) +
geom_boxplot() +
    labs(x=NULL) +
    coord_flip() +
  labs(
       y =  "Capital savings",
         title = "Figure 3: Capital savings by country") +
  theme_bw()

```

```{r savings by year, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
  dat <- data.frame(y = mean(x, na.rm = na.rm),
                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
  return(dat)
}

data.reduced %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time, y=savings)) +
 geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)+
labs(y =  "Capital savings",
     x = "Year",
     title = 'Figure 3: Average capital savings by year',
     subtitle = 'First and third quartile shown') +
  theme_bw()
```

## Basel formulas and Gini coefficient

As a second measure of the $PD$ distribution dispersion I compute Gini coefficients. Figure 4 shows the Lorenz curve of the $PD$ distributions  by country. Figure 5 shows the Lorenz curve of the $PD$ distributions by year. Figure 6 shows the Lorenz curve of the $PD$ distributions by Bank (names omitted for visibility). Finally Figure 7 plots the measure of capital savings and the Gini coefficients. \newline

```{r lorenz curve by country, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot( aes(x=PD, color = Country)) +
stat_lorenz() +
coord_fixed() +
geom_abline(linetype = "dashed") +
   labs(
       y =  "Lorenz curve",
         title = "Figure 4: Lorenz curve by country") +
theme_minimal()

```

```{r lorenz curve by year, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot( aes(x=PD, color = as.factor(year))) +
stat_lorenz() +
coord_fixed() +
geom_abline(linetype = "dashed") +
  labs(color = "",
       y =  "Lorenz curve",
         title = "Figure 5: Lorenz curve by year") +
theme_minimal()
```

```{r lorenz curve by bank, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot( aes(x=PD, color = Bank)) +
stat_lorenz() +
coord_fixed() +
geom_abline(linetype = "dashed") +
  theme_minimal()+
  labs(
       y =  "Lorenz curve",
         title = "Figure 6: Lorenz curve by bank") +
theme(legend.position = "none")
```

```{r gini savings total, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
data.gini <- data.pillar3 %>% 
  group_by(Country,Bank,year,`Portfolio - level 1`,`Portfolio - level 2`) %>%
  mutate(weight=EAD*LGD)%>%
  filter(weight>0, `Portfolio - level 1` %in% c("Wholesale","Retail"))%>%
  dplyr::summarize(gini = DescTools::Gini(PD,weight),
            K_calculated=sum(K_calculated),
            K_counterfactual=sum(K_counterfactual)) %>%
  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)

p <- ggplot(data = data.gini,
            mapping = aes(x = gini, y = savings, color=`Portfolio - level 1`))

 p + geom_point(alpha = 0.2)+
theme_minimal()+
   labs( title = "Figure 7: Capital savings and Gini coefficient",
         subtitle = 'Each point is a bank-year-portfolio observation',
         y =  "Inverted capital savings",
         x = "Gini coefficient",
         color = "")
```

```{r summary,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
bankscope %>% 
  subset(select = c("ROA.using.P.L.before.tax....","ROE.using.P.L.before.tax....","Non.Perf..Loans...Gross.Loans....","Total.Capital.Ratio....","Equity...Total.assets....","Employees","Total.assets","year","country","bvdid")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
          omit.summary.stat = c("p25","p75"))
```

### 2019-08-03 Analysis/Descriptive/ExtendedAbstract/ExtendedAbstract.Rmd
Add code for extended abstract

---
output: 
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
title: "Credit risk distribution and capital savings"
subtitle: ""
thanks: ""
date: # Today, unless specified. In smaller font if specified to allow for version numbering
author:
- name: Lucas Avezum
  affiliation: Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands

abstract: 
keywords: 
JEL classification: 
toc: false
geometry: margin= 1in
#fontfamily: mathpazo
fontsize: 12pt
#spacing: onehalf
bibliography: ../../../Auxiliar/mybibfile.bib
biblio-style: apsr
citecolor: black
urlcolor:
linkcolor: blue
anonymous: false # quickly sanitizes manuscript for blind review if 'true'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = F, 
                      error = F, 
                      tidy = T, 
                      tidy.opts=list(width.cutoff=50), 
                      collapse = T,
                      warning=FALSE,
                      error=FALSE,
                      message = FALSE, 
                      comment = "",
                      fig.cap = " ",
                      cache = TRUE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo
```

Extended abstract {-}
=================

This paper shows that higher-order moments of the distribution of banks' portfolio credit risk, in addition to its average, are important factors of risk-weights (RWs) variability across banks. More specifically, two distributions of probabilities of default (PD) with the same average but different variance result in different capital charges. Using a novel hand-collected dataset, I propose a measure that disantangles this PD distribution effect from the portfolio's average contribution to capital requirements. I estimate the impact of this measure on banks' performance.

Under the Internal ratings-based approach (IRB) banks supply their own risk estimates to regulatory formulas which determine the amount of capital banks have to hold for each asset. Importantly, these formulas are concave. This means, by Jensen’s inequality, that the sum of capital charges for two assets is lower than the charge for their average (in terms of PD). More generally, the distribution of credit risk matters to calculate capital requirements with, for instance, disperse portfolios yielding relatively lower capital charges.

To calculate the PD distribution contribution on capital requirements net from the average effect, one needs data on individual or grouped assets' regulatory risk parameters. However, data at such granularity is not available in the common balance sheet datasets. Although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information. Therefore, to overcome data limitations I collected information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports. The sample consists of 50 large banking groups, from 10 countries, that have adopted the IRB approach. For each bank, I collected consolidated information from the year the IRB approach was approved until 2018. The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate and sovereign for wholesale and real estate and qualifying revolving for retail). This level of portfolio breakdown is important because regulatory formulas vary across categories, in addition to allowing me to control for variation in capital requirements due to portfolio mix (sovereign portfolios require less capital than corporate). I augment my dataset with balance sheet information obtained from Orbis’ historical disks from 2001 to 2018 and include another 81 banks which calculate capital requirements only under the standardized approach (SA banks) until 2018, which act as a control group in our empirical analysis.

With this unique dataset, I start by analyzing the relative importance of the PD distribution effect to banks' overall RW, both in cross-section and time. Following the method proposed by @cannata2012inside I find that the PD distribution effect accounts for 25% of RW variation. Moreover, capital savings from the PD distribution effect are inversely correlated with the average PD contribution, which suggest that banks holding, high average PD portfolios reduce capital requirements by having, for instance, more disperse PD distributions. Conversely, the time decomposition analysis shows that banks, since introduction of the IRB approach, have decreased the amount of capital savings due to the PD distribution effect.

I then extend the analysis to find out whether banks adjust their portfolio risk distribution in response to capital requirement shocks. In the spirit of @gropp2018banks, I explore the 2011 EBA capital exercise as a quasi-natural experiment. This was a strong and unexpected test to the solvency of 63 European banks that resulted in a recommendation for these banks to build up their capital buffer to reach a 9.0% core tier 1 capital ratio by June 2012. Identification of the effect of higher capital requirements on banks’ portfolio risk distribution comes from correlating changes in the _capital savings_ measure around the exercise to the distance to achieve the 9.0% required ratio for banks that participated in the exercise. Once established that banks adjust their portfolio risk distribution to achieve regulatory capital requirements, a natural extension is to quantify the effect of these changes on capital ratio and evaluate if these changes impact other variables of interest, such as, return on equity, and risk-taking. To solve endogeneity concerns – for instance, worse capitalized banks have stronger incentives to adjust their portfolios – I use the Basel 2 adoption year across countries as an instrument for the PD distribution effect.

Taken together, the results show that strategic portfolio choice, in terms of credit risk, is an important way banks can reduce capital charges while maintaining their average portfolio risk and return on equity.

\newpage
\singlespacing

### JEL classification codes {-}
G21, G28, G11

### Keywords {-}
Internal ratings-based models, Risk-weighted assets dispersion, Regulatory arbitrage

# References

### 2019-08-03 Analysis/Draft/Draft.Rmd
Add decomposition and counterfactual ratio plots

@@
 thanks: Very preliminary draft
-abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
-  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
-  a concave mapping from probabilities of default to capital requiremets. As a consequence,
-  banks can reduce requirements or increase risk-taking by widening their portfolio
-  credit risk distribution.'
+#abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
+#  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
+#  a concave mapping from probabilities of default to capital requiremets. As a consequence,
+#  banks can reduce requirements or increase risk-taking by widening their portfolio
+#  credit risk distribution.'
 toc: no
@@
 load("../../Data/Datasets/BankData.Rda")
+load("../../Data/Datasets/CrossSectionDecomposition.Rda")
+load("../../Data/Datasets/timedecomposition.Rda")
+load("../../Data/Datasets/yeardecomposition.Rda")
+```
 
+```{r change data, eval= FALSE, include=FALSE}
 # Expand Pillar-III data frame to plot densities and lorenz curves
@@
 
-# Aggregate Pillar-III data by bank and merge to bankscope
-bank.data <- pillar3.data %>% 
-  group_by(name,bvdid,Country,year) %>%
+# Aggregate Pillar-III data by Portfolio
+portfolio.data <- pillar3.data %>% 
+  group_by(Country,Bank,year,Portfolio_1,Portfolio_2) %>%
   mutate(weight=EAD*LGD)%>%
-  filter(weight>0,
-         PD<1&PD>0,
-         Portfolio_1 %in% c("Wholesale","Retail"), 
-         Method2 %in% c("IRB"))%>%
+  filter(weight>0, Portfolio_1 %in% c("Wholesale","Retail"))%>%
   dplyr::summarize(gini             = DescTools::Gini(PD,weight),
-                   PD_sd            = sd(PD)*100,
-                   PD_mean          = mean(PD)*100,
                    K_calculated     = sum(K_calculated),
-                   K_original       = sum(K_original),
                    K_counterfactual = sum(K_counterfactual)) %>%
-  ungroup()%>% 
-  left_join(bank.data,by=c("bvdid","year"))%>%
+  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)
+
+bank.data <- bank.data %>%
 group_by(bvdid) %>% 
@@
 
-# Aggregate Pillar-III data by Portfolio
-portfolio.data <- pillar3.data %>% 
-  group_by(Country,Bank,year,Portfolio_1,Portfolio_2) %>%
-  mutate(weight=EAD*LGD)%>%
-  filter(weight>0, Portfolio_1 %in% c("Wholesale","Retail"))%>%
-  dplyr::summarize(gini             = DescTools::Gini(PD,weight),
-                   K_calculated     = sum(K_calculated),
-                   K_counterfactual = sum(K_counterfactual)) %>%
-  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)
-
 ```
 # Introduction
+
+Many studies have questioned the reliability of banks' calculations of risk-weighted assets (RWA) for prudential purpose.
+The significant divergences found across banks are taken as indicating excessive subjectivity in the current rules governning banks' risk measurement and capital requirement calculations.
+Importantly, this subjectivity is unintended since differences in RWA driven by it do not reflect differences in risk.
+This paper shows that the dispersion _within_ banks' portfolio credit risk, more specifically the probability of default (PD) parameter, explains substantially the variation in RWA _across_ banks.
+
+paragraph connecting to previous literature
+Basel II introduced a concave mapping from probabilities of default to capital requiremets.
+As a consequence, banks can reduce requirements or increase risk-taking by widening their portfolio credit risk distribution.
+
+To calculate equation X, or any other measure of capital savings arising from the concavity of the regulatory formula,  one needs data on individual, or group of assets' regulatory risk parameters.
+However, such granularity is not available in the common commercial balance sheet datasets, and, although very detailed, datasets from credit registers are restricted to one country and are not harmonized.
+Therefore, to analyze the drivers and implications of PD dispersion, I collected information on portfolio distribution from banks' individual Pillar-3 reports.\footnote{More specifically, I collected the information contained in tables that follow the CR5, CR6, CR10, CCR3, and CCR4 templates from the BCBS revised version of the Pillar 3 framework, or equivalent.} The sample consists of 50 large banking groups, from 10 countries, that have adopted the IRB approach.
+For each bank, I collected group consolidated information from the year the IRB approach was approved, which is early as 2007, until 2018.
+The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories.\footnote{The sub-categories for wholesale are, corporate, banks, sovereign, and specialized lending, for retail, real estate, qualifying revolving credit, and other retail, and for equity, portfolios under the PD/LGD approach, and under the simple approach.} This level of portfolio breakdown is important because the regulatory formula varies across categories.
+It also allows me to control for variation in capital requirements due to business mix (sovereign portfolios require less capital than corporate) that could  be otherwise captured by my capital savings measure.
+Information on the share of portfolio under the standardized approach is also gathered, which allow me to control for the roll-out effect, i.e., strategical reduction in capital requirements by moving specific portfolios from the standardized to the IRB approach.
+Finally
+
+paragraph on results
 
@@
 ```{r 1, fig.height = 4, fig.width = 5,results="hide",fig.cap= fig1_cap,fig.align = 'center'}
-bank.data %>% 
- ggplot(mapping = aes(x = reorder(name, savings_rate,FUN = median, na.rm=TRUE),
-                          y = savings_rate,
-                          color = Country,
+bank.data %>% filter(!is.na(name)) %>%
+ ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
+                          y = RW.s,
+                      color= Country, 
                       fill = Country)) +
@@
 ggplot(aes(x=time)) +
- geom_smooth(aes( y=savings_log, color = savings_log),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
+ geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
 labs(x="",y="",
@@
 
-plot2 <- bank.data %>% mutate(Tier.1.Ratio= Tier.1.Ratio/100) %>%
+plot2 <- bank.data %>% mutate(tier1ratio= tier1ratio/100) %>% filter(year>2001)%>%
   mutate(time = lubridate::ymd(year, truncated = 2L))%>%
 ggplot(aes(x=time)) +
-   geom_smooth(aes(y=Tier.1.Ratio, color = Tier.1.Ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
+   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
 labs(x="",y="",
@@
 
-plot3 <- bank.data %>% mutate(ROE.using.P.L.before.tax....= ROE.using.P.L.before.tax..../100) %>%
+plot3 <- bank.data %>% mutate(ROE = ROE/100) %>% filter(year>2001)%>%
   mutate(time = lubridate::ymd(year, truncated = 2L))%>%
 ggplot(aes(x=time)) +
-   geom_smooth(aes(y=ROE.using.P.L.before.tax...., color = ROE.using.P.L.before.tax....),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
+   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
 labs(x="",y="",
@@
 We restrict the analysis to banks for which we are able to collect information from Pillar-III reports.
+```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
+time.decomposition %>% 
+  select(time,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = relevel(condition,"dispersion.gain")) %>%
+  ggplot(mapping = aes(x = time,
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL)
+
+```
+
+banks. We restrict the analysis to banks for which we are able to collect information from Pillar-III reports. 
+```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
+year.decomposition %>% 
+  select(year,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = relevel(condition,"dispersion.gain")) %>%
+  ggplot(mapping = aes(x = year,
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL)
+
+```
 
@@
 The selection criteria is asset size.
+```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
+cross.section.decomposition %>% 
+  mutate(order.s = abs(dispersion.gain))%>%
+  select(name, Country, RW,order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = relevel(condition,"dispersion.gain")) %>%
+  ggplot(mapping = aes(x = reorder(name, order.s),
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL) +
+    coord_flip()
 
-```{r}
+```
+
+```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
+cross.section.decomposition %>% 
+  mutate(order.s = abs(dispersion.gain))%>%
+  select(name, Country, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(value = abs(value),
+         condition = relevel(condition,"dispersion.gain")) %>%
+  ggplot(mapping = aes(x = reorder(name, order.s),
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity',position = "fill")+
+  labs(x=NULL) +
+    coord_flip()
+
+```
+
+```{r,  fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
 fig1_cap = "The figure plots the linear relationship between the capital savings measure and,in descending order, the tier 1 capital ratio, the return on equity using PL before taxes, and the Z-score.
@@
 
-```{r relation, fig.height = 3, fig.width = 6, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-plot1 <- bank.data %>% group_by(name)%>%
+```{r relation, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
+ bank.data %>% 
+  select(name,RW.s, tier1ratio, tier.hat) %>%
+  group_by(name)%>%
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
- ggplot(mapping = aes(x = savings_log,
-                      y = Tier.1.Ratio)) +
+  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)%>%
+  filter(RW.s<0.4)%>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
 geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
   labs(x="Capital savings",y="CT1 capital ratio") +
@@
   theme_bw()
+```
 
-plot2 <- bank.data %>% group_by(name)%>%
+```{r relation, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
+bank.data %>% 
+  select(name,RW.s, totalcapitalratio, CAR.hat) %>%
+  group_by(name)%>%
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
- ggplot(mapping = aes(x = savings_log,
-                      y = ROE.using.P.L.before.tax....)) +
+  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
+  filter(RW.s<0.4)%>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
 geom_point( alpha = 0.5,size = 1) +
-  labs(x="Capital savings",y="ROE") +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="Capital savings",y="CT1 capital ratio") +
   geom_smooth(method = "lm",color="black") +
   theme_bw()
-
-plot3 <- bank.data %>% group_by(name)%>%
-  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
- ggplot(mapping = aes(x = savings_log,
-                      y = `Z-score`)) +
-geom_point( alpha = 0.5,size = 1) +
-  labs(x="Capital savings",y="Z-score") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()
-
-grid.arrange(plot1, plot2,plot3, ncol=3)
 

### 2019-08-03 Data/Clean.R
Simplify clean code (use RW from functions instead of K)

@@ -58,10 +58,10 @@ for (i in 1:nrow(IRB.data)){
   if (!IRB.data[i,"Portfolio_1"] %in% c("Wholesale","Equity")){
     next
   }
-  IRB.data[i,"K"]     <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"PD"],
-                                                             IRB.data[i,"LGD"])["K"]
-  IRB.data[i,"K.hat"] <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"w.avg.PD"],
-                                                             IRB.data[i,"LGD"])["K"]
+  IRB.data[i,"RWA"]     <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"PD"],
+                                                             IRB.data[i,"LGD"])["RW"]
+  IRB.data[i,"RWA.hat"] <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"w.avg.PD"],
+                                                             IRB.data[i,"LGD"])["RW"]
 }
 
 for (i in 1:nrow(IRB.data)){
@@ -71,12 +71,12 @@ for (i in 1:nrow(IRB.data)){
   if (!IRB.data[i,"Portfolio_2"] %in% c("Real estate")){
     next
   }
-  IRB.data[i,"K"]     <- IRB.data[i,"EAD"]*mapping_retail(IRB.data[i,"PD"],
+  IRB.data[i,"RWA"]     <- IRB.data[i,"EAD"]*mapping_retail(IRB.data[i,"PD"],
                                                           IRB.data[i,"LGD"],
-                                                                  0.15)["K"]
-  IRB.data[i,"K.hat"] <- IRB.data[i,"EAD"]*mapping_retail(IRB.data[i,"w.avg.PD"],
+                                                                  0.15)["RW"]
+  IRB.data[i,"RWA.hat"] <- IRB.data[i,"EAD"]*mapping_retail(IRB.data[i,"w.avg.PD"],
                                                           IRB.data[i,"LGD"],
-                                                                  0.15)["K"]
+                                                                  0.15)["RW"]
 }
 
 for (i in 1:nrow(IRB.data)){
@@ -86,12 +86,12 @@ for (i in 1:nrow(IRB.data)){
[…]

### 2019-08-14 Data/Import.R
Add code to convert variables to USD in Import.R

@@ -120,10 +120,15 @@ bankscope <- bankscope %>%
          -c("companyname", "countryisocode","lastavail", "id")) %>% 
   # Turn numeric variables to numeric format
   mutate_at(c(grep("usd", names(bankscope))[1]:ncol(bankscope)),as.numeric) %>%
+  # Convert to USD
+  rowwise() %>%
+  mutate_at(c(grep("lcu$", names(bankscope))[1]:ncol(bankscope)),
+            funs(prod(.,exchangeratefromoriginalcurrencyusd, na.rm = TRUE))) %>%
   # Add Basel 2 country level introduction year
-  left_join(basel.indicator, by = c("Country"))
+  ungroup() %>%
+  left_join(basel.indicator, by = c("Country")) %>%
   # Add EBA capital exercise information
-bankscope <- bankscope %>% left_join(EBA.indicator, by = c("bvdid"))
+  left_join(EBA.indicator, by = c("bvdid"))
 
 # Save data frame
 save(bankscope,file=paste0("Data/Temp/BankScope.Rda"))

### 2019-08-14 Auxiliar/Function.R
Add code for all dependet variables, code for dif-in-dif data and adjust regulatory formulas and code to match reported information

@@ -12,14 +12,15 @@ mapping_wholesale <- function(PD, LGD = 0.45, M = 2.5, alpha = 0.999){
   LGD <- ifelse(is.na(LGD), 0.45, LGD)
   R   <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
   b   <- (0.11852-0.05478*log(PD))^2
-  K   <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
+  f   <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
+  K   <- f*((1-1.5*b)^-1*(1+(M-2.5)*b))
   RW  <- K*12.5
-  data <- data.frame(K,RW,PD,LGD,M,R)
+  data <- data.frame(K,RW,PD,LGD,M,R,b,f)
   return(data)
 }
 
 mapping_retail <- function(PD, LGD, R, alpha = 0.999){
-  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
+  K  <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
   RW <- K*12.5
   data <- data.frame(K,RW,PD,LGD,R)
   return(data)
@@ -27,7 +28,7 @@ mapping_retail <- function(PD, LGD, R, alpha = 0.999){
 
 mapping_other_retail <- function(PD, LGD, alpha = 0.999){
   R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
-  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
+  K  <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
   RW <- K*12.5
   data <- data.frame(K,RW,PD,LGD,R)
   return(data)

### 2019-08-14 Data/Clean.R

@@ -11,14 +11,16 @@
 rm(list = ls())  
 library(tidyverse)                ## Data manipulation, pipe operator
 library(zoo)                      ## Apply function to rolling margins of data
+library(DescTools)                ## Command to winsorize variables
 source("Auxiliar/Function.R")     ## Capital requirements functions
+
+# Open dirt data frame
+load("Data/Temp/BankScope.Rda")
+load("Data/Temp/Pillar3Data.Rda")
 
 ##============================================================================##
 ## Pillar-III reports                                                         ##
 ##============================================================================##
-
-# Open dirt data frame
-load("Data/Temp/Pillar3Data.Rda")
 
 #------------------------------------------------------------------------------#
 # Basic cleaning and data creation of raw datasets                             #
@@ -37,7 +39,15 @@ pillar3.data <- as.data.frame(pillar3.data) %>%
          Portfolio_1 = as.factor(Portfolio_1),
          Portfolio_2 = as.factor(Portfolio_2))
 
-IRB.data <- pillar3.data %>%
+# Conver currency values to USD
+pillar3.data <- bankscope %>%
+  select(bvdid,year,exchangeratefromoriginalcurrencyusd) %>%
+  right_join(pillar3.data,by=c("bvdid","year")) %>%
+  mutate(EAD = EAD*exchangeratefromoriginalcurrencyusd,
+         RWA = RWA*exchangeratefromoriginalcurrencyusd)
+  
+
+IRB.data <- as.data.frame(pillar3.data) %>%
   filter(Method2 %in% c("IRB"),
          PD  < 1,  PD  > 0, EAD > 0) %>%
   mutate(LGD         = ifelse(is.na(LGD), 0.45, LGD),
@@ -47,7 +57,7 @@ IRB.data <- pillar3.data %>%
          w.sum.PD    = ave(w.PD, bvdid, year, Portfolio_1, Portfolio_2, FUN = function(x) sum(x, na.rm = TRUE)),
          w.avg.PD    = w.sum.PD/sum.EAD) 
   
-SA.data <- pillar3.data %>%
+SA.data <- as.data.frame(pillar3.data) %>%
   filter(!Method2 %in% c("IRB"),
          EAD > 0)
 
@@ -59,10 +69,17 @@ for (i in 1:nrow(IRB.data)){
[…]

### 2019-08-14 Analysis/Descriptive/ExtendedAbstract/ExtendedAbstract.Rmd
Update extended abstract before submitting to Copfir conference

@@
 - name: Lucas Avezum
-  affiliation: Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands
+  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands
 
@@
 Using a novel hand-collected dataset, I propose a measure that disantangles this PD distribution effect from the portfolio's average contribution to capital requirements.
-I estimate the impact of this measure on banks' performance.
+I then estimate the impact of this measure on banks' performance.
 
-Under the Internal ratings-based approach (IRB) banks supply their own risk estimates to regulatory formulas which determine the amount of capital banks have to hold for each asset.
+Several studies document significant RW variability across otherwise similar banks.
+Many take this evidence as indicative of either excessive subjectivity in the current capital requirement rules or risk measurement manipulation by banks to decrease capital charges [see,  @mariathasan2014manipulation; @ferribank].
+Others are more cautious and emphasize that part of RWA variability is desirable, since it reflects differences in risk among banks [see, @cannata2012inside;@BCBS2016regulatory; @EBA2019result].
+I contribute to the debate by showing that the PD distribution effect is an important factor of the RW variation across banks.
+This implies that previous studies in the literature are potentially overestimating subjectivity and manipulation of RWs.
+Nonetheless, the PD distribution effect entails unintended effects to banks’ solvency as they can explore this feature to reduce capital requirements without altering their portfolio average risk.
+
+Under the internal ratings-based approach (IRB) banks supply their own risk estimates to regulatory formulas which determine the amount of capital banks have to hold for each asset.
 Importantly, these formulas are concave.
@@
 This was a strong and unexpected test to the solvency of 63 European banks that resulted in a recommendation for these banks to build up their capital buffer to reach a 9.0% core tier 1 capital ratio by June 2012.
-Identification of the effect of higher capital requirements on banks’ portfolio risk distribution comes from correlating changes in the _capital savings_ measure around the exercise to the distance to achieve the 9.0% required ratio for banks that participated in the exercise.
+Identification of the effect of higher capital requirements on banks’ portfolio risk distribution comes from correlating changes in the PD distribution effect measure around the exercise to the distance to achieve the 9.0% required ratio for banks that participated in the exercise.
 Once established that banks adjust their portfolio risk distribution to achieve regulatory capital requirements, a natural extension is to quantify the effect of these changes on capital ratio and evaluate if these changes impact other variables of interest, such as, return on equity, and risk-taking.
-To solve endogeneity concerns – for instance, worse capitalized banks have stronger incentives to adjust their portfolios – I use the Basel 2 adoption year across countries as an instrument for the PD distribution effect.
+To alleviate endogeneity concerns – for instance, worse capitalized banks have stronger incentives to adjust their portfolios – I use the Basel 2 adoption year across countries as an instrument for the PD distribution effect.
 
 Taken together, the results show that strategic portfolio choice, in terms of credit risk, is an important way banks can reduce capital charges while maintaining their average portfolio risk and return on equity.
+Differently from manipulation and subjectivity of RW under IRB, the PD distribution effect hereby identified does not entail any regulatory breach.
+However, the assumption that riskier assets have lower systematic risk, which is one of the reasons
+the regulatory formulas are concave, may be too simplistic to capture the true systematic component of each asset.
+If that is the case, this paper shows that the current capital requirements framework does not correctly equate capital charges to credit risk, consequently, undermining financial stability.
 
-\newpage
 \singlespacing
@@
 
+\newpage
+
 # References

### 2019-08-14 Analysis/Draft/Draft.Rmd
Draft version with first try of IV and dif-in-dif analysis, plus decomposition plots

@@
 library(DescTools)           ## Command to winsorize variables
+library(robustHD)
 
-p_w <- c(0.0, 1)
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
 ```
 
-```{r load data, include=FALSE}
+```{r load data, include=TRUE}
 load("../../Data/Datasets/Pillar3Data.Rda")
@@
 load("../../Data/Datasets/yeardecomposition.Rda")
+load("../../Data/Datasets/DifInDif.Rda")
 ```
@@
 
+test <- cross.section.decomposition %>% mutate(mean.total.gain = mean(total.gain,na.rm = TRUE)) %>%
+  mutate_at(vars(SA.gain:total.gain),  funs((. - mean.total.gain)^2))%>%
+summarize_at(vars(SA.gain:total.gain),  sum, na.rm =  TRUE)%>%
+  mutate_all( funs((. / total.gain)))
 ```
-# Introduction
 
-Many studies have questioned the reliability of banks' calculations of risk-weighted assets (RWA) for prudential purpose.
-The significant divergences found across banks are taken as indicating excessive subjectivity in the current rules governning banks' risk measurement and capital requirement calculations.
-Importantly, this subjectivity is unintended since differences in RWA driven by it do not reflect differences in risk.
-This paper shows that the dispersion _within_ banks' portfolio credit risk, more specifically the probability of default (PD) parameter, explains substantially the variation in RWA _across_ banks.
+```{r}
 
-paragraph connecting to previous literature
-Basel II introduced a concave mapping from probabilities of default to capital requiremets.
-As a consequence, banks can reduce requirements or increase risk-taking by widening their portfolio credit risk distribution.
+bank.data <- bank.data %>% ungroup() %>%
+  mutate(RW.hat=RW.hat*100,
+         RW.s=RW.s*100)
+```
 
-To calculate equation X, or any other measure of capital savings arising from the concavity of the regulatory formula,  one needs data on individual, or group of assets' regulatory risk parameters.
-However, such granularity is not available in the common commercial balance sheet datasets, and, although very detailed, datasets from credit registers are restricted to one country and are not harmonized.
-Therefore, to analyze the drivers and implications of PD dispersion, I collected information on portfolio distribution from banks' individual Pillar-3 reports.\footnote{More specifically, I collected the information contained in tables that follow the CR5, CR6, CR10, CCR3, and CCR4 templates from the BCBS revised version of the Pillar 3 framework, or equivalent.} The sample consists of 50 large banking groups, from 10 countries, that have adopted the IRB approach.
-For each bank, I collected group consolidated information from the year the IRB approach was approved, which is early as 2007, until 2018.
-The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories.\footnote{The sub-categories for wholesale are, corporate, banks, sovereign, and specialized lending, for retail, real estate, qualifying revolving credit, and other retail, and for equity, portfolios under the PD/LGD approach, and under the simple approach.} This level of portfolio breakdown is important because the regulatory formula varies across categories.
-It also allows me to control for variation in capital requirements due to business mix (sovereign portfolios require less capital than corporate) that could  be otherwise captured by my capital savings measure.
-Information on the share of portfolio under the standardized approach is also gathered, which allow me to control for the roll-out effect, i.e., strategical reduction in capital requirements by moving specific portfolios from the standardized to the IRB approach.
-Finally
+# List
+* Use the share under IRB to weight capital savings measure (share vs intensity). Include share in regressions plus interaction with capital savings
 
-paragraph on results
+* Explore alternative capital savings measures, such as, savings in terms of capital (mn),in terms of capital ratio, in terms of capital ratio savings (counterfactual vs actual)
 
-# Hypotheses and Data
-In this section I describe the hypotheses of the paper by explaining the Basel II regulatory formulae under the IRB approach to calculate RWA for credit risk.
+* Plot marginal increase of savings in time.
 
-## Capital savings measure
+* Decompose capital savings: share effect, size effect, etc...
+
+* Remove step effect from regressions: average across all IRB banks (add IRB dummy) or IRB specific dummy (average out the step for each bank).
+Idea is to focus on difference across IRB banks.
+it might not make sense to include SA banks anymore.
+
+* Plot capital savings measures to other risk variables
+
+* Plot cross section for different capital saving measures (replicate first plot)
+
+* check and better clean data:
+
+# Capital savings measure
 
@@
 \begin{equation}
-\text{Capital savings}_{i}= f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i),
+\text{Capital savings}_{i}= \frac{f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i)}{\sum_{i}EAD_i},
 \end{equation}
@@
 
-## Data
-
-Figure 1 shows that the log of _Capital savings_ varies considerably across banks.
-While most canadian banks seem to have not explore this possilibity to save capital, the european banks in our sample appear to have more than halved their capital requirement due to the distribution mechanism.
-Importantly, this comparison does not means that european banks are riskier than their canadian counterparties but only that their credit risk distribution is more disperse.
-The figure also shows that there capital savings have changed in time.
-However, the figure is not appropriated to see how this variable evolved in time.
-
 ```{r}
-fig1_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
+savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
 ```
 
-```{r 1, fig.height = 4, fig.width = 5,results="hide",fig.cap= fig1_cap,fig.align = 'center'}
-bank.data %>% filter(!is.na(name)) %>%
+```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
+bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
  ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
@@
 
-Figure 2 shows that both _Capital savings_ and total capital ratio have increased slighly on average in time.
-After a spike in 2010, the average return on equity appears to have remained constant in time.
-
-```{r}
-fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE.
-The solid line is the average across banks for each year.
-The first and third quartiles are shown in the gray area.
-"
-```
-
-```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth"}
-mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
-  dat <- data.frame(y = mean(x, na.rm = TRUE),
-                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
-                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
-  return(dat)
-}
-
-plot1 <- bank.data %>% 
-  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
-ggplot(aes(x=time)) +
- geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
-labs(x="",y="",
-     title = 'Capital savings') +
-  theme_bw()
-
-plot2 <- bank.data %>% mutate(tier1ratio= tier1ratio/100) %>% filter(year>2001)%>%
-  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
-ggplot(aes(x=time)) +
-   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
-labs(x="",y="",
-     title = 'Total capital ratio') +
-  theme_bw()
-
-plot3 <- bank.data %>% mutate(ROE = ROE/100) %>% filter(year>2001)%>%
-  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
-ggplot(aes(x=time)) +
-   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
-labs(x="",y="",
-     title = 'ROE') +
-  theme_bw()
-
-grid.arrange(plot1, plot2, plot3, ncol=3)
-```
-
-Figure 3 shows the linear relationship between our variable of interest, _Capital savings_ and three potential outcome variables: core tier 1 capital ratio, return on equity and, a measure of riskness, Z-score.
-
 ```{r}
@@
 
-```{r, fig.height = 10, fig.width = 7, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
+```{r, fig.height = 10, fig.width = 7, eval=FALSE, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
 pillar3.expanded %>% 
@@
 
-Finally, we supplement the pillar-3 report data with annual information on banks' balance sheets obtained from the Bureau van Dijk's BankFocus database.
-Table 1 show summary statistics of the final dataset.
-The sample includes 25 banks.
-We restrict the analysis to banks for which we are able to collect information from Pillar-III reports.
-```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-time.decomposition %>% 
-  select(time,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+# Decomposition
+
+\begin{equation}
+\begin{aligned}
+\Delta RW & = q_{B}^{SA} \Delta RW^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \hat{RW}^{j} \\ 
+& + (1-q_{B}^{SA})\sum_{j=1}^{P} \hat{RW}^{j} \Delta q^{j} -(1-q_{B}^{SA})\Delta RW^{S} + (RW^{SA}-RW^{IRB})\Delta q^{SA},
+\end{aligned}
+\end{equation}
+
+\newpage
+## Time decomposition
+
+```{r}
+time.dec_cap = "Time decompostion.
+The figure shows the contribution of each component compared to the time of IRB introduction (t=0).
+The figure and calculations include only banks with observations for at least 5 consecutive years.
+The black line is the net change.
+Year decompostion.
+The figure shows the contribution of each component compared to 2008.
+The figure and calculations include all banks in the sample.
+The black line is the net change."
+```
+
+```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
+time.lable <- c("t=0", "t=1", "t=2", "t=3", "t=4")
+plot1 <-time.decomposition %>% 
+  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = relevel(condition,"dispersion.gain"),
+         RW = RW - t0.RW,
+         time = as.factor(time)) %>%
+  ggplot(mapping = aes(x = time,
+                       y    = value, 
+                       fill = condition)) +
+geom_bar(stat = 'identity')+
+geom_line(aes(group=1, y = RW))  +
+  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
+  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
+  scale_x_discrete( name=NULL,breaks = c("0","1","2","3","4"),labels = c("t=0","t=1","t=2","t=3","t=4")) +
+ theme(legend.position = c(0.3, 0.2), 
+        legend.title = element_blank(), 
+        legend.background=element_blank())
+
+year.decomposition <- year.decomposition %>%
+  mutate(RW_2 = RW - t0.RW,
+         year = as.factor(year)) 
+plot2 <- year.decomposition %>% 
+  select(year,RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
   gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
   mutate(condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = time,
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL)
+  ggplot(mapping = aes(x = year)) +
+geom_bar(aes(y = value, 
+          fill = condition),
+          stat = 'identity')+
+geom_line(data = year.decomposition, 
+          aes(group=1,y = RW_2))  +
+  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
+  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
+  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
+theme(legend.position='none')
+grid.arrange(plot1, plot2, ncol=2)
 
@@
 
-banks. We restrict the analysis to banks for which we are able to collect information from Pillar-III reports. 
-```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-year.decomposition %>% 
-  select(year,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = year,
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL)
+\newpage
 
+## Cross-section decomposition
+
+```{r}
+cross.dec_cap = "Cross-section decompostion.
+The figure shows the contribution of each component compared to the average bank.
+The figure and calculations include all banks in the sample.
+The black line is the total RW.
+Relative cross-section decompostion.
+The figure shows the relative contribution of each component compared to the average bank.
+The figure and calculations include all banks in the sample.
+"
 ```
 
-The current geographical composition of the sample is: 1 norwegian, 1 italian, 1 spanish, 2 british, 4 australian, 6 canadian, and 10 american banks.
-The target geographical composition is: 1 norwegian, 1 austrian, 2 finish, 3 belgium, 3 danish, 4 italian, 4 spanish banks, 4 dutch, 4 australian, 4 swedish, 5 french, 5 british, 6 canadian, 7 german, and 10 american banks.
-This sums up to 63 IRB banks.
-The selection criteria is asset size.
-```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-cross.section.decomposition %>% 
+```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
+plot1 <- cross.section.decomposition %>% 
   mutate(order.s = abs(dispersion.gain))%>%
-  select(name, Country, RW,order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
+  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
   gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
@@
 geom_bar(stat = 'identity')+
-  labs(x=NULL) +
-    coord_flip()
+  geom_line(aes(y = total,group=1))  +
+  labs(x=NULL, y= NULL,subtitle = "Plot a") +
+    coord_flip() +
+  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
+ theme(legend.position='none')
 
-```
-
-```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-cross.section.decomposition %>% 
+plot2 <- cross.section.decomposition %>% 
   mutate(order.s = abs(dispersion.gain))%>%
@@
 geom_bar(stat = 'identity',position = "fill")+
-  labs(x=NULL) +
-    coord_flip()
+  labs(x=NULL, y= NULL,subtitle = "Plot b") +
+    coord_flip() +
+  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
+   theme(legend.title = element_blank(), 
+        legend.background=element_blank(),
+       legend.text=element_text(size=7),
+       axis.text.y = element_blank())
+
+grid.arrange(plot1, plot2, ncol=2)
 
@@
 
-```{r,  fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-fig1_cap = "The figure plots the linear relationship between the capital savings measure and,in descending order, the tier 1 capital ratio, the return on equity using PL before taxes, and the Z-score.
-In each graph, a point is the average across time for each bank."
+# Counterfactual analysis
+
+```{r}
+tier1_cap = "The figure shows scatter plots and the linear relationship between the capital savings measure and capital ratios.
+The y-axis variable on plots a and b is the tier 1 capital ratio, and on plots c and d is the total capital ratio.
+Plots a and c plots all data points in the sample.
+Plots b and d plots the average across time for bank using the IRB approach."
 ```
 
-```{r relation, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
- bank.data %>% 
-  select(name,RW.s, tier1ratio, tier.hat) %>%
-  group_by(name)%>%
- summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)%>%
-  filter(RW.s<0.4)%>%
+```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
+ plot1 <- bank.data %>%  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% 
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
+  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>% 
  ggplot(mapping = aes(x = RW.s,
@@
   scale_y_continuous( limits=c(5, 20))+
-  labs(x="Capital savings",y="CT1 capital ratio") +
+  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
   geom_smooth(method = "lm",color="black") +
-  theme_bw()
-```
+  theme_bw()+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
+  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
+theme(legend.position='none')
 
-```{r relation, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
-bank.data %>% 
-  select(name,RW.s, totalcapitalratio, CAR.hat) %>%
+plot2 <- bank.data %>%  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
+  filter(!is.na(RW.s), IRB == 1) %>%
   group_by(name)%>%
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
-  filter(RW.s<0.4)%>%
+  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>%
  ggplot(mapping = aes(x = RW.s,
@@
   scale_y_continuous( limits=c(5, 20))+
-  labs(x="Capital savings",y="CT1 capital ratio") +
+  labs(x="",y="",subtitle = "Plot b") +
   geom_smooth(method = "lm",color="black") +
-  theme_bw()
+  theme_bw()+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
+  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
+  theme(legend.position = c(0.8, 0.85), 
+        legend.title = element_blank(), 
+        legend.background=element_blank())
 
+plot3 <- bank.data %>%  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
+  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
+geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
+  geom_smooth(method = "lm",color="black") +
+  theme_bw()+
+  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  theme(legend.position='none')
+
+plot4 <- bank.data %>% filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
+  filter(!is.na(RW.s), IRB == 1) %>%
+  group_by(name)%>%
+ summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
+  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
+geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="Capital savings",y="",  subtitle = "Plot d") +
+  geom_smooth(method = "lm",color="black") +
+  theme_bw()+
+  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  theme(legend.position='none')
+
+grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
 ```
-Banks that only implement the standardized approach (SA) for credit risk most of the sample period will also be included in the dataset.
-The current geographical composition of SA banks is: 1 portuguese, 1 dutch, 1 belgium, 1 finish, 1 hungarian, 2 maltese, 2 british, 2 greek, 2 polish, 3 french, 3 slovenian, 4 cypriot, 5 german, 5 italian, 8 spanish, 37 american.
-This sums up to 78 SA banks.
-The selection criteria is the same, asset size.
 
-```{r summary, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-bank.data %>% 
-  subset(select = c("savings_log","PD_mean","gini","Tier.1.Ratio","ROE.using.P.L.before.tax....","Z-score","asset_log","tier.1_log")) %>%
+```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
+  subset(select = c("RW.s","tier1ratio","tier.hat","totalcapitalratio","CAR.hat","ROE","Zscore")) %>%
   as.data.frame() %>% 
@@
           omit.summary.stat = c("p25","p75"),
-          covariate.labels = c("log Savings","Portfolio mean PD","Gini coefficient","CT1 ratio","ROE","Z-score", "log Total asset","log CT1 capital"),
+          covariate.labels = c("Capital savings","Tier 1 ratio","Counterfactual tier 1 ratio","Total ratio","Counterfactual total ratio","ROE","Z-score"),
           title = "Summary statistics",
@@
 
-# Empirical strategy
-In this section I briefly describe the empirical strategy of the paper.
+```{r}
+fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE.
+The solid line is the average across banks for each year.
+The first and third quartiles are shown in the gray area."
+```
 
-## EBA capital exercise and the effect on capital savings
+```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth", eval=FALSE, include=FALSE}
+mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
+  dat <- data.frame(y = mean(x, na.rm = TRUE),
+                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
+                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
+  return(dat)
+}
 
-To find out if banks adjust their portfolio risk distribution in response to capital requirement shock, I intend to explore the 2011 EBA capital exercise.
-This was a strong and unexpected test to the solvency of 63 european banks that resulted in a recommendation for these banks to build up their capital buffert to reach a 9.0% core tier 1 capital ratio by June 2012.
-Identification of the effect of higher capital requirements on banks'portfolio risk distribution comes from comparing the change in the _Capital savings_ measure (or other measure of the distribution such as the Gini coefficient) around the exercise between banks that participated in the exercise and banks that did not.
-In particular, I (will) estimate the following specification::
+plot1 <- bank.data %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu) %>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  mutate(time = lubridate::ymd(year, truncated = 2L))%>% 
+ggplot(aes(x=time)) +
+ geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
+labs(x="",y="",
+     title = 'Capital savings') +
+  theme_bw()
 
-\begin{equation}
-\Delta\text{Capital savings}_{i}= \beta\text{CEB}_{i} + \varepsilon_{i},
-\end{equation}
+plot2 <- bank.data  %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu) %>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
+ggplot(aes(x=time)) +
+   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
+labs(x="",y="",
+     title = 'ROE') +
+  theme_bw()
 
-where $\text{CEB}_{i}$ is an indicator function that equals one if bank $i$ was included in the exercise and $\beta$ is the treatment effect of higher requirements on portfolio risk distribution.
+plot3 <- bank.data  %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
+ggplot(aes(x=time)) +
+   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
+labs(x="",y="",
+     title = 'Tier 1 ratio') +
+  theme_bw()
 
-##  Basel II and the effect of capital savings
+grid.arrange(plot1, plot2, plot3, ncol=3)
+```
 
-Once we know that banks adjust their portolio risk distribution to achieve regulatory capital requirements, an natural extension is to quantify the effect of these changes on capital ratio and evaluate if these changes impact other variables of interest, for instance, return on equity and risk-taking.
-Endogeneity is an important issue to consider.
-For instance, worse capitalized banks have a stronger incentive to adjust their portfolios compared to their better capitalized peers resulting in downward biased estimates.
-Conversely, better-capitalized banks may have more leeway to achieve a portfolio distribution that maximizes return by increasing risk-taking.
-In this case the estimated effects of capital savings on return on equity and risk-taking will have a positive bias.
+```{r, include=FALSE}
 
-To solve the endogeneity concerns and estimate the causal effect of portfolio risk distribution, I (will) explore the time differences in introduction of Basel II across countries.
-Only after Basel II that the IRB models were allowed to be used and, consequently, the distribution of PD matter for the calculation of RWA.
-In particular, I (will) estimate the following specification:
+DiD.data <- DiD.data %>% ungroup() %>% 
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
 
-\begin{equation}
-Y_{it}= \beta\widehat{\text{Capital savings}}_{it}+\varepsilon_{i},
-\end{equation}
+DiD.intensity <- felm(outcome ~ Shortfall.RW | Country |0|0,
+                  data = DiD.data)
 
-where $Y_{it} \in \{\text{CT1 ratio},\text{ROE},\text{Z-score}\}$ and $\widehat{\text{Capital savings}}_{it}$ is the fitted values of the first stage equation:
+DiD.intensity.2 <- felm(outcome ~ Shortfall.RW.2 | Country |0|0,
+                  data = DiD.data)
 
-\begin{equation}
-\text{Capital savings}_{it}= \alpha\text{Basel II}_{it}+\epsilon_{i},
-\end{equation}
-
-where $\text{Basel II}_{it}$ is an indicator function that equals one if bank $i$ in year $t$ is hosted in country that has adopted Basel II and has its own IRB model approved.
-
-# (Preliminary) Results
-
-```{r, echo=FALSE}
-
-reg.hdfe1 <- felm(Tier.1.Ratio  ~ savings_log  | 0  |0|0,
-                  data = bank.data)
-
-reg.hdfe1.1 <- felm(Tier.1.Ratio  ~ savings_log + asset_log + tier.1_log +PD_mean | 0  |0|0,
-                  data = bank.data)
-
-reg.hdfe1.2 <- felm(Tier.1.Ratio  ~ savings_log + asset_log + tier.1_log +PD_mean | year +name  |0|0,
-                  data = bank.data)
-
-reg.hdfe2 <- felm(ROE.using.P.L.before.tax....~ savings_log  | 0 |0|0,
-                  data = bank.data)
-reg.hdfe2.1 <- felm(ROE.using.P.L.before.tax....~ savings_log + asset_log + tier.1_log +PD_mean | 0 |0|0,
-                  data = bank.data)
-reg.hdfe2.2 <- felm(ROE.using.P.L.before.tax....~ savings_log + asset_log + tier.1_log +PD_mean | year +name |0|0,
-                  data = bank.data)
-
-reg.hdfe3 <- felm(`Z-score`~ savings_log  | 0 |0|0,
-                  data = bank.data)
-reg.hdfe3.1 <- felm(`Z-score`~ savings_log + asset_log + tier.1_log +PD_mean | 0 |0|0,
-                  data = bank.data)
-reg.hdfe3.2 <- felm(`Z-score`~ savings_log + asset_log + tier.1_log +PD_mean | year+Country |0|0,
-                  data = bank.data)
+DiD.dummy <- felm(outcome ~ EBAbank | Country |0|0,
+                  data = subset(DiD.data, subset = EBAcountry == 1))
 
@@
 
-stargazer(reg.hdfe1,reg.hdfe1.1,reg.hdfe1.2,
+stargazer( DiD.intensity.2, DiD.dummy,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "EBA capital exercise and Capital Savings",
+          table.layout = "n=d#-t-as-",
+          notes = "\\parbox[t]{7cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\text{Capital savings}_i = \\beta X_i+\\varepsilon_i.
+\\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
+In column (1) the regressor is the capital shortfall during the capital exercise and in column (2) a dummy that indicates if bank $i$ was part of the exercise.}",
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          dep.var.labels = c("Capital savings"), column.sep.width = "2pt",
+          covariate.labels = c( "Shortfall","EBA dummy"),
+          add.lines = list(c("Country FE","Yes","Yes")))
+
+```
+
+```{r, echo=FALSE}
+
+bank.data <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+stage1 <- felm(data = bank.data , RW.s ~ basel + IRB  |  year +bvdid  |0|0)
+se.s1 <- summary(stage1, robust=TRUE)$coefficients[,2]
+
+bank.data <- mutate(bank.data, f.RW.s = stage1$fitted.values )
+
+iv.t1 <- felm(data = bank.data ,tier1ratio  ~ f.RW.s  | year + bvdid|0|0)
+#se.s1 <- summary(iv.t1$stage1,robust=TRUE)$coefficients[,2]
+se.iv.t1 <- summary(iv.t1,robust=TRUE)$coefficients[,2]
+
+iv.roe <- felm(data = bank.data ,ROE  ~ f.RW.s  | year + bvdid|0|0)
+se.iv.roe <- summary(iv.roe,robust=TRUE)$coefficients[,2]
+
+iv.z <- felm(data = bank.data , Zscore  ~ f.RW.s  | year + bvdid|0|0)
+se.iv.z <- summary(iv.z,robust=TRUE)$coefficients[,2]
+
+iv.roa <- felm(data = bank.data , ROA  ~ f.RW.s  | year + bvdid|0|0)
+se.iv.roa <- summary(iv.roa,robust=TRUE)$coefficients[,2]
+
+iv.sroa <- felm(data = bank.data , r.sd.ROA  ~ f.RW.s  | year + bvdid|0|0)
+se.iv.sroa <- summary(iv.sroa,robust=TRUE)$coefficients[,2]
+
+iv.lev <- felm(data = bank.data , leverage  ~ f.RW.s  | year + bvdid|0|0)
+se.iv.lev <- summary(iv.lev,robust=TRUE)$coefficients[,2]
+
+F.test <- round(t(sapply(stage1$lhs, function(lh) waldtest(stage1, ~ basel + IRB, lhs=lh)))[,5],1)
+
+F.test <- c("F-test",as.character(F.test),"","","","","","","")
+
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(stage1, iv.t1, iv.roe, iv.roa, iv.z, iv.sroa, iv.lev, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Instrumental variable analysis",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{15cm}{The table shows the relationship between the measure of capital savings and several outcome variables instrumented by the introduction of Basel (at country level) and IRB (at bank level).}",
+          se = list(se.s1, se.iv.t1, se.iv.roe, se.iv.z, se.iv.roa, se.iv.sroa, se.iv.lev),
+          covariate.labels = c("Basel","IRB","Savings"),
+          dep.var.labels = c("Savings","CET1","ROE","ROA", "Z-score", "sd(ROA)","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.labels   = c("First stage", "Second stage"), 
+          column.separate = c(1, 6),
+          column.sep.width = "3pt",
+          add.lines = list(c("Bank FE", "Yes", "Yes","Yes", "Yes", "Yes", "Yes", "Yes"),
+                           c("Year FE", "Yes", "Yes","Yes", "Yes", "Yes", "Yes", "Yes"),
+                           F.test))
+
+```
+```{r, echo=FALSE, include=FALSE}
+
+reg.hdfe1 <- felm(tier1ratio  ~ RW.s | 0  |0|0,
+                  data = bank.data)
+
+reg.hdfe2 <- felm(tier1ratio  ~ RW.s  | bvdid + year  |0|0,
+                  data = bank.data)
+
+reg.hdfe3 <- felm(Zscore  ~ RW.s  | 0  |0|0,
+                  data = bank.data)
+
+reg.hdfe4 <- felm(Zscore  ~ RW.s  | bvdid + year  |0|0,
+                  data = bank.data)
+
+```
+
+```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
+
+stargazer(reg.hdfe1,reg.hdfe2,reg.hdfe3,reg.hdfe4,
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
@@
           dep.var.labels = c("CT1 ratio"), column.sep.width = "2pt",
-          covariate.labels = c("log Savings", "log Total asset","log CT1 capital","Portfolio mean PD"),
           add.lines = list(c("Year FE", "No", "No","Yes"),
@@
 ```
-
-```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}
-
-stargazer(reg.hdfe2,reg.hdfe2.1,reg.hdfe2.2,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small", 
-          title = "Capital Savings and Return on Equity",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and return on equity.
-The dependent variable in each regression is the return on equity using PL before taxes.
-In column (2) and (3) I control for asset size, tier one capital size, and mean PD.
-In regressions (3) I also include bank and year fixed effects.}",
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          label = "",
-          style = "qje",
-          dep.var.labels = c( "ROE"), column.sep.width = "2pt",
-          covariate.labels = c("log Savings", "log Total asset","log CT1 capital","Portfolio mean PD"),
-          add.lines = list(c("Year FE", "No", "No","Yes"),
-                           c("Bank FE","No", "No","Yes")))
-
-```
-
-```{r, fig.height = 10, fig.width = 7, results="asis",  out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}
-
-stargazer(reg.hdfe3,reg.hdfe3.1,reg.hdfe3.2,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small", 
-          title = "Capital Savings and Risk",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and overall risk.
-The dependent variable in each regression is the Z-score measure.
-In column (2) and (3) I control for asset size, tier one capital size, and mean PD.
-In regressions (3) I also include country and year fixed effects.}",
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          label = "",
-          style = "qje",
-          dep.var.labels = c("Z-score"), column.sep.width = "2pt",
-          covariate.labels = c("log Savings", "log Total asset","log CT1 capital","Portfolio mean PD"),
-          add.lines = list(c("Year FE", "No", "No","Yes"),
-                           c("Country FE","No", "No","Yes")))
-
-```  
-## Robustness
-
-## Conclusion
-
-## References

### 2019-08-19 Analysis/Draft/Draft.Rmd
First version of paper structure with to-do list

@@
 ---
-title: "Distribution of portfolio credit risk and capital savings"
+title: "Credit risk distribution and capital savings"
 author:
-- affiliation: Tilburg University; l.avezum@uvt.nl
-  name: Lucas Avezum
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+- name: Lucas Avezum
+  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
+date: 
+header-includes:
+   - \usepackage{rotating, graphicx, dcolumn}
 output:
@@
   word_document: default
-bibliography: null
-citecolor: null
+bibliography: ../../Auxiliar/mybibfile.bib
+citecolor: black
 anonymous: no
@@
                       comment   = "",
-                      cache     = TRUE)
+                      cache     = FALSE)
 # These options are tuned for manuscript/presentation. 
@@
 library(DescTools)           ## Command to winsorize variables
-library(robustHD)
 
 p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+
 ```
@@
 load("../../Data/Datasets/yeardecomposition.Rda")
-load("../../Data/Datasets/DifInDif.Rda")
 ```
@@
 
-```{r}
+# To-Do List {-}
 
-bank.data <- bank.data %>% ungroup() %>%
-  mutate(RW.hat=RW.hat*100,
-         RW.s=RW.s*100)
-```
+\begin{itemize}
 
-# List
-* Use the share under IRB to weight capital savings measure (share vs intensity). Include share in regressions plus interaction with capital savings
+\item Data collection
 
-* Explore alternative capital savings measures, such as, savings in terms of capital (mn),in terms of capital ratio, in terms of capital ratio savings (counterfactual vs actual)
+   \begin{itemize}
 
-* Plot marginal increase of savings in time.
+   \item[$-$] Collect information for 32 IRB banks that participate on EBA exercise out of 52. Currently, we have information for 10 banks (and 35 SA banks).
 
-* Decompose capital savings: share effect, size effect, etc...
+   \end{itemize}
 
-* Remove step effect from regressions: average across all IRB banks (add IRB dummy) or IRB specific dummy (average out the step for each bank).
-Idea is to focus on difference across IRB banks.
-it might not make sense to include SA banks anymore.
+\item Write data section:
 
-* Plot capital savings measures to other risk variables
+   \begin{itemize}
 
-* Plot cross section for different capital saving measures (replicate first plot)
+   \item[$-$] Use decomposition to show how the intensity and extensity of PD distribution effect varies in time and across banks.
 
-* check and better clean data:
+   \item[$-$] Use decomposition to calculate how much of the PD distribution effect variation explains total RW variation (both in time and across banks).
 
-# Capital savings measure
+   \item[$-$] Explain capital savings measure
+
+   \item[$-$] Summary statistics
+
+   \end{itemize}
+
+\item Analysis
+
+   \begin{itemize}
+
+   \item[$-$] EBA vs non-EBA:Implement matching strategy following \cite{gropp2018banks}.
+
+   \item[$-$] Dif-on-Dif for EBA sample. Use capital shortfall as treatment intensity indicator.
+
+   \item[$-$] OLS relating capital savings measure to, capital ratio, RWA, ROA, ROE, Z-score, leverage.
+
+   \end{itemize}
+
+\item Robustness
+   \begin{itemize}
+
+   \item[$-$] Previous analysis with alternative capital savings measures (Gini coefficient, RW)
+
+   \end{itemize}
+
+\item Extension
+
+   \begin{itemize}
+
+   \item[$-$] Use Barth measures to study interaction with regulation.
+
+   \item[$-$] Use corporate governance \citep{anginer2016corporate} and subordinated debt \citep{gropp2018banks} to study different incentive to raise capital.
+
+   \item[$-$] Use NPL ratios to check if PD dispersion is associated with better risk management practices [@cucinelli2018credit].
+
+   \item[$-$] Check portfolio-level effects, for instance, corporate vs real estate.
+
+   \end{itemize}
+\end{itemize}
+\newpage
+
+# Introduction
+
+# Background
 
@@
 \begin{equation}
-\text{Capital savings}_{i}= \frac{f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i)}{\sum_{i}EAD_i},
+\label{eq1}
+\text{RWA}= 12.5 \sum_{i}EAD_iLGD_i f(PD_i),
+\end{equation}
+
+\begin{equation}
+\label{eq2}
+\widehat{\text{RWA}}= 12.5 f(\overline{PD})\sum_{i}EAD_iLGD_i,
 \end{equation}
@@
 
+\begin{equation*}
+\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
+\end{equation*}
+
+The capital savings measure is the difference between equations \ref{eq1} and \ref{eq2}:
+
 \begin{equation}
-\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
+\label{eq3}
+\text{RWA savings}=\widehat{\text{RWA}}-\text{RWA},
 \end{equation}
 
+\newpage
+
+# Data
+
 ```{r}
-savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
+savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation \\ref{eq3}. Each point in the plot is a bank-year observation."
 ```
@@
 bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
- ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
-                          y = RW.s,
+ ggplot(mapping = aes(x = reorder(name, log.RWA.s, FUN = median, na.rm=TRUE),
+                          y = log.RWA.s,
                       color= Country, 
@@
     coord_flip() +
-  labs(y =  "Capital savings") +
-
+  labs(y =  "Capital savings (USD mn)") +
   theme_bw()
-
-```
-
-```{r}
-fig3_cap = "PD distribution by bank. The figure shows smoothed PD distributions for each bank in our data. Banks are ranked by median PD and colored by host-country"
-```
-
-```{r, fig.height = 10, fig.width = 7, eval=FALSE, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
-pillar3.expanded %>% 
-  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
-ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
-  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
-                       scale = 2 , from = 0, to = 0.05) +
-  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
-  scale_y_discrete(expand = c(0.01, 0)) +
-  labs(
-    x = 'Probability of default') +
-  theme_ridges(font_size = 13, grid = TRUE) + 
-  theme(axis.title.y = element_blank())
-```
-
-# Decomposition
-
-\begin{equation}
-\begin{aligned}
-\Delta RW & = q_{B}^{SA} \Delta RW^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \hat{RW}^{j} \\ 
-& + (1-q_{B}^{SA})\sum_{j=1}^{P} \hat{RW}^{j} \Delta q^{j} -(1-q_{B}^{SA})\Delta RW^{S} + (RW^{SA}-RW^{IRB})\Delta q^{SA},
-\end{aligned}
-\end{equation}
-
-\newpage
-## Time decomposition
-
-```{r}
-time.dec_cap = "Time decompostion.
-The figure shows the contribution of each component compared to the time of IRB introduction (t=0).
-The figure and calculations include only banks with observations for at least 5 consecutive years.
-The black line is the net change.
-Year decompostion.
-The figure shows the contribution of each component compared to 2008.
-The figure and calculations include all banks in the sample.
-The black line is the net change."
-```
-
-```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
-time.lable <- c("t=0", "t=1", "t=2", "t=3", "t=4")
-plot1 <-time.decomposition %>% 
-  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain"),
-         RW = RW - t0.RW,
-         time = as.factor(time)) %>%
-  ggplot(mapping = aes(x = time,
-                       y    = value, 
-                       fill = condition)) +
-geom_bar(stat = 'identity')+
-geom_line(aes(group=1, y = RW))  +
-  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
-  scale_x_discrete( name=NULL,breaks = c("0","1","2","3","4"),labels = c("t=0","t=1","t=2","t=3","t=4")) +
- theme(legend.position = c(0.3, 0.2), 
-        legend.title = element_blank(), 
-        legend.background=element_blank())
-
-year.decomposition <- year.decomposition %>%
-  mutate(RW_2 = RW - t0.RW,
-         year = as.factor(year)) 
-plot2 <- year.decomposition %>% 
-  select(year,RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = year)) +
-geom_bar(aes(y = value, 
-          fill = condition),
-          stat = 'identity')+
-geom_line(data = year.decomposition, 
-          aes(group=1,y = RW_2))  +
-  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
-  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
-theme(legend.position='none')
-grid.arrange(plot1, plot2, ncol=2)
 
@@
 
-## Cross-section decomposition
+## Capital Ratio Savings
+
+\begin{equation}
+\label{eq4}
+\text{Counterfactual capital ratio}=\frac{\text{Capital}}{\widehat{\text{RWA}}},
+\end{equation}
+
+\begin{equation}
+\label{eq5}
+\text{Capital ratio savings}=\text{Counterfactual capital ratio}- \text{Observed capital ratio},
+\end{equation}
 
 ```{r}
-cross.dec_cap = "Cross-section decompostion.
-The figure shows the contribution of each component compared to the average bank.
-The figure and calculations include all banks in the sample.
-The black line is the total RW.
-Relative cross-section decompostion.
-The figure shows the relative contribution of each component compared to the average bank.
-The figure and calculations include all banks in the sample.
-"
+savings_cap = "The figure plots the capital ratio savings measure. This measure is calculated using equation \\ref{eq5}. Each point in the plot is a bank-year observation."
 ```
 
-```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
-plot1 <- cross.section.decomposition %>% 
-  mutate(order.s = abs(dispersion.gain))%>%
-  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = reorder(name, order.s),
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity')+
-  geom_line(aes(y = total,group=1))  +
-  labs(x=NULL, y= NULL,subtitle = "Plot a") +
+```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
+bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
+ ggplot(mapping = aes(x = reorder(name, CAR.dif, FUN = median, na.rm=TRUE),
+                          y = CAR.dif,
+                      color= Country, 
+                      fill = Country)) +
+geom_point(alpha = 0.5,size = 2) +
+    labs(x=NULL) +
     coord_flip() +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
- theme(legend.position='none')
-
-plot2 <- cross.section.decomposition %>% 
-  mutate(order.s = abs(dispersion.gain))%>%
-  select(name, Country, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(value = abs(value),
-         condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = reorder(name, order.s),
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity',position = "fill")+
-  labs(x=NULL, y= NULL,subtitle = "Plot b") +
-    coord_flip() +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
-   theme(legend.title = element_blank(), 
-        legend.background=element_blank(),
-       legend.text=element_text(size=7),
-       axis.text.y = element_blank())
-
-grid.arrange(plot1, plot2, ncol=2)
+  labs(y =  "Capital ratio savings (%)") +
+  theme_bw()
 
 ```
-
-# Counterfactual analysis
 
@@
 ```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
- plot1 <- bank.data %>%  filter(!is.na(RW.s),
+ plot1 <- bank.data %>%  ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
          tier1capitalmillcu < totalcapitalmillcu)%>% 
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
   select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
@@
 
-plot2 <- bank.data %>%  filter(!is.na(RW.s),
+plot2 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
   select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
@@
 
-plot3 <- bank.data %>%  filter(!is.na(RW.s),
+plot3 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
   select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
@@
 
-plot4 <- bank.data %>% filter(!is.na(RW.s),
+plot4 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
+         tier1capitalmillcu < totalcapitalmillcu)%>% 
+  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
   select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
@@
 
+```{r}
+fig3_cap = "PD distribution by bank. The figure shows smoothed PD distributions for each bank in our data. Banks are ranked by median PD and colored by host-country"
+```
+
+```{r, fig.height = 10, fig.width = 7, eval=FALSE, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
+pillar3.expanded %>% 
+  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
+ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
+  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
+                       scale = 2 , from = 0, to = 0.05) +
+  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
+  scale_y_discrete(expand = c(0.01, 0)) +
+  labs(
+    x = 'Probability of default') +
+  theme_ridges(font_size = 13, grid = TRUE) + 
+  theme(axis.title.y = element_blank())
+```
+
+\newpage
+
+## Decomposition
+
+Rearranging \ref{eq3} and dividing by EAD we have:
+
+\begin{equation}
+\begin{aligned}
+RW & =  \widehat{\text{RW}} +\text{RWA savings} \\
+   & =  \widehat{\text{RW}} +\text{RW}^s 
+\end{aligned}
+\end{equation}
+
+I use the method proposed by @cannata2012inside to decompose the components of risk-weights.
+
+\begin{equation}
+\begin{aligned}
+\Delta \text{RW} & = q_{B}^{SA} \Delta \text{RW}^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \widehat{\text{RW}}^{j} \\ 
+& + (1-q_{B}^{SA})\sum_{j=1}^{P} \widehat{\text{RW}}^{j} \Delta q^{j}  + (\text{RW}^{SA}-\widehat{\text{RW}})\Delta q^{SA} -(1-q_{B}^{SA})\Delta \text{RW}^{S}+ \text{RW}^{S}\Delta q^{SA}\\
+& = \text{Other effects} -\underbrace{(1-q_{B}^{SA})\Delta \text{RW}^{S}}_\text{Intensive margin}+ \underbrace{\text{RW}^{S}\Delta q^{SA}}_\text{Extensive margin}
+\end{aligned}
+\end{equation}
+
+```{r}
+time.dec_cap = "Time decompostion.
+The figure shows the contribution of each component compared to the time of IRB introduction (t=0).
+The figure and calculations include only banks with observations for at least 5 consecutive years.
+The black line is the net change.
+Year decompostion.
+The figure shows the contribution of each component compared to 2008.
+The figure and calculations include all banks in the sample.
+The black line is the net change."
+```
+
+```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
+
+plot1 <-time.decomposition %>% 
+  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain")),
+         RW = RW - t0.RW,
+         time = as.factor(time)) %>%
+  ggplot(mapping = aes(x = time,
+                       y    = value, 
+                       fill = condition)) +
+geom_bar(stat = 'identity')+
+geom_line(aes(group=1, y = RW))  +
+  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
+  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
+  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
+ theme(legend.position = c(0.3, 0.1), 
+        legend.title = element_blank(), 
+        legend.background=element_blank())
+
+year.decomposition <- year.decomposition %>%
+  mutate(RW_2 = RW - t0.RW,
+         year = as.factor(year)) 
+plot2 <- year.decomposition %>% 
+  select(year,RW,  SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
+  ggplot(mapping = aes(x = year)) +
+geom_bar(aes(y = value, 
+          fill = condition),
+          stat = 'identity')+
+geom_line(data = year.decomposition, 
+          aes(group=1,y = RW_2))  +
+  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
+  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
+  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
+theme(legend.position='none')
+grid.arrange(plot1, plot2, ncol=2)
+
+```
+
+```{r}
+cross.dec_cap = "Cross-section decompostion.
+The figure shows the contribution of each component compared to the average bank.
+The figure and calculations include all banks in the sample.
+The black line is the total RW.
+Relative cross-section decompostion.
+The figure shows the relative contribution of each component compared to the average bank.
+The figure and calculations include all banks in the sample.
+"
+```
+
+```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
+plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
+  mutate(order.s = abs(int.gain))%>%
+  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
+  ggplot(mapping = aes(x = reorder(name, total),
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity')+
+  geom_line(aes(y = total,group=1))  +
+  labs(x=NULL, y= NULL,subtitle = "Plot a") +
+    coord_flip() +
+  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
+ theme(legend.position='none')
+
+plot2 <- cross.section.decomposition %>% filter(q.SA<1)%>%
+  mutate(order.s = abs(int.gain))%>%
+  select(name, Country,total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(value = abs(value),
+         condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
+  ggplot(mapping = aes(x = reorder(name, total),
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity',position = "fill")+
+  labs(x=NULL, y= NULL,subtitle = "Plot b") +
+    coord_flip() +
+  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
+   theme(legend.title = element_blank(), 
+        legend.background=element_blank(),
+       legend.text=element_text(size=7),
+       axis.text.y = element_blank())
+
+grid.arrange(plot1, plot2, ncol=2)
+
+```
+
+\newpage
+
 ```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
 bank.data %>% ungroup()%>%
@@
   mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
-  subset(select = c("RW.s","tier1ratio","tier.hat","totalcapitalratio","CAR.hat","ROE","Zscore")) %>%
+  subset(select = c("log.RWA.s","tier1ratio","tier.hat","tier.dif","totalcapitalratio","CAR.hat","CAR.dif","ROE","ROA","Zscore","leverage")) %>%
   as.data.frame() %>% 
@@
           omit.summary.stat = c("p25","p75"),
-          covariate.labels = c("Capital savings","Tier 1 ratio","Counterfactual tier 1 ratio","Total ratio","Counterfactual total ratio","ROE","Z-score"),
+          covariate.labels = c("Log RWA savings","Tier 1 ratio","Counterfactual tier 1 ratio","Tier 1 ratio savings","Total ratio","Counterfactual total ratio","Capital ratio savings","ROE","ROA","Z-score","Leverage ratio"),
           title = "Summary statistics",
           table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{11cm}{The table shows summary statistics for the sample of banks collected.
+          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected.
 Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports.
@@
 
+# Results
+
+```{r}
+
+pre.treat.out <- bank.data %>%
+  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
+  filter(year %in% c("2009","2010"),
+         !is.na(IRB)) %>%
+  select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
+         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
+  group_by(bvdid) %>%
+  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
+
+post.treat.out <- bank.data %>%
+  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
+  filter(year %in% c("2012","2013"),
+         !is.na(IRB)) %>%
+  select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
+         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
+  group_by(bvdid) %>%
+  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
+
+DiD.data <- bank.data %>%
+  filter(year %in% c("2010")) %>%
+  mutate(Shortfall.RW.2 = ifelse(EBAcountry == 1,
+                                 ifelse(EBAbank == 1,Shortfall.RW, 0), NA))%>%
+  right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
+  right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
+    filter(EBAcountry == 1,
+           name != "UniCredit",
+           name != "Bank of Cyprus") %>%
+  mutate(Shortfall.mn   = log(1+Shortfall.mn), 
+         Shortfall.mn.2 = ifelse(EBAbank == 1,Shortfall.mn, 0),
+         Shortfall.RW.2 = ifelse(EBAbank == 1,Shortfall.RW, 0),
+         out.RWA.total  = log.RWA.total.post - log.RWA.total.pre,
+         out.RWA.other  = log.RWA.other.post - log.RWA.other.pre,
+         out.RWA.credit = log.RWA.credit.post - log.RWA.credit.pre,
+         out.RWA.IRB    = log.RWA.IRB.post - log.RWA.IRB.pre,
+         out.RWA.SA     = log.RWA.SA.post - log.RWA.SA.pre,
+         out.RWA.hat    = log.RWA.hat.post - log.RWA.hat.pre,
+         out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
+         out.K          = log.K.post - log.K.pre,
+         out.tier       = log.tier.post - log.tier.pre,
+         out.CAR.ratio  = totalcapitalratio.post - totalcapitalratio.pre,
+         out.tier.ratio = tier1ratio.post - tier1ratio.pre,
+         out.CAR.dif    = CAR.dif.post - CAR.dif.pre,
+         out.tier.dif   = tier.dif.post - tier.dif.pre,
+         Country        = as.factor(Country)) %>% 
+  ungroup() %>% 
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+```
+```{r, include=FALSE}
+ggplot(DiD.data, aes(x= EBAbank, y =  out.tier.ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
 ```{r, include=FALSE}
 
-DiD.data <- DiD.data %>% ungroup() %>% 
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+reg.1 <- felm(out.RWA.s ~ Shortfall.mn.2 | 0 |0|0,
+                  data = DiD.data)
+se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
 
-DiD.intensity <- felm(outcome ~ Shortfall.RW | Country |0|0,
+reg.2 <- felm(out.RWA.s ~ Shortfall.mn.2 
+                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| 0 |0|0,
                   data = DiD.data)
+se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
 
-DiD.intensity.2 <- felm(outcome ~ Shortfall.RW.2 | Country |0|0,
+reg.3 <- felm(out.RWA.s ~ Shortfall.mn.2 
+                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| Country |0|0,
                   data = DiD.data)
+se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
 
-DiD.dummy <- felm(outcome ~ EBAbank | Country |0|0,
-                  data = subset(DiD.data, subset = EBAcountry == 1))
+reg.4 <- felm(out.RWA.total ~ Shortfall.mn.2 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(out.RWA.hat ~ Shortfall.mn.2 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(out.CAR.ratio ~ Shortfall.mn.2 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(out.tier.ratio ~ Shortfall.mn.2 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
 
@@
 
-stargazer( DiD.intensity.2, DiD.dummy,
+stargazer(reg.1, reg.4, reg.5, reg.6, reg.7,
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
@@
           table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{7cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\text{Capital savings}_i = \\beta X_i+\\varepsilon_i.
-\\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
-In column (1) the regressor is the capital shortfall during the capital exercise and in column (2) a dummy that indicates if bank $i$ was part of the exercise.}",
+          notes = "\\parbox[t]{20cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
+The regressor is the capital shortfall during the capital exercise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels = c("$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA average", "$\\Delta$ Capital ratio", "$\\Delta$ CET1 ratio"),
+          se = list(se.1, se.4, se.5, se.6, se.7),
+          covariate.labels = c("Log Capital Shortfall"),
           notes.label = "",
@@
           style = "qje",
-          dep.var.labels = c("Capital savings"), column.sep.width = "2pt",
-          covariate.labels = c( "Shortfall","EBA dummy"),
-          add.lines = list(c("Country FE","Yes","Yes")))
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Country FE", "\\multicolumn{1}{c}{No}", "\\multicolumn{1}{c}{No}",    
+                             "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```   
+```{r, include=FALSE}
+
+reg.1 <- felm(out.RWA.s ~ EBAbank | 0 |0|0,
+                  data = DiD.data)
+se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(out.RWA.s ~ EBAbank 
+                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| 0 |0|0,
+                  data = DiD.data)
+se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(out.RWA.s ~ EBAbank 
+                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio| Country |0|0,
+                  data = DiD.data)
+se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(out.RWA.total ~ EBAbank 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(out.RWA.hat ~ EBAbank 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(out.CAR.ratio ~ EBAbank 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(out.tier.ratio ~ EBAbank 
+                 | 0 |0|0,
+                  data = DiD.data)
+se.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
 
@@
 
-```{r, echo=FALSE}
+```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}
 
-bank.data <- bank.data %>% ungroup()%>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>%
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
-
-stage1 <- felm(data = bank.data , RW.s ~ basel + IRB  |  year +bvdid  |0|0)
-se.s1 <- summary(stage1, robust=TRUE)$coefficients[,2]
-
-bank.data <- mutate(bank.data, f.RW.s = stage1$fitted.values )
-
-iv.t1 <- felm(data = bank.data ,tier1ratio  ~ f.RW.s  | year + bvdid|0|0)
-#se.s1 <- summary(iv.t1$stage1,robust=TRUE)$coefficients[,2]
-se.iv.t1 <- summary(iv.t1,robust=TRUE)$coefficients[,2]
-
-iv.roe <- felm(data = bank.data ,ROE  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.roe <- summary(iv.roe,robust=TRUE)$coefficients[,2]
-
-iv.z <- felm(data = bank.data , Zscore  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.z <- summary(iv.z,robust=TRUE)$coefficients[,2]
-
-iv.roa <- felm(data = bank.data , ROA  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.roa <- summary(iv.roa,robust=TRUE)$coefficients[,2]
-
-iv.sroa <- felm(data = bank.data , r.sd.ROA  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.sroa <- summary(iv.sroa,robust=TRUE)$coefficients[,2]
-
-iv.lev <- felm(data = bank.data , leverage  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.lev <- summary(iv.lev,robust=TRUE)$coefficients[,2]
-
-F.test <- round(t(sapply(stage1$lhs, function(lh) waldtest(stage1, ~ basel + IRB, lhs=lh)))[,5],1)
-
-F.test <- c("F-test",as.character(F.test),"","","","","","","")
-
-```
-
-```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(stage1, iv.t1, iv.roe, iv.roa, iv.z, iv.sroa, iv.lev, 
+stargazer(reg.1, reg.4, reg.5, reg.6, reg.7,
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Instrumental variable analysis",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{15cm}{The table shows the relationship between the measure of capital savings and several outcome variables instrumented by the introduction of Basel (at country level) and IRB (at bank level).}",
-          se = list(se.s1, se.iv.t1, se.iv.roe, se.iv.z, se.iv.roa, se.iv.sroa, se.iv.lev),
-          covariate.labels = c("Basel","IRB","Savings"),
-          dep.var.labels = c("Savings","CET1","ROE","ROA", "Z-score", "sd(ROA)","Leverage"), 
+          title = "EBA capital exercise and Capital Savings: Exercise vs Non-exercise banks",
+          table.layout = "n=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i =\\alpha +  \\beta \\text{Capital Shortfall}_i+\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
+The regressor is a dummy that indicates if bank $i$ was part of the exercise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels = c("$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA average", "$\\Delta$ Capital ratio", "$\\Delta$ CET1 ratio"),
+          se = list(se.1, se.4, se.5, se.6, se.7),
+          covariate.labels = c("EBA"),
           notes.label = "",
@@
           style = "qje",
-          column.labels   = c("First stage", "Second stage"), 
-          column.separate = c(1, 6),
           column.sep.width = "3pt",
-          add.lines = list(c("Bank FE", "Yes", "Yes","Yes", "Yes", "Yes", "Yes", "Yes"),
-                           c("Year FE", "Yes", "Yes","Yes", "Yes", "Yes", "Yes", "Yes"),
-                           F.test))
-
-```
-```{r, echo=FALSE, include=FALSE}
-
-reg.hdfe1 <- felm(tier1ratio  ~ RW.s | 0  |0|0,
-                  data = bank.data)
-
-reg.hdfe2 <- felm(tier1ratio  ~ RW.s  | bvdid + year  |0|0,
-                  data = bank.data)
-
-reg.hdfe3 <- felm(Zscore  ~ RW.s  | 0  |0|0,
-                  data = bank.data)
-
-reg.hdfe4 <- felm(Zscore  ~ RW.s  | bvdid + year  |0|0,
-                  data = bank.data)
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Country FE", "\\multicolumn{1}{c}{No}", "\\multicolumn{1}{c}{No}",    
+                             "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
 
@@
 
-```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier,log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
 
-stargazer(reg.hdfe1,reg.hdfe2,reg.hdfe3,reg.hdfe4,
+reg.1 <- felm(totalcapitalratio  ~ CAR.dif | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ CAR.dif | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
            header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Capital Savings and Capital Requirement Ratio",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and risk-weighted capital ratio.
-The dependent variable in each regression is the Core Tier 1 ratio.
-In column (2) and (3) I control for asset size, tier one capital size, and mean PD.
-In regressions (3) I also include bank and year fixed effects.}",
+          title = "Capital Ratio Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Capital ratio savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{CT1 ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
+$\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
           notes.label = "",
@@
           style = "qje",
-          dep.var.labels = c("CT1 ratio"), column.sep.width = "2pt",
-          add.lines = list(c("Year FE", "No", "No","Yes"),
-                           c("Bank FE","No", "No","Yes")))
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ CAR.dif  | year +  IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ CAR.dif | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ CAR.dif  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ CAR.dif  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
 
 ```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Capital Ratio Savings and Bank Characteristics: Bank $\\times$ IRB Fixed effects",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta \\times\\text{Capital ratio savings}_{it}+\\alpha_{i}\\times\\text{IRB}_{it}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
+$\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ CAR.dif + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ CAR.dif + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ CAR.dif + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Capital Ratio Savings and Bank Characteristics: IRB vs Non-IRB Banks",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta \\times\\text{Capital ratio savings}_{it}+\\gamma\\text{IRB}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
+$\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Capital ratio savings", "IRB","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif, tier.dif, 
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ CAR.dif | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ CAR.dif  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ CAR.dif  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Capital Ratio Savings and Bank Characteristics: Only IRB Banks",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta \\times \\text{Capital ratio savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
+Only banks that use the IRB approach are included in these regressions.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "RWA Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Log RWA savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K ,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ log.RWA.s  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K   ~ log.RWA.s  | year +  IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ log.RWA.s  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ log.RWA.s  | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ log.RWA.s  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ log.RWA.s  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ log.RWA.s  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "RWA Savings and Bank Characteristics: Bank $\\times$ IRB Fixed effects",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\alpha_{i}\\times IRB_{it}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings.
+$\\text{IRB}_{it}$ is a dummy variable equal to 1 if bank i at year t used the IRB approach to calculate capital requirements and zero otherwise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Log RWA savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K ,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ log.RWA.s + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ log.RWA.s + IRB  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ log.RWA.s + IRB  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "RWA Savings and Bank Characteristics: IRB vs Non-IRB Banks",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\gamma\\times\\text{IRB}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings.
+$\\text{IRB}_{it}$ is a dummy variable equal to 1 if bank i at year t used the IRB approach to calculate capital requirements and zero otherwise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Log RWA savings", "IRB","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r, echo=FALSE}
+p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, 
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage, log.RWA.total, log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(totalcapitalratio  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(log.K  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(log.RWA.total  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(ROE  ~ log.RWA.s | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(ROA  ~ log.RWA.s  | year + bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(Zscore  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+
+reg.7 <- felm(leverage  ~ log.RWA.s  | year + bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.7 <- summary(reg.7,robust=TRUE)$coefficients[,2]
+
+```
+```{r, include=FALSE}
+ggplot(bank.data, aes(x= log.RWA.s, y =  tier1ratio)) + geom_point() +geom_smooth(method = "lm")
+```
+
+```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6, reg.7, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "RWA Savings and Bank Characteristics: only IRB",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{it} = \\beta\\times \\text{Log RWA savings}_{it}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{it}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Log RWA savings}_{it}$ is the natural logarithm of risk-weighted assets savings.
+The sample is restricted to banks that use the IRB approach to compute their capital requirements.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5, se.reg.6, se.reg.7),
+          covariate.labels = c("Log RWA savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```

### 2019-08-19 Analysis/Simulation/Formula_sensitivity.Rmd
First version of paper structure with to-do list

@@
   b  <- (0.11852-0.05478*log(PD))^2
-  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
+  K  <- LGD*(punif(((1-R)^(-0.5))*qunif(PD)+((R/(1-R))^(0.5))*qunif(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
   RW <- K*12.5
@@
 mapping_retail <- function(PD,LGD,R,alpha,portfolio){
-  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
+  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha)))
+  RW <- K*12.5
+  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
+  return(data)
+}
+
+mapping_retail_alt <- function(PD,LGD,R,alpha,portfolio){
+  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qunif(alpha)))
   RW <- K*12.5
@@
 ```
+
+```{r wholesale vs retail, echo=FALSE}
+data <- mapping_retail(grid,0.45,0.15,0.999,factor("Mortgage"))
+data <- rbind(data,mapping_retail_alt(grid,0.45,0.15,0.999,factor("Mortgage_alt")))
+
+p <- ggplot(data = subset(data,subset = PD <1),
+            mapping = aes(x = PD, y=K, colour = portfolio)) 
+p + geom_line(alpha = 1) + 
+  labs(x = "PD", 
+         y = "Capital requirement per unit of exposure",title = "Figure 1: Wholesale vs Retail") +
+  theme_bw()
+```
+
+Concavity of RW formula: highly driven by Normality assumption.
+
+Findings so far:
+
+- Fatter-tailed distribution of systematic component lead to less concavity (uniform distribution used, tes with others)
 

### 2019-08-19 Analysis/Simulation/report.Rmd
First version of paper structure with to-do list

@@
 author: "Lucas Avezum"
-date: "May 14, 2019"
-output: pdf_document
+date: "June 6, 2019"
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
+  word_document: default
 ---
@@
 library(gglorenz)
+library(ggplot2)
+library(psych)
+library(table1)
 
 ```
-```{r, include=FALSE}
+```{r load data, include=FALSE}
 load("../../Data/Temp/Raw_data_2.Rda")
@@
 lc.retail <- Lc(lc.retail$PD,lc.retail$EAD*lc.retail$LGD)
+
+bankscope <- read_xlsx("../../Data/Temp/BankFocus_Export.xlsx",sheet = 2)
+collection <- read_xlsx("../../Data/Temp/First_List_banks.xlsx",sheet = 5)
+
+bankscope <- mutate(bankscope,guo_id=ifelse(is.na(guo_id),id,guo_id))
+bankscope <- mutate(bankscope,cap_req1=ifelse(cap_req1=="n.a.",cap_req2,cap_req1))
+bankscope <- mutate(bankscope,cap_req2=ifelse(cap_req2=="n.a.",cap_req1,cap_req2))
+
+collection <- mutate(collection,cap_req1=ifelse(cap_req1=="n.a.",cap_req2,cap_req1))
+collection <- mutate(collection,cap_req2=ifelse(cap_req2=="n.a.",cap_req1,cap_req2))
+
+bankscope <- filter(bankscope,guo_id==id)
+bankscope <- filter(bankscope,cap_req1!="n.a.")
+collection <- filter(collection,cap_req1!="n.a.")
+
+bankscope <- mutate(bankscope,total_asset=as.numeric(total_asset),
+                    cap_req1 = as.numeric(cap_req1),
+                    cap_req2 = as.numeric(cap_req2))
+collection <- mutate(collection,total_asset=as.numeric(total_asset),
+                    cap_req1 = as.numeric(cap_req1),
+                    cap_req2 = as.numeric(cap_req2))
+
+bankscope <- mutate(bankscope,total_asset=total_asset/1000)
+collection <- mutate(collection,total_asset=total_asset/1000)
+
 ```
 
-- Average Capital savings using $\hat{P}$ and LGD original distribution: 32.8%
+```{r data analysis, echo=FALSE,include=FALSE}
 
-- Average Gini coefficient $PD$ weighted by $LGD$: 0.67
+bankscope.size <- summarise_if(bankscope,is.numeric,funs(sum),na.rm=TRUE)
+collection.size <- summarise_if(collection,is.numeric,funs(sum),na.rm=TRUE)
+collection.IRB <- filter(collection,SA==0)
+collection.SA <- filter(collection,SA==1)
+collection.IRB.collected <- filter(collection,SA==0,collected==1)
 
-- Correlation between Gini coefficient and capital savings: 0.79
+```
 
-- $\hat{P}=\frac{\sum_i^nEAD_iLGD_iPD_i}{\sum_i^nEAD_iLGD_i}$ \newline
+- Banks in BankScope: 1048
 
-```{r lorenz curve, echo=FALSE,}
+- Banks in target sample: 124
+
+- Relative asset size (target/BankScope): 78.4%
+
+- IRB banks in target sample: 48
+
+- IRB banks collected: 22
+
+- SA banks in sample: 52
+
+- Unindentified banks in sample: 24
+
+```{r summary stats, echo=FALSE}
+table1::label(collection$total_asset) <- "Total assets (bn USD)"
+table1::label(collection$cap_req1) <- "Capital ratio (%)"
+table1::table1(~ total_asset + cap_req1  | country, data = collection,transpose = TRUE)
+```
+```{r summary stats 2 , echo=FALSE}
+table1::label(collection.IRB.collected$total_asset) <- "Total assets (bn USD)"
+table1::label(collection.IRB.collected$cap_req1) <- "Capital ratio (%)"
+table1::table1(~ total_asset + cap_req1  | country, data = collection.IRB.collected,transpose = TRUE)
+```
+```{r summary stats 3 , echo=FALSE}
+collection <- mutate(collection, Type=ifelse(SA==0,"SA","IRB"))
+table1::label(collection$total_asset) <- "Total assets (bn USD)"
+table1::label(collection$cap_req1) <- "Capital ratio (%)"
+table1::table1(~ total_asset + cap_req1 | Type, data = collection,transpose = TRUE)
+```
+
+```{r sample distribution, echo=FALSE,include=FALSE}
+p.total <- ggplot(collection, aes(x="", y=total_asset, fill=country))+
+geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0)
+p.total
+
+```
+
+```{r lorenz curve, echo=FALSE,include=FALSE}
 plot(lc.wholesale,

### 2019-08-19 Auxiliar/Function.R

@@ -13,7 +13,7 @@ mapping_wholesale <- function(PD, LGD = 0.45, M = 2.5, alpha = 0.999){
   R   <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
   b   <- (0.11852-0.05478*log(PD))^2
   f   <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
-  K   <- f*((1-1.5*b)^-1*(1+(M-2.5)*b))
+  K   <- f*((1+(M-2.5)*b)/(1-1.5*b))
   RW  <- K*12.5
   data <- data.frame(K,RW,PD,LGD,M,R,b,f)
   return(data)

### 2019-08-19 Data/Clean.R

@@ -12,11 +12,15 @@ rm(list = ls())
 library(tidyverse)                ## Data manipulation, pipe operator
 library(zoo)                      ## Apply function to rolling margins of data
 library(DescTools)                ## Command to winsorize variables
+library(readxl)       ## Command to open xlsx files                                                     
+
 source("Auxiliar/Function.R")     ## Capital requirements functions
 
 # Open dirt data frame
 load("Data/Temp/BankScope.Rda")
 load("Data/Temp/Pillar3Data.Rda")
+
+
 
 ##============================================================================##
 ## Pillar-III reports                                                         ##
@@ -69,17 +73,12 @@ for (i in 1:nrow(IRB.data)){
   if (!IRB.data[i,"Portfolio_1"] %in% c("Wholesale","Equity")){
     next
   }
-  if (is.na(IRB.data[i,"RW"])){
+ 
+  IRB.data[i,"RWA"]     <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"PD"],
+                                                               IRB.data[i,"LGD"])["RW"]
+ 
   IRB.data[i,"RWA.hat"] <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"w.avg.PD"],
-                                                                 IRB.data[i,"LGD"])["RW"]
-  } else {
-  IRB.data[i,"b"]       <- mapping_wholesale(IRB.data[i,"PD"],IRB.data[i,"LGD"])["b"]
-  IRB.data[i,"M"]       <- 2.5+(IRB.data[i,"RW"]*(1-1.5*(IRB.data[i,"b"]))/(12.5*mapping_wholesale(IRB.data[i,"PD"],IRB.data[i,"LGD"])["f"])-1)/IRB.data[i,"b"]
-  
-  IRB.data[i,"RWA.hat"] <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"w.avg.PD"],
-                                                               IRB.data[i,"LGD"],
-                                                              IRB.data[i,"M"])["RW"]
-  }
+                                                               IRB.data[i,"LGD"])["RW"]
 }
 
 for (i in 1:nrow(IRB.data)){
@@ -89,17 +88,15 @@ for (i in 1:nrow(IRB.data)){
[…]

### 2019-08-19 Data/Import.R

@@ -122,8 +122,8 @@ bankscope <- bankscope %>%
   mutate_at(c(grep("usd", names(bankscope))[1]:ncol(bankscope)),as.numeric) %>%
   # Convert to USD
   rowwise() %>%
-  mutate_at(c(grep("lcu$", names(bankscope))[1]:ncol(bankscope)),
-            funs(prod(.,exchangeratefromoriginalcurrencyusd, na.rm = TRUE))) %>%
+  mutate_at(c(grep("lcu$", names(bankscope))),
+            funs(.*exchangeratefromoriginalcurrencyusd)) %>%
   # Add Basel 2 country level introduction year
   ungroup() %>%
   left_join(basel.indicator, by = c("Country")) %>%

### 2019-09-03 Analysis/Draft/Draft.Rmd
Add portfolio breakdown of capital savings (wholesale, retail, corporate...)

@@
 
-p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+p_w <- c(0.0, 1)         ## Set winsorizing limits
 
@@
   select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
-         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
+         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
+         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
   group_by(bvdid) %>%
@@
   filter(year %in% c("2010")) %>%
-  mutate(Shortfall.RW.2 = ifelse(EBAcountry == 1,
-                                 ifelse(EBAbank == 1,Shortfall.RW, 0), NA))%>%
   right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
   right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
-    filter(EBAcountry == 1,
-           name != "UniCredit",
-           name != "Bank of Cyprus") %>%
   mutate(Shortfall.mn   = log(1+Shortfall.mn), 
-         Shortfall.mn.2 = ifelse(EBAbank == 1,Shortfall.mn, 0),
-         Shortfall.RW.2 = ifelse(EBAbank == 1,Shortfall.RW, 0),
+         Shortfall.mn.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.mn, 0), NA),
+         Shortfall.RW.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.RW, 0), NA),
          out.RWA.total  = log.RWA.total.post - log.RWA.total.pre,
@@
          out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
+         out.RWA.s.Retail = log.RWA.s.Retail.post - log.RWA.s.Retail.pre,
+         out.RWA.s.Equity = log.RWA.s.Equity.post - log.RWA.s.Equity.pre,
+         out.RWA.s.Wholesale = log.RWA.s.Wholesale.post - log.RWA.s.Wholesale.pre,
+         out.RWA.s.Corporate = log.RWA.s.Corporate.post - log.RWA.s.Corporate.pre,
+         out.RWA.s.Sovereign = log.RWA.s.Sovereign.post - log.RWA.s.Sovereign.pre,
+         out.RWA.s.Banks = log.RWA.s.Banks.post - log.RWA.s.Banks.pre,
          out.K          = log.K.post - log.K.pre,
@@
   ungroup() %>% 
+  filter(EBAcountry == 1,
+         name != "Bank of Cyprus") %>%
   mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
@@
 ```{r, include=FALSE}
-ggplot(DiD.data, aes(x= EBAbank, y =  out.tier.ratio)) + geom_point() +geom_smooth(method = "lm")
-```
-
+ggplot(DiD.data, aes(x= Shortfall.mn, y =  out.RWA.s.Corporate)) + geom_point() +geom_smooth(method = "lm")
+```
+```{r, include=FALSE}
+
+reg.1 <- felm(out.RWA.s ~ Shortfall.mn | 0 |0|0,
+                  data = DiD.data)
+se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(out.RWA.s.Retail ~ Shortfall.mn | 0 |0|0,
+                  data = DiD.data)
+se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(out.RWA.s.Wholesale ~ Shortfall.mn | 0 |0|0,
+                  data = DiD.data)
+se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(out.RWA.s.Corporate ~ Shortfall.mn | 0 |0|0,
+                  data = DiD.data)
+se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(out.RWA.s.Sovereign ~ Shortfall.mn | 0 |0|0,
+                  data = DiD.data)
+se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(out.RWA.s.Banks ~ Shortfall.mn | 0 |0|0,
+                  data = DiD.data)
+se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+```
+
+```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "EBA capital exercise and Capital Savings",
+          table.layout = "n=d#-t-as-",
+          notes = "\\parbox[t]{20cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
+The regressor is the capital shortfall during the capital exercise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels = c("$\\Delta \\text{Log RWA}^S$", "$\\Delta \\text{Log RWA}^S_R$", "$\\Delta \\text{Log RWA}^S_W$", "$\\Delta \\text{Log RWA}^S_C$", "$\\Delta \\text{Log RWA}^S_S$", "$\\Delta \\text{Log RWA}^S_B$"),
+          se = list(se.1, se.2, se.3, se.4, se.5, se.6),
+          covariate.labels = c("Log Capital Shortfall"),
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Country FE", "\\multicolumn{1}{c}{No}", "\\multicolumn{1}{c}{No}",    
+                             "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```  
 ```{r, include=FALSE}
@@
 
-reg.4 <- felm(out.RWA.total ~ Shortfall.mn.2 
+reg.4 <- felm(out.RWA.total ~ Shortfall.mn 
                  | 0 |0|0,

### 2019-09-03 Data/Clean.R

@@ -198,6 +198,31 @@ bank.data <- IRB.data %>%
   right_join(bank.data,by=c("bvdid","name","Country","year"))
 }
 
+## Aggregated IRB exposures at bank-portfolio-year level 
+for(i in c("Corporate","Sovereign","Banks")) {
+  bank.data <- IRB.data %>% 
+    group_by(name,bvdid,Country,year,Portfolio_1,Portfolio_2) %>%
+    filter(Portfolio_2 %in% c(i)) %>%
+    summarize(!!(paste(       "gini",        i, sep=".")) := DescTools::Gini(PD,w.EAD, na.rm = TRUE),
+              !!(paste("sd"  ,"PD"  ,        i, sep=".")) := sd(PD)*100,
+              !!(paste("mean","PD"  ,        i, sep=".")) := weighted.mean(PD,w.EAD, na.rm = TRUE)*100,
+              !!(paste(       "RWA" ,        i, sep=".")) := sum(RWA, na.rm = TRUE),
+              !!(paste(       "RWA" , "hat", i, sep=".")) := sum(RWA.hat, na.rm = TRUE),
+              !!(paste(       "EAD" ,        i, sep=".")) := sum(EAD, na.rm = TRUE))%>%
+    ungroup()%>%
+    full_join(sample, by=c("bvdid","name","Country","year"))%>%
+    mutate(!!(paste("RW" ,"hat"    , i , sep=".")) := 100*get(paste("RWA","hat", i , sep="."))/get(paste("EAD", i , sep=".")),
+           !!(paste("RWA","s", i , sep=".")) := get(paste("RWA","hat", i, sep="."))-get(paste("RWA", i , sep=".")),
+           !!(paste("RW" ,"s", i , sep=".")) := 100*ifelse(get(paste("EAD", i, sep=".")) == 0, 0,
+                                                           get(paste("RWA","s", i , sep="."))/get(paste("EAD", i , sep="."))),
+           !!(paste("mean","RWA",        i, sep=".")) := mean(get(paste(        "RWA",        i, sep=".")), na.rm = TRUE),
+           !!(paste("mean","RWA", "hat", i, sep=".")) := mean(get(paste(        "RWA", "hat", i, sep=".")), na.rm = TRUE),
+           !!(paste("mean","EAD",        i, sep=".")) := mean(get(paste(        "EAD",        i, sep=".")), na.rm = TRUE),
[…]

### 2019-11-11 Auxiliar/Function.R
Fix IRB formulas to return 0's when PD=0.

@@ -9,27 +9,43 @@
 ##=============================================================================##
 
 mapping_wholesale <- function(PD, LGD = 0.45, M = 2.5, alpha = 0.999){
+  if(PD==0){
+    R  <- NA
+    K  <- 0
+    RW <- 0
+  } else{
   LGD <- ifelse(is.na(LGD), 0.45, LGD)
   R   <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
   b   <- (0.11852-0.05478*log(PD))^2
-  f   <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
-  K   <- f*((1+(M-2.5)*b)/(1-1.5*b))
+  K   <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1+(M-2.5)*b)/(1-1.5*b))
   RW  <- K*12.5
-  data <- data.frame(K,RW,PD,LGD,M,R,b,f)
+  data <- data.frame(K,RW,PD,LGD,M,R)
   return(data)
+}
 }
 
 mapping_retail <- function(PD, LGD, R, alpha = 0.999){
+  if(PD==0){
+    K  <- 0
+    RW <- 0
+  } else{
   K  <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
   RW <- K*12.5
   data <- data.frame(K,RW,PD,LGD,R)
   return(data)
+  }
 }
 
 mapping_other_retail <- function(PD, LGD, alpha = 0.999){
+  if(PD==0){
+    R  <- NA
+    K  <- 0
+    RW <- 0
+  } else{
   R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
   K  <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
   RW <- K*12.5
   data <- data.frame(K,RW,PD,LGD,R)
   return(data)
-}  
+  }  
+}

### 2019-11-11 Data/Import.R
Add Bank Regulation and Supervision dataset plus small changes to Import.R

@@ -36,28 +36,51 @@ EBA.indicator <- read_xlsx("Data/Raw/Pillar3/Auxiliar/Auxiliar.xlsx",sheet = 3) 
   select(bvdid, EBAbank, EBAcountry, Shortfall.mn, Shortfall.RW)
 
 # Macro variables
-
 country.code <- read.csv("Data/Raw/WDI/WDICountry.csv") %>%
   rename_all(tolower)%>%
   select(country.code = ï..country.code,
          Country = x2.alpha.code)
 
 WDI <- read.csv("Data/Raw/WDI/WDIData.csv") %>%
-  filter(Indicator.Code %in% c("NY.GDP.DEFL.ZS","NY.GDP.PCAP.KD")) %>%
+  filter(Indicator.Code %in% c("NY.GDP.DEFL.ZS","NY.GDP.PCAP.KD","NY.GDP.PCAP.KN", "PA.NUS.FCRF", "PX.REX.REER")) %>%
   rename_all(funs(str_replace_all(.,"([\\w])([0-9]+)$", "\\1\\.\\2"))) %>%
   reshape(varying   = c(grep("X.", names(.))),
                       direction = 'long', 
                       timevar   = 'year')%>%
-  select(Country.Code,Indicator.Name,X,year) %>%
+  select(Country.Code,Indicator.Code,X,year) %>%
   reshape(direction = 'wide',
           idvar     = c('Country.Code','year'), 
-          timevar   = 'Indicator.Name') %>%
+          timevar   = 'Indicator.Code') %>%
   rename_all(tolower) %>%
   inner_join(country.code)%>%
-  select(deflator = `x.gdp deflator (base year varies by country)`,
-         gdppc    = `x.gdp per capita (constant 2010 us$)`,
+  select(deflator = x.ny.gdp.defl.zs,
+         gdppc.us = x.ny.gdp.pcap.kd,
+         gdppc    = x.ny.gdp.pcap.kn,
+         er       = x.pa.nus.fcrf,
+         reer     = x.px.rex.reer,
          year,
          Country)
+
+# Bank Regulation and Supervision
+country.code <- read.csv("Data/Raw/WDI/WDICountry.csv") %>%
+  rename_all(tolower)%>%
+  select(country.name = short.name,
+         Country      = x2.alpha.code)
+
+BRSS <- read_xls("Data/Raw/BRSS/BCL_Sup_Reg_Data_13JAN2013.xls",sheet = 4) %>%
[…]

### 2019-11-11 Data/Clean.R
Add several controls, moment calculation, RW decomposition without Capital Savings plus small changes to clean code

@@ -12,15 +12,16 @@ rm(list = ls())
 library(tidyverse)                ## Data manipulation, pipe operator
 library(zoo)                      ## Apply function to rolling margins of data
 library(DescTools)                ## Command to winsorize variables
-library(readxl)       ## Command to open xlsx files                                                     
-
+library(readxl)                   ## Command to open xlsx files                                                     
+library(Hmisc)                    ## Weighted variance and mean functions
+library(moments)                  ## Moments functions
+library(splitstackshape)          ## Command to expand data
 source("Auxiliar/Function.R")     ## Capital requirements functions
 
 # Open dirt data frame
 load("Data/Temp/BankScope.Rda")
 load("Data/Temp/Pillar3Data.Rda")
-
-
+load("Data/Temp/AuxiliarData.Rda")
 
 ##============================================================================##
 ## Pillar-III reports                                                         ##
@@ -30,36 +31,15 @@ load("Data/Temp/Pillar3Data.Rda")
 # Basic cleaning and data creation of raw datasets                             #
 #------------------------------------------------------------------------------#
 
-pillar3.data <- as.data.frame(pillar3.data) %>%
-  mutate_at(vars(EAD:PD),as.numeric)%>%
-  rename(Portfolio_1 = `Portfolio - level 1`,
-         Portfolio_2 = `Portfolio - level 2`) %>%
-  mutate(month       = as.numeric(sapply(Date, str_sub, start= 4, end=5)),
-         year        = as.numeric(sapply(Date, str_sub, start= -4)),
-         year        = ifelse(is.na(month), year, ifelse(month<6, year-1, year)),
-         bvdid       = as.factor(bvdid),
-         Method      = as.factor(Method),
[…]

### 2019-11-11 Analysis/Draft/Draft.Rmd
Paper version before presentation and just after adding the R-squared decomposiont of RW components

--- 2019-08-14 Analysis/Draft/Draft.Rmd
@@
 ---
-title: "Distribution of portfolio credit risk and capital savings"
+title: "Credit risk distribution and capital savings"
 author:
-- affiliation: Tilburg University; l.avezum@uvt.nl
-  name: Lucas Avezum
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+- name: Lucas Avezum
+  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
+date: 
+header-includes:
+   - \usepackage{rotating, graphicx, dcolumn, longtable}
 output:
@@
     template: tex-ms.tex
+    fig_caption: yes
   html_document:
@@
   word_document: default
-bibliography: null
-citecolor: null
+bibliography: ../../Auxiliar/mybibfile.bib
+citecolor: black
 anonymous: no
-fontsize: 12pt
+fontsize: 11pt
 geometry: margin= 1in
-keywords: Key1, Key2, Key3
-linkcolor: blue
+keywords: Internal ratings-based models, Risk-weighted assets dispersion, Regulatory arbitrage
+JEL: G21, G28, G11
+linkcolor: black
 biblio-style: apsr
@@
 thanks: Very preliminary draft
-#abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
-#  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
-#  a concave mapping from probabilities of default to capital requiremets. As a consequence,
-#  banks can reduce requirements or increase risk-taking by widening their portfolio
-#  credit risk distribution.'
+abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requiremets for banks adopting the internal ratings-based (IRB) approach.
+As a consequence, total capital requirement depends not only on the average PD but on the shape of the PD distribution.
+Using a novel hand-collected dataset, I propose a measure that disantangles the contribution of the PD distribution shape from the average PD effect to capital requirements.
+I document that the latter is an important factor of risk-weight variability across banks.
+I further investigate the prior ambiguous direction of its relationship with the business cycle and find it to be countercyclical.
+Altogether, my results show that issues related to RW manipulation and procyclicality of the IRB framework are not as severe as the literature suggests since the effect of the PD distribution shape was not taken into account."
 toc: no
@@
 library(tidyverse)            ## Data manipulation, pipe operator
-library(splitstackshape)     ## Command to expand data
 library(data.table)           ## Command to bind lists
@@
 library(lfe)                  ## High-dimensional fixed effects
-library(gridExtra)           ## Plot side-by-side
+library(gridExtra)            ## Create grid for plots and tables
+library(cowplot)              ## Create grid for plots and tables
 library(DescTools)            ## Command to winsorize variables
-library(robustHD)
+library(ggpubr)               ## More plots
+#library(relaimpo)            ## Calculate relative importance measures
 
-p_w <- c(0.01, 0.99)         ## Set winsorizing limits
+source("../../Auxiliar/Function.R") ## Capital requirements functions
+
 ```
@@
 load("../../Data/Datasets/yeardecomposition.Rda")
-load("../../Data/Datasets/DifInDif.Rda")
 ```
 
-```{r change data, eval= FALSE, include=FALSE}
-# Expand Pillar-III data frame to plot densities and lorenz curves
-expanded.list <- list()
-for(i in c("EAD","EL")) {
-  expanded.list[[i]] <- pillar3.data %>%
-  filter(Method2 %in% c("IRB")) %>%
-  select(name,Country,Method,Portfolio_1,EAD,LGD,PD,year)%>%
-  mutate(EL  = round(EAD*LGD/1000),
-         EAD = round(EAD/1000))%>%
-    na.omit(cols=c(i))%>%
-    expandRows(i) %>%
-    mutate(freq_type = as.factor(i))
-}
-pillar3.expanded <-  rbindlist(expanded.list, use.names=TRUE,fill = TRUE)
+# Introduction
 
-# Aggregate Pillar-III data by Portfolio
-portfolio.data <- pillar3.data %>% 
-  group_by(Country,Bank,year,Portfolio_1,Portfolio_2) %>%
-  mutate(weight=EAD*LGD)%>%
-  filter(weight>0, Portfolio_1 %in% c("Wholesale","Retail"))%>%
-  dplyr::summarize(gini             = DescTools::Gini(PD,weight),
-                   K_calculated     = sum(K_calculated),
-                   K_counterfactual = sum(K_counterfactual)) %>%
-  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)
+Bank capital adequacy requirements have evolved substancially since the first international standard, the 1988 Basel Accord.
+Arguably, at the center of these changes is the link of capital charges to asset risk [@basel2004international; @BCBS2017basel; @EBA2019policy].
+Risk sensitivity of capital requirements was present in the first accord and it was greatly enhanced with the Basel II framework by giving banks the option to calculate capital requirements using risk estimates obtained from their own models.
+The rationale for risk-linked capital charges is that regulating capital via a simple capital to asset ratio can incentivize banks to take on risk [@koehn1980regulation; @kim1988risk; @gordy2010risk].
+The rationale for allowing internal models are the better alignment of risk estimates to actual portfolio risk [@barakova2014banks] and the promotion of better risk management practices [@cucinelli2018credit].
+However, these same features may have adverse consequences.
+First, linking capital charges to asset risk exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
+Second, the inherent flexibility of the internal ratings-based (IRB) approach may allow for differences in capital requirements that do not reflect portfolio risk [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits; @ferribank].
+This paper contributes to the understanding of cyclicality and unintended variability of capital requirements by exploring the consequences of a feature of the IRB framework, which, to the best of my knowledge, I am the first to identify.
 
-bank.data <- bank.data %>%
-group_by(bvdid) %>% 
-  mutate(savings_log  = log(K_counterfactual-K_calculated),
-         Total.assets = Total.assets/10^6,
-         Total.Capital= Total.Capital/10^6,
-         Tier.1.Capital=Tier.1.Capital/10^6,
-         asset_log    = log(Total.assets),
-         capital_log  = log(Total.Capital),
-         tier.1_log   = log(Tier.1.Capital),
-         savings_rate = (K_counterfactual-K_calculated)/K_calculated,
-         savings_alt  = (K_counterfactual-K_original)/K_original,
-         error        = (K_calculated - K_original)/K_original,
-         asset_d      = Winsorize(log(Total.assets)-lag(log(Total.assets),order_by=year), 
-                                  probs = p_w ,na.rm = TRUE),
-         capital_d    = Winsorize(log(Total.Capital)-lag(log(Total.Capital),order_by=year), 
-                                  probs = p_w ,na.rm = TRUE),
-         PD_mean_d    = Winsorize(PD_mean-lag(PD_mean,order_by=year), 
-                                  probs = p_w ,na.rm = TRUE),
-         savings_d    = Winsorize(savings_log-lag(savings_log,order_by=year), 
-                                  probs = p_w ,na.rm = TRUE),
-         car_d  = Winsorize(Total.Capital.Ratio/100-lag(Total.Capital.Ratio/100,order_by=year), 
-                            probs = p_w ,na.rm = TRUE),
-         roe_d  = Winsorize(ROE.using.P.L.before.tax..../100-lag(ROE.using.P.L.before.tax..../100,order_by=year),
-                            probs = p_w ,na.rm = TRUE)) %>%
-  ungroup()
+Under the IRB approach banks supply their own risk estimates to regulatory formulas which determine the amount of capital banks have to hold for each asset.
+Importantly, the mapping from probability of default (PD) and capital requirements is concave and capital requirements are determined on the asset level with total requirement being simply the sum of assets' individual requirement.
+This means, by Jensen’s inequality, that the sum of capital charges for two assets is lower than the charge for their average (in terms of PD).
+More generally, total capital requirement depends not only on banks' portfolio average PD but also on the shape of the PD distribution.
+For instance, given an average PD, more disperse portfolios yield lower capital charges.
+I call the difference in capital requirements stemming from differences in the shape of the portfolio PD distribution, while holding the location parameter constant, the PD distribution effect.
 
-test <- cross.section.decomposition %>% mutate(mean.total.gain = mean(total.gain,na.rm = TRUE)) %>%
-  mutate_at(vars(SA.gain:total.gain),  funs((. - mean.total.gain)^2))%>%
-summarize_at(vars(SA.gain:total.gain),  sum, na.rm =  TRUE)%>%
-  mutate_all( funs((. / total.gain)))
+The first contribution of this paper is to introduce a dataset and method to quantify the PD distribution effect.
+The method is straightforward: compare the actual capital requirement, which depend on the PD distribution, to a counterfactual case where the requirement for the entire portfolio is calculated using the average PD.
+*Capital savings* is the difference between the two values and it measures the intensity of the PD distribution effect.
+
+To calculate the *Capital savings* measure, one needs information on the distribution of banks' regulatory risk parameters.
+However, data at such granularity is not available in the common balance sheet datasets, and although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
+Therefore, to overcome data limitations, I hand-collected information on the distribution of credit risk parameters from banks' Pillar-3 reports.
+The sample consists of 49 large banking groups that have adopted the IRB approach.
+The dataset covers 10 countries and goes from 2007 to 2018.
+
+With this unique dataset, I start the analysis by investigating the contribution of the PD distribution effect to banks' overall risk-weight (RW).
+Several studies document significant RW variability across otherwise similar banks.
+Some take this evidence as indicative of either excessive subjectivity in the current capital requirement rules or risk measurement manipulation by banks to decrease capital charges [see, @mariathasan2014manipulation; @ferribank].
+Others are more cautious and emphasize that part of RW variability is desirable since it reflects differences in risk among banks [@cannata2012inside; @BCBS2016regulatory; @EBA2019result].
+I contribute to the literature by showing that the PD distribution effect is an important factor of the RW variation across banks.
+By extending the decomposition method proposed by @cannata2012inside I find that the PD distribution effect accounts for 28% of RW variation across banks.
+Moreover, the inclusion of the PD distribution effect in the decomposition reduces the variation explained by the component most susceptible of manipulation by 17.7 percentages points.
+This result implies that previous studies in the literature are potentially overestimating subjectivity and manipulation of RWs.
+
+I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
+A consequence of the shape of the regulatory formula is that it affect the cyclicality of capital requirements.
+@behn2016procyclical document that IRB portfolios are more procyclical than portfolios under the standardized approach.
+During economic downturns credit risk increases which raises capital requirements under the IRB approach.
+This characterizes the IRB approach as procyclical since banks have to cut down lending exactly when borrowers need it the most.
+We can interpret this evidence as the result of shifts to the location parameter of the PD distribution.
+However, shocks to credit risk can also affect the shape of the distribution.
+For instance, riskier borrowers prior to the shock can be more adversely affected compared to safer borrowers causing the PD distribution to become more fat-tailed.
+If that is the case, the amount of *Capital savings* will be affected.
+
+I find the PD distribution effect to be countercyclical.
+I estimate that a 2.1% GDP growth rate (one standard deviation) decreases *Capital savings* by 5.9% on average.
+In other words, banks have to increase capital by 5.9% to keep their capital ratio constant.
+The effect is economically significant as it represents a quarter of the *Capital savings* standard deviation in the sample.
+I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases.
+This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
+
+Several cross‐sectional tests further strengthen our results.
+Increases in capital requirements are more problematic for less well capitalized banks, which consequently adjust their IRB loans more in response to the credit risk shock.
+This suggests that our findings would be less pronounced if banks kept larger buffers over the regulatory minimum.
+Furthermore, the IRB effect is more pronounced for less profitable firms, which are likely to be more affected by the credit risk shock, and for firms to which IRB banks have a large exposure.
+As robustness, I calculate the gini coefficient for the PD distribution as an alternative measure of the PD distribution effect and I find similar results.
+
+The remainder of this paper is organized as follows.
+Section 2 describes the capital requirement formula, the capital savings measure, and the hypothesis concerning this measure.
+Section 3 describes the data collected, and analyses the relative importance of capital savings in explaining differences in risk-weights across banks and time.
+Section 4 explains the empirical strategy.
+Section 5 presents evidence on the anticyclical feature of capital savings while Section 6 shows evidence of banks using this feature strategically to reduce capital charges.
+Finally, section 7 concludes.
+
+# Institutional Background
+
+## Capital Requirements under the IRB approach
+
+In June 2004, the Basel Committee issued a revised framework on international convergence of capital measurements and capital standards [@basel2004international], which serves as the basis for national rulemaking and implementation processes.
+Since then, financial institutions may choose between two approaches to calculate the capital requirement for credit risk; the standardized approach (essentially a slightly modified version of the first accord) and the IRB approach.
+
+In the IRB approach, capital requirements are derived from risk-weight formulas that take as input banks' own estimates of credit risk drivers.
+The basic regulatory formula derives from a model characterized by the property that the risk-weight of each asset depends only on the estimated credit risk parameters for this asset and not on the composition of the portfolio, i.e.
+the model is portfolio invariant.
+This feature leads to a bottom-up approach where capital requirements are determined on the asset level and total requirement is simply the sum of asset individual requirement.
+
+Equation \ref{eq1} is the basic capital requirement formulas:
+
+\begin{equation}
+\label{eq1}
+\text{K}= \sum_{i}EAD_iLGD_i N\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(PD_i)+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg) ,
+\end{equation}
+
+where the summation is over the assets and $N$ is the standard normal cumulative distribution.
+Basel accords segment credit portfolio into several categories, each having different additional features (maturity adjustments, risk-dependent correlation coefficient, among others) but without altering the basic structure shown in equation \ref{eq1}.
+The credit risk parameters estimated by banks, which are used as input on equation \ref{eq1}, are: Exposure at default (EAD), loss given default (LGD), and probability of default (PD).
+
+## Probability of Default Distribution and Capital Savings
+
+For the purpose of this paper, I focus on the mapping from PD to capital requirements.
+Figure \ref{fig:pdmap} shows that capital requirements is a concave function of PD.
+This means, by Jensen’s inequality, that the sum of capital charges for two assets is lower than the charge for their average (in terms of PD).
+More generally, the distribution of credit risk matters to calculate capital requirements with, for instance, disperse portfolios yielding relatively lower capital charges.
+
+```{r parameters and data, echo=FALSE, include=FALSE}
+set.seed(123)
+LGD <- 1
+M   <- 2.5
+grid  <- seq(0.0001, 1, by=0.00001) 
+data <- mapping_retail(grid,1,0.15,0.999)
 ```
 
-```{r}
+```{r, results="hide"}
+pd_map_cap <-  "\\label{fig:pdmap}The figure plots the mapping from probabilities of default to capital requirementes per unit of exposure according to equation \\ref{eq1}.
+$EAD$ and $LGD$ are set equal to one."
 
-bank.data <- bank.data %>% ungroup() %>%
-  mutate(RW.hat=RW.hat*100,
-         RW.s=RW.s*100)
 ```
 
-# List
-* Use the share under IRB to weight capital savings measure (share vs intensity). Include share in regressions plus interaction with capital savings
+```{r, fig.height = 3, fig.width = 4,results="hide", fig.cap = pd_map_cap, fig.align = 'center', echo=FALSE, message=FALSE}
+p <- ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K)) 
+p + geom_line(alpha = 1) + 
+  labs(x = "PD", 
+         y = "Capital requirement per unit of exposure") +
+  theme_bw()
 
-* Explore alternative capital savings measures, such as, savings in terms of capital (mn),in terms of capital ratio, in terms of capital ratio savings (counterfactual vs actual)
+```
 
-* Plot marginal increase of savings in time.
-
-* Decompose capital savings: share effect, size effect, etc...
-
-* Remove step effect from regressions: average across all IRB banks (add IRB dummy) or IRB specific dummy (average out the step for each bank).
-Idea is to focus on difference across IRB banks.
-it might not make sense to include SA banks anymore.
-
-* Plot capital savings measures to other risk variables
-
-* Plot cross section for different capital saving measures (replicate first plot)
-
-* check and better clean data:
-
-# Capital savings measure
-
-In this section I describe the chosen method to measure differences in risk distributions:
+The first contribution of this paper is a measure of capital requirement reduction due to the concavity of the regulatory formula.
+The method behind the measure is simple: compare actual capital requirement to a counterfactual case where the requirement for the entire portfolio is calculated using the average PD.
+In practice, I calculate differences in risk-weighted assets (RWA) rather than capital requirements because it is the former that is used to calculate capital ratios.
+I define the counterfactual risk-weighted assets as:
 
 \begin{equation}
-\text{Capital savings}_{i}= \frac{f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i)}{\sum_{i}EAD_i},
+\label{eq2}
+\text{RWA}^c= 12.5\sum_{i}EAD_iLGD_iN\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\overline{PD})+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg),
 \end{equation}
 
-where $i$ is one asset of the portfolio, $f()$ is the regulatory formula and,
+where the probability of default parameter for all assets is replaced by the portfolio weighted average:
+
+\begin{equation*}
+\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
+\end{equation*}
+
+The *Capital savings* measure is the difference between equation \ref{eq1} (multiplied by 12.5 to obtain RWA) and equation \ref{eq2}:
 
 \begin{equation}
-\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
+\label{eq3}
+\text{RWA}^s=\text{RWA}^c-\text{RWA}.
 \end{equation}
 
+*Capital savings* ($\text{RWA}^s$) is the amount that banks saves in capital in monetary units.
+A value of zero for this measures means that there is no capital savings, and implies a portfolio with a degenerate PD distribution.
+Any PD distribution with positive variance will yield a positive value of *Capital savings*.
+To avoid size effects and dealing with currency conversion I also consider a second measure, the *Capital savings ratio*, defined as follows:
+
+\begin{equation}
+\label{eq4}
+\text{RWA}^r=\frac{\text{RWA}^c}{\text{RWA}}.
+\end{equation}
+
+*Capital savings ratio* ($\text{RWA}^r$) is the rate at which a bank saves capital.
+For instance, a value of 0.5 means this portfolio would require 50% more capital in the abscense of the PD distribution effect to keep the same capital ratio.
+As with *Capital savings*, a value of zero means no capital savings, and any PD distribution with positive variance will result in a positive value.
+
+## Capital Savings and Cyclicality
+
+A consequence of the shape of the regulatory formula is that it affect the cyclicality of capital requirements.
+@behn2016procyclical document that IRB portfolios are more procyclical than portfolios under the standardized approach.
+During economic downturns credit risk increases which raises capital requirements under the IRB approach.
+This characterizes the IRB approach as procyclical since banks have to cut down lending exactly when borrowers need it the most.
+We can interpret this mechanism as shifts to the location parameter of the PD distribution.
+However, shocks to credit risk can also affect the shape of the distribution.
+For instance, riskier borrowers prior to the shock can be more adversely affected compared to safer borrowers, causing the PD distribution to become more fat-tailed.
+As a result, the amount of capital savings will be affected.
+
+Whether capital savings are pro- or anticyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the credit risk distribution.
+I use a simple example to illustrate this ambiguity.
+Consider a bank with the asset side of its balance sheet composed of \$100 of a safer asset and \$100 of a riskier asset and three states of the economy.
+Without loss of generality, assume $LGD=1$ and $R=0.15$.
+In the initial state, safer and riskier assets have a 0.1% and 0.2% probability of default, respectively.
+According to equation \ref{eq1}, the actual RWA of this bank is \$63.9.
+The weighted average of this portfolio in terms of $PD$ is 0.15% which yields a counterfactual RWA of \$64.7 according to equation \ref{eq2}.
+Hence, in the initial state capital savings equals \$0.8.
+
+Consider now a credit risk shock.
+In state A, both safer and riskier assets are affected, having their probabilities of default raised to 0.2% and 0.3%, respectively.
+In state B, only the riskier asset is affected, now with a probability of default equal to 0.4%.
+Note that the average $PD$ in both states are equal at 0.25%.
+However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B.
+The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%).
+Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states.
+This is the result of IRB procyclicality shown by @behn2016procyclical.
+However, due to the increased portfolio dispersion, capital savings are anticyclical in state B.
+Conversely, while higher-order moments remained constant in state A, the PD distribution moved to a less concave part of the regulatory function consequently reducing capital savings.
+Therefore, capital savings are procyclical in state A.
+Table \ref{tab1} summarizes the example.
+
+\begin{table}[!htbp] \centering 
+  \caption{Cyclicality of Capital Savings} 
+  \label{tab1} 
+\small 
+\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}} 
+\hline 
+\hline \\[-1.8ex] 
+\\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6}
+\\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\ 
+\hline \\[-1.8ex] 
+ $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\ 
+  $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\ 
+  $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\ 
+  \hline \\[-1.8ex]
+ Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\ 
+  S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\ 
+\hline \\[-1.8ex]
+
+\end{tabular} 
+\end{table}
+
+\newpage
+
+# Data
+
+To calculate equation the proposed measures of capital savings, one needs information on the distribution of banks’ regulatory risk parameters.
+However, data at such granularity is not available in the common balance sheet datasets.
+Although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
+Therefore, to overcome data limitations I collected information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports.
+The sample consists of 50 large banking groups, from 12 countries, that have adopted the IRB approach.
+For each bank, I collected consolidated information from the year the IRB approach was approved until 2018.
+The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate and sovereign for wholesale, and real estate and qualifying revolving credit for retail).
+This level of portfolio breakdown is important because regulatory formulas vary across categories, and consequently, more precise knowledge of portfolio composition reduces measurements errors.
+
+Figure \ref{fig:savingsbank} shows that the *Capital savings ratio* calculated with the compiled dataset.
+Each point is a bank-year observations and the bars indicates the average value across time for each bank.
+First thing to note is that there is substancial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
+Second, variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
+Lastly, *Capital savings ratio* also varies between and within countries.
+This is better shown by figure \ref{fig:savingscountry} which aggregate the dataset by country.
+In the figure, each point is a bank observation and the bars are the average of these points.
+Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).
+
 ```{r}
-savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
+savings_cap = "\\label{fig:savingsbank}The figure plots the log of banks capital savings.
+This measure is calculated using equation \\ref{eq3}.
+Each point in the plot is a bank-year observation and bars are the averages values for each bank."
 ```
 
-```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
-bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
- ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
-                          y = RW.s,
-                      color= Country, 
-                      fill = Country)) +
+```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
+bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
+  select(name, Country, RWA.r)%>%
+  group_by(name, Country) %>%
+#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
+  group_by(Country) %>%
+ # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
+  #ungroup()%>%
+  #mutate(ordering = order_aux+RWA.r,
+  #       name = as.factor(name),
+  #       name = fct_reorder(name, ordering),
+  #       Country = fct_reorder(Country, ordering)) %>%
+ ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
+                      y = RWA.r)) +
 geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
     labs(x=NULL) +
     coord_flip() +
-  labs(y =  "Capital savings") +
-
+  labs(y =  "RWA savings ratio") +
+  scale_fill_grey()+
     theme_bw()
@@
 ```{r}
-fig3_cap = "PD distribution by bank. The figure shows smoothed PD distributions for each bank in our data. Banks are ranked by median PD and colored by host-country"
+savings_cap_country = "\\label{fig:savingscountry}The figure plots the banks capital savings ratio.
+This measure is calculated using equation \\ref{eq4}.
+For each bank I used the average across time such that each point in the plot is a bank observation and bars are the averages values for each country."
 ```
 
-```{r, fig.height = 10, fig.width = 7, eval=FALSE, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
-pillar3.expanded %>% 
-  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
-ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
-  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
-                       scale = 2 , from = 0, to = 0.05) +
-  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
-  scale_y_discrete(expand = c(0.01, 0)) +
-  labs(
-    x = 'Probability of default') +
-  theme_ridges(font_size = 13, grid = TRUE) + 
-  theme(axis.title.y = element_blank())
-```
-
-# Decomposition
-
-\begin{equation}
-\begin{aligned}
-\Delta RW & = q_{B}^{SA} \Delta RW^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \hat{RW}^{j} \\ 
-& + (1-q_{B}^{SA})\sum_{j=1}^{P} \hat{RW}^{j} \Delta q^{j} -(1-q_{B}^{SA})\Delta RW^{S} + (RW^{SA}-RW^{IRB})\Delta q^{SA},
-\end{aligned}
-\end{equation}
-
-\newpage
-## Time decomposition
-
-```{r}
-time.dec_cap = "Time decompostion.
-The figure shows the contribution of each component compared to the time of IRB introduction (t=0).
-The figure and calculations include only banks with observations for at least 5 consecutive years.
-The black line is the net change.
-Year decompostion.
-The figure shows the contribution of each component compared to 2008.
-The figure and calculations include all banks in the sample.
-The black line is the net change."
-```
-
-```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
-time.lable <- c("t=0", "t=1", "t=2", "t=3", "t=4")
-plot1 <-time.decomposition %>% 
-  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain"),
-         RW = RW - t0.RW,
-         time = as.factor(time)) %>%
-  ggplot(mapping = aes(x = time,
-                       y    = value, 
-                       fill = condition)) +
-geom_bar(stat = 'identity')+
-geom_line(aes(group=1, y = RW))  +
-  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
-  scale_x_discrete( name=NULL,breaks = c("0","1","2","3","4"),labels = c("t=0","t=1","t=2","t=3","t=4")) +
- theme(legend.position = c(0.3, 0.2), 
-        legend.title = element_blank(), 
-        legend.background=element_blank())
-
-year.decomposition <- year.decomposition %>%
-  mutate(RW_2 = RW - t0.RW,
-         year = as.factor(year)) 
-plot2 <- year.decomposition %>% 
-  select(year,RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = year)) +
-geom_bar(aes(y = value, 
-          fill = condition),
-          stat = 'identity')+
-geom_line(data = year.decomposition, 
-          aes(group=1,y = RW_2))  +
-  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
-  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
-theme(legend.position='none')
-grid.arrange(plot1, plot2, ncol=2)
+```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
+bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
+  select(Country,name, year, RWA.r)%>%
+  group_by(Country,name) %>%
+ summarise_if(is.numeric,mean,na.rm= TRUE)%>%
+ ggplot(mapping = aes(x = reorder(Country, RWA.r, FUN = mean, na.rm=TRUE),
+                          y = RWA.r)) +
+geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+    labs(x=NULL) +
+  labs(y =  "RWA savings ratio") +
+  coord_flip() +
+  theme_bw()
 
 ```
-
-\newpage
-
-## Cross-section decomposition
-
-```{r}
-cross.dec_cap = "Cross-section decompostion.
-The figure shows the contribution of each component compared to the average bank.
-The figure and calculations include all banks in the sample.
-The black line is the total RW.
-Relative cross-section decompostion.
-The figure shows the relative contribution of each component compared to the average bank.
-The figure and calculations include all banks in the sample.
-"
-```
-
-```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
-plot1 <- cross.section.decomposition %>% 
-  mutate(order.s = abs(dispersion.gain))%>%
-  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = reorder(name, order.s),
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity')+
-  geom_line(aes(y = total,group=1))  +
-  labs(x=NULL, y= NULL,subtitle = "Plot a") +
-    coord_flip() +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
- theme(legend.position='none')
-
-plot2 <- cross.section.decomposition %>% 
-  mutate(order.s = abs(dispersion.gain))%>%
-  select(name, Country, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(value = abs(value),
-         condition = relevel(condition,"dispersion.gain")) %>%
-  ggplot(mapping = aes(x = reorder(name, order.s),
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity',position = "fill")+
-  labs(x=NULL, y= NULL,subtitle = "Plot b") +
-    coord_flip() +
-  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
-   theme(legend.title = element_blank(), 
-        legend.background=element_blank(),
-       legend.text=element_text(size=7),
-       axis.text.y = element_blank())
-
-grid.arrange(plot1, plot2, ncol=2)
-
-```
-
-# Counterfactual analysis
 
@@
 
-```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
- plot1 <- bank.data %>%  filter(!is.na(RW.s),
+```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth", include=FALSE}
+ plot1 <- bank.data %>%  ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
          tier1capitalmillcu < totalcapitalmillcu)%>% 
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
-  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>% 
+  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, tier, tier.hat, IRB) %>%
+  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>% 
  ggplot(mapping = aes(x = RW.s,
@@
 
-plot2 <- bank.data %>%  filter(!is.na(RW.s),
+plot2 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, tier, tier.hat, IRB) %>%
   filter(!is.na(RW.s), IRB == 1) %>%
@@
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>%
+  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>%
  ggplot(mapping = aes(x = RW.s,
@@
 
-plot3 <- bank.data %>%  filter(!is.na(RW.s),
+plot3 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
-  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, CAR, CAR.hat, IRB) %>%
+  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
  ggplot(mapping = aes(x = RW.s,
@@
 
-plot4 <- bank.data %>% filter(!is.na(RW.s),
+plot4 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
          !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
+         tier1capitalmillcu < totalcapitalmillcu)%>% 
+  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, CAR, CAR.hat, IRB) %>%
   filter(!is.na(RW.s), IRB == 1) %>%
@@
  summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
+  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
  ggplot(mapping = aes(x = RW.s,
@@
 
+\newpage
+
+## Decomposition
+
+In this section I extend the method proposed by @cannata2012inside to decompose RW in its components.
+To get capital savins in terms of RWs, we only need to divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
+Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
+We can further breakdown each RW into their portfolio components.
+I choose to do so only for the counterfactual RW ($\text{RW}^c$), such that I can account for the contribution of the IRB portfolio mix while letting the components associated with capital savings as minimum as possible.
+
+\begin{equation}
+\begin{aligned}
+\label{eq5}
+\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j=1}^{P} \text{RW}_{j}^{c} \Delta q^{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
+&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin}
+\end{aligned}
+\end{equation}
+
+```{r}
+cross.dec_cap = "\\label{fig:crossdecomposition} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+The figure and calculations include all banks in the sample.
+The black line is the total RW relative to the average total RW.
+Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
+Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
+```
+
+```{r old_cross_decomposition, eval = FALSE, fig.height = 6.5, fig.width = 6, results="hide",fig.cap= cross.dec_cap, include=TRUE}
+ cross.section.decomposition %>% filter(q.SA<1)%>%
+  select(name, Country, total= total.gain, SA.gain,IRB.gain,mix.gain, rollout.gain, int.gain, ext.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
+  ggplot(mapping = aes(x = reorder(name, total),
+                       y = value, fill = condition)) +
+geom_bar(stat = 'identity')+
+  geom_line(aes(y = total,group=1))  +
+  labs(x=NULL, y= NULL) +
+    coord_flip() +theme_bw()+
+  scale_fill_grey(na.value = "lightgrey",start = 0.1, end = 0.5,
+                  labels = c("Intensive margin", "Extensive margin", "Other components"), name = "Component")
+
+```
+
+```{r}
+cross.dec_cap2 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+The figure and calculations include all banks in the sample.
+The first plot shows the total RW for each bank relative to the average total RW.
+Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
+Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
+```
+
+```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
+ plot1 <- cross.section.decomposition %>% 
+  filter(q.SA<1)%>%
+  mutate(total= -PD.gain)%>%
+  select(name, Country,total, int.gain, ext.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
+                                       labels = c("Intensive margin", "Extensive margin")))%>%
+ggplot(mapping = aes(x = reorder(name, total),
+                       y = value)) +
+geom_bar(stat = 'identity')+
+  #geom_line(aes(y = -total,group=1))  +
+  labs(x=NULL, y= NULL) +
+    coord_flip() +
+  theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  facet_wrap(~condition)
+
+plot2 <- cross.section.decomposition %>% 
+  filter(q.SA<1)%>%
+  select(name, Country, int.gain, ext.gain) %>%
+  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
+  ggplot(mapping = aes(x = int.gain,
+                       y = ext.gain)) +
+    geom_point()+
+    geom_smooth(method = "lm", color = "black")  +
+  labs(x=NULL, y= NULL) +
+   theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  scale_fill_grey()+
+  stat_cor()
+
+  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
+            ncol = 1, nrow = 2, align = "v", axis = "lr")
+```
+```{r}
+cross.dec_cap3 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+The figure and calculations include all banks in the sample.
+The first plot shows the total RW for each bank relative to the average total RW.
+Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
+Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
+```
+
+```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap3, include=TRUE}
+plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
+  mutate(total= -PD.gain)%>%
+  select(name, Country,total, other.gain, PD.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
+                                       labels = c("Capital savings", "Other components")))%>%
+ggplot(mapping = aes(x = reorder(name, total),
+                       y = value)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL, y= NULL) +
+    coord_flip() +theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  facet_wrap(~condition)
+
+plot2 <- cross.section.decomposition %>% 
+  filter(q.SA<1)%>%
+  select(name, Country, other.gain, PD.gain) %>%
+  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
+  ggplot(mapping = aes(x = PD.gain,
+                       y = other.gain)) +
+    geom_point()+
+    geom_smooth(method = "lm", color = "black")  +
+  labs(x=NULL, y= NULL) +
+   theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  scale_fill_grey()+
+  stat_cor(label.y= -0.1)
+
+  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
+            ncol = 1, nrow = 2, align = "v", axis = "lr")
+```
+
+\newpage
+```{r}
+r_sqrt_decomp  = "\\label{fig:rsqrtdecomp} The figure shows the $R^2$ share for each of the RW components derived using equation Equation \\ref{eq5}." 
+```
+```{r,fig.height = 5, fig.width = 6, results="hide",fig.cap= r_sqrt_decomp , include=TRUE}
+
+unrestrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
+  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
+         total.gain = total.gain+epsilon)%>% 
+  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
+  relaimpo::calc.relimp(type = c("lmg", "last"), rela = FALSE )
+
+ unrestrict.data <- data.frame(share.lmg  = unrestrict@lmg,
+                               share.last = unrestrict@last,
+                               decomposition = "unrestrict") %>%
+   add_rownames(var = "component")
+
+ restrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
+  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
+         total.gain = total.gain+epsilon) %>%
+  mutate(IRB.gain     = IRB.gain.alt,
+         mix.gain     = mix.gain.alt,
+         rollout.gain = rollout.gain.alt)%>% 
+ lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + SA.gain)%>%
+ relaimpo::calc.relimp(type = c("lmg", "last"), rela = FALSE )
+
+data.frame(share.lmg  = restrict@lmg,
+           share.last = restrict@last,
+           decomposition = "restrict") %>%
+   add_rownames(var = "component")%>%
+    full_join(unrestrict.data)%>%
+  pivot_longer(cols = -c("decomposition","component"),
+               names_to = c("method"),
+               names_pattern = "share.(.*)",
+               values_to = "share")%>%
+  mutate(decomposition = factor(decomposition, 
+                                levels = c("unrestrict", "restrict"),
+                                labels = c("Unrestrict", "Restrict")),
+         method = factor(method, 
+                                levels = c("lmg", "last"),
+                                labels = c("LMG", "Last")),
+         component = factor(component,
+                      levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
+                      labels = c("PD avg.", "PD ext.", "PD int.", "SA", "Mix", "Rollout"))) %>%
+   arrange(decomposition,method)%>%
+ggplot(mapping = aes(x = component,
+                       y = share)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL, y = expression(paste("Share of ", R^{2}))) +
+    theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+ facet_wrap(method~decomposition)+
+  theme(strip.background = element_blank(),
+       # strip.text = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))+
+  scale_fill_grey()+
+scale_y_continuous(labels = scales::percent_format(accuracy = 1,
+                                 decimal.mark = '.'))
+
+```
+
+```{r}
+time.dec_cap = "\\label{fig:timedecomposition}The figure shows the evolution of average RW components in time.
+Plot a shows the average contribution of each RW component relative to the time of IRB introduction (t=0).
+The figure and calculations include only banks with observations for at least 9 consecutive years.
+Plot b shows the average contribution of each component relative to 2008.
+The figure and calculations include all banks in the sample.
+The black line in both plots is the average change in RW relative to the initial period."
+```
+
+```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
+
+plot1 <-time.decomposition %>% 
+  select(time, RW, t0.RW,  int.gain, ext.gain, other.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain")),
+         RW = RW - t0.RW,
+         time = as.factor(time)) %>%
+  ggplot(mapping = aes(x = time,
+                       y    = value, 
+                       fill = condition)) +
+geom_bar(stat = 'identity')+
+geom_line(aes(group=1, y = RW))  +
+  theme_bw()+
+  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
+  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
+ theme(legend.position = c(0.3, 0.2), 
+        legend.title = element_blank(), 
+        legend.background=element_blank())+
+  scale_fill_grey(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")
+
+year.decomposition <- year.decomposition %>%
+  mutate(RW_2 = RW - t0.RW,
+         year = as.factor(year)) 
+plot2 <- year.decomposition %>% 
+  select(year,RW, int.gain, ext.gain, other.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain"))) %>%
+  ggplot(mapping = aes(x = year)) +
+geom_bar(aes(y = value, 
+          fill = condition),
+          stat = 'identity')+
+geom_line(data = year.decomposition, 
+          aes(group=1,y = RW_2))  +
+  theme_bw()+
+  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
+  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
+theme(legend.position='none')+
+  scale_fill_grey()
+grid.arrange(plot1, plot2, ncol=2)
+
+```
+
+```{r}
+time.var_cap = "\\label{fig:timevariables}The figure shows the time series of banks' capital ratio, portfolio average PD of their portfolio, and PD standard deviation of their portfolio.
+The black line is average value across all banks in the sample and the shaded area shows the first and third quartile."
+```
+
+```{r, fig.height = 5, fig.width = 5, results="hide",fig.cap= time.var_cap}
+
+mean_cl_quantile <- function(x, q=c(0.25, 0.75), na.rm=TRUE){
+  dat <- data.frame(y=mean(x, na.rm = TRUE),
+                    ymin = quantile(x, probs = q[1],na.rm = na.rm),
+                    ymax = quantile(x, probs = q[2],na.rm = na.rm))
+  return(dat)
+}
+scaleFUN <- function(x) sprintf("%.1f", x)
+
+plot1 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year, capital.ratio) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=capital.ratio)) +
+geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'Capital Ratio (%)', x=NULL) + scale_y_continuous(labels=scaleFUN)
+
+plot2 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year,  mean.PD) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=mean.PD)) +
+geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+   labs(y = 'Average PD (%)', x=NULL)
+
+plot3 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year, var.PD) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
+geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+   labs(y = 'SD PD (%)', x=NULL)
+
+plot_grid(plot1, plot2, plot3, hjust = 0.5,
+            ncol = 1, nrow = 3, align = "v", axis = "lr")
+```
+
+\newpage
+
+## Control Variables and Summary Statistics
+
+I include a number of bank-specific control variables in the analysis.
+Several works relate bank-specific characteristics to either total RWA or average RW (Vallascas and Hagendorff (2013) ...).
+In this section I use these evidence to construct my priors on how each of these variables might affect capital savings.
+
+The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
+First, I control for bank size using the log of total assets (in millions of US dollars).
+The expectation regarding the impact of bank size is ambiguous.
+On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices in order to reduce capital savings variability.
+On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
+I also include the regulatory capital ratio as control.
+Better capitalized banks could be subject to less regulatory scrutiny (Calem and Rob 1999) and, consequently, in a better position to engage in regulatory arbitrage.
+However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
+Thus, the expected effect of capital ratio on capital savings is also ambiguous.I also control for .
+Profitable banks face fewer incentives to understate the value of risky assets and to engage in regulatory arbitrage more generally.
+
+The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
+For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
+Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
+As a result, banks with higher shares of these variables should have higher values of capital savings.
+Finally, I control for bank funding using the ratio of total deposits to total assets and bank profitability using net income over total assets.
+Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
+Banks profitability might be the result of higher risk-taking.
+The effect of deposits funding and profitability on capital savings depends on how risk-taking is manifested by these variables.
+If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then capital savings decrease.
+Conversely, capital savings increases if banks take risk by having fatter upper-tails in the PD distribution.
+
+Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the control variables used throughout the analysis.
+
 ```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-bank.data %>% ungroup()%>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>%
+p_w <- c(0.05, 0.95)         ## Set winsorizing limits
+bank.data.win <- bank.data %>% 
+  ungroup()%>%
+  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
   mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>% 
-  subset(select = c("RW.s","tier1ratio","tier.hat","totalcapitalratio","CAR.hat","ROE","Zscore")) %>%
+  group_by(bvdid)
+
+bank.data.win %>% select(
+  log.CAR_gr, 
+  log.K_gr, 
+  log.RWA.IRB_gr, 
+  log.RWA.hat_gr, 
+  log.RWA.r_gr, 
+  log.RWA.s_gr, 
+  log.RWA.s.Retail_gr,
+  log.RWA.s.Wholesale_gr, 
+  log.RWA.s.Corporate_gr, 
+  log.RWA.s.Sovereign_gr, 
+  log.RWA.s.Banks_gr, 
+  mean.PD_gr, 
+  var.PD_gr, 
+  skew.PD_gr,
+  kurt.PD_gr, 
+  gini.PD_gr,
+  log.gdp_gr,
+ #log.asset_gr, 
+ #capital.ratio_gr, 
+ #loans.ratio_gr, 
+ #NII.ratio_gr,  
+ #deposit.ratio_gr,
+ #income.ratio_gr,
+  mean.PD, 
+  log.asset, 
+  capital.ratio,
+  loans.ratio, 
+  NII.ratio,  
+  deposit.ratio, 
+  income.ratio) %>%
+  as.data.frame() %>% 
+stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
+          label = "tab:summary",
+          omit.summary.stat = c("p25","p75"),
+          covariate.labels = 
+            c("$\\Delta \\text{Log Capital Ratio}$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log capital}$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}$",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
+              "\\hspace{1.0cm}$\\Delta \\text{Log 1+RWA}^r$",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Retail",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Wholesale",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Corporate",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Sovereign", 
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Banks", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD Mean",
+              "\\hspace{1.0cm}$\\Delta$ Log PD Variance",
+              "\\hspace{1.0cm}$\\Delta$ Log PD Skewness", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD Kurtosis", 
+              "\\hspace{1.0cm}$\\Delta$ Log Gini",
+              "$\\Delta$ Log GDP",
+              #"$\\Delta$ Log Total Assets",
+              #"$\\Delta$ Capital ratio",
+              #"$\\Delta$ Loans / Total Assets",
+              #"$\\Delta$ Net Interest Income / Oper. Rev.",
+              #"$\\Delta$ Deposits / Total Assets", 
+              #"$\\Delta$ Net Income / Total Assets", 
+              "PD Mean",
+              "Log Total Assets",
+              "Capital Ratio",
+              "Loans / Total Assets",
+              "Net Interest Income / Oper. Rev.",
+              "Deposits / Total Assets", 
+              "Net Income / Total Assets"),
+          title = "Summary statistics",
+          table.layout = "n=d#-t-as-",
+          notes = "", 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE)
+```
+
+```{r, eval=FALSE}
+p_w <- c(0.0, 1)         ## Set winsorizing limits
+pre.treat.out <- bank.data %>%
+  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
+  filter(year %in% c("2009","2010"),
+         !is.na(IRB)) %>%
+  select(bvdid, log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
+         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
+         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
+  group_by(bvdid) %>%
+  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
+
+post.treat.out <- bank.data %>%
+  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
+  filter(year %in% c("2012","2013"),
+         !is.na(IRB)) %>%
+  select(bvdid,  log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
+         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
+         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
+  group_by(bvdid) %>%
+  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
+
+DiD.data <- bank.data %>%
+  filter(year %in% c("2010")) %>%
+  right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
+  right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
+  mutate(Shortfall.mn   = log(1+Shortfall.mn), 
+         Shortfall.mn.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.mn, 0), NA),
+         Shortfall.RW.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.RW, 0), NA),
+         out.RWA        = log.RWA.post - log.RWA.pre,
+         out.RWA.IRB    = log.RWA.IRB.post - log.RWA.IRB.pre,
+         out.RWA.SA     = log.RWA.SA.post - log.RWA.SA.pre,
+         out.RWA.hat    = log.RWA.hat.post - log.RWA.hat.pre,
+         out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
+         out.RWA.s.Retail = log.RWA.s.Retail.post - log.RWA.s.Retail.pre,
+         out.RWA.s.Equity = log.RWA.s.Equity.post - log.RWA.s.Equity.pre,
+         out.RWA.s.Wholesale = log.RWA.s.Wholesale.post - log.RWA.s.Wholesale.pre,
+         out.RWA.s.Corporate = log.RWA.s.Corporate.post - log.RWA.s.Corporate.pre,
+         out.RWA.s.Sovereign = log.RWA.s.Sovereign.post - log.RWA.s.Sovereign.pre,
+         out.RWA.s.Banks = log.RWA.s.Banks.post - log.RWA.s.Banks.pre,
+         out.K          = log.K.post - log.K.pre,
+         out.tier       = log.tier.post - log.tier.pre,
+         out.CAR.ratio  = totalcapitalratio.post - totalcapitalratio.pre,
+         out.tier.ratio = tier1ratio.post - tier1ratio.pre,
+         out.CAR.dif    = CAR.dif.post - CAR.dif.pre,
+         out.tier.dif   = tier.dif.post - tier.dif.pre,
+         Country        = as.factor(Country)) %>% 
+  ungroup() %>% 
+  filter(EBAcountry == 1,
+         name != "Bank of Cyprus") %>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+```
+
+```{r , eval=FALSE, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth", include=FALSE}
+p_w <- c(0.0, 1)         ## Set winsorizing limits
+DiD.data %>% ungroup()%>%
+  subset(select = c("Shortfall.mn","out.RWA.s","out.RWA.s.Retail","out.RWA.s.Wholesale","out.RWA.s.Corporate","out.RWA.s.Sovereign","out.RWA.s.Banks","out.RWA", "out.K","out.CAR.ratio","out.CAR.dif")) %>%
   as.data.frame() %>% 
@@
           omit.summary.stat = c("p25","p75"),
-          covariate.labels = c("Capital savings","Tier 1 ratio","Counterfactual tier 1 ratio","Total ratio","Counterfactual total ratio","ROE","Z-score"),
+          covariate.labels = c("Log Shortfall","$\\Delta \\text{Log RWA}^S$", "$\\Delta \\text{Log RWA}^S_R$", "$\\Delta \\text{Log RWA}^S_W$", "$\\Delta \\text{Log RWA}^S_C$", "$\\Delta \\text{Log RWA}^S_S$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}$", "$\\Delta \\text{Log Capital}$", "$\\Delta \\text{CAR}^S$"  ),
           title = "Summary statistics",
           table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{11cm}{The table shows summary statistics for the sample of banks collected.
+          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected.
 Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports.
@@
 
-```{r}
-fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE.
-The solid line is the average across banks for each year.
-The first and third quartiles are shown in the gray area."
-```
+\newpage
 
-```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth", eval=FALSE, include=FALSE}
-mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
-  dat <- data.frame(y = mean(x, na.rm = TRUE),
-                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
-                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
-  return(dat)
-}
+# Capital Savings and Business Cycle
 
-plot1 <- bank.data %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu) %>%
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  mutate(time = lubridate::ymd(year, truncated = 2L))%>% 
-ggplot(aes(x=time)) +
- geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
-labs(x="",y="",
-     title = 'Capital savings') +
-  theme_bw()
+```{r reg_portfolio, echo=FALSE}
 
-plot2 <- bank.data  %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu) %>%
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
-ggplot(aes(x=time)) +
-   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
-labs(x="",y="",
-     title = 'ROE') +
-  theme_bw()
+dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+aux      <- paste0("mean.PD", dep.vars)
+ind.vars <- paste0("log.gdp_gr")
 
-plot3 <- bank.data  %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>%
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
-  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
-ggplot(aes(x=time)) +
-   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
-labs(x="",y="",
-     title = 'Tier 1 ratio') +
-  theme_bw()
+# Creating formulas and running the models
+formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
 
-grid.arrange(plot1, plot2, plot3, ncol=3)
-```
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data =  mutate(bank.data.win, mean.PD := get(paste(aux))))
+})
 
-```{r, include=FALSE}
-
-DiD.data <- DiD.data %>% ungroup() %>% 
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
-
-DiD.intensity <- felm(outcome ~ Shortfall.RW | Country |0|0,
-                  data = DiD.data)
-
-DiD.intensity.2 <- felm(outcome ~ Shortfall.RW.2 | Country |0|0,
-                  data = DiD.data)
-
-DiD.dummy <- felm(outcome ~ EBAbank | Country |0|0,
-                  data = subset(DiD.data, subset = EBAcountry == 1))
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
 
@@
 
-```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE}
+```{r table_portfolio, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
-stargazer( DiD.intensity.2, DiD.dummy,
+stargazer(reg.list,
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "EBA capital exercise and Capital Savings",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{7cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\text{Capital savings}_i = \\beta X_i+\\varepsilon_i.
-\\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
-In column (1) the regressor is the capital shortfall during the capital exercise and in column (2) a dummy that indicates if bank $i$ was part of the exercise.}",
+          title = "Cyclicality of Capital Savings",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{s} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+\pagebreak
+## Alternative measures
+```{r reg_unit, echo=FALSE}
+
+dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+aux      <- paste0("mean.PD", dep.vars)
+ind.vars <- paste0("log.gdp_gr")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data =  mutate(bank.data.win, mean.PD := get(paste(aux))))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_unit, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Alternative Measures: Capital Savings Ratio",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^r = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+```{r reg_gini, echo=FALSE}
+
+dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+aux      <- paste0("mean.PD", dep.vars)
+ind.vars <- paste0("log.gdp_gr")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data =  mutate(bank.data.win, mean.PD := get(paste(aux))))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Alternative Measures: Gini Coefficient",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Gini}_{i,j,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+\pagebreak
+
+## Effect on Capital Ratio procyclicality
+
+```{r reg_contribution, eval=TRUE, echo=FALSE}
+# Variables in the model 
+dep.vars <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat_gr", "log.RWA.r_gr")
+ind.vars <- paste0("log.gdp_gr")
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = subset(bank.data, subset = !is.na(log.CAR_gr)))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+# Effect 
+effect.list  <-  lapply(reg.list, function(x){
+  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
+})
+effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
+effect.list <-  sprintf("%.1f \\%%",effect.list)
+
+```
+
+```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Cyclicality of Capital Ratio Components",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{15cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,j,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$\\Delta\\text{Log capital Ratio}_{i,t}$","$\\Delta\\text{Capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}^{c}_{i,t}$", "$\\Delta\\text{Log RWA}^{r}_{i,t}$"),
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Effect", effect.list),
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+\pagebreak
+
+## Moments
+
+```{r reg_moments, echo=FALSE}
+
+# Variables in the model 
+dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr")
+ind.vars <- paste0("log.gdp_gr")
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = bank.data.win)
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Cyclicality of PD distribution Moments",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{11cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Moment}_{i,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = c("$\\Delta\\text{Log GDP}_{i,t}$", "$\\overline{\\text{PD}}_{i,t-1}$","$\\text{Log Total Assets}_{i,t-1}$", "$\\text{Deposits / Total Assets}_{i,t-1}$", "$\\text{Loans / Total Assets}_{i,t-1}$", "$\\text{Net Income / Total Assets}_{i,t-1}$", "$\\text{Net Interest Income / Oper.
+Rev.}_{i,t-1}$"),
+          keep = c("log.gdp_gr"),
+          dep.var.labels = c("Mean", "Variance", "Skewness", "Kurtosis"), 
           notes.label = "",
@@
           style = "qje",
-          dep.var.labels = c("Capital savings"), column.sep.width = "2pt",
-          covariate.labels = c( "Shortfall","EBA dummy"),
-          add.lines = list(c("Country FE","Yes","Yes")))
+          column.sep.width = "3pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+\pagebreak
+
+## Extensions
+
+```{r reg_extension, echo=FALSE}
+
+# Variables in the model 
+# Dependent variables
+dep.vars <- c("log.RWA.s")
+dep.text <- c("$\\text{Log RWA}^{s}_{i,t}$")
+# Independent variables
+aux.1 <- c("lag(log.asset , order_by = year)",
+           "lag(deposit.ratio, order_by = year)",
+           "lag(loans.ratio, order_by = year)",
+           "lag(NII.ratio, order_by = year)",
+           "lag(income.ratio, order_by = year)",
+           "lag(ROA, order_by = year)",
+           "lag(subordinated, order_by = year)",
+           "lag(NPL.ratio, order_by = year)",
+           "lag(capital.ratio, order_by = year)")
+aux.2 <- c("lag(cap_reg, order_by = year)",
+           "lag(sup_power, order_by = year)")
+aux.3 <- paste(aux.1, collapse=" + ")
+ind.vars <- c(aux.1, aux.2, aux.3)
+ind.text <- c("$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{ROA}_{i,t-1}$",
+              "$\\text{Subordinated Debt}_{i,t-1}$",
+              "$\\text{NPL Ratio}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Capital Strigency}_{i,t-1}$",
+              "$\\text{Supervisory Power}_{i,t-1}$")
+# Fixed effects
+fe       <- c(rep("year + bvdid"  , length(aux.1)),
+              rep("year + Country", length(aux.2)),
+              rep("year + bvdid"  , length(aux.3)))
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
+                                rep("\\multicolumn{1}{c}{No}" , length(aux.2)),
+                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))),
+                c("Country FE", rep("\\multicolumn{1}{c}{-}"  , length(aux.1)),
+                                rep("\\multicolumn{1}{c}{Yes}", length(aux.2)),
+                                rep("\\multicolumn{1}{c}{-}"  , length(aux.3))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
 
@@
 
-```{r, echo=FALSE}
+```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
-bank.data <- bank.data %>% ungroup()%>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>%
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+stargazer(reg.list,
+          title = "capital Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta\\text{Log RWA}^{s}_{i,t} = \\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
 
-stage1 <- felm(data = bank.data , RW.s ~ basel + IRB  |  year +bvdid  |0|0)
-se.s1 <- summary(stage1, robust=TRUE)$coefficients[,2]
-
-bank.data <- mutate(bank.data, f.RW.s = stage1$fitted.values )
-
-iv.t1 <- felm(data = bank.data ,tier1ratio  ~ f.RW.s  | year + bvdid|0|0)
-#se.s1 <- summary(iv.t1$stage1,robust=TRUE)$coefficients[,2]
-se.iv.t1 <- summary(iv.t1,robust=TRUE)$coefficients[,2]
-
-iv.roe <- felm(data = bank.data ,ROE  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.roe <- summary(iv.roe,robust=TRUE)$coefficients[,2]
-
-iv.z <- felm(data = bank.data , Zscore  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.z <- summary(iv.z,robust=TRUE)$coefficients[,2]
-
-iv.roa <- felm(data = bank.data , ROA  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.roa <- summary(iv.roa,robust=TRUE)$coefficients[,2]
-
-iv.sroa <- felm(data = bank.data , r.sd.ROA  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.sroa <- summary(iv.sroa,robust=TRUE)$coefficients[,2]
-
-iv.lev <- felm(data = bank.data , leverage  ~ f.RW.s  | year + bvdid|0|0)
-se.iv.lev <- summary(iv.lev,robust=TRUE)$coefficients[,2]
-
-F.test <- round(t(sapply(stage1$lhs, function(lh) waldtest(stage1, ~ basel + IRB, lhs=lh)))[,5],1)
-
-F.test <- c("F-test",as.character(F.test),"","","","","","","")
-
-```
-
-```{r,eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(stage1, iv.t1, iv.roe, iv.roa, iv.z, iv.sroa, iv.lev, 
-           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Instrumental variable analysis",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{15cm}{The table shows the relationship between the measure of capital savings and several outcome variables instrumented by the introduction of Basel (at country level) and IRB (at bank level).}",
-          se = list(se.s1, se.iv.t1, se.iv.roe, se.iv.z, se.iv.roa, se.iv.sroa, se.iv.lev),
-          covariate.labels = c("Basel","IRB","Savings"),
-          dep.var.labels = c("Savings","CET1","ROE","ROA", "Z-score", "sd(ROA)","Leverage"), 
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          se               = se.list,
+          add.lines        = fe.list,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
           notes.label      = "",
@@
           style            = "qje",
-          column.labels   = c("First stage", "Second stage"), 
-          column.separate = c(1, 6),
-          column.sep.width = "3pt",
-          add.lines = list(c("Bank FE", "Yes", "Yes","Yes", "Yes", "Yes", "Yes", "Yes"),
-                           c("Year FE", "Yes", "Yes","Yes", "Yes", "Yes", "Yes", "Yes"),
-                           F.test))
+          column.sep.width = "-15pt",
+          float.env        = "sidewaystable",
+          align            = TRUE)
+```
 
-```
-```{r, echo=FALSE, include=FALSE}
+```{r,eval=FALSE, echo=FALSE}
 
-reg.hdfe1 <- felm(tier1ratio  ~ RW.s | 0  |0|0,
-                  data = bank.data)
+bank.data.2 <- bank.data %>% ungroup()%>%
+ # filter(!is.na(RW.s),
+  #       !is.na(basel),
+   #      tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA, log.RWA.s, log.RWA.hat, CAR.dif, tier.dif, 
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  RW, log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+#  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
 
-reg.hdfe2 <- felm(tier1ratio  ~ RW.s  | bvdid + year  |0|0,
-                  data = bank.data)
+reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
 
-reg.hdfe3 <- felm(Zscore  ~ RW.s  | 0  |0|0,
-                  data = bank.data)
+reg.2 <- felm(log.K  ~ CAR.dif  | year + IRB.bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
 
-reg.hdfe4 <- felm(Zscore  ~ RW.s  | bvdid + year  |0|0,
-                  data = bank.data)
+reg.3 <- felm(log.RWA  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(log.asset  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(RW  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
+se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
 
@@
 
-```{r, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
+```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}
 
-stargazer(reg.hdfe1,reg.hdfe2,reg.hdfe3,reg.hdfe4,
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
            header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Capital Savings and Capital Requirement Ratio",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{10cm}{The table shows the relationship between the measure of capital savings and risk-weighted capital ratio.
-The dependent variable in each regression is the Core Tier 1 ratio.
-In column (2) and (3) I control for asset size, tier one capital size, and mean PD.
-In regressions (3) I also include bank and year fixed effects.}",
+          title = "Capital Savings and Capital Ratio Components",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times \\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5),
+          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
+          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","Log Assets","RW"), 
           notes.label = "",
@@
           style = "qje",
-          dep.var.labels = c("CT1 ratio"), column.sep.width = "2pt",
-          add.lines = list(c("Year FE", "No", "No","Yes"),
-                           c("Bank FE","No", "No","Yes")))
+          column.sep.width = "3pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r,eval=FALSE, echo=FALSE}
+
+bank.data.2 <- bank.data %>% ungroup()%>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>%
+  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
+         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  log.tier, log.K,
+         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
+  group_by(bvdid) %>%
+#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
+ ungroup() %>%
+ mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
+
+reg.1 <- felm(ROE  ~ CAR.dif | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
+se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(ROA  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(Zscore  ~ CAR.dif  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(leverage  ~ CAR.dif  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
+se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
 
 ```
+
+```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, 
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Capital Savings, Performance, and Risk-Taking",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{15cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times\\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
+$\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4),
+          covariate.labels = c("Capital ratio savings"),
+          dep.var.labels = c("ROE","ROA","Z-score","Leverage"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r, eval=FALSE, echo=FALSE}
+
+reg.1 <- felm(out.tier.ratio ~ Shortfall.mn
+              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
+                  data = DiD.data)
+se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(out.RWA ~ Shortfall.mn 
+                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
+                  data = DiD.data)
+se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(out.RWA.s ~ Shortfall.mn 
+                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
+                  data = DiD.data)
+se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(out.RWA.hat ~ Shortfall.mn 
+              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
+                  data = DiD.data)
+se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(out.tier ~ Shortfall.mn
+              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
+                  data = DiD.data)
+se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+```
+
+```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "EBA capital exercise and Capital Savings",
+          table.layout = "n=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
+The regressor is the capital shortfall during the capital exercise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels = c("$\\Delta$ CET1 ratio", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA Counterfactual", "$\\Delta$ Log CET1" ),
+          se = list(se.1, se.2, se.3, se.4, se.5),
+          covariate.labels = c("Log Capital Shortfall"),
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          #float.env = "sidewaystable",
+          align = TRUE)
+
+```
+
+```{r,eval=FALSE, echo=FALSE}
+
+reg.1 <- felm(out.RWA.s ~ Shortfall.mn 
+               |0|0|0,
+                  data = DiD.data)
+se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
+
+reg.2 <- felm(out.RWA.s.Retail ~ Shortfall.mn 
+                |0|0|0,
+                  data = DiD.data)
+se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
+
+reg.3 <- felm(out.RWA.s.Wholesale ~ Shortfall.mn 
+                |0|0|0,
+                  data = DiD.data)
+se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
+
+reg.4 <- felm(out.RWA.s.Corporate ~ Shortfall.mn
+                |0|0|0,
+                  data = DiD.data)
+se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
+
+reg.5 <- felm(out.RWA.s.Sovereign ~ Shortfall.mn
+                |0|0|0,
+                  data = DiD.data)
+se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
+
+reg.6 <- felm(out.RWA.s.Banks ~ Shortfall.mn 
+               |0|0|0,
+                  data = DiD.data)
+se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
+```
+
+```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}
+
+stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "EBA capital exercise and Capital Savings: Portfolio Breakdown",
+          table.layout = "n=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}^s_{i,j} = \\alpha + \\beta \\text{Log Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
+The regressor is the capital shortfall during the capital exercise.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels = c("$j = \\text{Total}$", "$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"),
+          #se = list(se.1, se.2, se.3, se.4, se.5, se.6),
+          covariate.labels = c("Log Shortfall"),
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+         # float.env = "sidewaystable",
+          align = TRUE)
+```  
+\pagebreak
+# Conclusion and Discussion
+
+Taken together, the results show that strategic portfolio choice, in terms of credit risk, is an important way banks can reduce capital charges while maintaining their average portfolio risk and return on equity.
+Differently from manipulation and subjectivity of RW under IRB, the PD distribution effect hereby identified does not entail any regulatory breach.
+However, the assumption that riskier assets have lower systematic risk, which is one of the reasons the regulatory formulas are concave, may be too simplistic to capture the true systematic component of each asset.
+If that is the case, this paper shows that the current capital requirements framework does not correctly equate capital charges to credit risk, consequently, undermining financial stability.

### 2020-01-07 Auxiliar/Function.R
Add counterfactual exercise to IRB formula

@@ -8,14 +8,14 @@
 ## Basel II/III capital requirements formulas                                  ##
 ##=============================================================================##
 
-mapping_wholesale <- function(PD, LGD = 0.45, M = 2.5, alpha = 0.999){
+mapping_wholesale <- function(PD, LGD = 0.45, M = 2.5, alpha = 0.999, a = -50, R_l = 0.12, R_u = 0.24){
   if(PD==0){
     R  <- NA
     K  <- 0
     RW <- 0
   } else{
   LGD <- ifelse(is.na(LGD), 0.45, LGD)
-  R   <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
+  R   <- R_l*(1-exp(a*PD))/(1-exp(a))+R_u*(1-(1-exp(a*PD))/(1-exp(a)))
   b   <- (0.11852-0.05478*log(PD))^2
   K   <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1+(M-2.5)*b)/(1-1.5*b))
   RW  <- K*12.5
@@ -36,13 +36,13 @@ mapping_retail <- function(PD, LGD, R, alpha = 0.999){
   }
 }
 
-mapping_other_retail <- function(PD, LGD, alpha = 0.999){
+mapping_other_retail <- function(PD, LGD, alpha = 0.999, a = -35, R_l = 0.03, R_u = 0.16){
   if(PD==0){
     R  <- NA
     K  <- 0
     RW <- 0
   } else{
-  R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
+  R  <- R_l*(1-exp(a*PD))/(1-exp(a))+R_u*(1-(1-exp(a*PD))/(1-exp(a)))
   K  <- 1.06*LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
   RW <- K*12.5
   data <- data.frame(K,RW,PD,LGD,R)

### 2020-01-07 Data/Clean.R

@@ -9,6 +9,7 @@
 ##=============================================================================##
 
 rm(list = ls())  
+ptm <- proc.time()
 library(tidyverse)                ## Data manipulation, pipe operator
 library(zoo)                      ## Apply function to rolling margins of data
 library(DescTools)                ## Command to winsorize variables
@@ -59,6 +60,18 @@ for (i in 1:nrow(IRB.data)){
  
   IRB.data[i,"RWA.hat"] <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"w.avg.PD"],
                                                                IRB.data[i,"LGD"])["RW"]
+  
+  IRB.data[i,"RWA.til"]     <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"PD"],
+                                                               IRB.data[i,"LGD"],
+                                                               a = -50,
+                                                               R_l = 0.10,
+                                                               R_u = 0.28)["RW"]
+  
+  IRB.data[i,"RWA.bar"] <- IRB.data[i,"EAD"]*mapping_wholesale(IRB.data[i,"w.avg.PD"],
+                                                               IRB.data[i,"LGD"],
+                                                               a = -50,
+                                                               R_l = 0.10,
+                                                               R_u = 0.28)["RW"]
 }
 
 for (i in 1:nrow(IRB.data)){
@@ -151,18 +164,21 @@ for(i in c("Wholesale","Retail","Equity")) {
 bank.data <- IRB.data %>% 
   group_by(name,bvdid,Country,year,Portfolio_1) %>%
   filter(Portfolio_1 %in% c(i)) %>%
-  summarise(!!(paste("gini","PD"  ,        i, sep=".")) := Gini(PD,w.EAD, na.rm = TRUE),
-            !!(paste("sd"  ,"PD"  ,        i, sep=".")) := sqrt(wtd.var(PD,w.EAD, na.rm = TRUE))*100,
[…]

### 2020-01-07 Data/Import.R

@@ -12,7 +12,6 @@ rm(list = ls())
 library(tidyverse)    ## Data manipulation, pipe operator                                                             
 library(readxl)       ## Command to open xlsx files                                                     
 library(data.table)   ## Command to bind lists
-library(reshape2)     ## Command to transform data from wide to long
 library(stringr)      ## Command to clean variables names in Orbis
 
 ##============================================================================##

### 2020-01-07 poster/poster.Rmd
Rename test_poster_2.Rmd to poster.Rmd

---
main_topsize: 0.13 #percent coverage of the poster
main_bottomsize: 0.05
#ESSENTIALS
title: '**Credit Risk Distribution and Capital Savings**'
author:
  - name: '**Lucas Avezum**'
    affil: 1, 2
    main: true
    email: l.avezum@uvt.nl
affiliation:
  - num: 1
    address: CentER, Tilburg University
  - num: 2
    address: EBC
main_findings:
  - "The concavity of IRB formulas creates **countercyclical** capital savings. "
main_fontfamily: "Montserrat"
logoleft_name: Pictures/Tilburg_Logo_Horizontal.png
logoleft_width: "100%"
logoright_name: Pictures/ebc_logo.png
logoright_width: "5cm"
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: packages.bib
link-citations: true
primary_colour:  "#003366"
secondary_colour: "#008EC6"
accent_colour: "#000000"
---

<style>
#main-img-left {
 width: 20%;
}
#main-img-right {
 width: 7%;
}
p.caption {
  font-size: 0.8em;
}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html")

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
#library(relaimpo)            ## Calculate relative importance measures

source("../Auxiliar/Function.R") ## Capital requirements functions

```

```{r load data, include=TRUE}
load("../Data/Datasets/Pillar3Data.Rda")
load("../Data/Datasets/BankData.Rda")
load("../Data/Datasets/CrossSectionDecomposition.Rda")
load("../Data/Datasets/timedecomposition.Rda")
load("../Data/Datasets/yeardecomposition.Rda")
```

```{r , fig.height = 10, fig.width = 7, results="asis"}
p_w <- c(0.05, 0.95)         ## Set winsorizing limits
bank.data.win <- bank.data %>% 
  ungroup()%>%
  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>% 
  group_by(bvdid)

year.decomposition <- year.decomposition %>%
  mutate(RW_2 = RW - t0.RW,
         year = as.factor(year))

``` 
# Introduction

The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach. Therefore, total capital requirement depends not only on the average PD but also on the shape of the PD distribution. Using a hand-collected dataset, I propose a measure that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component. I document that the latter is an important factor of risk-weight variability across banks. I further investigate the prior ambiguous direction of its relationship with the business cycle and find it to be countercyclical. Altogether, the results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements.

# The Measure

Consider two portfolios. The first contains a low PD asset and a high PD asset in equal proportions. The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the arithmetic mean of low PD and high PD.

Figure \@ref(fig:measure) shows the amount of risk-weighted assets (RWA) calculated for each portfolio under the IRB approach: $\text{RWA}$ for the first and $\text{RWA}^c$ for the second. Because of the concavity of the IRB formula $\text{RWA}^c$ is greater than $\text{RWA}$.

```{r parameters curve, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 1
M   <- 2.5
grid  <- seq(0.0001, 1, by=0.00001) 
data <- mapping_retail(grid,1,0.15,0.999)
PD.l <- 0.01
PD.h <- 0.15
PD.a <- (PD.l+PD.h)/2
PD.h2 <- 0.17
PD.h3 <- 0.16
PD.l3 <- 0.02
PD.a2 <- (PD.l+PD.h2)/2
K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
```

```{r}
measure_cap = "The figure plots the method to measure capital savings."
```

```{r,measure, fig.height = 3, fig.width = 4,results="hide",out.width = "75%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= measure_cap}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=1)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=5)+
  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.5)+

  labs(x = "PD", 
         y = "Risk-weight") +
  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        axis.title.x = element_blank(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous(breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("RWA",expression(paste(RWA^{c}))),expand = c(0, 0)) +
  annotate("text", x = (PD.l+PD.a)/3 , y = (K.s+K.a)/2, label = "Capital \n Savings", size = 3)

```

I use this idea to construct a measure that reflects differences in RWA coming from differences on the shape of the PD distribution. The measure is straightforward: compare the actual capital requirement ($\text{RWA}$) to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio's average PD ($\text{RWA}^c$). The measure of capital savings ($\text{RWA}^s$) is defined as follows:

$$
\text{Capital Savings} = \text{RWA}^s = \frac{\text{RWA}^c}{\text{RWA}}
$$

# Data

The paper introduces a hand-collected dataset to quantify the intensity of the capital savings across banks and time.

```{r}
savings_cap = "The figure plots capital savings across banks. Each point is a bank-year observation and bars are the averages values for each bank."
```

```{r, savingsbank, include = TRUE, fig.height = 7, fig.width = 5,out.width = "85%",results="hide",fig.cap= savings_cap, fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
  select(name, Country, RWA.r)%>%
  group_by(name, Country) %>%
#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
  group_by(Country) %>%
 # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
  #ungroup()%>%
  #mutate(ordering = order_aux+RWA.r,
  #       name = as.factor(name),
  #       name = fct_reorder(name, ordering),
  #       Country = fct_reorder(Country, ordering)) %>%
 ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
                      y = RWA.r)) +
   stat_summary(fun.y="mean", geom="bar",alpha = 0.5, fill = "#C3BCB2" ) +
geom_point(alpha = 0.7,size = 2, colour = "#003366") +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital Savings") +
  #scale_fill_grey()+
   theme_bw()

```

I include the capital savings measure in a decomposition analysis of risk-weights' (RW). The extensive component measures the contribution of capital savings as banks move their portfolio from the SA to the IRB approach while holding the shape of the PD distribution constant. The intensive component measures the contribution of capital savings from changes of the shape of the PD distribution while holding the extensive component constant.

```{r}
savings_time_cap = "The figure plots the cross-bank average contribution of the several RW components and average RW (solid line) for each year relative to 2008."
```

```{r, savingstime, include=TRUE, fig.height = 4, fig.width = 5,out.width = "75%", results="hide", fig.cap= savings_time_cap, fig.align = 'center'}

year.decomposition %>%
  select(year,RW, int.gain, ext.gain, other.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain"))) %>%
  ggplot(mapping = aes(x = year)) +
geom_bar(aes(y = value, 
          fill = condition),
          stat = 'identity')+
geom_line(data = year.decomposition, 
          aes(group=1,y = RW_2))  +
  theme_bw()+
  labs(y = 'Change from 2008', x=NULL) +
  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
 theme(legend.position = c(0.3, 0.2), 
        legend.title = element_blank(), 
        legend.background=element_blank())+
  scale_fill_manual(values = c("#003366","#CC9933","#C3BCB2"),
                    labels = c("Intensive component", "Extensive component", "Other components"), name = "Component")
  #scale_fill_grey(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")

```

```{r}
rsqrt_cap = "The figure plots the percentage of total RW variation explained by each of its components in the cross-section."
```

```{r,rsqrt, fig.height = 4, fig.width = 5,out.width = "75%", results="hide", fig.cap= rsqrt_cap, include=TRUE,fig.align = 'center'}

unrestrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
         total.gain = total.gain+epsilon)%>% 
  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )

  data.frame(share  = unrestrict@lmg) %>%
   add_rownames(var = "component")%>%
  mutate(component = factor(component,
                      levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
                      labels = c("PD avg.", "Ext. comp.", "Int. comp.", "SA", "Mix", "Rollout"))) %>%
ggplot(mapping = aes(x = reorder(component, -share),
                       y = share, fill = component)) +
geom_bar(stat = 'identity')+
  labs(x=NULL, y = expression(paste("Share of ", R^{2}))) +
    theme_bw()+
  theme(panel.grid.minor = element_blank(),
        strip.background = element_blank(),
       # strip.text = element_blank()
       legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values = c("#C3BCB2","#CC9933",  "#003366", "#C3BCB2","#C3BCB2", "#C3BCB2"))+
scale_y_continuous(labels = scales::percent_format(accuracy = 1,
                                 decimal.mark = '.'))

```

# Results

Next, I test if capital savings are pro- or countercyclical. The expected direction is ambiguous. On the one hand, if the entire portfolio is equally affected, capital savings are procyclical. On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical. Figure \@ref(fig:cyclicality) illustrates the two possible outcomes.  
```{r parameters curve 2, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 1
M   <- 2.5
grid  <- seq(0.0001, 1, by=0.00001) 
data <- mapping_retail(grid,1,0.15,0.999)
PD.l <- 0.01
PD.h <- 0.13
PD.a <- (PD.l+PD.h)/2
PD.h2 <- 0.19
PD.h3 <- 0.16
PD.l3 <- 0.04
PD.a2 <- (PD.l+PD.h2)/2
K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
```

```{r}
cyclicality_cap = "The figure shows the two possible outcomes of a negative credit risk shock on a portfolio. The left-hand side plot shows the case where capital savings are procyclical. The right-hand side plot shows the case where capital savings are countercyclical. "
```

```{r, cyclicality, fig.height = 3, fig.width = 6,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap}

p1 <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K))+ 
  geom_line(alpha = 1) + 
  #geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  #geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
  #geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5, colour = "#008EC6")+
  #geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5, colour = "#008EC6")+
  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5, colour = "#008EC6")+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5, colour = "#008EC6")+
  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5, colour = "#008EC6")+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5, colour = "#008EC6")+
  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        #axis.text.x = element_blank(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
 scale_x_continuous(breaks=c(PD.h2),labels=c("PD"), expand = c(0, 0))+
  scale_y_continuous(breaks=c(K.h2),labels=c("RWA"),expand = c(0, 0))

p2 <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K))+ 
  geom_line(alpha = 1) + 
  #geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  #geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
  #geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5, colour = "#CC9933")+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5, colour = "#CC9933")+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5, colour = "#CC9933")+
  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=5, colour = "#CC9933")+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5, colour = "#CC9933")+
 theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        #axis.text.x = element_blank(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
 scale_x_continuous(breaks=c(PD.h2),labels=c("PD"), expand = c(0, 0))+
  scale_y_continuous(breaks=c(K.h2),labels=c("RWA"),expand = c(0, 0))

grid.arrange(p1, p2, ncol=2)
```

```{r , echo=FALSE}

dep.vars <- c("log.RWA.r_gr",
              "log.RWA.r.Retail_gr",
              "log.RWA.r.Wholesale_gr",
              "log.RWA.r.Corporate_gr")
dep.text <- c("$j = \\text{Total} <span>&nbsp;</span> $",
              "$j = \\text{Retail} <span>&nbsp;</span>$", 
              "$j = \\text{Wholesale} <span>&nbsp;</span>$", 
              "$j = \\text{Corporate}<span>&nbsp;</span>$")
# Independent variables
aux.1 <- c("log.gdp_gr")
aux.2 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.2)
ind.text <- c("$<span>&#916;</span> Log GDP_{i,t}$")
# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = filter(bank.data.win, RWA.s>0))
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

```

Which of the two effects dominates is an empirical question. Table 1 provides evidence supporting the second scenario, i.e., countercyclical capital savings.

\centering 
```{r , table2, eval=TRUE, echo=FALSE, results='asis' ,warning=FALSE, message=FALSE}
stargazer(reg.list,
          title = "Table 1: Capital Savings and Cyclicality",
          table.layout = "n=c-!d#-t-as-",
          omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = fe.list,
          column.labels    = c("Dependent variable: $<span>&#916;</span>\\text{Log RWA}^{s}_{i,j,t}$"),
          column.separate  = c(5),
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "html",
          font.size        = "tiny",
          notes.label      = "",
          #label            = "table2",
          notes.align      = "c", 
          notes.append     = FALSE,
          style            = "qje",
          column.sep.width = "10pt",
          #float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```

```{r , echo=FALSE}

dep.vars <- c("log.CAR_gr", 
              "log.K_gr", 
              "log.RWA.IRB_gr", 
              "log.RWA.hat_gr", 
              "log.RWA.r_gr")
dep.text <- c("$<span>&#916;</span>\\text{Log Cap. Ratio}_{i,t}$",
              "$<span>&#916;</span>\\text{Log Capital}_{i,t}$", 
              "$<span>&#916;</span>\\text{Log RWA}_{i,t}$", 
              "$<span>&#916;</span>\\text{Log RWA}^{c}_{i,t}$", 
              "$<span>&#916;</span>\\text{Log RWA}^{s}_{i,t}$")
# Independent variables
aux.1 <- c("log.gdp_gr")
aux.2 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.2)
ind.text <- c("$<span>&#916;</span> \\text{Log GDP}_{i,t}$")
# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = subset(bank.data, subset = !is.na(log.CAR_gr)))
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

# Effect 
effect  <-  lapply(reg.list, function(x){
  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
})
effect[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
effect.list <-  list(c("Effect", sprintf("%.1f \\%%",effect)))

#additional lines

add.list <- append(fe.list, effect.list, after = 0)
```

Table 2 uses the following equality to calculate the contribution of capital savings (and other components) to the cyclicality of capital ratios:

$$
\small
\begin{aligned}
\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^s_{i,t} 
\end{aligned}
\small
$$

\centering 
```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE,out.width = "1\\textwidth"}
stargazer(reg.list,
          title = "Table 2: Cyclicality of Capital Ratio Components",
          table.layout = "n=c-!d#-t-as-",
          omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = add.list,
          #column.labels    = c("Panel A: Actual"),
          #column.separate  = c(5),
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "html",
          font.size        = "tiny",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          #label            = "tab1",
          style            = "qje",
          column.sep.width = "0pt",
          #float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```

# Conclusion

Capital savings due to the shape of PD distribution and the concavity of the IRB formulas are:

1. **economically relevant**. They explain 27% of the RW variation across banks. 
2. **countercyclical**. Reduce the procyclicality of RWA by 11.6 percentage points.

Hence, the concavity of IRB formulas could be explored to reduce procyclicality of capital regulation.

### 2020-01-07 docs/poster/poster.Rmd
change folder of poster to docs

--- 2020-01-07 poster/poster.Rmd
@@
 
-source("../Auxiliar/Function.R") ## Capital requirements functions
+source("../../Auxiliar/Function.R") ## Capital requirements functions
 
@@
 ```{r load data, include=TRUE}
-load("../Data/Datasets/Pillar3Data.Rda")
-load("../Data/Datasets/BankData.Rda")
-load("../Data/Datasets/CrossSectionDecomposition.Rda")
-load("../Data/Datasets/timedecomposition.Rda")
-load("../Data/Datasets/yeardecomposition.Rda")
+load("../../Data/Datasets/Pillar3Data.Rda")
+load("../../Data/Datasets/BankData.Rda")
+load("../../Data/Datasets/CrossSectionDecomposition.Rda")
+load("../../Data/Datasets/timedecomposition.Rda")
+load("../../Data/Datasets/yeardecomposition.Rda")
 ```

### 2020-03-16 Data/Clean.R
Add updated BRSS dataset

@@ -24,6 +24,7 @@ load("Data/Temp/BankScope.Rda")
 load("Data/Temp/Pillar3Data.Rda")
 load("Data/Temp/AuxiliarData.Rda")
 
+
 ##============================================================================##
 ## Pillar-III reports                                                         ##
 ##============================================================================##
@@ -233,10 +234,10 @@ bank.data <- bank.data%>%
   # Add macro variables
   left_join(WDI, by = c("Country", "year")) %>%
   # Add bank regulation variables
-  mutate(survey = ifelse(year<2001, 1, 
-                         ifelse(between(year, 2001, 2004), 2, 
-                                ifelse(between(year, 2005, 2007), 3, 4))))%>%
-  left_join(BRSS, by = c("Country", "survey"))
+  # mutate(survey = ifelse(year<2001, 1, 
+  #                        ifelse(between(year, 2001, 2004), 2, 
+  #                               ifelse(between(year, 2005, 2007), 3, 4))))%>%
+  left_join(BRSS, by = c("Country", "year"))
 
 #------------------------------------------------------------------------------#
 # Basic cleaning and variable calculation                                      #
@@ -313,7 +314,7 @@ bank.data <- bank.data %>% ungroup() %>% rowwise() %>%
           Zscore     = (r.mean.ROA + r.mean.CAR)/r.sd.ROA)%>%
    # Growth rates
    ungroup() %>% group_by(bvdid) %>%
-   mutate_at(vars(matches("log.|.ratio")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
+   mutate_at(vars(matches("_|log.|.ratio")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
                                                                     . - lag(., order_by = year), NA))) %>%
    mutate_at(vars(matches(".PD")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
                                                             log(.) - log(lag(., order_by = year)), NA))) %>%

### 2020-03-16 Data/Import.R

@@ -66,17 +66,18 @@ country.code <- read.csv("Data/Raw/WDI/WDICountry.csv") %>%
   select(country.name = short.name,
          Country      = x2.alpha.code)
 
-BRSS <- read_xls("Data/Raw/BRSS/BCL_Sup_Reg_Data_13JAN2013.xls",sheet = 4) %>%
-  select(-c("Number","Question response"),-contains("value"))%>%
-  pivot_longer(-c('Name','Survey'), names_to = "Country", values_to = "value")%>%
-  pivot_wider(names_from = Name, values_from = value)%>%
-  filter(!is.na(Survey))%>%
-  rename_at(vars(matches("[|]|*")),funs(str_replace_all(., "\\(|\\)|\\[|\\]|\\*|[0-9]", "")))%>%
-  rename_all(tolower) %>%
-  mutate_at(vars(-country),as.numeric)%>%
-  select_if(~!all(is.na(.)))%>%
-  select(survey, country.name = country, ovr_cap_string, init_cap_strin, cap_reg, sup_power)%>%
-  inner_join(country.code)
+# BRSS <- read_xls("Data/Raw/BRSS/BCL_Sup_Reg_Data_13JAN2013.xls",sheet = 4) %>%
+#   select(-c("Number","Question response"),-contains("value"))%>%
+#   pivot_longer(-c('Name','Survey'), names_to = "Country", values_to = "value")%>%
+#   pivot_wider(names_from = Name, values_from = value)%>%
+#   filter(!is.na(Survey))%>%
+#   rename_at(vars(matches("[|]|*")),funs(str_replace_all(., "\\(|\\)|\\[|\\]|\\*|[0-9]", "")))%>%
+#   rename_all(tolower) %>%
+#   mutate_at(vars(-country),as.numeric)%>%
+#   select_if(~!all(is.na(.)))%>%
+#   select(survey, country.name = country, ovr_cap_string, init_cap_strin, cap_reg, sup_power)%>%
+#   inner_join(country.code)
+load("Data/Raw/BRSS/Output/BRSS.Rdata")
   
   save(list = c("basel.indicator","EBA.indicator","WDI","BRSS"),
        file=paste0("Data/Temp/AuxiliarData.Rda"))

### 2020-03-16 Analysis/Presentation/Presentation.Rmd
Add presentation code

---
title: "Credit Risk Distribution and Capital Savings"
author: | 
  | Lucas Avezum
  | Tilburg University
date: "November 19, 2019"
header-includes:
   - \usepackage{bbm,dcolumn,transparent}
output: 
  beamer_presentation:
    theme: "Boadilla"
    colortheme: "beaver"

bibliography: ../../Auxiliar/mybibfile.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE)

rm(list = ls())

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
#library(relaimpo)            ## Calculate relative importance measures

source("../../Auxiliar/Function.R") ## Capital requirements functions

```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

## Banks and Regulation

Banks:

- Pool short-term liabilities to fund long-term assets.

- Perform other important activities (payment process).

- Central role to monetary policy transmission [@bernanke1995inside].

- unstable, prone to runs [@diamond1983bank].

## Capital Requirements

- Rationale: Compensate for public guarantees and externalities that result in banks having less capital than socially desirable.

- However, there is no consensus on appropriate level and implementation [@diamond2000theory; @diamond2001liquidity; @admati2014bankers].

- Hence, regulation on capital requirements have changed substantially since first international standard.

- Arguably, risk sensitivity is at the center of the changes [@danielsson2001academic; @basel2004international; @EBA2019policy]

## Capital Requirements and Risk Sensitivity

- Basel I (July 1988): Sets fixed risk-weights for assets within same class.

\begin{equation*}
\begin{aligned}
\label{eq5}
\text{Capital Ratio} = \frac{K}{\sum_j\alpha_j\sum_i  x_{i,j}} \geq k
\end{aligned}
\end{equation*}

$K$: equity capital \newline
$x_i$: asset $i$ from class $j \in \{\text{Corporate, Retail, Sovereign, etc...} \}$ \newline
$\alpha_j$: weight for assets in class $j$ \newline
$k$: capital ratio requirement \newline

\pause

- Problem: simple rule may incentivize banks to hold more risky portfolios and increase probability of failure [@koehn1980regulation]

## Capital Requirements and Risk sensitivity

Basel II (2004): Introduces the possibility to use internal-based rating (IRB) models to calculate risk-weights

\begin{equation*}
\begin{aligned}
\label{eq5}
\text{Capital Ratio} = \frac{K}{\sum_j\sum_i \alpha_{i,j}  x_{i,j}} \geq k
\end{aligned}
\end{equation*}

$\alpha_{i,j}$: weight for assets $i$, estimated using banks own risk models.

\pause

Rationale:

- Riskier banks will hold more capital [@kim1988risk].
- Alignment of risk estimates to actual risk [@barakova2014banks].

\pause

Problems:

- Enhances procyclical nature of lending [@behn2016procyclical].
- Allow differences for reasons other than risk [@mariathasan2014manipulation; @behn2016limits].

## Credit Risk Distribution and Capital Savings

- This paper contributes to the understanding of risk-sensitive capital requirements ($\alpha_{i,j}$).

- I identify a novel effect stemming from the design of the IRB framework: banks save capital requirements because mapping from risk (PD) is concave.

- I collect data and propose a measure of this effect.

- Key findings:
  - Reduces the procyclicality of RWA by 11.2 percentage points. 
  - Explains 27% of RW variation across banks.

## Concavity of Capital Requirements

```{r parameters and data, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 1
M   <- 2.5
grid  <- seq(0.0001, 1, by=0.00001) 
data <- mapping_retail(grid,1,0.15,0.999)
PD.l <- 0.01
PD.h <- 0.15
PD.a <- (PD.l+PD.h)/2
PD.h2 <- 0.17
PD.h3 <- 0.16
PD.l3 <- 0.02
PD.a2 <- (PD.l+PD.h2)/2
K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
```
```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) +
  labs(x = "PD", 
         y = "Risk-weight") +
  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.h+0.03),labels=c("PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.h),labels=c("RWA"),expand = c(0, 0))

```
## Concavity of Capital Requirements
```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=1)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("",""),expand = c(0, 0))

```

## Concavity of Capital Requirements
```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=1)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=5)+
  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.5)+

  labs(x = "PD", 
         y = "Risk-weight") +
  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("RWA",expression(paste(RWA^{c}))),expand = c(0, 0)) +
  annotate("text", x = (PD.l+PD.a)/3 , y = (K.s+K.a)/2, label = "Capital \n Savings", size = 3)

```
## Capital Savings Measure

\begin{equation*}
\begin{aligned}
\label{eq5}
\text{RWA}^s = \frac{\text{RWA}^c}{\text{RWA}} 
\end{aligned}
\end{equation*}

\pause

\begin{equation*}
\begin{aligned}
\label{eq5}
\text{RWA}^{\$} = \text{RWA}^c-\text{RWA} 
\end{aligned}
\end{equation*}

## Data and Pillar-III Disclosures
```{r , echo=FALSE,  out.width = '100%'}
knitr::include_graphics("../../Auxiliar/Pictures/Citi_table.png")
```

## Capital Savings Across Banks

```{r, include = TRUE, fig.height = 5, fig.width = 8,out.width = "90%",results="hide", fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
  select(name, Country, RWA.r)%>%
  group_by(name, Country) %>%
#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
  group_by(Country) %>%
 # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
  #ungroup()%>%
  #mutate(ordering = order_aux+RWA.r,
  #       name = as.factor(name),
  #       name = fct_reorder(name, ordering),
  #       Country = fct_reorder(Country, ordering)) %>%
 ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
                      y = RWA.r)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
  labs(y =  "RWA savings ratio") +
   theme_bw()+
  theme(panel.grid = element_blank(),
         axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_grey()

```

## Capital Savings Across Countries

```{r, include=TRUE, fig.height = 4, fig.width = 3, out.width = "50%", results="hide", fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  select(Country,name, year, RWA.r)%>%
  group_by(Country,name) %>%
 summarise_if(is.numeric,mean,na.rm= TRUE)%>%
 ggplot(mapping = aes(x = reorder(Country, RWA.r, FUN = mean, na.rm=TRUE),
                          y = RWA.r)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
  labs(y =  "RWA savings ratio") +
  coord_flip() +
  theme_bw()

```

## Decomposition

\begin{equation*}
\begin{aligned}
\label{eq5}
\text{RWA} & = \text{RWA}_{SA}+\text{RWA}_{IRB} \\
\pause
& = \text{RWA}_{SA}+\text{RWA}^{c}-\text{RWA}^{\$}\\
\pause
& = \text{RWA}_{SA}+\sum_{j=1}^P\text{RWA}^{c}_j-\text{RWA}^{\$}\\
\end{aligned}
\end{equation*}

## Decomposition

\begin{equation*}
\begin{aligned}
\label{eq5}
\frac{\text{RWA}}{\text{EAD}}&= \frac{\text{RWA}_{SA}}{\text{EAD}_{SA}}\frac{\text{EAD}_{SA}}{\text{EAD}}+\frac{\text{EAD}_{IRB}}{\text{EAD}}\sum_{j=1}^P\frac{\text{EAD}_{IRB,j}}{\text{EAD}_{IRB}}\frac{\text{RWA}^{c}_j}{\text{EAD}_{IRB}}+\frac{\text{RWA}^{\$}}{\text{EAD}_{IRB}}\frac{\text{EAD}_{IRB}}{\text{EAD}}\\
\pause
\text{RW} &= \text{RW}_{SA}q_{SA}+(1-q_{SA})\sum_{j=1}^P q_{j}\text{RW}^{c}_j +(1-q_{SA})\text{RW}^{s}\\
\end{aligned}
\end{equation*}

## Decomposition

\begin{equation*}
\begin{aligned}
\label{eq5}
\text{RW}-\text{RW}^B =& \text{RW}_{SA}q_{SA}+(1-q_{SA})\sum_{j=1}^P q_{j}\text{RW}^{c}_j +(1-q_{SA})\text{RW}^{s}\\
&- \text{RW}_{SA}^Bq_{SA}^B+(1-q_{SA}^B)\sum_{j=1}^P q_{j}^B\text{RW}^{c,B}_j +(1-q_{SA})\text{RW}^{s,B}
\end{aligned}
\end{equation*}

## Decomposition

\begin{equation*}
\begin{aligned}
\label{eq5}
\Delta \text{RW} & = \underbrace{q^{B}_{SA} \Delta \text{RW}_{SA}}_\text{SA mix}+\underbrace{(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j}_\text{Mean IRB Risk}\\
&\\
&+ \underbrace{(1-q^{B}_{SA})\sum_{j=1}^{P} \text{RW}_{j}^{c} \Delta q^{j}}_\text{IRB mix}  + \underbrace{(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}}_\text{Roll-out}\\
&\\
&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin}
\end{aligned}
\end{equation*}

## Cross-sectional Decomposition

```{r, fig.height = 5, fig.width = 8,out.width = "90%", results="hide", include=TRUE,fig.align = 'center'}
cross.section.decomposition %>% filter(q.SA<1)%>%
  mutate(total= -PD.gain)%>%
  select(name, Country,total, other.gain, PD.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
                                       labels = c("Capital savings", "Other components")))%>%
ggplot(mapping = aes(x = reorder(name, total),
                       y = value)) +
geom_bar(stat = 'identity')+
  labs(x=NULL, y= NULL) +
    theme_bw()+
  theme(panel.grid = element_blank(),
         axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~condition, nrow = 2, ncol = 1)

```

## Cross-sectional Decomposition

```{r, fig.height = 3, fig.width = 4, results="hide",out.width = "80%", include=TRUE,fig.align = 'center'}
cross.section.decomposition %>% 
  filter(q.SA<1)%>%
  select(name, Country, other.gain, PD.gain) %>%
  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
  ggplot(mapping = aes(x = PD.gain,
                       y = other.gain)) +
    geom_point()+
    geom_smooth(method = "lm", color = "black")  +
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey()+
  stat_cor(label.y= -0.1)
```

## Cross-sectional Decomposition

```{r, fig.height = 5, fig.width = 8,out.width = "90%", results="hide", include=TRUE,fig.align = 'center'}
cross.section.decomposition %>% filter(q.SA<1)%>%
   mutate(total= -PD.gain)%>%
  select(name, Country,total, int.gain, ext.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
                                       labels = c("Intensive margin", "Extensive margin")))%>%
ggplot(mapping = aes(x = reorder(name, total),
                       y = value)) +
geom_bar(stat = 'identity')+
  labs(x=NULL, y= NULL) +
    theme_bw()+
  theme(panel.grid = element_blank(),
         axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~condition, nrow = 2, ncol = 1)

```

## Cross-sectional Decomposition

```{r, fig.height = 3, fig.width = 4, results="hide",out.width = "80%", include=TRUE,fig.align = 'center'}
cross.section.decomposition %>% 
  filter(q.SA<1)%>%
  select(name, Country, int.gain, ext.gain) %>%
  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
  ggplot(mapping = aes(x = int.gain,
                       y = ext.gain)) +
    geom_point()+
    geom_smooth(method = "lm", color = "black")  +
  labs(x=NULL, y= NULL) +
   theme_bw() +
  theme(panel.grid.minor = element_blank())+
  scale_fill_grey()+
  stat_cor(label.y= -0.1)
```

## R-squared decompostion

```{r,fig.height = 4, fig.width = 6,out.width = "70%", results="hide", include=TRUE,fig.align = 'center'}

unrestrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
         total.gain = total.gain+epsilon)%>% 
  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )

  data.frame(share  = unrestrict@lmg) %>%
   add_rownames(var = "component")%>%
  mutate(component = factor(component,
                      levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
                      labels = c("PD avg.", "PD ext.", "PD int.", "SA", "Mix", "Rollout"))) %>%
ggplot(mapping = aes(x = reorder(component, -share),
                       y = share)) +
geom_bar(stat = 'identity')+
  labs(x=NULL, y = expression(paste("Share of ", R^{2}))) +
    theme_bw()+
  theme(panel.grid.minor = element_blank())+
  theme(strip.background = element_blank(),
       # strip.text = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_grey()+
scale_y_continuous(labels = scales::percent_format(accuracy = 1,
                                 decimal.mark = '.'))

```

## Time decomposition

```{r, fig.height = 4, fig.width = 7, results="hide",out.width = "1\\textwidth",fig.align = 'center'}

plot1 <-time.decomposition %>% 
  select(time, RW, t0.RW,  int.gain, ext.gain, other.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain")),
         RW = RW - t0.RW,
         time = as.factor(time)) %>%
  ggplot(mapping = aes(x = time,
                       y    = value, 
                       fill = condition)) +
geom_bar(stat = 'identity')+
geom_line(aes(group=1, y = RW))  +
  theme_bw()+
  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
 theme(legend.position = c(0.3, 0.2), 
        legend.title = element_blank(), 
        legend.background=element_blank())+
  scale_fill_grey(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")

year.decomposition <- year.decomposition %>%
  mutate(RW_2 = RW - t0.RW,
         year = as.factor(year)) 
plot2 <- year.decomposition %>% 
  select(year,RW, int.gain, ext.gain, other.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain"))) %>%
  ggplot(mapping = aes(x = year)) +
geom_bar(aes(y = value, 
          fill = condition),
          stat = 'identity')+
geom_line(data = year.decomposition, 
          aes(group=1,y = RW_2))  +
  theme_bw()+
  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
theme(legend.position='none')+
  scale_fill_grey()
grid.arrange(plot1, plot2, ncol=2)

```

## Regulatory Pressure and Reduction in Overall Risk

```{r, fig.height = 4, fig.width = 5,out.width = "80%", results="hide",fig.align = 'center'}

mean_cl_quantile <- function(x, q=c(0.25, 0.75), na.rm=TRUE){
  dat <- data.frame(y=mean(x, na.rm = TRUE),
                    ymin = quantile(x, probs = q[1],na.rm = na.rm),
                    ymax = quantile(x, probs = q[2],na.rm = na.rm))
  return(dat)
}
scaleFUN <- function(x) sprintf("%.1f", x)

plot1 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year, capital.ratio) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=capital.ratio)) +
geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  labs(y = 'Capital Ratio (%)', x=NULL) + scale_y_continuous(labels=scaleFUN)

plot2 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year,  mean.PD) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=mean.PD)) +
geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
   labs(y = 'Average PD (%)', x=NULL)

plot3 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year, var.PD) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
   labs(y = 'SD PD (%)', x=NULL)

plot_grid(plot1, plot2, plot3, hjust = 0.5,
            ncol = 1, nrow = 3, align = "v", axis = "lr")
```

```{r , fig.height = 10, fig.width = 7, results="asis"}
p_w <- c(0.05, 0.95)         ## Set winsorizing limits
bank.data.win <- bank.data %>% 
  ungroup()%>%
  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>% 
  group_by(bvdid)

```

## Capital Savings and Cyclicality
```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("",""),expand = c(0, 0))

```
## Capital Savings and Cyclicality
```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+

  geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5)+

  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("",""),expand = c(0, 0))

```
## Capital Savings and Cyclicality
```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+

  geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=6)+

  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("",""),expand = c(0, 0))

```

## Capital Savings and Cyclicality

```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K)) 
p + geom_line(alpha = 1) + 
  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=6)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=6)+

  geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5)+
  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5)+

  theme_bw()+
  theme(axis.ticks = element_blank(),
        axis.line = element_line(),
        panel.grid = element_blank())+
  expand_limits(x = 0, y = 0)+
  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("",""),expand = c(0, 0))

```
## Capital Savings and Cyclicality

```{r , echo=FALSE}

dep.vars <- c("log.RWA.r_gr",
              "log.RWA.r.Retail_gr",
              "log.RWA.r.Wholesale_gr",
              "log.RWA.r.Corporate_gr",
              "log.RWA.r.Sovereign_gr",
              "log.RWA.r.Banks_gr")
dep.text <- c("$j = \\text{Total}$",
              "$j = \\text{Retail}$", 
              "$j = \\text{Wholesale}$", 
              "$j = \\text{Corporate}$", 
              "$j = \\text{Sovereign}$", 
              "$j = \\text{Banks}$")
# Independent variables
aux.1 <- c("log.gdp_gr")
aux.2 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.2)
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")
# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = filter(bank.data.win, RWA.s>0))
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

```

\centering 
```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
stargazer(reg.list,
          #title = "Capital Savings and Cyclicality",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{s} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*}}",
          #omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = fe.list,
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "tiny",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "0pt",
          float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```

## Effect on Capital Ratio

```{r , echo=FALSE}

dep.vars <- c("log.CAR_gr", 
              "log.K_gr", 
              "log.RWA.IRB_gr", 
              "log.RWA.hat_gr", 
              "log.RWA.r_gr")
dep.text <- c("$\\Delta\\text{Log Capital Ratio}_{i,t}$",
              "$\\Delta\\text{Log Capital}_{i,t}$", 
              "$\\Delta\\text{Log RWA}_{i,t}$", 
              "$\\Delta\\text{Log RWA}^{c}_{i,t}$", 
              "$\\Delta\\text{Log RWA}^{s}_{i,t}$")
# Independent variables
aux.1 <- c("log.gdp_gr")
aux.2 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.2)
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")
# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank and year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = subset(bank.data, subset = !is.na(log.CAR_gr)))
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

# Effect 
effect  <-  lapply(reg.list, function(x){
  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
})
effect[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
effect.list <-  list(c("Effect (a)", sprintf("%.1f \\%%",effect)))

#additional lines

add.list <- append(fe.list, effect.list, after = 0)
```

\centering 
```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE,out.width = "1\\textwidth"}
stargazer(reg.list,
          #title = "Capital Savings and Cyclicality",
          table.layout = "n=c-!d#-t-as-",
          notes = "\\parbox[t]{5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Component}_{i,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*}}",
          #omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = add.list,
          column.labels    = c("Panel A: Actual"),
          column.separate  = c(5),
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "tiny",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "0pt",
          float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```
\pause

```{r , echo=FALSE}

dep.vars <- c("log.CAR.til_gr", 
              "log.K_gr", 
              "log.RWA.til_gr", 
              "log.RWA.bar_gr", 
              "log.RWA.c_gr")
dep.text <- c("$\\Delta\\text{Log Capital Ratio}_{i,t}$",
              "$\\Delta\\text{Log Capital}_{i,t}$", 
              "$\\Delta\\text{Log RWA}_{i,t}$", 
              "$\\Delta\\text{Log RWA}^{c}_{i,t}$", 
              "$\\Delta\\text{Log RWA}^{s}_{i,t}$")
# Independent variables
aux.1 <- c("log.gdp_gr")
aux.2 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.2)
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")
# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank and year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = subset(bank.data, subset = !is.na(log.CAR.til_gr)))
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

# Effect 
effect.1  <-  lapply(reg.list, function(x){
  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
})
effect.1[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
effect.list <-  list(c("Effect (b)", sprintf("%.1f \\%%",effect.1)))
change      <- mapply("-", effect.1, effect, SIMPLIFY = FALSE)
change.list <-  list(c("(a) - (b)", sprintf("%.1f \\%%",change)))
#additional lines

add.list <- append(fe.list, change.list, after = 0)
add.list <- append(add.list, effect.list, after = 0)
```

\centering 
```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE,out.width = "1\\textwidth"}
stargazer(reg.list,
          #title = "Capital Savings and Cyclicality",
          table.layout = "c-!d#-t-as-",
          notes = "\\parbox[t]{5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,j,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*}}",
          omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = add.list,
          column.labels    = c("Panel B: Counterfactual"),
          column.separate  = c(5),
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "tiny",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "0pt",
          float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```

## Moments

```{r , echo=FALSE}

dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr",  "kurt.PD_gr")
dep.text <- c("Mean"      , "Variance" , "Skewness"  , "Kurtosis")
bank.data <- mutate(bank.data, crisis= ifelse(log.gdp_gr>0,0,1))
# Independent variables
aux.1 <- c("log.gdp_gr")
aux.2 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.1)
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")
# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = bank.data.win)
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

```

\centering 
```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, fig.align = 'center'}
stargazer(reg.list,
          #title = "Capital Savings and Cyclicality",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Moment}_{i,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*}}",
          #omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = fe.list,
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "tiny",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "0pt",
          float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```
## Capital Savings and Bank Characteristics
```{r, echo=FALSE}

dep.vars <- c("RWA.r",
              "RWA.r.Retail",
              "RWA.r.Wholesale",
              "RWA.r.Corporate",
              "RWA.r.Sovereign",
              "RWA.r.Banks")
dep.text <- c("$j = \\text{Total}$",
              "$j = \\text{Retail}$", 
              "$j = \\text{Wholesale}$", 
              "$j = \\text{Corporate}$", 
              "$j = \\text{Sovereign}$", 
              "$j = \\text{Banks}$")
# Independent variables
aux.1 <- c("lag(log.asset , order_by = year)",
           "lag(NPL.ratio, order_by = year)",
           "lag(capital.ratio, order_by = year)")
aux.3 <- paste(aux.1, collapse=" + ")
ind.vars <- c(aux.3)
ind.text <- c("$\\text{Log Total Assets}_{i,t-1}$",
              "$\\text{NPL Ratio}_{i,t-1}$",
              "$\\text{Capital Ratio}_{i,t-1}$",
              "$\\text{ROE}_{i,t-1}$",
              "$\\text{Capital Strigency}_{i,t-1}$",
              "$\\text{Supervisory Power}_{i,t-1}$")

# Fixed effects
fe       <- c(rep("year + bvdid"  , length(dep.vars)))
fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))

# Model formulas
formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
  felm(x, data = bank.data.win)
})

# Robust standard errors
se.list  <-  lapply(reg.list, function(x){
  summary(x, robust=TRUE)$coefficients[,2]
})

```

\centering 
```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
stargazer(reg.list,
          #title = "Capital Savings and Bank Characteristics",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{RWA}_{i,j,t}^{s} = \\beta\\times X_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*}}",
          #omit.table.layout = "n",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          se               = se.list,
          add.lines        = fe.list,
          #keep            = c("log.gdp_gr"),
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "tiny",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "-5pt",
          float            = FALSE,
          #float.env        = "sidewaystable",
          align            = TRUE,
          no.space         = FALSE)
```

## Conclusion

- Capital Savings due to the concavity of  are economically relevant.

- Countercyclical nature may offer another policy tool to reduce procyclicality of capital regulation.

- Potential risk: banks can use same effect to increase average risk while keeping capital requirements constant.

## References
\tiny

### 2020-03-16 Analysis/Presentation/Presentation_BdP.Rmd
Add presentation code

--- 2020-03-16 Analysis/Presentation/Presentation.Rmd
@@
   | Tilburg University
-date: "November 19, 2019"
+date: "February 5, 2019"
 header-includes:
@@
 
-## Banks and Regulation
+## Motivation
 
-Banks:
+- The regulation on bank capital requirements have evolved substantially since the first international standard, the 1988 Basel Accord.
 
-- Pool short-term liabilities to fund long-term assets.
+- At the heart of these changes is the link of capital charges to asset risk [@BCBS2017basel; @EBA2019policy].
 
-- Perform other important activities (payment process).
+- Basel II introduced the internal-ratings based (IRB) approach [@basel2004international].
 
-- Central role to monetary policy transmission [@bernanke1995inside].
-
-- unstable, prone to runs [@diamond1983bank].
-
-## Capital Requirements
-
-- Rationale: Compensate for public guarantees and externalities that result in banks having less capital than socially desirable.
-
-- However, there is no consensus on appropriate level and implementation [@diamond2000theory; @diamond2001liquidity; @admati2014bankers].
-
-- Hence, regulation on capital requirements have changed substantially since first international standard.
-
-- Arguably, risk sensitivity is at the center of the changes [@danielsson2001academic; @basel2004international; @EBA2019policy]
-
-## Capital Requirements and Risk Sensitivity
-
-- Basel I (July 1988): Sets fixed risk-weights for assets within same class.
-
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\text{Capital Ratio} = \frac{K}{\sum_j\alpha_j\sum_i  x_{i,j}} \geq k
-\end{aligned}
-\end{equation*}
-
-$K$: equity capital \newline
-$x_i$: asset $i$ from class $j \in \{\text{Corporate, Retail, Sovereign, etc...} \}$ \newline
-$\alpha_j$: weight for assets in class $j$ \newline
-$k$: capital ratio requirement \newline
-
-\pause
-
-- Problem: simple rule may incentivize banks to hold more risky portfolios and increase probability of failure [@koehn1980regulation]
-
-## Capital Requirements and Risk sensitivity
-
-Basel II (2004): Introduces the possibility to use internal-based rating (IRB) models to calculate risk-weights
-
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\text{Capital Ratio} = \frac{K}{\sum_j\sum_i \alpha_{i,j}  x_{i,j}} \geq k
-\end{aligned}
-\end{equation*}
-
-$\alpha_{i,j}$: weight for assets $i$, estimated using banks own risk models.
-
-\pause
-
-Rationale:
-
-- Riskier banks will hold more capital [@kim1988risk].
-- Alignment of risk estimates to actual risk [@barakova2014banks].
-
-\pause
-
+## Motivation
 Problems:
 
-- Enhances procyclical nature of lending [@behn2016procyclical].
-- Allow differences for reasons other than risk [@mariathasan2014manipulation; @behn2016limits].
+ - Linking capital charges to asset risks exacerbates the **procyclicality** of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
 
-## Credit Risk Distribution and Capital Savings
+ - The inherent flexibility of the IRB approach may allow for differences in capital requirements that do not reflect portfolio risk but rather modelling choices [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits; @berg2017analysis; @ferribank].
 
-- This paper contributes to the understanding of risk-sensitive capital requirements ($\alpha_{i,j}$).
+## Contribution
 
-- I identify a novel effect stemming from the design of the IRB framework: banks save capital requirements because mapping from risk (PD) is concave.
+- This paper contributes to understanding the cyclicality and variability of capital requirements.
 
-- I collect data and propose a measure of this effect.
+- I identify the effects of an unexplored feature of the IRB framework on capital requirements: 
+  - Banks save capital requirements because mapping from probability of default (PD) to capital requirements is concave.
 
-- Key findings:
-  - Reduces the procyclicality of RWA by 11.2 percentage points. 
-  - Explains 27% of RW variation across banks.
-
-## Concavity of Capital Requirements
+## Actual Capital Requirements
 
@@
 ```
+
 ```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}
@@
 p + geom_line(alpha = 1) + 
+  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=1)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=5)+
   labs(x = "PD", 
@@
   expand_limits(x = 0, y = 0)+
-  scale_x_continuous("",breaks=c(PD.h+0.03),labels=c("PD"),expand = c(0, 0))+
-  scale_y_continuous("",breaks=c(K.h),labels=c("RWA"),expand = c(0, 0))
+  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h,PD.h+0.045),labels=c("Low PD","Mean PD","High PD","PD"),expand = c(0, 0))+
+  scale_y_continuous("",breaks=c(K.s,K.h+0.03),labels=c(expression(paste(RW^{a})),"RW"),expand = c(0, 0))
 
 ```
-## Concavity of Capital Requirements
+
+## Counterfactual Capital Requirements
 ```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}
@@
   geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=5)+
+  labs(x = "PD", 
+         y = "Risk-weight") +
   theme_bw()+
@@
   expand_limits(x = 0, y = 0)+
-  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
-  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("",""),expand = c(0, 0))
+  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h,PD.h+0.045),labels=c("Low PD","Mean PD","High PD","PD"),expand = c(0, 0))+
+  scale_y_continuous("",breaks=c(K.a,K.h+0.03),labels=c(expression(paste(RW^{c})),"RW"),expand = c(0, 0))
 
@@
 
-## Concavity of Capital Requirements
+## Capital Savings
 ```{r, fig.height = 3, fig.width = 4,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE}
@@
   expand_limits(x = 0, y = 0)+
-  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h),labels=c("Low PD","Mean PD","High PD"),expand = c(0, 0))+
-  scale_y_continuous("",breaks=c(K.s,K.a),labels=c("RWA",expression(paste(RWA^{c}))),expand = c(0, 0)) +
+  scale_x_continuous("",breaks=c(PD.l,PD.a,PD.h,PD.h+0.045),labels=c("Low PD","Mean PD","High PD","PD"),expand = c(0, 0))+
+  scale_y_continuous("",breaks=c(K.s,K.a,K.h+0.03),labels=c(expression(paste(RW^{a})),expression(paste(RW^{c})),"RW"),expand = c(0, 0)) +
   annotate("text", x = (PD.l+PD.a)/3 , y = (K.s+K.a)/2, label = "Capital \n Savings", size = 3)
@@
 ```
-## Capital Savings Measure
+
+## Results
+
+- Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component (*Capital savings*).
 
@@
 \label{eq5}
-\text{RWA}^s = \frac{\text{RWA}^c}{\text{RWA}} 
+\text{RW}^s = \frac{\text{RW}^c}{\text{RW}^c} 
 \end{aligned}
@@
 
-\pause
+- Key findings:
+  - *Capital savings* explain 27% of RW variation across banks.
+  - 1 p.p. lower GDP growth increases *Capital savings* growth by 0.49 p.p. on average.
+  - *Capital savings* reduce the procyclicality of capital ratio by 11.2 p.p. on average.
 
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\text{RWA}^{\$} = \text{RWA}^c-\text{RWA} 
-\end{aligned}
-\end{equation*}
+## Outline
+
+- Data
+
+- Relative importance of *Capital Savings* to RWs
+
+- Countercyclicality of *Capital Savings*
+
+- Conclusion
 
@@
     labs(x=NULL) +
-  labs(y =  "RWA savings ratio") +
+  labs(y =  expression(paste(RW^{s})-1)) +
    theme_bw()+
@@
     labs(x=NULL) +
-  labs(y =  "RWA savings ratio") +
+  labs(y =  expression(paste(RW^{s})-1)) +
   coord_flip() +
@@
 
-## Decomposition
+## RW Decomposition
 
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\text{RWA} & = \text{RWA}_{SA}+\text{RWA}_{IRB} \\
-\pause
-& = \text{RWA}_{SA}+\text{RWA}^{c}-\text{RWA}^{\$}\\
-\pause
-& = \text{RWA}_{SA}+\sum_{j=1}^P\text{RWA}^{c}_j-\text{RWA}^{\$}\\
-\end{aligned}
-\end{equation*}
+- I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
 
-## Decomposition
+- Components of RW:
+  - *Capital savings* components:
+    - Intensive margin
+    - Extensive margin (roll-out)
+  - Average IRB components:
+    - Intensive margin
+    - Extensive margin (roll-out)
+    - Portfolio mix
+  - Standardized approach component
 
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\frac{\text{RWA}}{\text{EAD}}&= \frac{\text{RWA}_{SA}}{\text{EAD}_{SA}}\frac{\text{EAD}_{SA}}{\text{EAD}}+\frac{\text{EAD}_{IRB}}{\text{EAD}}\sum_{j=1}^P\frac{\text{EAD}_{IRB,j}}{\text{EAD}_{IRB}}\frac{\text{RWA}^{c}_j}{\text{EAD}_{IRB}}+\frac{\text{RWA}^{\$}}{\text{EAD}_{IRB}}\frac{\text{EAD}_{IRB}}{\text{EAD}}\\
-\pause
-\text{RW} &= \text{RW}_{SA}q_{SA}+(1-q_{SA})\sum_{j=1}^P q_{j}\text{RW}^{c}_j +(1-q_{SA})\text{RW}^{s}\\
-\end{aligned}
-\end{equation*}
+## Cross-sectional Decomposition
 
-## Decomposition
+```{r, fig.height = 5, fig.width = 8,out.width = "90%", results="hide", include=TRUE,fig.align = 'center'}
+cross.section.decomposition %>% filter(q.SA<1)%>%
+   mutate(total= -PD.gain)%>%
+  select(name, Country,total, int.gain, ext.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
+                                       labels = c("Intensive margin", "Extensive margin")))%>%
+ggplot(mapping = aes(x = reorder(name, total),
+                       y = value)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL, y= NULL) +
+    theme_bw()+
+  theme(panel.grid = element_blank(),
+         axis.text.x = element_text(angle = 45, hjust = 1))+
+  facet_wrap(~condition, nrow = 2, ncol = 1)
 
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\text{RW}-\text{RW}^B =& \text{RW}_{SA}q_{SA}+(1-q_{SA})\sum_{j=1}^P q_{j}\text{RW}^{c}_j +(1-q_{SA})\text{RW}^{s}\\
-&- \text{RW}_{SA}^Bq_{SA}^B+(1-q_{SA}^B)\sum_{j=1}^P q_{j}^B\text{RW}^{c,B}_j +(1-q_{SA})\text{RW}^{s,B}
-\end{aligned}
-\end{equation*}
+```
 
-## Decomposition
+## Cross-sectional Decomposition
 
-\begin{equation*}
-\begin{aligned}
-\label{eq5}
-\Delta \text{RW} & = \underbrace{q^{B}_{SA} \Delta \text{RW}_{SA}}_\text{SA mix}+\underbrace{(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j}_\text{Mean IRB Risk}\\
-&\\
-&+ \underbrace{(1-q^{B}_{SA})\sum_{j=1}^{P} \text{RW}_{j}^{c} \Delta q^{j}}_\text{IRB mix}  + \underbrace{(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}}_\text{Roll-out}\\
-&\\
-&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin}
-\end{aligned}
-\end{equation*}
+```{r, fig.height = 3, fig.width = 4, results="hide",out.width = "80%", include=TRUE,fig.align = 'center'}
+cross.section.decomposition %>% 
+  filter(q.SA<1)%>%
+  select(name, Country, int.gain, ext.gain) %>%
+  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
+  ggplot(mapping = aes(x = int.gain,
+                       y = ext.gain)) +
+    geom_point()+
+    geom_smooth(method = "lm", color = "black")  +
+  labs(x=NULL, y= NULL) +
+   theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  scale_fill_grey()+
+  stat_cor(label.y= -0.1)
+```
 
@@
 
-## Cross-sectional Decomposition
-
-```{r, fig.height = 5, fig.width = 8,out.width = "90%", results="hide", include=TRUE,fig.align = 'center'}
-cross.section.decomposition %>% filter(q.SA<1)%>%
-   mutate(total= -PD.gain)%>%
-  select(name, Country,total, int.gain, ext.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
-                                       labels = c("Intensive margin", "Extensive margin")))%>%
-ggplot(mapping = aes(x = reorder(name, total),
-                       y = value)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL, y= NULL) +
-    theme_bw()+
-  theme(panel.grid = element_blank(),
-         axis.text.x = element_text(angle = 45, hjust = 1))+
-  facet_wrap(~condition, nrow = 2, ncol = 1)
-
-```
-
-## Cross-sectional Decomposition
-
-```{r, fig.height = 3, fig.width = 4, results="hide",out.width = "80%", include=TRUE,fig.align = 'center'}
-cross.section.decomposition %>% 
-  filter(q.SA<1)%>%
-  select(name, Country, int.gain, ext.gain) %>%
-  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
-  ggplot(mapping = aes(x = int.gain,
-                       y = ext.gain)) +
-    geom_point()+
-    geom_smooth(method = "lm", color = "black")  +
-  labs(x=NULL, y= NULL) +
-   theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  scale_fill_grey()+
-  stat_cor(label.y= -0.1)
-```
-
 ## R-squared decompostion
@@
                       levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
-                      labels = c("PD avg.", "PD ext.", "PD int.", "SA", "Mix", "Rollout"))) %>%
+                      labels = c("PD avg.", "Ext. comp.", "Int. comp.", "SA", "Mix", "Rollout"))) %>%
 ggplot(mapping = aes(x = reorder(component, -share),
-                       y = share)) +
+                       y = share, fill = component)) +
 geom_bar(stat = 'identity')+
@@
     theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  theme(strip.background = element_blank(),
-       # strip.text = element_blank(),
+  theme(panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+       # strip.text = element_blank()
+       legend.position = "none",
         axis.text.x = element_text(angle = 45, hjust = 1))+
-  scale_fill_grey()+
+  scale_fill_manual(values = c("#C3BCB2","#353535",  "#353535", "#C3BCB2","#C3BCB2", "#C3BCB2"))+
 scale_y_continuous(labels = scales::percent_format(accuracy = 1,
@@
 ```
-## Capital Savings and Bank Characteristics
-```{r, echo=FALSE}
-
-dep.vars <- c("RWA.r",
-              "RWA.r.Retail",
-              "RWA.r.Wholesale",
-              "RWA.r.Corporate",
-              "RWA.r.Sovereign",
-              "RWA.r.Banks")
-dep.text <- c("$j = \\text{Total}$",
-              "$j = \\text{Retail}$", 
-              "$j = \\text{Wholesale}$", 
-              "$j = \\text{Corporate}$", 
-              "$j = \\text{Sovereign}$", 
-              "$j = \\text{Banks}$")
-# Independent variables
-aux.1 <- c("lag(log.asset , order_by = year)",
-           "lag(NPL.ratio, order_by = year)",
-           "lag(capital.ratio, order_by = year)")
-aux.3 <- paste(aux.1, collapse=" + ")
-ind.vars <- c(aux.3)
-ind.text <- c("$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{ROE}_{i,t-1}$",
-              "$\\text{Capital Strigency}_{i,t-1}$",
-              "$\\text{Supervisory Power}_{i,t-1}$")
-
-# Fixed effects
-fe       <- c(rep("year + bvdid"  , length(dep.vars)))
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-\centering 
-```{r , eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-stargazer(reg.list,
-          #title = "Capital Savings and Bank Characteristics",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{RWA}_{i,j,t}^{s} = \\beta\\times X_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*}}",
-          #omit.table.layout = "n",
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "tiny",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "tab1",
-          style            = "qje",
-          column.sep.width = "-5pt",
-          float            = FALSE,
-          #float.env        = "sidewaystable",
-          align            = TRUE,
-          no.space         = FALSE)
-```
 
@@
 
-- Capital Savings due to the concavity of  are economically relevant.
+- *Capital savings* due to the concavity are economically relevant and countercyclical.
 
-- Countercyclical nature may offer another policy tool to reduce procyclicality of capital regulation.
+- Policy implication: Countercyclical nature of *Capital savings* offers another policy tool to reduce procyclicality of capital regulation.
 
-- Potential risk: banks can use same effect to increase average risk while keeping capital requirements constant.
+- Potential risk: banks can use the same effect to increase average risk while keeping capital requirements constant.
 

### 2020-03-16 Analysis/Draft/Draft.Rmd
Update paper draft

@@
 ---
-title: "Credit risk distribution and capital savings"
+title: "Credit Risk Distribution and Capital Savings"
 author:
@@
     template: tex-ms.tex
-    fig_caption: yes
+    fig_caption: true
   html_document:
@@
 citecolor: black
-anonymous: no
+anonymous: false
 fontsize: 11pt
+spacing: "double"
 geometry: margin= 1in
-keywords: Internal ratings-based models, Risk-weighted assets dispersion, Regulatory arbitrage
+keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
 JEL: G21, G28, G11
@@
 subtitle: ''
-thanks: Very preliminary draft
-abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requiremets for banks adopting the internal ratings-based (IRB) approach.
-As a consequence, total capital requirement depends not only on the average PD but on the shape of the PD distribution.
-Using a novel hand-collected dataset, I propose a measure that disantangles the contribution of the PD distribution shape from the average PD effect to capital requirements.
-I document that the latter is an important factor of risk-weight variability across banks.
-I further investigate the prior ambiguous direction of its relationship with the business cycle and find it to be countercyclical.
-Altogether, my results show that issues related to RW manipulation and procyclicality of the IRB framework are not as severe as the literature suggests since the effect of the PD distribution shape was not taken into account."
-toc: no
-urlcolor: null
+thanks: Preliminary draft.  
+abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach.
+Therefore, total capital requirements depends not only on the portfolio's average PD but also on the shape of its PD distribution.
+Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
+While a larger average component increases capital charges, a larger shape component generates capital savings.
+I document that these capital savings are important sources of risk-weight variability across banks.
+I further investigate the relationship of capital savings with the business cycle and find it to be countercyclical.
+The results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements."
+urlcolor: black
 ---
@@
 library(ggpubr)               ## More plots
+library(textablr)
+library(sandwich)
+library(lmtest)
 #library(relaimpo)            ## Calculate relative importance measures
@@
 
-Bank capital adequacy requirements have evolved substancially since the first international standard, the 1988 Basel Accord.
-Arguably, at the center of these changes is the link of capital charges to asset risk [@basel2004international; @BCBS2017basel; @EBA2019policy].
-Risk sensitivity of capital requirements was present in the first accord and it was greatly enhanced with the Basel II framework by giving banks the option to calculate capital requirements using risk estimates obtained from their own models.
-The rationale for risk-linked capital charges is that regulating capital via a simple capital to asset ratio can incentivize banks to take on risk [@koehn1980regulation; @kim1988risk; @gordy2010risk].
-The rationale for allowing internal models are the better alignment of risk estimates to actual portfolio risk [@barakova2014banks] and the promotion of better risk management practices [@cucinelli2018credit].
-However, these same features may have adverse consequences.
-First, linking capital charges to asset risk exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
-Second, the inherent flexibility of the internal ratings-based (IRB) approach may allow for differences in capital requirements that do not reflect portfolio risk [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits; @ferribank].
-This paper contributes to the understanding of cyclicality and unintended variability of capital requirements by exploring the consequences of a feature of the IRB framework, which, to the best of my knowledge, I am the first to identify.
-
-Under the IRB approach banks supply their own risk estimates to regulatory formulas which determine the amount of capital banks have to hold for each asset.
-Importantly, the mapping from probability of default (PD) and capital requirements is concave and capital requirements are determined on the asset level with total requirement being simply the sum of assets' individual requirement.
-This means, by Jensen’s inequality, that the sum of capital charges for two assets is lower than the charge for their average (in terms of PD).
-More generally, total capital requirement depends not only on banks' portfolio average PD but also on the shape of the PD distribution.
-For instance, given an average PD, more disperse portfolios yield lower capital charges.
-I call the difference in capital requirements stemming from differences in the shape of the portfolio PD distribution, while holding the location parameter constant, the PD distribution effect.
-
-The first contribution of this paper is to introduce a dataset and method to quantify the PD distribution effect.
-The method is straightforward: compare the actual capital requirement, which depend on the PD distribution, to a counterfactual case where the requirement for the entire portfolio is calculated using the average PD.
-*Capital savings* is the difference between the two values and it measures the intensity of the PD distribution effect.
-
-To calculate the *Capital savings* measure, one needs information on the distribution of banks' regulatory risk parameters.
-However, data at such granularity is not available in the common balance sheet datasets, and although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
-Therefore, to overcome data limitations, I hand-collected information on the distribution of credit risk parameters from banks' Pillar-3 reports.
-The sample consists of 49 large banking groups that have adopted the IRB approach.
-The dataset covers 10 countries and goes from 2007 to 2018.
-
-With this unique dataset, I start the analysis by investigating the contribution of the PD distribution effect to banks' overall risk-weight (RW).
-Several studies document significant RW variability across otherwise similar banks.
+The regulation on bank capital requirements have evolved substantially since the first international standard, the 1988 Basel Accord.
+At the heart of these changes is the link of capital charges to asset risk [@basel2004international; @BCBS2017basel; @EBA2019policy].
+Risk sensitivity of capital requirements was present in the first accord and it was greatly enhanced in the Basel II framework by giving banks the option to calculate capital requirements using risk estimates obtained from their own models, known as the internal ratings-based (IRB) approach.
+The rationale for risk-linked capital charges is that regulating capital via a simple capital to asset ratio can incentivize banks to take on excessive risk [@koehn1980regulation; @kim1988risk; @gordy2010risk].
+The rationale for allowing internal models is the better alignment of risk estimates to actual portfolio risk [@barakova2014banks] and the promotion of better risk management practices [@cucinelli2018credit].
+
+However, both risk sensitivity and the use of internal models may have adverse consequences.
+First, linking capital charges to asset risks exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
+As the overall credit quality detiorates during economic downturns, capital requirements tighten, forcing banks to cut down lending.
+Second, the inherent flexibility of the IRB approach may allow for differences in capital requirements that do not reflect portfolio risk but rather modelling choices [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits;@berg2017analysis; @ferribank].
+
+This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements.
+Under the IRB approach, banks supply their own risk estimates, such as probability of default (PD), to formulas set by the regulation which determine the amount of capital banks have to hold.
+Two characteristics of the framework are important for this analysis.
+First, the mapping from PD estimates to capital requirements is a concave function.
+Second capital requirements are calculated for each asset separately and then added up to reach total capital charges.
+Together, these characteristics imply that actual capital requirements are lower than if they were calculated using the average PD to the entire portfolio.
+For clarity, consider two portfolios.
+The first contains a low PD asset and a high PD asset in equal proportions.
+The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD.
+Because of the concavity of the IRB formula, capital requirements for the second portfolio are greater than the first.
+More generally, total capital requirements depend not only on banks' portfolio average PD but also on the shape of the PD distribution.
+
+I propose a measure that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
+The method is straightforward: for each bank-year combination, compare actual capital requirements to the counterfactual case where requirements for the entire portfolio are calculated using the portfolio’s average PD.
+I label the shape component *Capital savings* because it reduces total capital requirements.
+
+To calculate measures of *Capital savings*, information on the distribution of banks' regulatory risk parameters is needed.
+I hand-collected information on the distribution of credit risk parameters from banks' Pillar-3 reports.
+As is common in the literature, I calculate *Capital savings* in terms of risk-weighted assets (RWA), which are simply capital requirements times 12.5, or risk-weights (RW), which are RWA divided by assets.
+The sample consists of 52 large banking groups that have adopted the IRB approach.
+The dataset covers 13 countries and goes from 2007 to 2018.
+Data at such granularity is not available in the common balance sheet datasets, and although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
+
+With this novel dataset, I start by documenting the relative importance of *Capital savings* to banks' overall RW.
+By extending the decomposition method proposed by @cannata2012inside, I find that *Capital savings* accounts for 27% of the RW variation across banks.
+I also find that *Capital savings* measures are negatively correlated to other RW components.
+This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
+Lastly, although banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach (known as the roll-out effect), the intensity of *Capital savings* has, on average, decreased in time.
+That is, for the same average PD, capital requirements have increased from 2008 to 2018.
+
+I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
+Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
+However, as credit conditions deteriorate (and banks react to these new credit conditions) the PD distribution and, consequently, the amount of *Capital savings*, also change.
+A priori, the direction of the business cycle effect on *Capital savings* is ambiguous.
+At one extreme, if the entire portfolio is equally affected, then there is only a shift of the location parameter of the PD distribution.
+Because the PD-RW formula elasticity is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
+On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
+The ambiguity arises because *Capital savings* can only increase during a recession if the shape of the PD distribution is affected during the cycle.
+
+I find *Capital savings* to be countercyclical.
+I estimate that a 1.7% GDP growth rate (one standard deviation) decreases *Capital savings* by 0.83 percentage points on average.
+In other words, banks have to increase capital by 0.83 percentage points to keep their capital ratio constant.
+The effect is economically significant as it represents a quarter of the capital ratio standard deviation in the sample.
+In a capital ratio decomposition exercise, I also find that the the cyclicality of *Capital savings* reduces the procyclicality of capital ratios by 11.6%.
+The results are robust to different measures of the shape component, such as, *Capital savings* in terms of ratios, *Capital savings* in monetary units, and using Gini coefficients.
+
+I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic expasions average PD and PD variance decrease while the PD kurtosis and PD skewness increase.
+This evidence implies that distributions become more symmetric and with less extreme values during recessions, suggesting a flight to safety behaviour by banks.
+As the recession reduces their existing portfolio credit quality (evidenced by the negative coefficient of average PD) banks respond by reducing risk-taking.
+
+This papers contributes to two strands of the banking literature.
+First, several studies document significant RW variability across otherwise similar banks.
 Some take this evidence as indicative of either excessive subjectivity in the current capital requirement rules or risk measurement manipulation by banks to decrease capital charges [see, @mariathasan2014manipulation; @ferribank].
-Others are more cautious and emphasize that part of RW variability is desirable since it reflects differences in risk among banks [@cannata2012inside; @BCBS2016regulatory; @EBA2019result].
-I contribute to the literature by showing that the PD distribution effect is an important factor of the RW variation across banks.
-By extending the decomposition method proposed by @cannata2012inside I find that the PD distribution effect accounts for 28% of RW variation across banks.
-Moreover, the inclusion of the PD distribution effect in the decomposition reduces the variation explained by the component most susceptible of manipulation by 17.7 percentages points.
-This result implies that previous studies in the literature are potentially overestimating subjectivity and manipulation of RWs.
-
-I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
-A consequence of the shape of the regulatory formula is that it affect the cyclicality of capital requirements.
-@behn2016procyclical document that IRB portfolios are more procyclical than portfolios under the standardized approach.
-During economic downturns credit risk increases which raises capital requirements under the IRB approach.
-This characterizes the IRB approach as procyclical since banks have to cut down lending exactly when borrowers need it the most.
-We can interpret this evidence as the result of shifts to the location parameter of the PD distribution.
-However, shocks to credit risk can also affect the shape of the distribution.
-For instance, riskier borrowers prior to the shock can be more adversely affected compared to safer borrowers causing the PD distribution to become more fat-tailed.
-If that is the case, the amount of *Capital savings* will be affected.
-
-I find the PD distribution effect to be countercyclical.
-I estimate that a 2.1% GDP growth rate (one standard deviation) decreases *Capital savings* by 5.9% on average.
-In other words, banks have to increase capital by 5.9% to keep their capital ratio constant.
-The effect is economically significant as it represents a quarter of the *Capital savings* standard deviation in the sample.
-I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases.
-This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
-
-Several cross‐sectional tests further strengthen our results.
-Increases in capital requirements are more problematic for less well capitalized banks, which consequently adjust their IRB loans more in response to the credit risk shock.
-This suggests that our findings would be less pronounced if banks kept larger buffers over the regulatory minimum.
-Furthermore, the IRB effect is more pronounced for less profitable firms, which are likely to be more affected by the credit risk shock, and for firms to which IRB banks have a large exposure.
-As robustness, I calculate the gini coefficient for the PD distribution as an alternative measure of the PD distribution effect and I find similar results.
+Others are more cautious and emphasize that part of RW variability is desirable since it reflects differences in risk among banks [@cannata2012inside; @BCBS2016regulatory;@berg2017analysis; @EBA2019result].
+I contribute to the literature by showing that *Capital savings* are an important source of the RW variation across banks.
+
+Second, while the literature shows that capital requirements, especially for portfolios under the IRB approach, are on average procyclical [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical], I identify a countercyclical component of the current IRB framework.
+A better understanding of this countercyclical component can help policy makers designing a less procyclical regulation in the future.
 
 The remainder of this paper is organized as follows.
-Section 2 describes the capital requirement formula, the capital savings measure, and the hypothesis concerning this measure.
-Section 3 describes the data collected, and analyses the relative importance of capital savings in explaining differences in risk-weights across banks and time.
-Section 4 explains the empirical strategy.
-Section 5 presents evidence on the anticyclical feature of capital savings while Section 6 shows evidence of banks using this feature strategically to reduce capital charges.
-Finally, section 7 concludes.
+Section 2 describes the capital requirement formula, the *Capital savings* measure, and the hypothesis concerning this measure.
+Section 3 describes the collected data and analyses the relative importance of *Capital savings* in explaining differences in risk-weights across banks and time.
+Section 4 presents evidence on the countercyclical feature of *Capital savings*.
+Finally, section 5 concludes.
 
@@
 
-In the IRB approach, capital requirements are derived from risk-weight formulas that take as input banks' own estimates of credit risk drivers.
+In the IRB approach, capital requirements are derived from risk-weight formulas that take as input banks' own estimates of credit risk parameters.
 The basic regulatory formula derives from a model characterized by the property that the risk-weight of each asset depends only on the estimated credit risk parameters for this asset and not on the composition of the portfolio, i.e.
@@
 
-Equation \ref{eq1} is the basic capital requirement formulas:
+Equation \ref{eq1} is the basic risk-weight formula:
 
@@
 \label{eq1}
-\text{K}= \sum_{i}EAD_iLGD_i N\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(PD_i)+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg) ,
+\text{RW}= 12.5\frac{\sum_{i}\text{EAD}_i\text{LGD}_i N\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\text{PD}_i)+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)}{\sum_{i}\text{EAD}_i} ,
 \end{equation}
 
-where the summation is over the assets and $N$ is the standard normal cumulative distribution.
-Basel accords segment credit portfolio into several categories, each having different additional features (maturity adjustments, risk-dependent correlation coefficient, among others) but without altering the basic structure shown in equation \ref{eq1}.
+where $i$ is an asset in the portfolio and $N$ is the standard normal cumulative distribution.
+Equation \ref{eq1} calculates the average RW for the porftolio.
+The total amount of RWA for this portfolio is $\text{RW}\times \sum_i\text{EAD}_i$ and total capital requirements $\text{RWA}\times 8\%$.
+
+Basel accords segment credit portfolio into several categories, each having different additional features (maturity adjustments, risk-dependent correlation coefficient, among others) while keeping the basic structure of equation \ref{eq1}.
 The credit risk parameters estimated by banks, which are used as input on equation \ref{eq1}, are: Exposure at default (EAD), loss given default (LGD), and probability of default (PD).
@@
 
-For the purpose of this paper, I focus on the mapping from PD to capital requirements.
-Figure \ref{fig:pdmap} shows that capital requirements is a concave function of PD.
-This means, by Jensen’s inequality, that the sum of capital charges for two assets is lower than the charge for their average (in terms of PD).
-More generally, the distribution of credit risk matters to calculate capital requirements with, for instance, disperse portfolios yielding relatively lower capital charges.
-
-```{r parameters and data, echo=FALSE, include=FALSE}
+Figure \ref{fig:pdmap} plots the mapping from PD to RW according to equation \ref{eq1}.
+For my analysis, the crucial characteristic is that this mapping is concave.
+The concavity of the PD-RW mapping together with the bottom-up method to aggregate capital requirements imply that actual capital requirements are lower than if they were calculated using the average PD to the entire portfolio.
+For clarity, consider two portfolios.
+The first contains a low PD asset and a high PD asset in equal proportions.
+The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD.
+Because of the concavity of the IRB formula, by Jensen’s inequality, the capital requirement for the second portfolio is greater than the first.
+More generally, total capital requirements depend not only on banks' portfolio average PD but also on the shape of the PD distribution.
+Figure \ref{fig:pdmap} illustrates the comparison.
+$\text{RW}^a$ and $\text{RW}^c$ are the average RW for the first portfolio and the second portfolio, respectively.
+The difference between  $\text{RW}^c$ and $\text{RW}^a$ is the monetary amount of *Capital savings* per unit of EAD due to the concavity of the IRB formula.
+
+```{r parameters curve, echo=FALSE, include=FALSE}
 set.seed(123)
@@
 data <- mapping_retail(grid,1,0.15,0.999)
+PD.l <- 0.01
+PD.h <- 0.15
+PD.a <- (PD.l+PD.h)/2
+PD.h2 <- 0.17
+PD.h3 <- 0.16
+PD.l3 <- 0.02
+PD.a2 <- (PD.l+PD.h2)/2
+PD.i  <- 0.19
+K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
+K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
+K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
+K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
+K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
+K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
+K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
+K.i  <- as.numeric(mapping_retail(PD.i,1,0.15,0.999)["K"])
+K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
+K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
+K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
+aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
+                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
+                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
 ```
@@
 p + geom_line(alpha = 1) + 
+  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=1)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=5)+
+  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.5)+
+
   labs(x = "PD", 
-         y = "Capital requirement per unit of exposure") +
-  theme_bw()
-
-```
-
-The first contribution of this paper is a measure of capital requirement reduction due to the concavity of the regulatory formula.
-The method behind the measure is simple: compare actual capital requirement to a counterfactual case where the requirement for the entire portfolio is calculated using the average PD.
-In practice, I calculate differences in risk-weighted assets (RWA) rather than capital requirements because it is the former that is used to calculate capital ratios.
-I define the counterfactual risk-weighted assets as:
+         y = "Risk-weight") +
+  theme_bw()+
+  theme(axis.ticks = element_blank(),
+        axis.line = element_line(),
+        axis.title.x = element_blank(),
+        panel.grid = element_blank())+
+  expand_limits(x = 0, y = 0)+
+  scale_x_continuous(breaks=c(PD.l,PD.a,PD.h,PD.i),labels=c("Low PD","Mean PD","High PD","PD"),expand = c(0, 0))+
+  scale_y_continuous("",breaks=c(K.s,K.a,K.i),labels=c(expression(paste(RW^{a})),expression(paste(RW^{c})),"RW"),expand = c(0, 0)) +
+  annotate("text", x = (PD.l+PD.a)/3 , y = (K.s+K.a)/2, label = "Capital \n Savings", size = 3)
+
+```
+
+I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
+The method is a generalization of the example in Figure \ref{fig:pdmap}: for each bank-year combination, compare the actual capital requirement to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio’s average PD.
+
+I define the counterfactual RW as:
 
@@
 \label{eq2}
-\text{RWA}^c= 12.5\sum_{i}EAD_iLGD_iN\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\overline{PD})+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg),
+\text{RW}^c= 12.5\frac{\sum_{i}\text{EAD}_i\text{LGD}_iN\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\overline{\text{PD}})+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)}{\sum_i \text{EAD}_i},
 \end{equation}
-
-where the probability of default parameter for all assets is replaced by the portfolio weighted average:
+in which the PD parameter for all assets is replaced by the portfolio weighted average:
 
 \begin{equation*}
-\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
+\overline{\text{PD}}=\frac{\sum_i\text{EAD}_i\text{LGD}_i\text{PD}_i}{\sum_i\text{EAD}_i\text{LGD}_i}.
 \end{equation*}
 
-The *Capital savings* measure is the difference between equation \ref{eq1} (multiplied by 12.5 to obtain RWA) and equation \ref{eq2}:
+The main measure of *Capital savings* used in this paper is the ratio of equation \ref{eq2} to equation \ref{eq1} as follows:
+
+\begin{equation}
+\label{eq4}
+\text{RW}^r=\frac{\text{RW}^c}{\text{RW}}.
+\end{equation}
+
+$\text{RW}^r$ is the rate at which a bank saves capital.
+For instance, a value of 0.5 means this portfolio would require 50% more capital in the abscense of *Capital savings* to keep the same capital ratio.
+A value of zero means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
+Any PD distribution with positive variance results in positive *Capital savings*.
+
+An alternative way to measure *Capital savings* is in monetary units as follows:
 
@@
 
-*Capital savings* ($\text{RWA}^s$) is the amount that banks saves in capital in monetary units.
-A value of zero for this measures means that there is no capital savings, and implies a portfolio with a degenerate PD distribution.
-Any PD distribution with positive variance will yield a positive value of *Capital savings*.
-To avoid size effects and dealing with currency conversion I also consider a second measure, the *Capital savings ratio*, defined as follows:
-
-\begin{equation}
-\label{eq4}
-\text{RWA}^r=\frac{\text{RWA}^c}{\text{RWA}}.
-\end{equation}
-
-*Capital savings ratio* ($\text{RWA}^r$) is the rate at which a bank saves capital.
-For instance, a value of 0.5 means this portfolio would require 50% more capital in the abscense of the PD distribution effect to keep the same capital ratio.
-As with *Capital savings*, a value of zero means no capital savings, and any PD distribution with positive variance will result in a positive value.
-
 ## Capital Savings and Cyclicality
 
-A consequence of the shape of the regulatory formula is that it affect the cyclicality of capital requirements.
-@behn2016procyclical document that IRB portfolios are more procyclical than portfolios under the standardized approach.
-During economic downturns credit risk increases which raises capital requirements under the IRB approach.
-This characterizes the IRB approach as procyclical since banks have to cut down lending exactly when borrowers need it the most.
-We can interpret this mechanism as shifts to the location parameter of the PD distribution.
-However, shocks to credit risk can also affect the shape of the distribution.
-For instance, riskier borrowers prior to the shock can be more adversely affected compared to safer borrowers, causing the PD distribution to become more fat-tailed.
-As a result, the amount of capital savings will be affected.
-
-Whether capital savings are pro- or anticyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the credit risk distribution.
-I use a simple example to illustrate this ambiguity.
-Consider a bank with the asset side of its balance sheet composed of \$100 of a safer asset and \$100 of a riskier asset and three states of the economy.
-Without loss of generality, assume $LGD=1$ and $R=0.15$.
-In the initial state, safer and riskier assets have a 0.1% and 0.2% probability of default, respectively.
-According to equation \ref{eq1}, the actual RWA of this bank is \$63.9.
-The weighted average of this portfolio in terms of $PD$ is 0.15% which yields a counterfactual RWA of \$64.7 according to equation \ref{eq2}.
-Hence, in the initial state capital savings equals \$0.8.
-
-Consider now a credit risk shock.
-In state A, both safer and riskier assets are affected, having their probabilities of default raised to 0.2% and 0.3%, respectively.
-In state B, only the riskier asset is affected, now with a probability of default equal to 0.4%.
-Note that the average $PD$ in both states are equal at 0.25%.
-However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B.
-The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%).
-Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states.
-This is the result of IRB procyclicality shown by @behn2016procyclical.
-However, due to the increased portfolio dispersion, capital savings are anticyclical in state B.
-Conversely, while higher-order moments remained constant in state A, the PD distribution moved to a less concave part of the regulatory function consequently reducing capital savings.
-Therefore, capital savings are procyclical in state A.
-Table \ref{tab1} summarizes the example.
+I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
+Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
+However, as credit conditions deteriorates (and banks react to these new credit conditions) the PD distribution and, consequently, the amount of *Capital savings* also change.
+A priori the direction of the economic cycle effect on *Capital savings* is ambiguous.
+At one extreme, if the entire portfolio is equally affected, then there is only a shift of the location parameter of the PD distribution.
+Because the PD-RW formula elasticity is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
+On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
+The ambiguity arises because *Capital savings* can only increase during a recession if the shape of the PD distribution is affect during the cycle.
 
@@
 
-\newpage
+I use a simple example to illustrate this ambiguity.
+Consider a bank with the asset side of its balance sheet composed of \$100 of a safer asset and \$100 of a riskier asset and three states of the economy.
+Without loss of generality, assume $\text{LGD}=1$ and $\text{R}=0.15$.
+In the initial state, safer and riskier assets have a 0.1% and 0.2% probability of default, respectively.
+According to equation \ref{eq1}, the actual RWA of this bank is \$63.9.
+The weighted average of this portfolio in terms of $\text{PD}$ is 0.15% which yields a counterfactual RWA of \$64.7 according to equation \ref{eq2}.
+Hence, in the initial state *Capital savings* equal \$0.8.
+
+```{r parameters curve 2, echo=FALSE, include=FALSE}
+set.seed(123)
+LGD <- 1
+M   <- 2.5
+grid  <- seq(0.0001, 1, by=0.00001) 
+data <- mapping_retail(grid,1,0.15,0.999)
+PD.l <- 0.01
+PD.h <- 0.13
+PD.a <- (PD.l+PD.h)/2
+PD.h2 <- 0.19
+PD.h3 <- 0.16
+PD.l3 <- 0.04
+PD.a2 <- (PD.l+PD.h2)/2
+K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
+K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
+K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
+K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
+K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
+K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
+K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
+K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
+K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
+K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
+aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
+                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
+                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
+```
+
+```{r}
+cyclicality_cap = "\\label{fig:pdmapcases}The figure shows the two possible outcomes of a negative credit risk shock on a portfolio.
+The left-hand side plot shows the case where capital savings are procyclical.
+The right-hand side plot shows the case where capital savings are countercyclical.
+"
+```
+
+```{r, cyclicality, fig.height = 3, fig.width = 6,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap}
+
+p1 <- ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K))+ 
+  geom_line(alpha = 1) + 
+  #geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
+  #geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
+  #geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5, colour = "#008EC6")+
+  #geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5, colour = "#008EC6")+
+  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5, colour = "#008EC6")+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5, colour = "#008EC6")+
+  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5, colour = "#008EC6")+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5, colour = "#008EC6")+
+  theme_bw()+
+  theme(axis.ticks = element_blank(),
+        axis.line = element_line(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        #axis.text.x = element_blank(),
+        panel.grid = element_blank())+
+  expand_limits(x = 0, y = 0)+
+ scale_x_continuous(breaks=c(PD.h2),labels=c("PD"), expand = c(0, 0))+
+  scale_y_continuous(breaks=c(K.h2),labels=c("RW"),expand = c(0, 0))
+
+p2 <- ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K))+ 
+  geom_line(alpha = 1) + 
+  #geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
+  #geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
+  #geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5, colour = "#CC9933")+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5, colour = "#CC9933")+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5, colour = "#CC9933")+
+  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=5, colour = "#CC9933")+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5, colour = "#CC9933")+
+ theme_bw()+
+  theme(axis.ticks = element_blank(),
+        axis.line = element_line(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        #axis.text.x = element_blank(),
+        panel.grid = element_blank())+
+  expand_limits(x = 0, y = 0)+
+ scale_x_continuous(breaks=c(PD.h2),labels=c("PD"), expand = c(0, 0))+
+  scale_y_continuous(breaks=c(K.h2),labels=c("RW"),expand = c(0, 0))
+
+grid.arrange(p1, p2, ncol=2)
+```
+
+Consider now a credit risk shock.
+In state A, both safer and riskier assets are affected, having their PDs raised to 0.2% and 0.3%, respectively.
+In state B, only the riskier asset is affected, having its PD equal to 0.4%.
+Note that the average PD in both states are equal at 0.25%.
+However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B.
+The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%).
+Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states.
+This net procyclical effect is in line with the results of @behn2016procyclical.
+However, due to the increased portfolio dispersion, capital savings are countercyclical in state B.
+Conversely, because the PD-RW formula elasticity is lower for higher PD estimates, in stata A *Capital savings* are procyclical.
+Therefore, capital savings are procyclical in state A.
+Table \ref{tab1} summarizes the example and Figure Figure \ref{fig:pdmapcases}
+
+Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the PD distribution.
 
@@
 
-To calculate equation the proposed measures of capital savings, one needs information on the distribution of banks’ regulatory risk parameters.
+To calculate the proposed measures of capital savings, one needs information on the distribution of banks’ regulatory risk parameters.
 However, data at such granularity is not available in the common balance sheet datasets.
@@
 This level of portfolio breakdown is important because regulatory formulas vary across categories, and consequently, more precise knowledge of portfolio composition reduces measurements errors.
-
-Figure \ref{fig:savingsbank} shows that the *Capital savings ratio* calculated with the compiled dataset.
-Each point is a bank-year observations and the bars indicates the average value across time for each bank.
-First thing to note is that there is substancial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
-Second, variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
-Lastly, *Capital savings ratio* also varies between and within countries.
-This is better shown by figure \ref{fig:savingscountry} which aggregate the dataset by country.
-In the figure, each point is a bank observation and the bars are the average of these points.
-Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).
 
@@
 
+Figure \ref{fig:savingsbank} shows that the *Capital savings ratio* calculated with the compiled dataset.
+Each point is a bank-year observations and the bars indicates the average value across time for each bank.
+There is substancial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
+Variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
+*Capital savings ratio* also varies between and within countries.
+Figure \ref{fig:savingscountry}  aggregates the dataset by country.
+In the figure, each point is a bank observation and the bars are the average of these points.
+Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).
+
 ```{r}
@@
 
-```{r}
-tier1_cap = "The figure shows scatter plots and the linear relationship between the capital savings measure and capital ratios.
-The y-axis variable on plots a and b is the tier 1 capital ratio, and on plots c and d is the total capital ratio.
-Plots a and c plots all data points in the sample.
-Plots b and d plots the average across time for bank using the IRB approach."
-```
-
-```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth", include=FALSE}
- plot1 <- bank.data %>%  ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% 
-  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, tier, tier.hat, IRB) %>%
-  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>% 
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
-  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
-theme(legend.position='none')
-
-plot2 <- bank.data %>% ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, tier, tier.hat, IRB) %>%
-  filter(!is.na(RW.s), IRB == 1) %>%
-  group_by(name)%>%
- summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>%
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="",y="",subtitle = "Plot b") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
-  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
-  theme(legend.position = c(0.8, 0.85), 
-        legend.title = element_blank(), 
-        legend.background=element_blank())
-
-plot3 <- bank.data %>% ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, CAR, CAR.hat, IRB) %>%
-  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  theme(legend.position='none')
-
-plot4 <- bank.data %>% ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% 
-  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, CAR, CAR.hat, IRB) %>%
-  filter(!is.na(RW.s), IRB == 1) %>%
-  group_by(name)%>%
- summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="Capital savings",y="",  subtitle = "Plot d") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  theme(legend.position='none')
-
-grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
-```
-
-\newpage
-
-## Decomposition
-
-In this section I extend the method proposed by @cannata2012inside to decompose RW in its components.
-To get capital savins in terms of RWs, we only need to divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
+\pagebreak
+
+## RW Decomposition
+
+In this section I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
+I divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
 Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
-We can further breakdown each RW into their portfolio components.
-I choose to do so only for the counterfactual RW ($\text{RW}^c$), such that I can account for the contribution of the IRB portfolio mix while letting the components associated with capital savings as minimum as possible.
+I further breakdown RW_{IRB} into its portfolio components, $\text{RW}_{IRB}=\sum_j q_{j}\text{RW}_{j}$, in which the $j$ subscript stands for the different portfolios under the IRB approach: Wholesale, retail and equity.
+Equation \ref{eq5} shows the final decomposition:
 
@@
 \label{eq5}
-\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j=1}^{P} \text{RW}_{j}^{c} \Delta q^{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
-&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin}
+\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j} \text{RW}_{j}^{c} \Delta q_{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
+&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin},
 \end{aligned}
 \end{equation}
+in which  $\Delta$ stands for changes relative to a benchmark.
+For the cross-section decomposition, the benchmark is the average bank.
+For the time decomposition, the benchmark is either the year when the IRB approach was first introduced at the bank level or 2008.
+The superscript $B$ stands for benchmark.
+For instance, $q_SA^B$ is the share of the portfolio under the standardized approach for the benchmark.
 
 ```{r}
-cross.dec_cap = "\\label{fig:crossdecomposition} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
-The figure and calculations include all banks in the sample.
-The black line is the total RW relative to the average total RW.
-Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
-Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
-```
-
-```{r old_cross_decomposition, eval = FALSE, fig.height = 6.5, fig.width = 6, results="hide",fig.cap= cross.dec_cap, include=TRUE}
- cross.section.decomposition %>% filter(q.SA<1)%>%
-  select(name, Country, total= total.gain, SA.gain,IRB.gain,mix.gain, rollout.gain, int.gain, ext.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
-  ggplot(mapping = aes(x = reorder(name, total),
-                       y = value, fill = condition)) +
-geom_bar(stat = 'identity')+
-  geom_line(aes(y = total,group=1))  +
-  labs(x=NULL, y= NULL) +
-    coord_flip() +theme_bw()+
-  scale_fill_grey(na.value = "lightgrey",start = 0.1, end = 0.5,
-                  labels = c("Intensive margin", "Extensive margin", "Other components"), name = "Component")
-
-```
-
-```{r}
-cross.dec_cap2 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+cross.dec_cap1 = "\\label{fig:crossdecomposition1} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
 The figure and calculations include all banks in the sample.
@@
 
-```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
+```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap1, include=TRUE}
  plot1 <- cross.section.decomposition %>% 
@@
 ```
+
+I decompose the variation of RW relative to the benchmark in six components, two of them relating to *Capital savings*.
+The first, $(1-q^{B}_{SA})\Delta \text{RW}^{s}$, measures the intensity of *Capital savings*.
+That is, I fix the share of the portfolio under the IRB approach ($q_{SA}^B$) and variation in the component comes only from changes of $RW^s$ relative to the benchmark.
+The second component, $\text{RW}^{s}\Delta q^{B}_{SA}$, measures the extensive margin of *Capital savings*, i.e., variation in this component comes only from changes of the share under the IRB approach relative to the benchmark.
+Note that the magnitude of the second component is moderated by how intense *Capital savings* are in absolute terms.
+
+The remaining components are extensively discussed by @cannata2012inside.
+The first on the right-hand side of equation \ref{eq5} is the intensity of RW under the standardized approach ($q^{B}_{SA} \Delta \text{RW}_{SA}$).
+The contribution of this component to $\Delta \text{RW}$ comes from variaton of RW in portfolios under the standardized approach while fixing the share of this portfolio for all banks.
+The second term measures the intensity of RW under the IRB approach excluding the *Capital savings*, or differently, the intensity of the average PD ($(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j$).
+The contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of RW in the three portfolios under the IRB approach relative to the respective benchmark RW while holding constant the share of each IRB portfolio and the share of total IRB.
+The third component measure the contribution of the IRB portfolio mix.
+Because some types of portfolios have on average lower RW, the composition within IRB across types of portfolios matter to how large overall RW is.
+For instance retail portfolios tend to have lower RW relative to wholesale portfolios because the former usually are colletarized loans, such as mortgages, while the latter is mostly composed of corporate loans.
+Hence, the contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of the three portfolios shares within the IRB portfolio relative to the respective benchmarks while holding constant the respective intensity of average PD ($RW_j^c$) and the share of total IRB.
+Lastly, $(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}$ is the roll-out effect, i.e., the contribution to $\Delta \text{RW}$ coming from moving the portfolio from the standardized approach to the IRB approach.
+
+Figure \ref{fig:crossdecomposition1} plots the intensive and extensive margin of each bank in my sample using the average bank as benchmark.
+The scatter plot shows that the correlation between the two *Capital savings* components is not statistically significant.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
+If Goldman Sachs is included the correlation becomes statistically significant.} The figure suggests that banks benefit from the two components separately, some benefting more of the extensive margin, such as Northern Trust, and others more of the intensive margin, such as BayernLB.
+
 ```{r}
-cross.dec_cap3 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+cross.dec_cap2 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
 The figure and calculations include all banks in the sample.
@@
 
-```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap3, include=TRUE}
+```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
 plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
@@
 
-\newpage
+Figure \ref{fig:crossdecomposition2} shows that the net effect of the intensive and extensive margin is negatively correlated to the net effect of the other components.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
+If Goldman Sachs is included the correlation becomes even more negative.} This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
+
 ```{r}
-r_sqrt_decomp  = "\\label{fig:rsqrtdecomp} The figure shows the $R^2$ share for each of the RW components derived using equation Equation \\ref{eq5}." 
-```
-```{r,fig.height = 5, fig.width = 6, results="hide",fig.cap= r_sqrt_decomp , include=TRUE}
+rsqrt_cap = "\\label{fig:rsqrtdecomp} The figure shows the $R^2$ share for each of the RW components derived using equation Equation \\ref{eq5}." 
+```
+
+```{r,rsqrt, fig.height = 4, fig.width = 5,out.width = "75%", results="hide", fig.cap= rsqrt_cap, include=TRUE,fig.align = 'center'}
 
@@
   lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
-  relaimpo::calc.relimp(type = c("lmg", "last"), rela = FALSE )
-
- unrestrict.data <- data.frame(share.lmg  = unrestrict@lmg,
-                               share.last = unrestrict@last,
-                               decomposition = "unrestrict") %>%
-   add_rownames(var = "component")
-
- restrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
-  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
-         total.gain = total.gain+epsilon) %>%
-  mutate(IRB.gain     = IRB.gain.alt,
-         mix.gain     = mix.gain.alt,
-         rollout.gain = rollout.gain.alt)%>% 
- lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + SA.gain)%>%
- relaimpo::calc.relimp(type = c("lmg", "last"), rela = FALSE )
-
-data.frame(share.lmg  = restrict@lmg,
-           share.last = restrict@last,
-           decomposition = "restrict") %>%
+  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )
+
+  data.frame(share  = unrestrict@lmg) %>%
    add_rownames(var = "component")%>%
-    full_join(unrestrict.data)%>%
-  pivot_longer(cols = -c("decomposition","component"),
-               names_to = c("method"),
-               names_pattern = "share.(.*)",
-               values_to = "share")%>%
-  mutate(decomposition = factor(decomposition, 
-                                levels = c("unrestrict", "restrict"),
-                                labels = c("Unrestrict", "Restrict")),
-         method = factor(method, 
-                                levels = c("lmg", "last"),
-                                labels = c("LMG", "Last")),
-         component = factor(component,
+  mutate(component = factor(component,
                       levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
-                      labels = c("PD avg.", "PD ext.", "PD int.", "SA", "Mix", "Rollout"))) %>%
-   arrange(decomposition,method)%>%
-ggplot(mapping = aes(x = component,
-                       y = share)) +
+                      labels = c("PD avg.", "Ext. comp.", "Int. comp.", "SA", "Mix", "Rollout"))) %>%
+ggplot(mapping = aes(x = reorder(component, share),
+                       y = share, fill = component)) +
 geom_bar(stat = 'identity')+
@@
     theme_bw()+
-  theme(panel.grid.minor = element_blank())+
- facet_wrap(method~decomposition)+
-  theme(strip.background = element_blank(),
-       # strip.text = element_blank(),
-        axis.text.x = element_text(angle = 45, hjust = 1))+
-  scale_fill_grey()+
+  theme(panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+       # strip.text = element_blank()
+       legend.position = "none",
+        axis.text.x = element_text(hjust = 1))+
+  scale_fill_manual(values = c("#C3BCB2","#353535",  "#353535", "#C3BCB2","#C3BCB2", "#C3BCB2"))+
 scale_y_continuous(labels = scales::percent_format(accuracy = 1,
-                                 decimal.mark = '.'))
-
-```
+                                 decimal.mark = '.')) +
+    coord_flip()
+
+```
+
+As Figure \ref{fig:rsqrtdecomp} shows *Capital savings* components account for 27% of the RW variation across banks.
+Each component accounts for around 13.5% of the RW variation and are second and third in rank, losing only to the average PD contribution.
+I derive the $\text{R}^2$ contribution of each RW component using the Shapley-Owen decomposition method.
+This method consider the contribution of each variable to total $\text{R}^2$ as the sum of $\text{R}^2$ from regressions using all possible variable combinations that include the variable of interest and weighting each regression by the number of variable permutations.
+In sum, figure \ref{fig:rsqrtdecomp} offers strong evindence of the relative importance of *Capital savings* as source of variation to RW across banks.
+
+Figure \ref{fig:timedecomposition} shows the evolution of RW components relative to the time of IRB introduction (left-hand side plot), and relative to 2008 (right-hand side plot).
+The negative values for the extensive margin implies that banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach.
+However, during the same period, the intensity of *Capital savings* has, on average, decreased in time.
+That is, for the same average PD, capital requirements have increased from 2008 to 2018 and since the year IRB approach was introduced.
+
+The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}.
+Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.\footnote{Although supervisors favored raising additional capital as means to achieve the required capital ratios \citep{EBA2011recommendation}, banks respond to capital requirements mostly by reducing RWA \citep{gropp2018banks}}  The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
+This distribution compression leads to a reduction in *Capital savings*.
+This explanation is in line with the hypothesis that *Capital savings* are countercyclical.
+Efforts to increase capital ratio by risk reduction are increasingly more difficult.
 
@@
   theme_bw()+
-  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
+  labs( y = 'Change from t=0') +
   scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
@@
   theme_bw()+
-  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
+  labs(y = 'Change from 2008', x=NULL) +
   scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
@@
 
-```{r, fig.height = 5, fig.width = 5, results="hide",fig.cap= time.var_cap}
+```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= time.var_cap}
 
@@
   ggplot(mapping = aes(x = year,y=capital.ratio)) +
+  geom_point(alpha = 0.3) +
   geom_smooth(stat = 'summary',
             fun.data = mean_cl_quantile,
-            color = "black")+
+            color = "black") +
   theme_bw()+
@@
   ggplot(mapping = aes(x = year,y=mean.PD)) +
+  geom_point(alpha = 0.3) +
   geom_smooth(stat = 'summary',
@@
   ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
+  geom_point(alpha = 0.3) +
   geom_smooth(stat = 'summary',
@@
 
-plot_grid(plot1, plot2, plot3, hjust = 0.5,
-            ncol = 1, nrow = 3, align = "v", axis = "lr")
-```
-
-\newpage
-
-## Control Variables and Summary Statistics
-
-I include a number of bank-specific control variables in the analysis.
-Several works relate bank-specific characteristics to either total RWA or average RW (Vallascas and Hagendorff (2013) ...).
-In this section I use these evidence to construct my priors on how each of these variables might affect capital savings.
-
-The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
-First, I control for bank size using the log of total assets (in millions of US dollars).
-The expectation regarding the impact of bank size is ambiguous.
-On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices in order to reduce capital savings variability.
-On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
-I also include the regulatory capital ratio as control.
-Better capitalized banks could be subject to less regulatory scrutiny (Calem and Rob 1999) and, consequently, in a better position to engage in regulatory arbitrage.
-However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
-Thus, the expected effect of capital ratio on capital savings is also ambiguous.I also control for .
-Profitable banks face fewer incentives to understate the value of risky assets and to engage in regulatory arbitrage more generally.
-
-The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
-For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
-Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
-As a result, banks with higher shares of these variables should have higher values of capital savings.
-Finally, I control for bank funding using the ratio of total deposits to total assets and bank profitability using net income over total assets.
-Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
-Banks profitability might be the result of higher risk-taking.
-The effect of deposits funding and profitability on capital savings depends on how risk-taking is manifested by these variables.
-If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then capital savings decrease.
-Conversely, capital savings increases if banks take risk by having fatter upper-tails in the PD distribution.
-
-Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the control variables used throughout the analysis.
+plot4 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year, pillarI_stringency) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=pillarI_stringency)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'GDP growth (%)', x=NULL)
+
+plot_grid(plot1, plot2, plot3, plot4, hjust = 0.5,
+            ncol = 2, nrow = 2, align = "v", axis = "lr")
+```
+```{r test, fig.height = 5, fig.width = 7, results="hide",}
+bank.data %>% filter(year>2007, bvdid %in% pillar3.data$bvdid)%>%
+  select(bvdid, year, log.gdp_gr, Country, pillarI_stringency,cap_reg) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year)) +
+  geom_line(mapping = aes(y=pillarI_stringency))+
+  geom_line(mapping = aes(y=cap_reg))+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'GDP growth (%)', x=NULL) +
+  facet_wrap(~Country)
+```
+
+\pagebreak
+
+## Other Determinants of Capital Savings and Summary Statistics
+
+Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
+Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks ].
+In this section, I use these prior evidence to develop my expectations on how each of these variables may affect capital savings.
+
+Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
 
@@
   log.RWA.hat_gr, 
+  log.RWA.s_gr,
   log.RWA.r_gr,
-  log.RWA.s_gr, 
-  log.RWA.s.Retail_gr,
-  log.RWA.s.Wholesale_gr, 
-  log.RWA.s.Corporate_gr, 
-  log.RWA.s.Sovereign_gr, 
-  log.RWA.s.Banks_gr, 
+  log.RWA.r.Retail_gr,
+  log.RWA.r.Wholesale_gr, 
+  log.RWA.r.Corporate_gr, 
+  log.RWA.r.Sovereign_gr, 
+  log.RWA.r.Banks_gr, 
   mean.PD_gr, 
@@
  #income.ratio_gr,
-  mean.PD, 
+  log.gdp,
   log.asset,
   capital.ratio,
+  income.ratio,
   loans.ratio, 
   NII.ratio, 
-  deposit.ratio, 
-  income.ratio) %>%
+  NPL.ratio,
+  deposit.ratio) %>%
   as.data.frame() %>% 
@@
               "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
-              "\\hspace{1.0cm}$\\Delta \\text{Log 1+RWA}^r$",
               "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Retail",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Wholesale",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Corporate",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Sovereign", 
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Banks", 
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^r$",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^r$ Retail",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^r$ Wholesale",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Corporate",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Sovereign", 
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Banks", 
               "\\hspace{1.0cm}$\\Delta$ Log PD Mean",
@@
               #"$\\Delta$ Net Income / Total Assets", 
-              "PD Mean",
               "Log Total Assets",
               "Capital Ratio",
+              "Net Income / Total Assets",
               "Loans / Total Assets",
               "Net Interest Income / Oper. Rev.",
-              "Deposits / Total Assets", 
-              "Net Income / Total Assets"),
+              "Non-performing Loans / Loans",
+              "Deposits / Total Assets"),
           title = "Summary statistics",
           table.layout = "n=d#-t-as-",
-          notes = "", 
+          notes = "\\parbox[t]{14cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\\%}", 
           notes.label = "",
@@
 
-```{r, eval=FALSE}
+The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
+First, I control for bank size using the log of total assets (in millions of US dollars).
+The expectation regarding the impact of bank size is ambiguous.
+On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
+On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
+I also include the regulatory capital ratio as control.
+Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
+However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
+Thus, the expected effect of capital ratio on capital savings is also ambiguous.
+Next, I add bank profitability using net income over total assets.
+Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
+
+The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
+For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
+Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
+As a result, I expect banks with higher shares of these variables to have more *Capital savings*.
+
+Finally, I study the relationship of *Capital savings* to risk-taking.
+The RW decomposition analysis suggests that riskier banks benefit more from *Capital savings*.
+I add the ratio of non-performing loans to total loans to test this proposition.
+I also include bank funding using the ratio of total deposits to total assets.
+Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
+The effect of deposits funding on *Capital savings* depends on how risk-taking is manifested by the deposit funding.
+If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then deposits to total assets are negatively correlated to *Capital savings*.
+Conversely, *Capital savings* increase if banks take risk by having fatter upper-tails in the PD distribution.
+
+# Capital Savings and Business Cycle
+
+In this section I investigate the relationship between *Capital savings* and the business cycle.
+Overall, capital requirements are procyclical: during economic downturns credit risk in the economy increases which raises capital requirements, under the IRB approach forcing banks to cut down lending.
+However, as credit conditions changes (and banks react to it) the amount of *Capital savings* also changes.
+The expected direction of *Capital savings* cyclicality is ambiguous.
+On the one hand, if the entire portfolio is equally affected, capital savings are procyclical.
+On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
+The ambiguity arises because *Capital savings* can only increase, when credit conditions deteriorates, if the credit distribution widens.
+In the first situation, there is only a shift of the location parameters of the distribution and, consequently, a reduction of *Capital savings*.
+
+```{r reg_portfolio, echo=FALSE}
+
+dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+aux      <- paste0("mean.PD", dep.vars)
+ind.vars <- paste0("log.gdp_gr")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+   felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_portfolio, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Cyclicality of Capital Savings",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{r} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "tab1",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+Table \ref{tab1} shows that total *Capital savings* are countercyclical.
+According to the regression (1) a 1.7% GDP growth rate (one standard deviation) decreases on average *Capital savings* growth rate by 0.83 percentage points.
+The effect is economically significant as it represents a 17% of *Capital savings* standard deviation in the sample.
+The evidence in Table \ref{tab1} suggests that the effect is driven by wholesale portfolios, especially corporate portfolios.
+
+## Alternative measures
+
+The results in Table \ref{tab1} are robust to alternative measures of *Capital savings*.
+First, I replace the ratio measure of *Capital savings* by the measure in monetary units (equation \ref{eq3}).
+The results shown in Table \ref{tab2} reinforce the evidence of coutercyclicality of *Capital savings* is likely driven by the corporate portfolio.
+
+```{r reg_unit, echo=FALSE}
+
+dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+aux      <- paste0("mean.PD", dep.vars)
+ind.vars <- paste0("log.gdp_gr")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+ felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_unit, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Alternative Measures: Capital Savings in Monetary Units",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^s = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "tab2",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+Second, I use Gini coefficient to measure the dispersion in the PD distributions. Again, the results shown in Table \ref{tab3} suppport the evidence from the last two tables.
+
+```{r reg_gini, echo=FALSE}
+
+dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+aux      <- paste0("mean.PD", dep.vars)
+ind.vars <- paste0("log.gdp_gr")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Alternative Measures: Gini Coefficient",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Gini}_{i,j,t} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "tab3",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+## Effect on Capital Ratio procyclicality
+
+To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
+The capital ratio decomposition is shown in equation \ref{eq5}.
+The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
+To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
+Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab4} shows that *Capital savings* are the only countercyclical component of capital ratios.
+The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
+
+\begin{equation}
+\begin{aligned}
+\label{eq5}
+\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
+&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^r_{i,t} 
+\end{aligned}
+\end{equation}
+
+```{r reg_contribution, eval=TRUE, echo=FALSE}
+# Variables in the model 
+dep.vars <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat_gr", "log.RWA.r_gr")
+ind.vars <- paste0("log.gdp_gr")
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = subset(bank.data, subset = !is.na(log.CAR_gr)))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+# Effect 
+effect.list  <-  lapply(reg.list, function(x){
+  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
+})
+effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
+effect.list <-  sprintf("%.1f \\%%",effect.list)
+
+```
+
+```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Cyclicality of Capital Ratio Components",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,j,t} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = 
+            c("$\\Delta\\text{Log GDP}_{i,t}$", 
+              #"$\\Delta\\text{Mean PD}_{i,t}$",
+              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
+              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
+              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
+              "$\\text{PD Mean}_{i,j,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
+              "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+          #keep = c("log.gdp_gr"),
+          dep.var.labels = c("$\\Delta\\text{Log capital Ratio}_{i,t}$","$\\Delta\\text{Capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}^{c}_{i,t}$", "$\\Delta\\text{Log RWA}^{r}_{i,t}$"),
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          float = TRUE,
+          label = "tab4",
+          style = "qje",
+          column.sep.width = "2pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(c("Effect", effect.list),
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+
+```
+
+## Moments
+
+I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases.
+This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
+
+```{r reg_moments, echo=FALSE}
+
+# Variables in the model 
+dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr")
+ind.vars <- paste0("log.gdp_gr")
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = bank.data.win)
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
+          title = "Cyclicality of PD distribution Moments",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{11cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Moment}_{i,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          se = se.list,
+          covariate.labels = c("$\\Delta\\text{Log GDP}_{i,t}$", "$\\overline{\\text{PD}}_{i,t-1}$","$\\text{Log Total Assets}_{i,t-1}$", "$\\text{Deposits / Total Assets}_{i,t-1}$", "$\\text{Loans / Total Assets}_{i,t-1}$", "$\\text{Net Income / Total Assets}_{i,t-1}$", "$\\text{Net Interest Income / Oper.
+Rev.}_{i,t-1}$"),
+          keep = c("log.gdp_gr"),
+          dep.var.labels = c("Mean", "Variance", "Skewness", "Kurtosis"), 
+          notes.label = "",
+          notes.align = "c", 
+          notes.append = FALSE,
+          label = "",
+          style = "qje",
+          column.sep.width = "3pt",
+          #float.env = "sidewaystable",
+          align = TRUE,
+          add.lines = list(
+                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
+                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
+```
+
+```{r reg_extension1, echo=FALSE, include=FALSE}
+
+# Variables in the model 
+# Dependent variables
+dep.vars <- c("log.RWA.r","log.RWA.r.Retail", "log.RWA.r.Wholesale", "log.RWA.r.Corporate", "log.RWA.r.Sovereign", "log.RWA.r.Banks")
+dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
+# Independent variables
+aux.1 <- c("log.gdp",
+           #"log.gdp_gr",
+           "NPL.ratio",
+           "capital.ratio",
+           "log.asset",
+           "pillarI_stringency",
+           #"cap_reg",
+           "Sup_Power",
+           "Zscore",
+           "ROE",
+           "ROA")
+ind.vars  <- paste(aux.1, collapse=" + ")
+ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
+              #"$\\Delta\\text{ Log GDP}_{i,t-1}$",
+              "$\\text{NPL Ratio}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$", 
+              "$\\text{Pillar I stringency}_{i,t-1}$", 
+              #"$\\text{Capital stringency}_{i,t-1}$", 
+              "$\\text{Official Supervisory power}_{i,t-1}$",
+              "$\\text{Z-score}_{i,t-1}$",
+              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              #"$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
+             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{ROE}_{i,t-1}$",
+              "$\\text{ROA}_{i,t-1}$")
+
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_extension1, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
+
+stargazer(reg.list,
+          title = "capital Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          se               = se.list,
+          add.lines        = fe.list,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab6",
+          style            = "qje",
+          column.sep.width = "-15pt",
+          #float.env        = "sidewaystable",
+          align            = TRUE)
+```
+```{r reg_extension2, echo=FALSE, include=FALSE}
+
+# Variables in the model 
+# Dependent variables
+dep.vars <- c("log.RWA.s")
+dep.text <- c("$\\text{Log RWA}^{s}_{i,t}$")
+# Independent variables
+aux.1 <- c("log.gdp",
+           "NPL.ratio",
+           "capital.ratio",
+           "log.asset",
+           "pillarI_stringency",
+           #"cap_reg",
+           "Sup_Power",
+           "Zscore",
+          "NII.ratio",
+          "ROE",
+          "ROA")
+aux.3 <- paste(aux.1, collapse=" + ")
+ind.vars <- c(aux.1, aux.3)
+ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
+              "$\\text{NPL Ratio}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$", 
+              "$\\text{Pillar I stringency}_{i,t-1}$", 
+              #"$\\text{Capital stringency}_{i,t-1}$", 
+              "$\\text{Official Supervisory power}_{i,t-1}$",
+              "$\\text{Z-score}_{i,t-1}$",
+              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
+             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{ROE}_{i,t-1}$",
+              "$\\text{ROA}_{i,t-1}$")
+# Fixed effects
+fe       <- c(rep("year + bvdid"  , length(aux.1)),
+              rep("year + bvdid"  , length(aux.3)))
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
+                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_extension2, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
+
+stargazer(reg.list,
+          title = "capital Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          se               = se.list,
+          add.lines        = fe.list,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "tab6",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "",
+          style            = "qje",
+          column.sep.width = "-15pt",
+          float.env        = "sidewaystable",
+          align            = TRUE)
+```
+
+```{r reg_extension3, echo=FALSE, include=FALSE}
+
+# Variables in the model 
+# Dependent variables
+dep.vars <- c("gini.PD")
+dep.text <- c("$\\text{Log RWA}^{r}_{i,t}$")
+# Independent variables
+aux.1 <- c("log.gdp",
+           "NPL.ratio",
+           "capital.ratio",
+           "log.asset",
+           "pillarI_stringency",
+           #"cap_reg",
+           "Sup_Power",
+           "Zscore",
+          "NII.ratio",
+          "ROE",
+          "ROA")
+aux.3 <- paste(aux.1, collapse=" + ")
+ind.vars <- c(aux.1, aux.3)
+ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
+              "$\\text{NPL Ratio}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$", 
+              "$\\text{Pillar I stringency}_{i,t-1}$", 
+              #"$\\text{Capital stringency}_{i,t-1}$", 
+              "$\\text{Official Supervisory power}_{i,t-1}$",
+              "$\\text{Z-score}_{i,t-1}$",
+              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
+             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{ROE}_{i,t-1}$",
+              "$\\text{ROA}_{i,t-1}$")
+# Fixed effects
+fe       <- c(rep("year + bvdid"  , length(aux.1)),
+              rep("year + bvdid"  , length(aux.3)))
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
+                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_extension3, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
+
+stargazer(reg.list,
+          title = "capital Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          se               = se.list,
+          add.lines        = fe.list,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "tab6",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "",
+          style            = "qje",
+          column.sep.width = "-15pt",
+          float.env        = "sidewaystable",
+          align            = TRUE)
+```
+
+```{r reg_extension4, echo=FALSE, include=FALSE}
+
+# Variables in the model 
+# Dependent variables
+dep.vars <- c("log.RWA.r.Wholesale")
+dep.text <- c("$\\text{Log RWA}^{r}_{i,t}$")
+# Independent variables
+aux.1 <- c("log.gdp",
+           "NPL.ratio",
+           "capital.ratio",
+           "log.asset",
+           "pillarI_stringency",
+           #"cap_reg",
+           "Sup_Power",
+           "Zscore",
+          "NII.ratio",
+          "ROE",
+          "ROA")
+aux.3 <- paste(aux.1, collapse=" + ")
+ind.vars <- c(aux.1, aux.3)
+ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
+              "$\\text{NPL Ratio}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$", 
+              "$\\text{Pillar I stringency}_{i,t-1}$", 
+              #"$\\text{Capital stringency}_{i,t-1}$", 
+              "$\\text{Official Supervisory power}_{i,t-1}$",
+              "$\\text{Z-score}_{i,t-1}$",
+              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
+              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
+             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+              "$\\text{ROE}_{i,t-1}$",
+              "$\\text{ROA}_{i,t-1}$")
+# Fixed effects
+fe       <- c(rep("year + bvdid"  , length(aux.1)),
+              rep("year + bvdid"  , length(aux.3)))
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
+                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_extension4, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
+
+stargazer(reg.list,
+          title = "capital Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          se               = se.list,
+          add.lines        = fe.list,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "tab6",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "",
+          style            = "qje",
+          column.sep.width = "-15pt",
+          float.env        = "sidewaystable",
+          align            = TRUE)
+```
+```{r reg_extension, echo=FALSE, include=FALSE}
+
+# Variables in the model 
+# Dependent variables
+dep.vars <- c("log.RWA.s")
+dep.text <- c("$\\text{Log RWA}^{r}_{i,t}$")
+# Independent variables
+aux.1 <- c("lag(log.gdp , order_by = year)",
+           "lag(NPL.ratio, order_by = year)",
+           "lag(capital.ratio, order_by = year)",
+           "lag(log.asset, order_by = year)",
+           "lag(pillarI_stringency, order_by = year)",
+          # "lag(loans.ratio, order_by = year)",
+          # "lag(NII.ratio, order_by = year)",
+          # "lag(income.ratio, order_by = year)",
+          # "lag(ROA, order_by = year)",
+           "lag(subordinated, order_by = year)")
+aux.3 <- paste(aux.1, collapse=" + ")
+ind.vars <- c(aux.1, aux.3)
+ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
+              "$\\text{NPL Ratio}_{i,t-1}$",
+              "$\\text{Capital Ratio}_{i,t-1}$",
+              "$\\text{Log Total Assets}_{i,t-1}$", 
+              "$\\text{Pillar I strigency}_{i,t-1}$", 
+             # "$\\text{Loans / Total Assets}_{i,t-1}$", 
+             # "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
+             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
+             # "$\\text{ROA}_{i,t-1}$",
+              "$\\text{Subordinated Debt}_{i,t-1}$")
+# Fixed effects
+fe       <- c(rep("year + bvdid"  , length(aux.1)),
+              rep("year + bvdid"  , length(aux.3)))
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
+                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
+
+stargazer(reg.list,
+          title = "capital Savings and Bank Characteristics",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t-1}$ is a vector of bank control variables.
+Standard errors are adjusted for heteroskedasticity.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          se               = se.list,
+          add.lines        = fe.list,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "tab6",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "",
+          style            = "qje",
+          column.sep.width = "-15pt",
+          float.env        = "sidewaystable",
+          align            = TRUE)
+```
+
+# Conclusion and Discussion
+
+This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements.
+*Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are economically relevant.
+They explain 27% of the RW variation across banks.
+*Capital savings* are also countercyclical and reduce the procyclicality of RWA by 11.6%.
+Hence, the concavity of IRB formulas could be explored to reduce procyclicality of capital regulation.
+
+# References
+
+```{r did data, eval=FALSE, include=FALSE}
 p_w <- c(0.0, 1)         ## Set winsorizing limits
@@
 
-```{r , eval=FALSE, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth", include=FALSE}
+```{r did, eval=FALSE, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth", include=FALSE}
 p_w <- c(0.0, 1)         ## Set winsorizing limits
@@
 
-\newpage
-
-# Capital Savings and Business Cycle
-
-```{r reg_portfolio, echo=FALSE}
-
-dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
-aux      <- paste0("mean.PD", dep.vars)
-ind.vars <- paste0("log.gdp_gr")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data =  mutate(bank.data.win, mean.PD := get(paste(aux))))
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_portfolio, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Cyclicality of Capital Savings",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{s} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
-          #keep = c("log.gdp_gr"),
-          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          float = TRUE,
-          label = "",
-          style = "qje",
-          column.sep.width = "2pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-\pagebreak
-## Alternative measures
-```{r reg_unit, echo=FALSE}
-
-dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
-aux      <- paste0("mean.PD", dep.vars)
-ind.vars <- paste0("log.gdp_gr")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data =  mutate(bank.data.win, mean.PD := get(paste(aux))))
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_unit, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Alternative Measures: Capital Savings Ratio",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^r = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
-          #keep = c("log.gdp_gr"),
-          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          float = TRUE,
-          label = "",
-          style = "qje",
-          column.sep.width = "2pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-```{r reg_gini, echo=FALSE}
-
-dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
-aux      <- paste0("mean.PD", dep.vars)
-ind.vars <- paste0("log.gdp_gr")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data =  mutate(bank.data.win, mean.PD := get(paste(aux))))
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Alternative Measures: Gini Coefficient",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Gini}_{i,j,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
-          #keep = c("log.gdp_gr"),
-          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          float = TRUE,
-          label = "",
-          style = "qje",
-          column.sep.width = "2pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-\pagebreak
-
-## Effect on Capital Ratio procyclicality
-
-```{r reg_contribution, eval=TRUE, echo=FALSE}
-# Variables in the model 
-dep.vars <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat_gr", "log.RWA.r_gr")
-ind.vars <- paste0("log.gdp_gr")
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = subset(bank.data, subset = !is.na(log.CAR_gr)))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-# Effect 
-effect.list  <-  lapply(reg.list, function(x){
-  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
-})
-effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
-effect.list <-  sprintf("%.1f \\%%",effect.list)
-
-```
-
-```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Cyclicality of Capital Ratio Components",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{15cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,j,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
-          #keep = c("log.gdp_gr"),
-          dep.var.labels = c("$\\Delta\\text{Log capital Ratio}_{i,t}$","$\\Delta\\text{Capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}^{c}_{i,t}$", "$\\Delta\\text{Log RWA}^{r}_{i,t}$"),
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          float = TRUE,
-          label = "",
-          style = "qje",
-          column.sep.width = "2pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(c("Effect", effect.list),
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-\pagebreak
-
-## Moments
-
-```{r reg_moments, echo=FALSE}
-
-# Variables in the model 
-dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr")
-ind.vars <- paste0("log.gdp_gr")
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Cyclicality of PD distribution Moments",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{11cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Moment}_{i,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = c("$\\Delta\\text{Log GDP}_{i,t}$", "$\\overline{\\text{PD}}_{i,t-1}$","$\\text{Log Total Assets}_{i,t-1}$", "$\\text{Deposits / Total Assets}_{i,t-1}$", "$\\text{Loans / Total Assets}_{i,t-1}$", "$\\text{Net Income / Total Assets}_{i,t-1}$", "$\\text{Net Interest Income / Oper.
-Rev.}_{i,t-1}$"),
-          keep = c("log.gdp_gr"),
-          dep.var.labels = c("Mean", "Variance", "Skewness", "Kurtosis"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          label = "",
-          style = "qje",
-          column.sep.width = "3pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-```
-
-\pagebreak
-
-## Extensions
-
-```{r reg_extension, echo=FALSE}
-
-# Variables in the model 
-# Dependent variables
-dep.vars <- c("log.RWA.s")
-dep.text <- c("$\\text{Log RWA}^{s}_{i,t}$")
-# Independent variables
-aux.1 <- c("lag(log.asset , order_by = year)",
-           "lag(deposit.ratio, order_by = year)",
-           "lag(loans.ratio, order_by = year)",
-           "lag(NII.ratio, order_by = year)",
-           "lag(income.ratio, order_by = year)",
-           "lag(ROA, order_by = year)",
-           "lag(subordinated, order_by = year)",
-           "lag(NPL.ratio, order_by = year)",
-           "lag(capital.ratio, order_by = year)")
-aux.2 <- c("lag(cap_reg, order_by = year)",
-           "lag(sup_power, order_by = year)")
-aux.3 <- paste(aux.1, collapse=" + ")
-ind.vars <- c(aux.1, aux.2, aux.3)
-ind.text <- c("$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{ROA}_{i,t-1}$",
-              "$\\text{Subordinated Debt}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Capital Strigency}_{i,t-1}$",
-              "$\\text{Supervisory Power}_{i,t-1}$")
-# Fixed effects
-fe       <- c(rep("year + bvdid"  , length(aux.1)),
-              rep("year + Country", length(aux.2)),
-              rep("year + bvdid"  , length(aux.3)))
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
-                                rep("\\multicolumn{1}{c}{No}" , length(aux.2)),
-                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))),
-                c("Country FE", rep("\\multicolumn{1}{c}{-}"  , length(aux.1)),
-                                rep("\\multicolumn{1}{c}{Yes}", length(aux.2)),
-                                rep("\\multicolumn{1}{c}{-}"  , length(aux.3))))
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          title = "capital Savings and Bank Characteristics",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta\\text{Log RWA}^{s}_{i,t} = \\gamma\\times X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "",
-          style            = "qje",
-          column.sep.width = "-15pt",
-          float.env        = "sidewaystable",
-          align            = TRUE)
+```{r, eval = FALSE, fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth", include=FALSE}
+ plot1 <- bank.data %>%  ungroup() %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% 
+  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, tier, tier.hat, IRB) %>%
+  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>% 
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
+geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
+  geom_smooth(method = "lm",color="black") +
+  theme_bw()+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
+  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
+theme(legend.position='none')
+
+plot2 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, tier, tier.hat, IRB) %>%
+  filter(!is.na(RW.s), IRB == 1) %>%
+  group_by(name)%>%
+ summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
+  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
+geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="",y="",subtitle = "Plot b") +
+  geom_smooth(method = "lm",color="black") +
+  theme_bw()+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
+  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
+  theme(legend.position = c(0.8, 0.85), 
+        legend.title = element_blank(), 
+        legend.background=element_blank())
+
+plot3 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, CAR, CAR.hat, IRB) %>%
+  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
+geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
+  geom_smooth(method = "lm",color="black") +
+  theme_bw()+
+  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  theme(legend.position='none')
+
+plot4 <- bank.data %>% ungroup() %>%
+  filter(!is.na(RW.s),
+         !is.na(basel),
+         tier1capitalmillcu < totalcapitalmillcu)%>% 
+  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
+  select(name,RW.s, CAR, CAR.hat, IRB) %>%
+  filter(!is.na(RW.s), IRB == 1) %>%
+  group_by(name)%>%
+ summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
+  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
+ ggplot(mapping = aes(x = RW.s,
+                      y = value,
+                      color = condition,
+                      fill = condition)) +
+geom_point( alpha = 0.5,size = 1) +
+  scale_y_continuous( limits=c(5, 20))+
+  labs(x="Capital savings",y="",  subtitle = "Plot d") +
+  geom_smooth(method = "lm",color="black") +
+  theme_bw()+
+  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
+  theme(legend.position='none')
+
+grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
 ```
@@
 ```
-\pagebreak
-# Conclusion and Discussion
-
-Taken together, the results show that strategic portfolio choice, in terms of credit risk, is an important way banks can reduce capital charges while maintaining their average portfolio risk and return on equity.
-Differently from manipulation and subjectivity of RW under IRB, the PD distribution effect hereby identified does not entail any regulatory breach.
-However, the assumption that riskier assets have lower systematic risk, which is one of the reasons the regulatory formulas are concave, may be too simplistic to capture the true systematic component of each asset.
-If that is the case, this paper shows that the current capital requirements framework does not correctly equate capital charges to credit risk, consequently, undermining financial stability.

### 2020-04-02 Data/Import.R
Add country level credit variables (credit-to-GDP etc...)

@@ -8,11 +8,12 @@
 ## Code setup                                                                  ##
 ##=============================================================================##
 
-rm(list = ls()) 
+
 library(tidyverse)    ## Data manipulation, pipe operator                                                             
 library(readxl)       ## Command to open xlsx files                                                     
 library(data.table)   ## Command to bind lists
 library(stringr)      ## Command to clean variables names in Orbis
+rm(list = ls()) 
 
 ##============================================================================##
 ## Auxiliar data                                                              ##
@@ -41,7 +42,8 @@ country.code <- read.csv("Data/Raw/WDI/WDICountry.csv") %>%
          Country = x2.alpha.code)
 
 WDI <- read.csv("Data/Raw/WDI/WDIData.csv") %>%
-  filter(Indicator.Code %in% c("NY.GDP.DEFL.ZS","NY.GDP.PCAP.KD","NY.GDP.PCAP.KN", "PA.NUS.FCRF", "PX.REX.REER")) %>%
+  filter(Indicator.Code %in% c("NY.GDP.DEFL.ZS","NY.GDP.PCAP.KD","NY.GDP.PCAP.KN", "PA.NUS.FCRF", "PX.REX.REER",
+                               "FS.AST.DOMS.GD.ZS","FS.AST.PRVT.GD.ZS","FD.AST.PRVT.GD.ZS")) %>%
   rename_all(funs(str_replace_all(.,"([\\w])([0-9]+)$", "\\1\\.\\2"))) %>%
   reshape(varying   = c(grep("X.", names(.))),
                       direction = 'long', 
@@ -52,11 +54,14 @@ WDI <- read.csv("Data/Raw/WDI/WDIData.csv") %>%
           timevar   = 'Indicator.Code') %>%
   rename_all(tolower) %>%
   inner_join(country.code)%>%
-  select(deflator = x.ny.gdp.defl.zs,
-         gdppc.us = x.ny.gdp.pcap.kd,
-         gdppc    = x.ny.gdp.pcap.kn,
-         er       = x.pa.nus.fcrf,
-         reer     = x.px.rex.reer,
+  select(deflator        = x.ny.gdp.defl.zs,
+         gdppc.us        = x.ny.gdp.pcap.kd,
+         gdppc           = x.ny.gdp.pcap.kn,
+         er              = x.pa.nus.fcrf,
+         reer            = x.px.rex.reer,
[…]

### 2020-04-02 Data/Clean.R
Add RW decomposition at the bank-year level and keep only variables that are used in the paper.

@@ -276,9 +276,9 @@ bank.data <- bank.data %>% ungroup() %>% rowwise() %>%
           tier          = 100*(tier1capitalmillcu/RWA),
           CAR.hat       = 100*(totalcapitalmillcu/RWA.hat.total),
           tier.hat      = 100*(tier1capitalmillcu/RWA.hat.total),
-          CAR.dif.alt   = ifelse(EAD.IRB == 0 & IRB == 0, 0,totalcapitalratio - CAR.alt),
-          CAR.dif       = ifelse(EAD.IRB == 0 & IRB == 0, 0,CAR - CAR.hat),
-          tier.dif      = ifelse(EAD.IRB == 0 & IRB == 0, 0,tier - tier.hat),
+          CAR.dif.alt   = ifelse(EAD.IRB == 0 & IRB == 0, 0, totalcapitalratio - CAR.alt),
+          CAR.dif       = ifelse(EAD.IRB == 0 & IRB == 0, 0, CAR - CAR.hat),
+          tier.dif      = ifelse(EAD.IRB == 0 & IRB == 0, 0, tier - tier.hat),
           log.CAR       = log(CAR),
           log.CAR.til   = log(CAR.til),
           log.K         = log(totalcapitalmillcu),
@@ -314,20 +314,18 @@ bank.data <- bank.data %>% ungroup() %>% rowwise() %>%
           Zscore     = (r.mean.ROA + r.mean.CAR)/r.sd.ROA)%>%
    # Growth rates
    ungroup() %>% group_by(bvdid) %>%
-   mutate_at(vars(matches("_|log.|.ratio")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
+   mutate_at(vars(matches("log.|.ratio|RW.s|RW.SA|q.")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
                                                                     . - lag(., order_by = year), NA))) %>%
-   mutate_at(vars(matches(".PD")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
+   mutate_at(vars(matches("_gdp|.PD")), .funs = list(gr= ~ifelse(. > 0 & lag(., order_by = year)>0, 
                                                             log(.) - log(lag(., order_by = year)), NA))) %>%
   # Remove IRB not collected  
    mutate(IRB.indicator = mean(IRB),
           not.collect   = ifelse(IRB.indicator >0 & sample == "No", 1, 0)) %>%
[…]

### 2020-04-02 Analysis/Draft/Draft.Rmd
Last version of draft before switching to bookdown

@@
 header-includes:
-   - \usepackage{rotating, graphicx, dcolumn, longtable}
+   - \usepackage{rotating, graphicx, dcolumn, longtable, float}
 output:
   pdf_document:
-    citation_package: natbib
+    #citation_package: natbib
     latex_engine: pdflatex
@@
 bibliography: ../../Auxiliar/mybibfile.bib
+csl: ../../Auxiliar/cambridge-journal-of-economics.csl
 citecolor: black
@@
                       comment   = "",
-                      cache     = FALSE)
+                      cache     = FALSE,
+                      fig.pos   = 'H')
 # These options are tuned for manuscript/presentation. 
@@
 library(ggpubr)               ## More plots
-library(textablr)
-library(sandwich)
-library(lmtest)
+#library(textablr)
+#library(sandwich)
+#library(lmtest)
 #library(relaimpo)            ## Calculate relative importance measures
@@
 source("../../Auxiliar/Function.R") ## Capital requirements functions
+source("../../Auxiliar/longtable.R") ## longtable for stargazer
 
@@
 The difference between  $\text{RW}^c$ and $\text{RW}^a$ is the monetary amount of *Capital savings* per unit of EAD due to the concavity of the IRB formula.
+
+I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
+The method is a generalization of the example in Figure \ref{fig:pdmap}: for each bank-year combination, compare the actual capital requirement to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio’s average PD.
+
+I define the counterfactual RW as:
+
+\begin{equation}
+\label{eq2}
+\text{RW}^c= 12.5\frac{\sum_{i}\text{EAD}_i\text{LGD}_iN\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\overline{\text{PD}})+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)}{\sum_i \text{EAD}_i},
+\end{equation}
+in which the PD parameter for all assets is replaced by the portfolio weighted average:
+
+\begin{equation*}
+\overline{\text{PD}}=\frac{\sum_i\text{EAD}_i\text{LGD}_i\text{PD}_i}{\sum_i\text{EAD}_i\text{LGD}_i}.
+\end{equation*}
+
+The main measure of *Capital savings* used in this paper is the ratio of equation \ref{eq2} to equation \ref{eq1} as follows:
+
+\begin{equation}
+\label{measure1}
+\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}}.
+\end{equation}
+
+$\text{RW}^r$ is the rate at which a bank saves capital.
+For instance, a value of 0.5 means this portfolio would require 50% more capital in the abscense of *Capital savings* to keep the same capital ratio.
+A value of zero means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
+Any PD distribution with positive variance results in positive *Capital savings*.
+
+An alternative way to measure *Capital savings* is in monetary units as follows:
+
+\begin{equation}
+\label{measure2}
+\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}.
+\end{equation}
+
+An alternative way to measure *Capital savings* is in monetary units as follows:
+
+\begin{equation}
+\label{measure3}
+\text{RW}^s=\frac{\text{RWA}^{\$}}{\text{RWA}}.
+\end{equation}
+
+# Data
+
+To calculate the proposed measures of capital savings, one needs information on the distribution of banks’ regulatory risk parameters.
+However, data at such granularity is not available in the common balance sheet datasets.
+Although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
+Therefore, to overcome data limitations I collected information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports.
+The sample consists of 50 large banking groups, from 12 countries, that have adopted the IRB approach.
+For each bank, I collected consolidated information from the year the IRB approach was approved until 2018.
+The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate and sovereign for wholesale, and real estate and qualifying revolving credit for retail).
+This level of portfolio breakdown is important because regulatory formulas vary across categories, and consequently, more precise knowledge of portfolio composition reduces measurements errors.
+
+Figure \ref{fig:savingsbank} shows that the *Capital savings ratio* calculated with the compiled dataset.
+Each point is a bank-year observations and the bars indicates the average value across time for each bank.
+There is substancial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
+Variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
+*Capital savings ratio* also varies between and within countries.
+Figure \ref{fig:savingscountry}  aggregates the dataset by country.
+In the figure, each point is a bank observation and the bars are the average of these points.
+Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).
+
+## RW Decomposition
+
+In this section I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
+I divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
+Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
+I further breakdown RW_{IRB} into its portfolio components, $\text{RW}_{IRB}=\sum_j q_{j}\text{RW}_{j}$, in which the $j$ subscript stands for the different portfolios under the IRB approach: Wholesale, retail and equity.
+Equation \ref{eq5} shows the final decomposition:
+
+\begin{equation}
+\begin{aligned}
+\label{eq5}
+\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j} \text{RW}_{j}^{c} \Delta q_{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
+&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin},
+\end{aligned}
+\end{equation}
+in which  $\Delta$ stands for changes relative to a benchmark.
+For the cross-section decomposition, the benchmark is the average bank.
+For the time decomposition, the benchmark is either the year when the IRB approach was first introduced at the bank level or 2008.
+The superscript $B$ stands for benchmark.
+For instance, $q_SA^B$ is the share of the portfolio under the standardized approach for the benchmark.
+
+I decompose the variation of RW relative to the benchmark in six components, two of them relating to *Capital savings*.
+The first, $(1-q^{B}_{SA})\Delta \text{RW}^{s}$, measures the intensity of *Capital savings*.
+That is, I fix the share of the portfolio under the IRB approach ($q_{SA}^B$) and variation in the component comes only from changes of $RW^s$ relative to the benchmark.
+The second component, $\text{RW}^{s}\Delta q^{B}_{SA}$, measures the extensive margin of *Capital savings*, i.e., variation in this component comes only from changes of the share under the IRB approach relative to the benchmark.
+Note that the magnitude of the second component is moderated by how intense *Capital savings* are in absolute terms.
+
+The remaining components are extensively discussed by @cannata2012inside.
+The first on the right-hand side of equation \ref{eq5} is the intensity of RW under the standardized approach ($q^{B}_{SA} \Delta \text{RW}_{SA}$).
+The contribution of this component to $\Delta \text{RW}$ comes from variaton of RW in portfolios under the standardized approach while fixing the share of this portfolio for all banks.
+The second term measures the intensity of RW under the IRB approach excluding the *Capital savings*, or differently, the intensity of the average PD ($(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j$).
+The contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of RW in the three portfolios under the IRB approach relative to the respective benchmark RW while holding constant the share of each IRB portfolio and the share of total IRB.
+The third component measure the contribution of the IRB portfolio mix.
+Because some types of portfolios have on average lower RW, the composition within IRB across types of portfolios matter to how large overall RW is.
+For instance retail portfolios tend to have lower RW relative to wholesale portfolios because the former usually are colletarized loans, such as mortgages, while the latter is mostly composed of corporate loans.
+Hence, the contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of the three portfolios shares within the IRB portfolio relative to the respective benchmarks while holding constant the respective intensity of average PD ($RW_j^c$) and the share of total IRB.
+Lastly, $(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}$ is the roll-out effect, i.e., the contribution to $\Delta \text{RW}$ coming from moving the portfolio from the standardized approach to the IRB approach.
+
+Figure \ref{fig:crossdecomposition1} plots the intensive and extensive margin of each bank in my sample using the average bank as benchmark.
+The scatter plot shows that the correlation between the two *Capital savings* components is not statistically significant.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
+If Goldman Sachs is included the correlation becomes statistically significant.} The figure suggests that banks benefit from the two components separately, some benefting more of the extensive margin, such as Northern Trust, and others more of the intensive margin, such as BayernLB.
+
+Figure \ref{fig:crossdecomposition2} shows that the net effect of the intensive and extensive margin is negatively correlated to the net effect of the other components.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
+If Goldman Sachs is included the correlation becomes even more negative.} This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
+
+As Figure \ref{fig:rsqrtdecomp} shows *Capital savings* components account for 27% of the RW variation across banks.
+Each component accounts for around 13.5% of the RW variation and are second and third in rank, losing only to the average PD contribution.
+I derive the $\text{R}^2$ contribution of each RW component using the Shapley-Owen decomposition method.
+This method consider the contribution of each variable to total $\text{R}^2$ as the sum of $\text{R}^2$ from regressions using all possible variable combinations that include the variable of interest and weighting each regression by the number of variable permutations.
+In sum, figure \ref{fig:rsqrtdecomp} offers strong evindence of the relative importance of *Capital savings* as source of variation to RW across banks.
+
+Figure \ref{fig:timedecomposition} shows the evolution of RW components relative to the time of IRB introduction (left-hand side plot), and relative to 2008 (right-hand side plot).
+The negative values for the extensive margin implies that banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach.
+However, during the same period, the intensity of *Capital savings* has, on average, decreased in time.
+That is, for the same average PD, capital requirements have increased from 2008 to 2018 and since the year IRB approach was introduced.
+
+The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}.
+Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.
+[^1] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
+This distribution compression leads to a reduction in *Capital savings*.
+This explanation is in line with the hypothesis that *Capital savings* are countercyclical.
+Efforts to increase capital ratio by risk reduction are increasingly more difficult.
+
+[^1]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].
+
+## Other Determinants of Capital Savings and Summary Statistics
+
+Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
+Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks ].
+In this section, I use these prior evidence to develop my expectations on how each of these variables may affect capital savings.
+
+Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
+
+The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
+First, I control for bank size using the log of total assets (in millions of US dollars).
+The expectation regarding the impact of bank size is ambiguous.
+On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
+On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
+I also include the regulatory capital ratio as control.
+Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
+However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
+Thus, the expected effect of capital ratio on capital savings is also ambiguous.
+Next, I add bank profitability using net income over total assets.
+Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
+
+The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
+For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
+Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
+As a result, I expect banks with higher shares of these variables to have more *Capital savings*.
+
+Finally, I study the relationship of *Capital savings* to risk-taking.
+The RW decomposition analysis suggests that riskier banks benefit more from *Capital savings*.
+I add the ratio of non-performing loans to total loans to test this proposition.
+I also include bank funding using the ratio of total deposits to total assets.
+Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
+The effect of deposits funding on *Capital savings* depends on how risk-taking is manifested by the deposit funding.
+If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then deposits to total assets are negatively correlated to *Capital savings*.
+Conversely, *Capital savings* increase if banks take risk by having fatter upper-tails in the PD distribution.
+
+# Capital Savings and the Business Cycle
+
+In this section I investigate the relationship between *Capital savings* and the business cycle.
+I start by explaining the theoretical ambiguous relationship between *Capital savings* and the business cycle.
+Next I show that *Capital savings* are countercyclical, i.e., given an average portfolio PD, banks need more capital in good times than in bad times to reach the same capital ratio.
+Finally, I provide evidence that this relationship is not purely mechanical but depends on how banks realocate their assets through the business cycle.
+
+## A simple theoretical framework
+
+Overall, capital requirements are procyclical: during economic downturns credit risk in the economy increases which raises capital requirements, under the IRB approach forcing banks to cut down lending.
+However, as credit conditions changes (and banks react to it) the amount of *Capital savings* also changes.
+The expected direction of *Capital savings* cyclicality is ambiguous.
+On the one hand, if the entire portfolio is equally affected, capital savings are procyclical.
+On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
+The ambiguity arises because *Capital savings* can only increase, when credit conditions deteriorates, if the credit distribution widens.
+In the first situation, there is only a shift of the location parameters of the distribution and, consequently, a reduction of *Capital savings*.
+
+I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
+Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
+However, as credit conditions deteriorates (and banks react to these new credit conditions) the PD distribution and, consequently, the amount of *Capital savings* also change.
+A priori the direction of the economic cycle effect on *Capital savings* is ambiguous.
+At one extreme, if the entire portfolio is equally affected, then there is only a shift of the location parameter of the PD distribution.
+Because the PD-RW formula elasticity is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
+On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
+The ambiguity arises because *Capital savings* can only increase during a recession if the shape of the PD distribution is affect during the cycle.
+
+I use a simple example to illustrate this ambiguity.
+Consider a bank with the asset side of its balance sheet composed of \$100 of a safer asset and \$100 of a riskier asset and three states of the economy.
+Without loss of generality, assume $\text{LGD}=1$ and $\text{R}=0.15$.
+In the initial state, safer and riskier assets have a 0.1% and 0.2% probability of default, respectively.
+According to equation \ref{eq1}, the actual RWA of this bank is \$63.9.
+The weighted average of this portfolio in terms of $\text{PD}$ is 0.15% which yields a counterfactual RWA of \$64.7 according to equation \ref{eq2}.
+Hence, in the initial state *Capital savings* equal \$0.8.
+
+Consider now a credit risk shock.
+In state A, both safer and riskier assets are affected, having their PDs raised to 0.2% and 0.3%, respectively.
+In state B, only the riskier asset is affected, having its PD equal to 0.4%.
+Note that the average PD in both states are equal at 0.25%.
+However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B.
+The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%).
+Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states.
+This net procyclical effect is in line with the results of @behn2016procyclical.
+However, due to the increased portfolio dispersion, capital savings are countercyclical in state B.
+Conversely, because the PD-RW formula elasticity is lower for higher PD estimates, in stata A *Capital savings* are procyclical.
+Therefore, capital savings are procyclical in state A.
+Table \ref{tab1} summarizes the example and Figure \ref{fig:pdmapcases}
+
+Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the PD distribution.
+
+## Empirical Results
+
+In Table \ref{tab1}, I regress the change in the logarithm of *Capital savings* on the change in the logatithm of real GDP per capita.
+In all regressions the measure of *Capital savings* follows equation \ref{eq4}.
+The difference between regressions in the table is the portfolio used to calculate the *Capital savings*.
+The real GDP per capita used  comes from the country where the bank has its headquarter.
+All regressions include bank and year fixed effects and standard errors clustered at the country level.
+
+Column 1 shows that total *Capital savings* are countercyclical as the growth rate of tota l*Capital savings* is negatively associated to the GDP growth rate.
+According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average.
+The effect is economically sizeable representing 14% of total *Capital savings* standard deviation in the sample.
+
+The remaining columns of Table \ref{tab1} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios.
+The estimated coefficient in Column 2 is negative suggesting that *Capital savings* generated by retail portfolios are also countercyclical.
+However, this coefficient is not statistically significant.
+On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant.
+According to the regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation.
+Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios.
+Only in regression 4 the coefficient of GDP growth rate is negative and statistically significant.
+
+## Robustness to Alternative Measures
+
+A concern to the previous analysis is the possibility of that I could be overestimating the average impact of GDP growth rate on *Capital savings*.
+First, Because the measure used in Table \ref{tab1} is dimentionless, i.e., it is a measure of intensity of *Capital savings*, if the portfolio of small banks, in terms of asset, are more impacted during the business cycle than the portfolio of large banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank.
+To account for differences in *Capital savings* size, in Table \ref{tab2} the measure of *Capital savings* follows equation \ref{eq3}, which is a measure of *Capital savings* in monetary units.
+
+The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.
+
+Second, because both measures in Tables \ref{tab1} and \ref{tab2} are relative to the IRB portfolio, if the portfolio of banks with lower share of their portfolio under the IRB approach are more impacted during the business cycle than the portfolio of high IRB share banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank.
+The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.
+
+Lastly, the method that I proposed to measure the shape component of the PD distribution could be flawed.
+In Table \ref{tab3} I use the well-known Gini coefficient to measure the shape component of the PD distribution.
+Again, the results shown in Table \ref{tab3} suppport the evidence from the last three tables: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
+
+## Capital Savings and Portfolio Realocation
+
+The results from Tables \ref{tab1} to XX raises the questions of why it appears that only corporate portfolios are impacted during the business cycle. 
+I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases.
+This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
+
+## Economic Impact of Capital Savings Cyclicality
+
+# Policy Discussion: Using Capital Savings to Reduce Capital Ratio Procyclicality
+
+To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
+The capital ratio decomposition is shown in equation \ref{eq5}.
+The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
+To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
+Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab4} shows that *Capital savings* are the only countercyclical component of capital ratios.
+The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
+
+\begin{equation}
+\begin{aligned}
+\label{eq5}
+\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
+&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^r_{i,t} 
+\end{aligned}
+\end{equation}
+
+# Conclusion and Discussion
+
+This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements.
+*Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are economically relevant.
+They explain 27% of the RW variation across banks.
+*Capital savings* are also countercyclical and reduce the procyclicality of RWA by 11.6%.
+Hence, the concavity of IRB formulas could be explored to reduce procyclicality of capital regulation.
+
+# References {-}
+
+<!-- <div id="refs"></div> -->
+
+\newpage
+
+# Figures {-}
 
@@
 
-I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
-The method is a generalization of the example in Figure \ref{fig:pdmap}: for each bank-year combination, compare the actual capital requirement to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio’s average PD.
-
-I define the counterfactual RW as:
-
-\begin{equation}
-\label{eq2}
-\text{RW}^c= 12.5\frac{\sum_{i}\text{EAD}_i\text{LGD}_iN\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\overline{\text{PD}})+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)}{\sum_i \text{EAD}_i},
-\end{equation}
-in which the PD parameter for all assets is replaced by the portfolio weighted average:
-
-\begin{equation*}
-\overline{\text{PD}}=\frac{\sum_i\text{EAD}_i\text{LGD}_i\text{PD}_i}{\sum_i\text{EAD}_i\text{LGD}_i}.
-\end{equation*}
-
-The main measure of *Capital savings* used in this paper is the ratio of equation \ref{eq2} to equation \ref{eq1} as follows:
-
-\begin{equation}
-\label{eq4}
-\text{RW}^r=\frac{\text{RW}^c}{\text{RW}}.
-\end{equation}
-
-$\text{RW}^r$ is the rate at which a bank saves capital.
-For instance, a value of 0.5 means this portfolio would require 50% more capital in the abscense of *Capital savings* to keep the same capital ratio.
-A value of zero means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
-Any PD distribution with positive variance results in positive *Capital savings*.
-
-An alternative way to measure *Capital savings* is in monetary units as follows:
-
-\begin{equation}
-\label{eq3}
-\text{RWA}^s=\text{RWA}^c-\text{RWA}.
-\end{equation}
-
-## Capital Savings and Cyclicality
-
-I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
-Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
-However, as credit conditions deteriorates (and banks react to these new credit conditions) the PD distribution and, consequently, the amount of *Capital savings* also change.
-A priori the direction of the economic cycle effect on *Capital savings* is ambiguous.
-At one extreme, if the entire portfolio is equally affected, then there is only a shift of the location parameter of the PD distribution.
-Because the PD-RW formula elasticity is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
-On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
-The ambiguity arises because *Capital savings* can only increase during a recession if the shape of the PD distribution is affect during the cycle.
-
-\begin{table}[!htbp] \centering 
-  \caption{Cyclicality of Capital Savings} 
-  \label{tab1} 
-\small 
-\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}} 
-\hline 
-\hline \\[-1.8ex] 
-\\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6}
-\\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\ 
-\hline \\[-1.8ex] 
- $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\ 
-  $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\ 
-  $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\ 
-  \hline \\[-1.8ex]
- Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\ 
-  S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\ 
-\hline \\[-1.8ex]
-
-\end{tabular} 
-\end{table}
-
-I use a simple example to illustrate this ambiguity.
-Consider a bank with the asset side of its balance sheet composed of \$100 of a safer asset and \$100 of a riskier asset and three states of the economy.
-Without loss of generality, assume $\text{LGD}=1$ and $\text{R}=0.15$.
-In the initial state, safer and riskier assets have a 0.1% and 0.2% probability of default, respectively.
-According to equation \ref{eq1}, the actual RWA of this bank is \$63.9.
-The weighted average of this portfolio in terms of $\text{PD}$ is 0.15% which yields a counterfactual RWA of \$64.7 according to equation \ref{eq2}.
-Hence, in the initial state *Capital savings* equal \$0.8.
-
 ```{r parameters curve 2, echo=FALSE, include=FALSE}
@@
 
-Consider now a credit risk shock.
-In state A, both safer and riskier assets are affected, having their PDs raised to 0.2% and 0.3%, respectively.
-In state B, only the riskier asset is affected, having its PD equal to 0.4%.
-Note that the average PD in both states are equal at 0.25%.
-However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B.
-The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%).
-Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states.
-This net procyclical effect is in line with the results of @behn2016procyclical.
-However, due to the increased portfolio dispersion, capital savings are countercyclical in state B.
-Conversely, because the PD-RW formula elasticity is lower for higher PD estimates, in stata A *Capital savings* are procyclical.
-Therefore, capital savings are procyclical in state A.
-Table \ref{tab1} summarizes the example and Figure Figure \ref{fig:pdmapcases}
-
-Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the PD distribution.
-
-# Data
-
-To calculate the proposed measures of capital savings, one needs information on the distribution of banks’ regulatory risk parameters.
-However, data at such granularity is not available in the common balance sheet datasets.
-Although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
-Therefore, to overcome data limitations I collected information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports.
-The sample consists of 50 large banking groups, from 12 countries, that have adopted the IRB approach.
-For each bank, I collected consolidated information from the year the IRB approach was approved until 2018.
-The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate and sovereign for wholesale, and real estate and qualifying revolving credit for retail).
-This level of portfolio breakdown is important because regulatory formulas vary across categories, and consequently, more precise knowledge of portfolio composition reduces measurements errors.
-
 ```{r}
@@
 
-Figure \ref{fig:savingsbank} shows that the *Capital savings ratio* calculated with the compiled dataset.
-Each point is a bank-year observations and the bars indicates the average value across time for each bank.
-There is substancial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
-Variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
-*Capital savings ratio* also varies between and within countries.
-Figure \ref{fig:savingscountry}  aggregates the dataset by country.
-In the figure, each point is a bank observation and the bars are the average of these points.
-Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).
-
 ```{r}
@@
 
-\pagebreak
-
-## RW Decomposition
-
-In this section I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
-I divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
-Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
-I further breakdown RW_{IRB} into its portfolio components, $\text{RW}_{IRB}=\sum_j q_{j}\text{RW}_{j}$, in which the $j$ subscript stands for the different portfolios under the IRB approach: Wholesale, retail and equity.
-Equation \ref{eq5} shows the final decomposition:
-
-\begin{equation}
-\begin{aligned}
-\label{eq5}
-\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j} \text{RW}_{j}^{c} \Delta q_{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
-&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin},
-\end{aligned}
-\end{equation}
-in which  $\Delta$ stands for changes relative to a benchmark.
-For the cross-section decomposition, the benchmark is the average bank.
-For the time decomposition, the benchmark is either the year when the IRB approach was first introduced at the bank level or 2008.
-The superscript $B$ stands for benchmark.
-For instance, $q_SA^B$ is the share of the portfolio under the standardized approach for the benchmark.
-
 ```{r}
@@
  plot1 <- cross.section.decomposition %>% 
-  filter(q.SA<1)%>%
   mutate(total= -PD.gain)%>%
@@
 plot2 <- cross.section.decomposition %>% 
-  filter(q.SA<1)%>%
   select(name, Country, int.gain, ext.gain) %>%
@@
 
-I decompose the variation of RW relative to the benchmark in six components, two of them relating to *Capital savings*.
-The first, $(1-q^{B}_{SA})\Delta \text{RW}^{s}$, measures the intensity of *Capital savings*.
-That is, I fix the share of the portfolio under the IRB approach ($q_{SA}^B$) and variation in the component comes only from changes of $RW^s$ relative to the benchmark.
-The second component, $\text{RW}^{s}\Delta q^{B}_{SA}$, measures the extensive margin of *Capital savings*, i.e., variation in this component comes only from changes of the share under the IRB approach relative to the benchmark.
-Note that the magnitude of the second component is moderated by how intense *Capital savings* are in absolute terms.
-
-The remaining components are extensively discussed by @cannata2012inside.
-The first on the right-hand side of equation \ref{eq5} is the intensity of RW under the standardized approach ($q^{B}_{SA} \Delta \text{RW}_{SA}$).
-The contribution of this component to $\Delta \text{RW}$ comes from variaton of RW in portfolios under the standardized approach while fixing the share of this portfolio for all banks.
-The second term measures the intensity of RW under the IRB approach excluding the *Capital savings*, or differently, the intensity of the average PD ($(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j$).
-The contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of RW in the three portfolios under the IRB approach relative to the respective benchmark RW while holding constant the share of each IRB portfolio and the share of total IRB.
-The third component measure the contribution of the IRB portfolio mix.
-Because some types of portfolios have on average lower RW, the composition within IRB across types of portfolios matter to how large overall RW is.
-For instance retail portfolios tend to have lower RW relative to wholesale portfolios because the former usually are colletarized loans, such as mortgages, while the latter is mostly composed of corporate loans.
-Hence, the contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of the three portfolios shares within the IRB portfolio relative to the respective benchmarks while holding constant the respective intensity of average PD ($RW_j^c$) and the share of total IRB.
-Lastly, $(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}$ is the roll-out effect, i.e., the contribution to $\Delta \text{RW}$ coming from moving the portfolio from the standardized approach to the IRB approach.
-
-Figure \ref{fig:crossdecomposition1} plots the intensive and extensive margin of each bank in my sample using the average bank as benchmark.
-The scatter plot shows that the correlation between the two *Capital savings* components is not statistically significant.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
-If Goldman Sachs is included the correlation becomes statistically significant.} The figure suggests that banks benefit from the two components separately, some benefting more of the extensive margin, such as Northern Trust, and others more of the intensive margin, such as BayernLB.
-
 ```{r}
@@
 ```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
-plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
+plot1 <- cross.section.decomposition %>% 
   mutate(total= -PD.gain)%>%
@@
 plot2 <- cross.section.decomposition %>%
-  filter(q.SA<1)%>%
   select(name, Country, other.gain, PD.gain) %>%
@@
 
-Figure \ref{fig:crossdecomposition2} shows that the net effect of the intensive and extensive margin is negatively correlated to the net effect of the other components.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
-If Goldman Sachs is included the correlation becomes even more negative.} This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
-
 ```{r}
@@
 
-unrestrict <- cross.section.decomposition %>% filter(q.SA<1) %>%
+unrestrict <- cross.section.decomposition %>% 
   mutate(epsilon = rnorm(nrow(.),0,0.0000001),
@@
 
-As Figure \ref{fig:rsqrtdecomp} shows *Capital savings* components account for 27% of the RW variation across banks.
-Each component accounts for around 13.5% of the RW variation and are second and third in rank, losing only to the average PD contribution.
-I derive the $\text{R}^2$ contribution of each RW component using the Shapley-Owen decomposition method.
-This method consider the contribution of each variable to total $\text{R}^2$ as the sum of $\text{R}^2$ from regressions using all possible variable combinations that include the variable of interest and weighting each regression by the number of variable permutations.
-In sum, figure \ref{fig:rsqrtdecomp} offers strong evindence of the relative importance of *Capital savings* as source of variation to RW across banks.
-
-Figure \ref{fig:timedecomposition} shows the evolution of RW components relative to the time of IRB introduction (left-hand side plot), and relative to 2008 (right-hand side plot).
-The negative values for the extensive margin implies that banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach.
-However, during the same period, the intensity of *Capital savings* has, on average, decreased in time.
-That is, for the same average PD, capital requirements have increased from 2008 to 2018 and since the year IRB approach was introduced.
-
-The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}.
-Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.\footnote{Although supervisors favored raising additional capital as means to achieve the required capital ratios \citep{EBA2011recommendation}, banks respond to capital requirements mostly by reducing RWA \citep{gropp2018banks}}  The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
-This distribution compression leads to a reduction in *Capital savings*.
-This explanation is in line with the hypothesis that *Capital savings* are countercyclical.
-Efforts to increase capital ratio by risk reduction are increasingly more difficult.
-
 ```{r}
@@
 
-\pagebreak
-
-## Other Determinants of Capital Savings and Summary Statistics
-
-Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
-Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks ].
-In this section, I use these prior evidence to develop my expectations on how each of these variables may affect capital savings.
-
-Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
-
-```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-p_w <- c(0.05, 0.95)         ## Set winsorizing limits
+<!-- \begin{table}[!htbp] \centering  -->
+<!--   \caption{Cyclicality of Capital Savings}  -->
+<!--   \label{tab1}  -->
+<!-- \small  -->
+<!-- \begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}}  -->
+<!-- \hline  -->
+<!-- \hline \\[-1.8ex]  -->
+<!-- \\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6} -->
+<!-- \\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\  -->
+<!-- \hline \\[-1.8ex]  -->
+<!--  $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\  -->
+<!--   $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\  -->
+<!--   $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\  -->
+<!--   \hline \\[-1.8ex] -->
+<!--  Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\  -->
+<!--   S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\  -->
+<!-- \hline \\[-1.8ex]  -->
+<!-- \end{tabular}  -->
+<!-- \end{table}  -->
+```{r winsorized_data, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+## Set winsorizing limits
+p_w <- c(0.05, 0.95)        
 bank.data.win <- bank.data %>% 
@@
 
-bank.data.win %>% select(
-  log.CAR_gr, 
-  log.K_gr, 
-  log.RWA.IRB_gr, 
-  log.RWA.hat_gr, 
-  log.RWA.s_gr,
-  log.RWA.r_gr,
-  log.RWA.r.Retail_gr,
-  log.RWA.r.Wholesale_gr, 
-  log.RWA.r.Corporate_gr, 
-  log.RWA.r.Sovereign_gr, 
-  log.RWA.r.Banks_gr, 
-  mean.PD_gr, 
-  var.PD_gr, 
-  skew.PD_gr,
-  kurt.PD_gr, 
-  gini.PD_gr,
-  log.gdp_gr,
- #log.asset_gr, 
- #capital.ratio_gr, 
- #loans.ratio_gr, 
- #NII.ratio_gr,  
- #deposit.ratio_gr,
- #income.ratio_gr,
-  log.gdp,
-  log.asset,
-  capital.ratio,
-  income.ratio,
-  loans.ratio, 
-  NII.ratio, 
-  NPL.ratio,
-  deposit.ratio) %>%
-  as.data.frame() %>% 
-stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
-          label = "tab:summary",
-          omit.summary.stat = c("p25","p75"),
-          covariate.labels = 
-            c("$\\Delta \\text{Log Capital Ratio}$",
-              "\\hspace{0.5cm}$\\Delta \\text{Log capital}$",
-              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}$",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^r$",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^r$ Retail",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^r$ Wholesale",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Corporate",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Sovereign", 
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^r$ Banks", 
-              "\\hspace{1.0cm}$\\Delta$ Log PD Mean",
-              "\\hspace{1.0cm}$\\Delta$ Log PD Variance",
-              "\\hspace{1.0cm}$\\Delta$ Log PD Skewness", 
-              "\\hspace{1.0cm}$\\Delta$ Log PD Kurtosis", 
-              "\\hspace{1.0cm}$\\Delta$ Log Gini",
-              "$\\Delta$ Log GDP",
-              #"$\\Delta$ Log Total Assets",
-              #"$\\Delta$ Capital ratio",
-              #"$\\Delta$ Loans / Total Assets",
-              #"$\\Delta$ Net Interest Income / Oper. Rev.",
-              #"$\\Delta$ Deposits / Total Assets", 
-              #"$\\Delta$ Net Income / Total Assets", 
-              "Log Total Assets",
-              "Capital Ratio",
-              "Net Income / Total Assets",
-              "Loans / Total Assets",
-              "Net Interest Income / Oper. Rev.",
-              "Non-performing Loans / Loans",
-              "Deposits / Total Assets"),
-          title = "Summary statistics",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{14cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\\%}", 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE)
-```
-
-The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
-First, I control for bank size using the log of total assets (in millions of US dollars).
-The expectation regarding the impact of bank size is ambiguous.
-On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
-On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
-I also include the regulatory capital ratio as control.
-Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
-However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
-Thus, the expected effect of capital ratio on capital savings is also ambiguous.
-Next, I add bank profitability using net income over total assets.
-Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
-
-The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
-For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
-Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
-As a result, I expect banks with higher shares of these variables to have more *Capital savings*.
-
-Finally, I study the relationship of *Capital savings* to risk-taking.
-The RW decomposition analysis suggests that riskier banks benefit more from *Capital savings*.
-I add the ratio of non-performing loans to total loans to test this proposition.
-I also include bank funding using the ratio of total deposits to total assets.
-Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
-The effect of deposits funding on *Capital savings* depends on how risk-taking is manifested by the deposit funding.
-If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then deposits to total assets are negatively correlated to *Capital savings*.
-Conversely, *Capital savings* increase if banks take risk by having fatter upper-tails in the PD distribution.
-
-# Capital Savings and Business Cycle
-
-In this section I investigate the relationship between *Capital savings* and the business cycle.
-Overall, capital requirements are procyclical: during economic downturns credit risk in the economy increases which raises capital requirements, under the IRB approach forcing banks to cut down lending.
-However, as credit conditions changes (and banks react to it) the amount of *Capital savings* also changes.
-The expected direction of *Capital savings* cyclicality is ambiguous.
-On the one hand, if the entire portfolio is equally affected, capital savings are procyclical.
-On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
-The ambiguity arises because *Capital savings* can only increase, when credit conditions deteriorates, if the credit distribution widens.
-In the first situation, there is only a shift of the location parameters of the distribution and, consequently, a reduction of *Capital savings*.
-
-```{r reg_portfolio, echo=FALSE}
-
+``` 
+# Tables {-}
+
+```{r reg_cyclicality_main, echo=FALSE}
+
+# Variables in the model 
+# Dependent variables
 dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
-aux      <- paste0("mean.PD", dep.vars)
-ind.vars <- paste0("log.gdp_gr")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
+# Independent variables
+ind.vars <- c("log.gdp_gr")
+ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")
+
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
 
+# Run the models
 reg.list <- lapply(formulas, function(x) {
@@
 
+# Robust standard errors
 se.list  <-  lapply(reg.list, function(x){
@@
 
-```{r table_portfolio, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
+```{r table_cyclicality_main, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Cyclicality of Capital Savings",
+          title = "Capital Savings and the Business Cycle",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} for portfolio $j$.
+In all regressions the measure of \\textit{Capital savings} follows equation \\ref{measure1}.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "small",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab1",
+          style            = "qje",
+          column.sep.width = "1pt",
+          align            = TRUE)
+```
+
+\newpage
+
+\begin{table}[!htbp] \centering  
+\caption{Capital Savings and the Business Cycle: Robustness to Alternative Measures}  
+\label{tab2:reg}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{16cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}.
+\end{equation*} \newline
+The dependent variable is the change in the logarithm of \textit{Capital savings} for portfolio $j$.
+In Panel A the measure of \textit{Capital savings} follows equation \ref{eq5}.
+In Panel B the measure of \textit{Capital savings} follows equation \ref{eq6}.
+In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution.
+$\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
+[-1.8ex] \\
+\hline \\
+\end{tabular}
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}$} \\
+\end{tabular}
+
+```{r reg_cyclicality_alt, echo=FALSE}
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+ felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_alt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se              = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          float            = FALSE,
+          align            = TRUE)
+
+```
+
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+ \\
+\multicolumn{6}{c}{Panel B: $\text{Y}=\text{Log RW}^{s}$ } \\
+\end{tabular}
+```{r reg_cyclicality_rw, echo=FALSE}
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("RW.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+ felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_rw, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se              = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          float            = FALSE,
+          align            = TRUE)
+
+```
+
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\\
+\multicolumn{6}{c}{Panel C: $\text{Y}=\text{Log Gini}$ } \\
+\end{tabular}
+
+```{r reg_cyclicality_gini, echo=FALSE}
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          title = "Alternative measure: PD distribution Gini Coefficient",
+          table.layout = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab3",
+          style            = "qje",
+          column.sep.width = "2pt",
+          float            = FALSE, 
+          align            = TRUE)
+
+```
+
+\end{table}
+
+\newpage
+
+```{r reg_cyclicality_components, echo=FALSE}
+dep.vars2 <- c("log.RWA.SA_gr","RW.SA_gr ", "q.SA_gr", "q.Wholesale_gr", "q.Retail_gr")
+dep.text2 <- c("$\\Delta\\text{Log RWA}_{i,t}^{SA}$", "$\\Delta\\text{Log RW}_{i,t}^{SA}$",
+               "$\\Delta\\text{q}_{i,t}^{SA}$", "$\\Delta\\text{q}_{i,t}^{Wholesale}$", "$\\Delta\\text{q}_{i,t}^{Retail}$")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+ felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_components, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          title = "Other RWA Components and the Business Cycle",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{14cm}{The table examines the relationship between the business cycle and other components of RWA.
+The dependent variable is the change in the logarithm of risk-weighted assets for portfolios under the standardize approach in column 1, the change in the logarithm of risk-weights for portfolios under the standardize approach in column 2, the change in the logarithm of the ratio of exposure under the standardize approach to total exposure in column 3, the change in the logarithm of the share of the IRB portfolio as wholesale exposure in column 4, and the change in the logarithm of the share of the IRB portfolio as retail exposure in column 5.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels   = dep.text2,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "small",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE)
+
+```
+
+\newpage
+
+```{r reg_moments, echo=FALSE}
+
+# Variables in the model 
+dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr", "log.RWA.r_gr")
+aux.1 <- c("$\\Delta \\text{Mean(PD}_{i,t}$)", "$\\Delta \\text{Variance(PD}_{i,t}$)", "$\\Delta \\text{Skewness(PD}_{i,t}$)", "$\\Delta \\text{Kurtosis(PD}_{i,t}$)")
+aux.2 <- c( "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
+dep.text <- c(aux.1, aux.2)
+ind.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr")
+ind.vars  <- paste(ind.vars, collapse=" + ")
+ind.vars <- c(rep("log.gdp_gr", 4), ind.vars)
+ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$", aux.1)
+
+# Fixed effects
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, RWA.s>0))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          title = "Alternative measure: PD distribution Gini Coefficient",
           table.layout = "n=c=d#-t-as-",
@@
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-Standard errors are adjusted for heteroskedasticity.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
 *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
           #keep            = c("log.gdp_gr"),
-          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "small",
           notes.label      = "",
@@
           notes.append     = FALSE,
-          float = TRUE,
-          label = "tab1",
-          style = "qje",
-          column.sep.width = "2pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-Table \ref{tab1} shows that total *Capital savings* are countercyclical.
-According to the regression (1) a 1.7% GDP growth rate (one standard deviation) decreases on average *Capital savings* growth rate by 0.83 percentage points.
-The effect is economically significant as it represents a 17% of *Capital savings* standard deviation in the sample.
-The evidence in Table \ref{tab1} suggests that the effect is driven by wholesale portfolios, especially corporate portfolios.
-
-## Alternative measures
-
-The results in Table \ref{tab1} are robust to alternative measures of *Capital savings*.
-First, I replace the ratio measure of *Capital savings* by the measure in monetary units (equation \ref{eq3}).
-The results shown in Table \ref{tab2} reinforce the evidence of coutercyclicality of *Capital savings* is likely driven by the corporate portfolio.
-
-```{r reg_unit, echo=FALSE}
-
-dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
-aux      <- paste0("mean.PD", dep.vars)
-ind.vars <- paste0("log.gdp_gr")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("log.RWA.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-reg.list <- lapply(formulas, function(x) {
- felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_unit, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Alternative Measures: Capital Savings in Monetary Units",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^s = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
-          #keep = c("log.gdp_gr"),
-          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          float = TRUE,
-          label = "tab2",
-          style = "qje",
-          column.sep.width = "2pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-Second, I use Gini coefficient to measure the dispersion in the PD distributions. Again, the results shown in Table \ref{tab3} suppport the evidence from the last two tables.
-
-```{r reg_gini, echo=FALSE}
-
-dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
-aux      <- paste0("mean.PD", dep.vars)
-ind.vars <- paste0("log.gdp_gr")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Alternative Measures: Gini Coefficient",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Gini}_{i,j,t} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
-          #keep = c("log.gdp_gr"),
-          dep.var.labels = c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          float = TRUE,
           label            = "tab3",
@@
           #float.env        = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-## Effect on Capital Ratio procyclicality
-
-To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
-The capital ratio decomposition is shown in equation \ref{eq5}.
-The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
-To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
-Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab4} shows that *Capital savings* are the only countercyclical component of capital ratios.
-The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
-
-\begin{equation}
-\begin{aligned}
-\label{eq5}
-\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
-&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^r_{i,t} 
-\end{aligned}
-\end{equation}
+          align            = TRUE)
+
+```
+
+\newpage
 
@@
 # Variables in the model 
+# Dependent variables
 dep.vars <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat_gr", "log.RWA.r_gr")
+dep.text <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
+# Independent variables
 ind.vars <- paste0("log.gdp_gr")
+ind.text <- c("$\\Delta\\text{Log GDP}_{i,t-1}$")
 
 # Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = subset(bank.data, subset = !is.na(log.CAR_gr)))
+  felm(x, data = filter(bank.data.win, RWA.s>0 & !is.na(log.CAR_gr)))
 })
@@
 
+fe.list <- list(c("Effect", effect.list),
+                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
 ```
@@
 stargazer(reg.list,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Cyclicality of Capital Ratio Components",
+          title = "Alternative measure: PD distribution Gini Coefficient",
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,j,t} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{r} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
 \\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-Standard errors are adjusted for heteroskedasticity.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
 *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = 
-            c("$\\Delta\\text{Log GDP}_{i,t}$", 
-              #"$\\Delta\\text{Mean PD}_{i,t}$",
-              #"$\\Delta\\text{Log Total Assets}_{i,t}$",
-              #"$\\Delta\\text{Deposits / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Loans / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Income / Total Assets}_{i,t}$", 
-              #"$\\Delta\\text{Net Interest Income / Oper. Rev.}_{i,t}$",
-              "$\\text{PD Mean}_{i,j,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Deposits / Total Assets}_{i,t-1}$", 
-              "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$"),
+
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
           #keep            = c("log.gdp_gr"),
-          dep.var.labels = c("$\\Delta\\text{Log capital Ratio}_{i,t}$","$\\Delta\\text{Capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}^{c}_{i,t}$", "$\\Delta\\text{Log RWA}^{r}_{i,t}$"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "small",
           notes.label      = "",
@@
           notes.append     = FALSE,
-          float = TRUE,
-          label = "tab4",
+          label            = "tab3",
           style            = "qje",
@@
           #float.env        = "sidewaystable",
-          align = TRUE,
-          add.lines = list(c("Effect", effect.list),
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-
-```
-
-## Moments
-
-I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases.
-This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
-
-```{r reg_moments, echo=FALSE}
-
-# Variables in the model 
-dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr")
-ind.vars <- paste0("log.gdp_gr")
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Cyclicality of PD distribution Moments",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{11cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Moment}_{i,t} = \\beta\\times \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = se.list,
-          covariate.labels = c("$\\Delta\\text{Log GDP}_{i,t}$", "$\\overline{\\text{PD}}_{i,t-1}$","$\\text{Log Total Assets}_{i,t-1}$", "$\\text{Deposits / Total Assets}_{i,t-1}$", "$\\text{Loans / Total Assets}_{i,t-1}$", "$\\text{Net Income / Total Assets}_{i,t-1}$", "$\\text{Net Interest Income / Oper.
-Rev.}_{i,t-1}$"),
-          keep = c("log.gdp_gr"),
-          dep.var.labels = c("Mean", "Variance", "Skewness", "Kurtosis"), 
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          label = "",
-          style = "qje",
-          column.sep.width = "3pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(
-                           c("Bank FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-```
-
-```{r reg_extension1, echo=FALSE, include=FALSE}
+          align            = TRUE)
+
+```
+
+\newpage
+
+```{r reg_extension, echo=FALSE, include=FALSE}
 
@@
 aux.1 <- c("log.gdp",
-           #"log.gdp_gr",
+           "bank_credit_gdp",
+           #"pillarI_stringency",
+           "cap_reg",
+           "Sup_Power",
            "NPL.ratio",
-           "capital.ratio",
            "log.asset",
-           "pillarI_stringency",
-           #"cap_reg",
-           "Sup_Power",
            "Zscore",
@@
 ind.vars  <- paste(aux.1, collapse=" + ")
-ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
-              #"$\\Delta\\text{ Log GDP}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$", 
-              "$\\text{Pillar I stringency}_{i,t-1}$", 
-              #"$\\text{Capital stringency}_{i,t-1}$", 
-              "$\\text{Official Supervisory power}_{i,t-1}$",
-              "$\\text{Z-score}_{i,t-1}$",
+ind.text <- c("$\\text{Log GDP}_{i,t}$",
+              "$\\text{Credit-to-GDP}_{i,t}$",
+              #"$\\text{Pillar I stringency}_{i,t}$", 
+              "$\\text{Capital stringency}_{i,t}$", 
+              "$\\text{Official Supervisory power}_{i,t}$",
+               "$\\text{NPL Ratio}_{i,t}$",
+              "$\\text{Log Total Assets}_{i,t}$",
+              "$\\text{Z-score}_{i,t}$",
               # "$\\text{Loans / Total Assets}_{i,t-1}$", 
@@
              # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{ROE}_{i,t-1}$",
-              "$\\text{ROA}_{i,t-1}$")
-
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
+              "$\\text{ROE}_{i,t}$",
+              "$\\text{ROA}_{i,t}$")
+
+fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
 
 # Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | 0 ")
+formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
 formulas <- lapply(formula.text, formula)
@@
 
-```{r table_extension1, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
+```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
 
@@
 $X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
 *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
@@
           add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
           keep.stat        = c("n","rsq"),
@@
           style            = "qje",
-          column.sep.width = "-15pt",
+          column.sep.width = "-1pt",
           #float.env        = "sidewaystable",
@@
 ```
-```{r reg_extension2, echo=FALSE, include=FALSE}
-
-# Variables in the model 
-# Dependent variables
-dep.vars <- c("log.RWA.s")
-dep.text <- c("$\\text{Log RWA}^{s}_{i,t}$")
-# Independent variables
-aux.1 <- c("log.gdp",
-           "NPL.ratio",
-           "capital.ratio",
-           "log.asset",
-           "pillarI_stringency",
-           #"cap_reg",
-           "Sup_Power",
-           "Zscore",
-          "NII.ratio",
-          "ROE",
-          "ROA")
-aux.3 <- paste(aux.1, collapse=" + ")
-ind.vars <- c(aux.1, aux.3)
-ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$", 
-              "$\\text{Pillar I stringency}_{i,t-1}$", 
-              #"$\\text{Capital stringency}_{i,t-1}$", 
-              "$\\text{Official Supervisory power}_{i,t-1}$",
-              "$\\text{Z-score}_{i,t-1}$",
-              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
-             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{ROE}_{i,t-1}$",
-              "$\\text{ROA}_{i,t-1}$")
-# Fixed effects
-fe       <- c(rep("year + bvdid"  , length(aux.1)),
-              rep("year + bvdid"  , length(aux.3)))
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
-                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_extension2, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
-
-stargazer(reg.list,
-          title = "capital Savings and Bank Characteristics",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "tab6",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "",
-          style            = "qje",
-          column.sep.width = "-15pt",
-          float.env        = "sidewaystable",
-          align            = TRUE)
-```
-
-```{r reg_extension3, echo=FALSE, include=FALSE}
-
-# Variables in the model 
-# Dependent variables
-dep.vars <- c("gini.PD")
-dep.text <- c("$\\text{Log RWA}^{r}_{i,t}$")
-# Independent variables
-aux.1 <- c("log.gdp",
-           "NPL.ratio",
-           "capital.ratio",
-           "log.asset",
-           "pillarI_stringency",
-           #"cap_reg",
-           "Sup_Power",
-           "Zscore",
-          "NII.ratio",
-          "ROE",
-          "ROA")
-aux.3 <- paste(aux.1, collapse=" + ")
-ind.vars <- c(aux.1, aux.3)
-ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$", 
-              "$\\text{Pillar I stringency}_{i,t-1}$", 
-              #"$\\text{Capital stringency}_{i,t-1}$", 
-              "$\\text{Official Supervisory power}_{i,t-1}$",
-              "$\\text{Z-score}_{i,t-1}$",
-              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
-             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{ROE}_{i,t-1}$",
-              "$\\text{ROA}_{i,t-1}$")
-# Fixed effects
-fe       <- c(rep("year + bvdid"  , length(aux.1)),
-              rep("year + bvdid"  , length(aux.3)))
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
-                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_extension3, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
-
-stargazer(reg.list,
-          title = "capital Savings and Bank Characteristics",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "tab6",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "",
-          style            = "qje",
-          column.sep.width = "-15pt",
-          float.env        = "sidewaystable",
-          align            = TRUE)
-```
-
-```{r reg_extension4, echo=FALSE, include=FALSE}
-
-# Variables in the model 
-# Dependent variables
-dep.vars <- c("log.RWA.r.Wholesale")
-dep.text <- c("$\\text{Log RWA}^{r}_{i,t}$")
-# Independent variables
-aux.1 <- c("log.gdp",
-           "NPL.ratio",
-           "capital.ratio",
-           "log.asset",
-           "pillarI_stringency",
-           #"cap_reg",
-           "Sup_Power",
-           "Zscore",
-          "NII.ratio",
-          "ROE",
-          "ROA")
-aux.3 <- paste(aux.1, collapse=" + ")
-ind.vars <- c(aux.1, aux.3)
-ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$", 
-              "$\\text{Pillar I stringency}_{i,t-1}$", 
-              #"$\\text{Capital stringency}_{i,t-1}$", 
-              "$\\text{Official Supervisory power}_{i,t-1}$",
-              "$\\text{Z-score}_{i,t-1}$",
-              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
-             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{ROE}_{i,t-1}$",
-              "$\\text{ROA}_{i,t-1}$")
-# Fixed effects
-fe       <- c(rep("year + bvdid"  , length(aux.1)),
-              rep("year + bvdid"  , length(aux.3)))
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
-                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_extension4, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
-
-stargazer(reg.list,
-          title = "capital Savings and Bank Characteristics",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "tab6",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "",
-          style            = "qje",
-          column.sep.width = "-15pt",
-          float.env        = "sidewaystable",
-          align            = TRUE)
-```
-```{r reg_extension, echo=FALSE, include=FALSE}
-
-# Variables in the model 
-# Dependent variables
-dep.vars <- c("log.RWA.s")
-dep.text <- c("$\\text{Log RWA}^{r}_{i,t}$")
-# Independent variables
-aux.1 <- c("lag(log.gdp , order_by = year)",
-           "lag(NPL.ratio, order_by = year)",
-           "lag(capital.ratio, order_by = year)",
-           "lag(log.asset, order_by = year)",
-           "lag(pillarI_stringency, order_by = year)",
-          # "lag(loans.ratio, order_by = year)",
-          # "lag(NII.ratio, order_by = year)",
-          # "lag(income.ratio, order_by = year)",
-          # "lag(ROA, order_by = year)",
-           "lag(subordinated, order_by = year)")
-aux.3 <- paste(aux.1, collapse=" + ")
-ind.vars <- c(aux.1, aux.3)
-ind.text <- c("$\\text{Log GDP}_{i,t-1}$",
-              "$\\text{NPL Ratio}_{i,t-1}$",
-              "$\\text{Capital Ratio}_{i,t-1}$",
-              "$\\text{Log Total Assets}_{i,t-1}$", 
-              "$\\text{Pillar I strigency}_{i,t-1}$", 
-             # "$\\text{Loans / Total Assets}_{i,t-1}$", 
-             # "$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
-             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-             # "$\\text{ROA}_{i,t-1}$",
-              "$\\text{Subordinated Debt}_{i,t-1}$")
-# Fixed effects
-fe       <- c(rep("year + bvdid"  , length(aux.1)),
-              rep("year + bvdid"  , length(aux.3)))
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(aux.1)),
-                                rep("\\multicolumn{1}{c}{Yes}", length(aux.3))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(ind.vars))))
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars), " | ", paste0(fe), " | 0 | 0 ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
-
-stargazer(reg.list,
-          title = "capital Savings and Bank Characteristics",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "tab6",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "",
-          style            = "qje",
-          column.sep.width = "-15pt",
-          float.env        = "sidewaystable",
-          align            = TRUE)
-```
-
-# Conclusion and Discussion
-
-This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements.
-*Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are economically relevant.
-They explain 27% of the RW variation across banks.
-*Capital savings* are also countercyclical and reduce the procyclicality of RWA by 11.6%.
-Hence, the concavity of IRB formulas could be explored to reduce procyclicality of capital regulation.
-
-# References
-
-```{r did data, eval=FALSE, include=FALSE}
-p_w <- c(0.0, 1)         ## Set winsorizing limits
-pre.treat.out <- bank.data %>%
-  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
-  filter(year %in% c("2009","2010"),
-         !is.na(IRB)) %>%
-  select(bvdid, log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
-         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
-         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
-  group_by(bvdid) %>%
-  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
-
-post.treat.out <- bank.data %>%
-  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
-  filter(year %in% c("2012","2013"),
-         !is.na(IRB)) %>%
-  select(bvdid,  log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
-         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
-         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
-  group_by(bvdid) %>%
-  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
-
-DiD.data <- bank.data %>%
-  filter(year %in% c("2010")) %>%
-  right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
-  right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
-  mutate(Shortfall.mn   = log(1+Shortfall.mn), 
-         Shortfall.mn.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.mn, 0), NA),
-         Shortfall.RW.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.RW, 0), NA),
-         out.RWA        = log.RWA.post - log.RWA.pre,
-         out.RWA.IRB    = log.RWA.IRB.post - log.RWA.IRB.pre,
-         out.RWA.SA     = log.RWA.SA.post - log.RWA.SA.pre,
-         out.RWA.hat    = log.RWA.hat.post - log.RWA.hat.pre,
-         out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
-         out.RWA.s.Retail = log.RWA.s.Retail.post - log.RWA.s.Retail.pre,
-         out.RWA.s.Equity = log.RWA.s.Equity.post - log.RWA.s.Equity.pre,
-         out.RWA.s.Wholesale = log.RWA.s.Wholesale.post - log.RWA.s.Wholesale.pre,
-         out.RWA.s.Corporate = log.RWA.s.Corporate.post - log.RWA.s.Corporate.pre,
-         out.RWA.s.Sovereign = log.RWA.s.Sovereign.post - log.RWA.s.Sovereign.pre,
-         out.RWA.s.Banks = log.RWA.s.Banks.post - log.RWA.s.Banks.pre,
-         out.K          = log.K.post - log.K.pre,
-         out.tier       = log.tier.post - log.tier.pre,
-         out.CAR.ratio  = totalcapitalratio.post - totalcapitalratio.pre,
-         out.tier.ratio = tier1ratio.post - tier1ratio.pre,
-         out.CAR.dif    = CAR.dif.post - CAR.dif.pre,
-         out.tier.dif   = tier.dif.post - tier.dif.pre,
-         Country        = as.factor(Country)) %>% 
-  ungroup() %>% 
-  filter(EBAcountry == 1,
-         name != "Bank of Cyprus") %>%
-  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
-
-```
-
-```{r did, eval=FALSE, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth", include=FALSE}
-p_w <- c(0.0, 1)         ## Set winsorizing limits
-DiD.data %>% ungroup()%>%
-  subset(select = c("Shortfall.mn","out.RWA.s","out.RWA.s.Retail","out.RWA.s.Wholesale","out.RWA.s.Corporate","out.RWA.s.Sovereign","out.RWA.s.Banks","out.RWA", "out.K","out.CAR.ratio","out.CAR.dif")) %>%
+
+\newpage
+
+# Appendix {-}
+
+\begin{table}[!htbp] \centering  
+\caption{Summary Statistics}  
+\label{tab1:summary}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
+\end{tabular}
+
+```{r summary_components, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+bank.data.win %>% select(
+  capital.ratio_gr,
+  log.CAR_gr,
+  log.K_gr, 
+  log.RWA.IRB_gr,
+  log.RWA.r_gr,
+  log.RWA.r.Retail_gr,
+  log.RWA.r.Wholesale_gr, 
+  log.RWA.r.Corporate_gr, 
+  log.RWA.r.Sovereign_gr, 
+  log.RWA.r.Banks_gr,
+  log.RWA.hat_gr, 
+  log.RWA.hat.Retail_gr,
+  log.RWA.hat.Wholesale_gr, 
+  log.RWA.hat.Corporate_gr, 
+  log.RWA.hat.Sovereign_gr, 
+  log.RWA.hat.Banks_gr) %>%
   as.data.frame() %>% 
-stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
+stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
+          label = "tab:summary",
           omit.summary.stat = c("p25","p75"),
-          covariate.labels = c("Log Shortfall","$\\Delta \\text{Log RWA}^S$", "$\\Delta \\text{Log RWA}^S_R$", "$\\Delta \\text{Log RWA}^S_W$", "$\\Delta \\text{Log RWA}^S_C$", "$\\Delta \\text{Log RWA}^S_S$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}$", "$\\Delta \\text{Log Capital}$", "$\\Delta \\text{CAR}^S$"  ),
-          title = "Summary statistics",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected.
-Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports.
-The remaining variables come from BankFocus.}",
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE)
-```
-
-```{r, eval = FALSE, fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth", include=FALSE}
- plot1 <- bank.data %>%  ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% 
-  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, tier, tier.hat, IRB) %>%
-  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>% 
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
-  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
-theme(legend.position='none')
-
-plot2 <- bank.data %>% ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, tier, tier.hat, IRB) %>%
-  filter(!is.na(RW.s), IRB == 1) %>%
-  group_by(name)%>%
- summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>%
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="",y="",subtitle = "Plot b") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
-  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
-  theme(legend.position = c(0.8, 0.85), 
-        legend.title = element_blank(), 
-        legend.background=element_blank())
-
-plot3 <- bank.data %>% ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, CAR, CAR.hat, IRB) %>%
-  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  theme(legend.position='none')
-
-plot4 <- bank.data %>% ungroup() %>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>% 
-  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
-  select(name,RW.s, CAR, CAR.hat, IRB) %>%
-  filter(!is.na(RW.s), IRB == 1) %>%
-  group_by(name)%>%
- summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
-  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
- ggplot(mapping = aes(x = RW.s,
-                      y = value,
-                      color = condition,
-                      fill = condition)) +
-geom_point( alpha = 0.5,size = 1) +
-  scale_y_continuous( limits=c(5, 20))+
-  labs(x="Capital savings",y="",  subtitle = "Plot d") +
-  geom_smooth(method = "lm",color="black") +
-  theme_bw()+
-  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
-  theme(legend.position='none')
-
-grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
-```
-
-```{r,eval=FALSE, echo=FALSE}
-
-bank.data.2 <- bank.data %>% ungroup()%>%
- # filter(!is.na(RW.s),
-  #       !is.na(basel),
-   #      tier1capitalmillcu < totalcapitalmillcu)%>%
-  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA, log.RWA.s, log.RWA.hat, CAR.dif, tier.dif, 
-         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  RW, log.tier, log.K,
-         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
-#  group_by(bvdid) %>%
-#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
- ungroup() %>%
- mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
-
-reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
-se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
-
-reg.2 <- felm(log.K  ~ CAR.dif  | year + IRB.bvdid  | 0 | 0 , data = bank.data.2)
-se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
-
-reg.3 <- felm(log.RWA  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
-se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
-
-reg.4 <- felm(log.asset  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
-se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
-
-reg.5 <- felm(RW  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
-se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
-
-```
-
-```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}
-
-stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
-           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Capital Savings and Capital Ratio Components",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times \\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5),
-          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
-          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","Log Assets","RW"), 
+          covariate.labels = 
+            c("$\\Delta \\text{Log Capital Ratio}$",
+              "$\\Delta \\text{Log Adjusted Capital Ratio}$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log capital}$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}$",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^c$ Retail",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^c$ Wholesale",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^c$ Corporate",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^c$ Sovereign", 
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^c$ Banks",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Retail",
+              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Wholesale",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Corporate",
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Sovereign", 
+              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Banks"),
           notes.label = "",
@@
           notes.append = FALSE,
-          label = "",
-          style = "qje",
-          column.sep.width = "3pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-```
-
-```{r,eval=FALSE, echo=FALSE}
-
-bank.data.2 <- bank.data %>% ungroup()%>%
-  filter(!is.na(RW.s),
-         !is.na(basel),
-         tier1capitalmillcu < totalcapitalmillcu)%>%
-  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
-         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  log.tier, log.K,
-         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
-  group_by(bvdid) %>%
-#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
- ungroup() %>%
- mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)
-
-reg.1 <- felm(ROE  ~ CAR.dif | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
-se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
-
-reg.2 <- felm(ROA  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
-se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
-
-reg.3 <- felm(Zscore  ~ CAR.dif  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
-se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
-
-reg.4 <- felm(leverage  ~ CAR.dif  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
-se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
-
-```
-
-```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}
-
-stargazer(reg.1, reg.2, reg.3, reg.4, 
-           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "Capital Savings, Performance, and Risk-Taking",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{15cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times\\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio.
-$\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4),
-          covariate.labels = c("Capital ratio savings"),
-          dep.var.labels = c("ROE","ROA","Z-score","Leverage"), 
+          summary = TRUE,
+          digits = 3,
+          float = FALSE)
+```
+
+\end{table}
+
+\begin{table}[!htbp] \centering  
+\caption{Summary Statistics}  
+\label{tab1:summary}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
+\end{tabular}
+
+```{r summary_other_measures, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+bank.data.win %>% select(
+  RW.s_gr,
+  RW.s.Retail_gr,
+  RW.s.Wholesale_gr, 
+  RW.s.Corporate_gr, 
+  RW.s.Sovereign_gr, 
+  RW.s.Banks_gr,
+  log.RWA.s_gr,
+  log.RWA.s.Retail_gr,
+  log.RWA.s.Wholesale_gr, 
+  log.RWA.s.Corporate_gr, 
+  log.RWA.s.Sovereign_gr, 
+  log.RWA.s.Banks_gr,
+  gini.PD_gr,
+  gini.PD.Retail_gr,
+  gini.PD.Wholesale_gr, 
+  gini.PD.Corporate_gr, 
+  gini.PD.Sovereign_gr, 
+  gini.PD.Banks_gr,
+  log.RWA.r, 
+  log.RWA.r.Retail,
+  log.RWA.r.Wholesale, 
+  log.RWA.r.Corporate, 
+  log.RWA.r.Sovereign, 
+  log.RWA.r.Banks_gr) %>%
+  as.data.frame() %>% 
+stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
+          label = "tab:summary",
+          omit.summary.stat = c("p25","p75"),
+          covariate.labels = 
+            c("\\hspace{0.0cm}$\\Delta \\text{Log RW}^s$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RW}^s$ Retail",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RW}^s$ Wholesale",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RW}^s$ Corporate",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RW}^s$ Sovereign", 
+              "\\hspace{1.0cm}$\\Delta \\text{Log RW}^s$ Banks",
+              "\\hspace{0.0cm}$\\Delta \\text{Log RWA}^{\\$}$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ Retail",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ Wholesale",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ Corporate",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ Sovereign", 
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ Banks",
+              "\\hspace{0.0cm}$\\Delta$ Log Gini",
+              "\\hspace{0.5cm}$\\Delta$ Log Gini Retail",
+              "\\hspace{0.5cm}$\\Delta$ Log Gini Wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log Gini Corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log Gini Sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log Gini Banks",
+              "\\hspace{0.0cm}$ \\text{Log RWA}^r$",
+              "\\hspace{0.5cm}$ \\text{Log RWA}^r$ Retail",
+              "\\hspace{0.5cm}$ \\text{Log RWA}^r$ Wholesale",
+              "\\hspace{1.0cm}$ \\text{Log RWA}^r$ Corporate",
+              "\\hspace{1.0cm}$ \\text{Log RWA}^r$ Sovereign", 
+              "\\hspace{1.0cm}$ \\text{Log RWA}^r$ Banks"),
           notes.label = "",
@@
           notes.append = FALSE,
-          label = "",
-          style = "qje",
-          column.sep.width = "3pt",
-          #float.env = "sidewaystable",
-          align = TRUE,
-          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
-                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
-```
-
-```{r, eval=FALSE, echo=FALSE}
-
-reg.1 <- felm(out.tier.ratio ~ Shortfall.mn
-              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
-                  data = DiD.data)
-se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
-
-reg.2 <- felm(out.RWA ~ Shortfall.mn 
-                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
-                  data = DiD.data)
-se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
-
-reg.3 <- felm(out.RWA.s ~ Shortfall.mn 
-                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
-                  data = DiD.data)
-se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
-
-reg.4 <- felm(out.RWA.hat ~ Shortfall.mn 
-              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
-                  data = DiD.data)
-se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
-
-reg.5 <- felm(out.tier ~ Shortfall.mn
-              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
-                  data = DiD.data)
-se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
-
-```
-
-```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}
-
-stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "EBA capital exercise and Capital Savings",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
-The regressor is the capital shortfall during the capital exercise.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          dep.var.labels = c("$\\Delta$ CET1 ratio", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA Counterfactual", "$\\Delta$ Log CET1" ),
-          se = list(se.1, se.2, se.3, se.4, se.5),
-          covariate.labels = c("Log Capital Shortfall"),
+          summary = TRUE,
+          digits = 3,
+          float = FALSE)
+```
+
+\end{table}
+
+\begin{table}[!htbp] \centering  
+\caption{Summary Statistics}  
+\label{tab1:summary}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
+\end{tabular}
+
+```{r summary_portfolio, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+bank.data.win %>% select(
+  mean.PD_gr, 
+  mean.PD.Retail_gr,
+  mean.PD.Wholesale_gr, 
+  mean.PD.Corporate_gr, 
+  mean.PD.Sovereign_gr, 
+  mean.PD.Banks_gr,
+  var.PD_gr, 
+  skew.PD_gr,
+  kurt.PD_gr) %>%
+  as.data.frame() %>% 
+stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
+          label = "tab:summary",
+          omit.summary.stat = c("p25","p75"),
+          covariate.labels = 
+            c("$\\Delta$ Log PD Mean",
+              "\\hspace{0.5cm}$\\Delta$ Log PD Mean Retail",
+              "\\hspace{0.5cm}$\\Delta$ Log PD Mean Wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log PD Mean Corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log PD Mean Sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD Mean Banks", 
+              "$\\Delta$ Log PD Variance",
+              "$\\Delta$ Log PD Skewness", 
+              "$\\Delta$ Log PD Kurtosis"),
           notes.label = "",
@@
           notes.append = FALSE,
-          label = "",
-          style = "qje",
-          column.sep.width = "3pt",
-          #float.env = "sidewaystable",
-          align = TRUE)
-
-```
-
-```{r,eval=FALSE, echo=FALSE}
-
-reg.1 <- felm(out.RWA.s ~ Shortfall.mn 
-               |0|0|0,
-                  data = DiD.data)
-se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]
-
-reg.2 <- felm(out.RWA.s.Retail ~ Shortfall.mn 
-                |0|0|0,
-                  data = DiD.data)
-se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]
-
-reg.3 <- felm(out.RWA.s.Wholesale ~ Shortfall.mn 
-                |0|0|0,
-                  data = DiD.data)
-se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]
-
-reg.4 <- felm(out.RWA.s.Corporate ~ Shortfall.mn
-                |0|0|0,
-                  data = DiD.data)
-se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]
-
-reg.5 <- felm(out.RWA.s.Sovereign ~ Shortfall.mn
-                |0|0|0,
-                  data = DiD.data)
-se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]
-
-reg.6 <- felm(out.RWA.s.Banks ~ Shortfall.mn 
-               |0|0|0,
-                  data = DiD.data)
-se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
-```
-
-```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}
-
-stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6,
-          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
-          title = "EBA capital exercise and Capital Savings: Portfolio Breakdown",
-          table.layout = "n=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}^s_{i,j} = \\alpha + \\beta \\text{Log Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings.
-The regressor is the capital shortfall during the capital exercise.
-Standard errors are adjusted for heteroskedasticity.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          dep.var.labels = c("$j = \\text{Total}$", "$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"),
-          #se = list(se.1, se.2, se.3, se.4, se.5, se.6),
-          covariate.labels = c("Log Shortfall"),
+          digits = 3,
+          summary = TRUE,
+          float = FALSE)
+```
+
+\end{table}
+
+\begin{table}[!htbp] \centering  
+\caption{Summary Statistics}  
+\label{tab1:summary}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
+\end{tabular}
+
+```{r summary_control, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+bank.data.win %>% select(
+  log.gdp_gr,
+  log.gdp,
+  credit_gdp,
+  pillarI_stringency,
+  cap_reg,
+  Sup_Power,
+  log.asset,
+  NPL.ratio,
+  Zscore,
+  ROE,
+  ROA) %>%
+  as.data.frame() %>% 
+stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
+          label = "tab:summary",
+          omit.summary.stat = c("p25","p75"),
+          covariate.labels = 
+            c("$\\Delta \\text{Log GDP}$",
+              "Log GDP",
+              "Credit / GDP",
+              "Capital Stringency Index",
+              "Pillar I Stringency",
+              "Official Supervisory Power",
+              "Log Total Assets",
+              "Non-performing Loans / Loans",
+              "Z-score",
+              "ROE",
+              "ROA"),
           notes.label = "",
@@
           notes.append = FALSE,
-          label = "",
-          style = "qje",
-          column.sep.width = "3pt",
-         # float.env = "sidewaystable",
-          align = TRUE)
-```
+          summary = TRUE,
+          digits = 3,
+          float = FALSE)
+```
+
+\end{table}

### 2020-04-02 Analysis/Draft/Draft_bookdown.Rmd
Add new markdown file for draft (using bookdown)

--- 2020-04-02 Analysis/Draft/Draft.Rmd
@@
 author:
-- name: Lucas Avezum
-  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
-date: 
-header-includes:
-   - \usepackage{rotating, graphicx, dcolumn, longtable, float}
+- affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan
+    2, 5037 AB Tilburg, NL, The Netherlands
+  name: Lucas Avezum
+date: '\emph{\today}'
 output:
-  pdf_document:
-    #citation_package: natbib
+  bookdown::pdf_document2:
+    fig_caption: yes
     latex_engine: pdflatex
     template: tex-ms.tex
-    fig_caption: true
-  html_document:
-    df_print: paged
+  pdf_document: default
   word_document: default
+biblio-style: apsr
 bibliography: ../../Auxiliar/mybibfile.bib
+citecolor: black
 csl: ../../Auxiliar/cambridge-journal-of-economics.csl
-citecolor: black
-anonymous: false
+anonymous: no
 fontsize: 11pt
-spacing: "double"
 geometry: margin= 1in
+header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float}
 keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
-JEL: G21, G28, G11
 linkcolor: black
-biblio-style: apsr
+abstract: The Basel II accord introduced a concave mapping from probabilities of default
+  (PD) to capital requirements for banks adopting the internal ratings-based (IRB)
+  approach. Therefore, total capital requirements depends not only on the portfolio's
+  average PD but also on the shape of its PD distribution. Using a hand-collected
+  dataset, I propose a method that decomposes the contribution of the PD distribution
+  to capital requirements into an average component and a shape component. While a
+  larger average component increases capital charges, a larger shape component generates
+  capital savings. I document that these capital savings are important sources of
+  risk-weight variability across banks. I further investigate the relationship of
+  capital savings with the business cycle and find it to be countercyclical. The results
+  suggest that a recalibration of the current IRB framework can significantly reduce
+  the procyclicality of capital requirements.
+spacing: double
 subtitle: ''
 thanks: Preliminary draft.
-abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach.
-Therefore, total capital requirements depends not only on the portfolio's average PD but also on the shape of its PD distribution.
-Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
-While a larger average component increases capital charges, a larger shape component generates capital savings.
-I document that these capital savings are important sources of risk-weight variability across banks.
-I further investigate the relationship of capital savings with the business cycle and find it to be countercyclical.
-The results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements."
+JEL: G21, G28, G11
 urlcolor: black
+toc: FALSE
 ---
@@
 
-## Other Determinants of Capital Savings and Summary Statistics
-
-Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
-Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks ].
-In this section, I use these prior evidence to develop my expectations on how each of these variables may affect capital savings.
-
-Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
-
-The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
-First, I control for bank size using the log of total assets (in millions of US dollars).
-The expectation regarding the impact of bank size is ambiguous.
-On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
-On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
-I also include the regulatory capital ratio as control.
-Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
-However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
-Thus, the expected effect of capital ratio on capital savings is also ambiguous.
-Next, I add bank profitability using net income over total assets.
-Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
-
-The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
-For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
-Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
-As a result, I expect banks with higher shares of these variables to have more *Capital savings*.
-
-Finally, I study the relationship of *Capital savings* to risk-taking.
-The RW decomposition analysis suggests that riskier banks benefit more from *Capital savings*.
-I add the ratio of non-performing loans to total loans to test this proposition.
-I also include bank funding using the ratio of total deposits to total assets.
-Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
-The effect of deposits funding on *Capital savings* depends on how risk-taking is manifested by the deposit funding.
-If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then deposits to total assets are negatively correlated to *Capital savings*.
-Conversely, *Capital savings* increase if banks take risk by having fatter upper-tails in the PD distribution.
-
 # Capital Savings and the Business Cycle
@@
 
-Lastly, the method that I proposed to measure the shape component of the PD distribution could be flawed.
-In Table \ref{tab3} I use the well-known Gini coefficient to measure the shape component of the PD distribution.
+Lastly,  In Table \ref{tab3} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution.
 Again, the results shown in Table \ref{tab3} suppport the evidence from the last three tables: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
@@
 ## Capital Savings and Portfolio Realocation
+
+Ideally we want to fix the porfolio at some point in time and compare its and new loans PD distribution.
+Unfortunately, I cannot track individual assets of a portfolio.
+In this section I compare moments of the PD distribution between good and bad times without disantangling the direct effect of the business cycle on the distribution from the realocation effect.
 
@@
 ## Economic Impact of Capital Savings Cyclicality
-
-# Policy Discussion: Using Capital Savings to Reduce Capital Ratio Procyclicality
 
@@
 
+# Determinants of Capital Savings
+
+Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
+Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks ].
+In this section, I use these prior evidence to develop my expectations on how each of these variables may affect capital savings.
+
+Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
+
+The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
+First, I control for bank size using the log of total assets (in millions of US dollars).
+The expectation regarding the impact of bank size is ambiguous.
+On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
+On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
+I also include the regulatory capital ratio as control.
+Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
+However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
+Thus, the expected effect of capital ratio on capital savings is also ambiguous.
+Next, I add bank profitability using net income over total assets.
+Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
+
+The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
+For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
+Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
+As a result, I expect banks with higher shares of these variables to have more *Capital savings*.
+
+Finally, I study the relationship of *Capital savings* to risk-taking.
+The RW decomposition analysis suggests that riskier banks benefit more from *Capital savings*.
+I add the ratio of non-performing loans to total loans to test this proposition.
+I also include bank funding using the ratio of total deposits to total assets.
+Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
+The effect of deposits funding on *Capital savings* depends on how risk-taking is manifested by the deposit funding.
+If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then deposits to total assets are negatively correlated to *Capital savings*.
+Conversely, *Capital savings* increase if banks take risk by having fatter upper-tails in the PD distribution.
+
 # Conclusion and Discussion
@@
 
-<!-- <div id="refs"></div> -->
+<div id="refs"></div>
 
@@
 
+```{r}
+savings_cap = "\\label{fig:savingsbank}The figure plots the log of banks capital savings.
+This measure is calculated using equation \\ref{eq3}.
+Each point in the plot is a bank-year observation and bars are the averages values for each bank."
+```
+
+```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
+bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
+  select(name, Country, RWA.r)%>%
+  group_by(name, Country) %>%
+#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
+  group_by(Country) %>%
+ # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
+  #ungroup()%>%
+  #mutate(ordering = order_aux+RWA.r,
+  #       name = as.factor(name),
+  #       name = fct_reorder(name, ordering),
+  #       Country = fct_reorder(Country, ordering)) %>%
+ ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
+                      y = RWA.r)) +
+geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+    labs(x=NULL) +
+    coord_flip() +
+  labs(y =  "RWA savings ratio") +
+  scale_fill_grey()+
+    theme_bw()
+
+```
+
+```{r}
+savings_cap_country = "\\label{fig:savingscountry}The figure plots the banks capital savings ratio.
+This measure is calculated using equation \\ref{eq4}.
+For each bank I used the average across time such that each point in the plot is a bank observation and bars are the averages values for each country."
+```
+
+```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
+bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
+  select(Country,name, year, RWA.r)%>%
+  group_by(Country,name) %>%
+ summarise_if(is.numeric,mean,na.rm= TRUE)%>%
+ ggplot(mapping = aes(x = reorder(Country, RWA.r, FUN = mean, na.rm=TRUE),
+                          y = RWA.r)) +
+geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+    labs(x=NULL) +
+  labs(y =  "RWA savings ratio") +
+  coord_flip() +
+  theme_bw()
+
+```
+
+```{r}
+cross.dec_cap1 = "\\label{fig:crossdecomposition1} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+The figure and calculations include all banks in the sample.
+The first plot shows the total RW for each bank relative to the average total RW.
+Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
+Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
+```
+
+```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap1, include=TRUE}
+ plot1 <- cross.section.decomposition %>% 
+  mutate(total= -PD.gain)%>%
+  select(name, Country,total, int.gain, ext.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
+                                       labels = c("Intensive margin", "Extensive margin")))%>%
+ggplot(mapping = aes(x = reorder(name, total),
+                       y = value)) +
+geom_bar(stat = 'identity')+
+  #geom_line(aes(y = -total,group=1))  +
+  labs(x=NULL, y= NULL) +
+    coord_flip() +
+  theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  facet_wrap(~condition)
+
+plot2 <- cross.section.decomposition %>% 
+  select(name, Country, int.gain, ext.gain) %>%
+  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
+  ggplot(mapping = aes(x = int.gain,
+                       y = ext.gain)) +
+    geom_point()+
+    geom_smooth(method = "lm", color = "black")  +
+  labs(x=NULL, y= NULL) +
+   theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  scale_fill_grey()+
+  stat_cor()
+
+  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
+            ncol = 1, nrow = 2, align = "v", axis = "lr")
+```
+
+```{r}
+cross.dec_cap2 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
+The figure and calculations include all banks in the sample.
+The first plot shows the total RW for each bank relative to the average total RW.
+Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
+Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
+```
+
+```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
+plot1 <- cross.section.decomposition %>% 
+  mutate(total= -PD.gain)%>%
+  select(name, Country,total, other.gain, PD.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
+                                       labels = c("Capital savings", "Other components")))%>%
+ggplot(mapping = aes(x = reorder(name, total),
+                       y = value)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL, y= NULL) +
+    coord_flip() +theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  facet_wrap(~condition)
+
+plot2 <- cross.section.decomposition %>%
+  select(name, Country, other.gain, PD.gain) %>%
+  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
+  ggplot(mapping = aes(x = PD.gain,
+                       y = other.gain)) +
+    geom_point()+
+    geom_smooth(method = "lm", color = "black")  +
+  labs(x=NULL, y= NULL) +
+   theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  scale_fill_grey()+
+  stat_cor(label.y= -0.1)
+
+  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
+            ncol = 1, nrow = 2, align = "v", axis = "lr")
+```
+
+```{r}
+rsqrt_cap = "\\label{fig:rsqrtdecomp} The figure shows the $R^2$ share for each of the RW components derived using equation Equation \\ref{eq5}." 
+```
+
+```{r,rsqrt, fig.height = 4, fig.width = 5,out.width = "75%", results="hide", fig.cap= rsqrt_cap, include=TRUE,fig.align = 'center'}
+
+unrestrict <- cross.section.decomposition %>% 
+  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
+         total.gain = total.gain+epsilon)%>% 
+  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
+  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )
+
+  data.frame(share  = unrestrict@lmg) %>%
+   add_rownames(var = "component")%>%
+  mutate(component = factor(component,
+                      levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
+                      labels = c("PD avg.", "Ext. comp.", "Int. comp.", "SA", "Mix", "Rollout"))) %>%
+ggplot(mapping = aes(x = reorder(component, share),
+                       y = share, fill = component)) +
+geom_bar(stat = 'identity')+
+  labs(x=NULL, y = expression(paste("Share of ", R^{2}))) +
+    theme_bw()+
+  theme(panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+       # strip.text = element_blank()
+       legend.position = "none",
+        axis.text.x = element_text(hjust = 1))+
+  scale_fill_manual(values = c("#C3BCB2","#353535",  "#353535", "#C3BCB2","#C3BCB2", "#C3BCB2"))+
+scale_y_continuous(labels = scales::percent_format(accuracy = 1,
+                                 decimal.mark = '.')) +
+    coord_flip()
+
+```
+
+```{r}
+time.dec_cap = "\\label{fig:timedecomposition}The figure shows the evolution of average RW components in time.
+Plot a shows the average contribution of each RW component relative to the time of IRB introduction (t=0).
+The figure and calculations include only banks with observations for at least 9 consecutive years.
+Plot b shows the average contribution of each component relative to 2008.
+The figure and calculations include all banks in the sample.
+The black line in both plots is the average change in RW relative to the initial period."
+```
+
+```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
+
+plot1 <-time.decomposition %>% 
+  select(time, RW, t0.RW,  int.gain, ext.gain, other.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain")),
+         RW = RW - t0.RW,
+         time = as.factor(time)) %>%
+  ggplot(mapping = aes(x = time,
+                       y    = value, 
+                       fill = condition)) +
+geom_bar(stat = 'identity')+
+geom_line(aes(group=1, y = RW))  +
+  theme_bw()+
+  labs( y = 'Change from t=0') +
+  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
+ theme(legend.position = c(0.3, 0.2), 
+        legend.title = element_blank(), 
+        legend.background=element_blank())+
+  scale_fill_grey(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")
+
+year.decomposition <- year.decomposition %>%
+  mutate(RW_2 = RW - t0.RW,
+         year = as.factor(year)) 
+plot2 <- year.decomposition %>% 
+  select(year,RW, int.gain, ext.gain, other.gain) %>%
+  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
+  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain"))) %>%
+  ggplot(mapping = aes(x = year)) +
+geom_bar(aes(y = value, 
+          fill = condition),
+          stat = 'identity')+
+geom_line(data = year.decomposition, 
+          aes(group=1,y = RW_2))  +
+  theme_bw()+
+  labs(y = 'Change from 2008', x=NULL) +
+  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
+theme(legend.position='none')+
+  scale_fill_grey()
+grid.arrange(plot1, plot2, ncol=2)
+
+```
+
+```{r}
+time.var_cap = "\\label{fig:timevariables}The figure shows the time series of banks' capital ratio, portfolio average PD of their portfolio, and PD standard deviation of their portfolio.
+The black line is average value across all banks in the sample and the shaded area shows the first and third quartile."
+```
+
+```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= time.var_cap}
+
+mean_cl_quantile <- function(x, q=c(0.25, 0.75), na.rm=TRUE){
+  dat <- data.frame(y=mean(x, na.rm = TRUE),
+                    ymin = quantile(x, probs = q[1],na.rm = na.rm),
+                    ymax = quantile(x, probs = q[2],na.rm = na.rm))
+  return(dat)
+}
+scaleFUN <- function(x) sprintf("%.1f", x)
+
+plot1 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year, capital.ratio) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=capital.ratio)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black") +
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'Capital Ratio (%)', x=NULL) + scale_y_continuous(labels=scaleFUN)
+
+plot2 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year,  mean.PD) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=mean.PD)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'Average PD (%)', x=NULL)
+
+plot3 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year, var.PD) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'SD PD (%)', x=NULL)
+
+plot4 <-bank.data %>% filter(year>2007)%>%
+  select(bvdid, year, pillarI_stringency) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year,y=pillarI_stringency)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(stat = 'summary',
+            fun.data = mean_cl_quantile,
+            color = "black")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'GDP growth (%)', x=NULL)
+
+plot_grid(plot1, plot2, plot3, plot4, hjust = 0.5,
+            ncol = 2, nrow = 2, align = "v", axis = "lr")
+```
+```{r test, fig.height = 5, fig.width = 7, results="hide",}
+bank.data %>% filter(year>2007, bvdid %in% pillar3.data$bvdid)%>%
+  select(bvdid, year, log.gdp_gr, Country, pillarI_stringency,cap_reg) %>%
+  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
+  ggplot(mapping = aes(x = year)) +
+  geom_line(mapping = aes(y=pillarI_stringency))+
+  geom_line(mapping = aes(y=cap_reg))+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
+  labs(y = 'GDP growth (%)', x=NULL) +
+  facet_wrap(~Country)
+```
+
+<!-- \begin{table}[!htbp] \centering  -->
+<!--   \caption{Cyclicality of Capital Savings}  -->
+<!--   \label{tab1}  -->
+<!-- \small  -->
+<!-- \begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}}  -->
+<!-- \hline  -->
+<!-- \hline \\[-1.8ex]  -->
+<!-- \\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6} -->
+<!-- \\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\  -->
+<!-- \hline \\[-1.8ex]  -->
+<!--  $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\  -->
+<!--   $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\  -->
+<!--   $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\  -->
+<!--   \hline \\[-1.8ex] -->
+<!--  Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\  -->
+<!--   S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\  -->
+<!-- \hline \\[-1.8ex]  -->
+<!-- \end{tabular}  -->
+<!-- \end{table}  -->
+
 ```{r parameters curve 2, echo=FALSE, include=FALSE}
@@
 
-```{r}
-savings_cap = "\\label{fig:savingsbank}The figure plots the log of banks capital savings.
-This measure is calculated using equation \\ref{eq3}.
-Each point in the plot is a bank-year observation and bars are the averages values for each bank."
-```
-
-```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
-bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
-  select(name, Country, RWA.r)%>%
-  group_by(name, Country) %>%
-#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
-  group_by(Country) %>%
- # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
-  #ungroup()%>%
-  #mutate(ordering = order_aux+RWA.r,
-  #       name = as.factor(name),
-  #       name = fct_reorder(name, ordering),
-  #       Country = fct_reorder(Country, ordering)) %>%
- ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
-                      y = RWA.r)) +
-geom_point(alpha = 0.5,size = 2) +
-  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
-    labs(x=NULL) +
-    coord_flip() +
-  labs(y =  "RWA savings ratio") +
-  scale_fill_grey()+
-    theme_bw()
-
-```
-
-```{r}
-savings_cap_country = "\\label{fig:savingscountry}The figure plots the banks capital savings ratio.
-This measure is calculated using equation \\ref{eq4}.
-For each bank I used the average across time such that each point in the plot is a bank observation and bars are the averages values for each country."
-```
-
-```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
-bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
-  select(Country,name, year, RWA.r)%>%
-  group_by(Country,name) %>%
- summarise_if(is.numeric,mean,na.rm= TRUE)%>%
- ggplot(mapping = aes(x = reorder(Country, RWA.r, FUN = mean, na.rm=TRUE),
-                          y = RWA.r)) +
-geom_point(alpha = 0.5,size = 2) +
-  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
-    labs(x=NULL) +
-  labs(y =  "RWA savings ratio") +
-  coord_flip() +
-  theme_bw()
-
-```
-
-```{r}
-cross.dec_cap1 = "\\label{fig:crossdecomposition1} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
-The figure and calculations include all banks in the sample.
-The first plot shows the total RW for each bank relative to the average total RW.
-Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
-Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
-```
-
-```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap1, include=TRUE}
- plot1 <- cross.section.decomposition %>% 
-  mutate(total= -PD.gain)%>%
-  select(name, Country,total, int.gain, ext.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
-                                       labels = c("Intensive margin", "Extensive margin")))%>%
-ggplot(mapping = aes(x = reorder(name, total),
-                       y = value)) +
-geom_bar(stat = 'identity')+
-  #geom_line(aes(y = -total,group=1))  +
-  labs(x=NULL, y= NULL) +
-    coord_flip() +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  facet_wrap(~condition)
-
-plot2 <- cross.section.decomposition %>% 
-  select(name, Country, int.gain, ext.gain) %>%
-  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
-  ggplot(mapping = aes(x = int.gain,
-                       y = ext.gain)) +
-    geom_point()+
-    geom_smooth(method = "lm", color = "black")  +
-  labs(x=NULL, y= NULL) +
-   theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  scale_fill_grey()+
-  stat_cor()
-
-  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
-            ncol = 1, nrow = 2, align = "v", axis = "lr")
-```
-
-```{r}
-cross.dec_cap2 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
-The figure and calculations include all banks in the sample.
-The first plot shows the total RW for each bank relative to the average total RW.
-Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
-Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
-```
-
-```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
-plot1 <- cross.section.decomposition %>% 
-  mutate(total= -PD.gain)%>%
-  select(name, Country,total, other.gain, PD.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
-                                       labels = c("Capital savings", "Other components")))%>%
-ggplot(mapping = aes(x = reorder(name, total),
-                       y = value)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL, y= NULL) +
-    coord_flip() +theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  facet_wrap(~condition)
-
-plot2 <- cross.section.decomposition %>%
-  select(name, Country, other.gain, PD.gain) %>%
-  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
-  ggplot(mapping = aes(x = PD.gain,
-                       y = other.gain)) +
-    geom_point()+
-    geom_smooth(method = "lm", color = "black")  +
-  labs(x=NULL, y= NULL) +
-   theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  scale_fill_grey()+
-  stat_cor(label.y= -0.1)
-
-  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
-            ncol = 1, nrow = 2, align = "v", axis = "lr")
-```
-
-```{r}
-rsqrt_cap = "\\label{fig:rsqrtdecomp} The figure shows the $R^2$ share for each of the RW components derived using equation Equation \\ref{eq5}." 
-```
-
-```{r,rsqrt, fig.height = 4, fig.width = 5,out.width = "75%", results="hide", fig.cap= rsqrt_cap, include=TRUE,fig.align = 'center'}
-
-unrestrict <- cross.section.decomposition %>% 
-  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
-         total.gain = total.gain+epsilon)%>% 
-  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
-  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )
-
-  data.frame(share  = unrestrict@lmg) %>%
-   add_rownames(var = "component")%>%
-  mutate(component = factor(component,
-                      levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
-                      labels = c("PD avg.", "Ext. comp.", "Int. comp.", "SA", "Mix", "Rollout"))) %>%
-ggplot(mapping = aes(x = reorder(component, share),
-                       y = share, fill = component)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL, y = expression(paste("Share of ", R^{2}))) +
-    theme_bw()+
-  theme(panel.grid.minor = element_blank(),
-        strip.background = element_blank(),
-       # strip.text = element_blank()
-       legend.position = "none",
-        axis.text.x = element_text(hjust = 1))+
-  scale_fill_manual(values = c("#C3BCB2","#353535",  "#353535", "#C3BCB2","#C3BCB2", "#C3BCB2"))+
-scale_y_continuous(labels = scales::percent_format(accuracy = 1,
-                                 decimal.mark = '.')) +
-    coord_flip()
-
-```
-
-```{r}
-time.dec_cap = "\\label{fig:timedecomposition}The figure shows the evolution of average RW components in time.
-Plot a shows the average contribution of each RW component relative to the time of IRB introduction (t=0).
-The figure and calculations include only banks with observations for at least 9 consecutive years.
-Plot b shows the average contribution of each component relative to 2008.
-The figure and calculations include all banks in the sample.
-The black line in both plots is the average change in RW relative to the initial period."
-```
-
-```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
-
-plot1 <-time.decomposition %>% 
-  select(time, RW, t0.RW,  int.gain, ext.gain, other.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain")),
-         RW = RW - t0.RW,
-         time = as.factor(time)) %>%
-  ggplot(mapping = aes(x = time,
-                       y    = value, 
-                       fill = condition)) +
-geom_bar(stat = 'identity')+
-geom_line(aes(group=1, y = RW))  +
-  theme_bw()+
-  labs( y = 'Change from t=0') +
-  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
- theme(legend.position = c(0.3, 0.2), 
-        legend.title = element_blank(), 
-        legend.background=element_blank())+
-  scale_fill_grey(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")
-
-year.decomposition <- year.decomposition %>%
-  mutate(RW_2 = RW - t0.RW,
-         year = as.factor(year)) 
-plot2 <- year.decomposition %>% 
-  select(year,RW, int.gain, ext.gain, other.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain"))) %>%
-  ggplot(mapping = aes(x = year)) +
-geom_bar(aes(y = value, 
-          fill = condition),
-          stat = 'identity')+
-geom_line(data = year.decomposition, 
-          aes(group=1,y = RW_2))  +
-  theme_bw()+
-  labs(y = 'Change from 2008', x=NULL) +
-  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
-theme(legend.position='none')+
-  scale_fill_grey()
-grid.arrange(plot1, plot2, ncol=2)
-
-```
-
-```{r}
-time.var_cap = "\\label{fig:timevariables}The figure shows the time series of banks' capital ratio, portfolio average PD of their portfolio, and PD standard deviation of their portfolio.
-The black line is average value across all banks in the sample and the shaded area shows the first and third quartile."
-```
-
-```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= time.var_cap}
-
-mean_cl_quantile <- function(x, q=c(0.25, 0.75), na.rm=TRUE){
-  dat <- data.frame(y=mean(x, na.rm = TRUE),
-                    ymin = quantile(x, probs = q[1],na.rm = na.rm),
-                    ymax = quantile(x, probs = q[2],na.rm = na.rm))
-  return(dat)
-}
-scaleFUN <- function(x) sprintf("%.1f", x)
-
-plot1 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, capital.ratio) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=capital.ratio)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black") +
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'Capital Ratio (%)', x=NULL) + scale_y_continuous(labels=scaleFUN)
-
-plot2 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year,  mean.PD) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=mean.PD)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'Average PD (%)', x=NULL)
-
-plot3 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, var.PD) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'SD PD (%)', x=NULL)
-
-plot4 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, pillarI_stringency) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=pillarI_stringency)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'GDP growth (%)', x=NULL)
-
-plot_grid(plot1, plot2, plot3, plot4, hjust = 0.5,
-            ncol = 2, nrow = 2, align = "v", axis = "lr")
-```
-```{r test, fig.height = 5, fig.width = 7, results="hide",}
-bank.data %>% filter(year>2007, bvdid %in% pillar3.data$bvdid)%>%
-  select(bvdid, year, log.gdp_gr, Country, pillarI_stringency,cap_reg) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year)) +
-  geom_line(mapping = aes(y=pillarI_stringency))+
-  geom_line(mapping = aes(y=cap_reg))+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'GDP growth (%)', x=NULL) +
-  facet_wrap(~Country)
-```
-
-<!-- \begin{table}[!htbp] \centering  -->
-<!--   \caption{Cyclicality of Capital Savings}  -->
-<!--   \label{tab1}  -->
-<!-- \small  -->
-<!-- \begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}}  -->
-<!-- \hline  -->
-<!-- \hline \\[-1.8ex]  -->
-<!-- \\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6} -->
-<!-- \\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\  -->
-<!-- \hline \\[-1.8ex]  -->
-<!--  $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\  -->
-<!--   $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\  -->
-<!--   $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\  -->
-<!--   \hline \\[-1.8ex] -->
-<!--  Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\  -->
-<!--   S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\  -->
-<!-- \hline \\[-1.8ex]  -->
-<!-- \end{tabular}  -->
-<!-- \end{table}  -->
 ```{r winsorized_data, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
@@
 # Appendix {-}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
 

### 2020-04-02 Analysis/Simulation/DiD-analysis.Rmd
Add backup file for the DiD analysis using the EBA capital exercise

---
title: "Credit Risk Distribution and Capital Savings"
author:
- name: Lucas Avezum
  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
date: 
header-includes:
   - \usepackage{rotating, graphicx, dcolumn, longtable}
output:
  pdf_document:
    #citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
    fig_caption: true
  html_document:
    df_print: paged
  word_document: default
bibliography: ../../Auxiliar/mybibfile.bib
csl: ../../Auxiliar/journal-of-finance.csl
citecolor: black
anonymous: false
fontsize: 11pt
spacing: "double"
geometry: margin= 1in
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
JEL: G21, G28, G11
linkcolor: black
biblio-style: apsr
subtitle: ''
thanks: Preliminary draft.  
abstract: "The Basel II accord introduced a concave mapping from probabilities of default (PD) to capital requirements for banks adopting the internal ratings-based (IRB) approach. Therefore, total capital requirements depends not only on the portfolio's average PD but also on the shape of its PD distribution. Using a hand-collected dataset, I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component. While a larger average component increases capital charges, a larger shape component generates capital savings. I document that these capital savings are important sources of risk-weight variability across banks. I further investigate the relationship of capital savings with the business cycle and find it to be countercyclical. The results suggest that a recalibration of the current IRB framework can significantly reduce the procyclicality of capital requirements."
urlcolor: black
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

rm(list = ls())

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
#library(textablr)
#library(sandwich)
#library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

source("../../Auxiliar/Function.R") ## Capital requirements functions

```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

```{r did data, eval=FALSE, include=FALSE}
p_w <- c(0.0, 1)         ## Set winsorizing limits
pre.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2009","2010"),
         !is.na(IRB)) %>%
  select(bvdid, log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

post.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2012","2013"),
         !is.na(IRB)) %>%
  select(bvdid,  log.RWA, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.RWA.s.Wholesale, log.RWA.s.Retail, log.RWA.s.Equity, log.RWA.s.Corporate, log.RWA.s.Sovereign,    
         log.RWA.s.Banks, log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

DiD.data <- bank.data %>%
  filter(year %in% c("2010")) %>%
  right_join(pre.treat.out,  by = c("bvdid"), suffix = c("",".pre")) %>%
  right_join(post.treat.out, by = c("bvdid"), suffix = c("",".post")) %>%
  mutate(Shortfall.mn   = log(1+Shortfall.mn), 
         Shortfall.mn.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.mn, 0), NA),
         Shortfall.RW.2 = ifelse(EBAcountry == 1, ifelse(EBAbank == 1,Shortfall.RW, 0), NA),
         out.RWA        = log.RWA.post - log.RWA.pre,
         out.RWA.IRB    = log.RWA.IRB.post - log.RWA.IRB.pre,
         out.RWA.SA     = log.RWA.SA.post - log.RWA.SA.pre,
         out.RWA.hat    = log.RWA.hat.post - log.RWA.hat.pre,
         out.RWA.s      = log.RWA.s.post - log.RWA.s.pre,
         out.RWA.s.Retail = log.RWA.s.Retail.post - log.RWA.s.Retail.pre,
         out.RWA.s.Equity = log.RWA.s.Equity.post - log.RWA.s.Equity.pre,
         out.RWA.s.Wholesale = log.RWA.s.Wholesale.post - log.RWA.s.Wholesale.pre,
         out.RWA.s.Corporate = log.RWA.s.Corporate.post - log.RWA.s.Corporate.pre,
         out.RWA.s.Sovereign = log.RWA.s.Sovereign.post - log.RWA.s.Sovereign.pre,
         out.RWA.s.Banks = log.RWA.s.Banks.post - log.RWA.s.Banks.pre,
         out.K          = log.K.post - log.K.pre,
         out.tier       = log.tier.post - log.tier.pre,
         out.CAR.ratio  = totalcapitalratio.post - totalcapitalratio.pre,
         out.tier.ratio = tier1ratio.post - tier1ratio.pre,
         out.CAR.dif    = CAR.dif.post - CAR.dif.pre,
         out.tier.dif   = tier.dif.post - tier.dif.pre,
         Country        = as.factor(Country)) %>% 
  ungroup() %>% 
  filter(EBAcountry == 1,
         name != "Bank of Cyprus") %>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

```

```{r did, eval=FALSE, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth", include=FALSE}
p_w <- c(0.0, 1)         ## Set winsorizing limits
DiD.data %>% ungroup()%>%
  subset(select = c("Shortfall.mn","out.RWA.s","out.RWA.s.Retail","out.RWA.s.Wholesale","out.RWA.s.Corporate","out.RWA.s.Sovereign","out.RWA.s.Banks","out.RWA", "out.K","out.CAR.ratio","out.CAR.dif")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("Log Shortfall","$\\Delta \\text{Log RWA}^S$", "$\\Delta \\text{Log RWA}^S_R$", "$\\Delta \\text{Log RWA}^S_W$", "$\\Delta \\text{Log RWA}^S_C$", "$\\Delta \\text{Log RWA}^S_S$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}^S_B$", "$\\Delta \\text{Log RWA}$", "$\\Delta \\text{Log Capital}$", "$\\Delta \\text{CAR}^S$"  ),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

```{r, eval = FALSE, fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth", include=FALSE}
 plot1 <- bank.data %>%  ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier, tier.hat, IRB) %>%
  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>% 
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
theme(legend.position='none')

plot2 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier, tier.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier:tier.hat, factor_key=TRUE)  %>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="",subtitle = "Plot b") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  theme(legend.position = c(0.8, 0.85), 
        legend.title = element_blank(), 
        legend.background=element_blank())

plot3 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, CAR, CAR.hat, IRB) %>%
  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

plot4 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, CAR, CAR.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, CAR:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="",  subtitle = "Plot d") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
```

```{r,eval=FALSE, echo=FALSE}

bank.data.2 <- bank.data %>% ungroup()%>%
 # filter(!is.na(RW.s),
  #       !is.na(basel),
   #      tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA, log.RWA.s, log.RWA.hat, CAR.dif, tier.dif, 
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  RW, log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
#  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(totalcapitalratio  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(log.K  ~ CAR.dif  | year + IRB.bvdid  | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(log.RWA  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(log.asset  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(RW  ~ CAR.dif  | year + IRB.bvdid   | 0 | 0 , data = bank.data.2)
se.reg.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

```

```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Savings and Capital Ratio Components",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times \\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4, se.reg.5),
          covariate.labels = c("Capital ratio savings", "IRB share","CT1 savings $\\times$ IRB share"),
          dep.var.labels = c("Capital ratio","Log Capital","Log RWA","Log Assets","RW"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          #float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r,eval=FALSE, echo=FALSE}

bank.data.2 <- bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  select(bvdid, IRB.bvdid, year, IRB, Country, log.RWA.s, log.RWA.hat, CAR.dif,
         totalcapitalratio, tier1ratio, ROE, ROA, Zscore, leverage,  log.tier, log.K,
         income.ratio, deposit.ratio, loans.ratio, log.asset, NII.ratio)%>%
  group_by(bvdid) %>%
#  mutate_at(vars(log.RWA.s:NII.ratio), funs(. - lag(., order_by=year))) %>%
 ungroup() %>%
 mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

reg.1 <- felm(ROE  ~ CAR.dif | year +  IRB.bvdid  | 0 | 0 , data = bank.data.2)
se.reg.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(ROA  ~ CAR.dif  | year +  IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(Zscore  ~ CAR.dif  | year +IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(leverage  ~ CAR.dif  | year  + IRB.bvdid    | 0 | 0 , data = bank.data.2)
se.reg.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

```

```{r,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, 
           header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "Capital Savings, Performance, and Risk-Taking",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{15cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Y}_{i,t} = \\beta \\times\\text{Capital ratio savings}_{i,t}+\\alpha_{i,IRB}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The first line of the table show the dependent variable in each regression. $\\text{Capital ratio savings}_{it}$ is the difference between the counterfactual (without savings) and actual total capital ratio. $\\text{IRB share}_{it}$ is the share of bank i portfolio under the internal-ratings based approach.  Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          se = list(se.reg.1, se.reg.2, se.reg.3, se.reg.4),
          covariate.labels = c("Capital ratio savings"),
          dep.var.labels = c("ROE","ROA","Z-score","Leverage"), 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          #float.env = "sidewaystable",
          align = TRUE,
          add.lines = list(c("Bank $\\times$ IRB FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}"),
                           c("Year FE", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}", "\\multicolumn{1}{c}{Yes}")))
```

```{r, eval=FALSE, echo=FALSE}

reg.1 <- felm(out.tier.ratio ~ Shortfall.mn
              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(out.RWA ~ Shortfall.mn 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(out.RWA.s ~ Shortfall.mn 
                 + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(out.RWA.hat ~ Shortfall.mn 
              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(out.tier ~ Shortfall.mn
              + log.asset + deposit.ratio + loans.ratio + NII.ratio + income.ratio |0|0|0,
                  data = DiD.data)
se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

```

```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "EBA capital exercise and Capital Savings",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta Y_i = \\alpha + \\beta \\text{Capital Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings. The regressor is the capital shortfall during the capital exercise. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}", 
          dep.var.labels = c("$\\Delta$ CET1 ratio", "$\\Delta$ Log RWA", "$\\Delta$ Log RWA savings", "$\\Delta$ Log RWA Counterfactual", "$\\Delta$ Log CET1" ),
          se = list(se.1, se.2, se.3, se.4, se.5),
          covariate.labels = c("Log Capital Shortfall"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
          #float.env = "sidewaystable",
          align = TRUE)

```

```{r,eval=FALSE, echo=FALSE}

reg.1 <- felm(out.RWA.s ~ Shortfall.mn 
               |0|0|0,
                  data = DiD.data)
se.1 <- summary(reg.1,robust=TRUE)$coefficients[,2]

reg.2 <- felm(out.RWA.s.Retail ~ Shortfall.mn 
                |0|0|0,
                  data = DiD.data)
se.2 <- summary(reg.2,robust=TRUE)$coefficients[,2]

reg.3 <- felm(out.RWA.s.Wholesale ~ Shortfall.mn 
                |0|0|0,
                  data = DiD.data)
se.3 <- summary(reg.3,robust=TRUE)$coefficients[,2]

reg.4 <- felm(out.RWA.s.Corporate ~ Shortfall.mn
                |0|0|0,
                  data = DiD.data)
se.4 <- summary(reg.4,robust=TRUE)$coefficients[,2]

reg.5 <- felm(out.RWA.s.Sovereign ~ Shortfall.mn
                |0|0|0,
                  data = DiD.data)
se.5 <- summary(reg.5,robust=TRUE)$coefficients[,2]

reg.6 <- felm(out.RWA.s.Banks ~ Shortfall.mn 
               |0|0|0,
                  data = DiD.data)
se.6 <- summary(reg.6,robust=TRUE)$coefficients[,2]
```

```{r,eval=FALSE, fig.height = 10, fig.width = 7, results="asis", out.width = "1\\textwidth", cache = FALSE, warning=FALSE, message=FALSE, include=FALSE}

stargazer(reg.1, reg.2, reg.3, reg.4, reg.5, reg.6,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small",
          title = "EBA capital exercise and Capital Savings: Portfolio Breakdown",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{17cm}{The table shows the estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}^s_{i,j} = \\alpha + \\beta \\text{Log Shortfall}_i +\\varepsilon_i \\end{equation*} \\newline The dependent variable in each regression is the measure of capital savings. The regressor is the capital shortfall during the capital exercise. Standard errors are adjusted for heteroskedasticity. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}", 
          dep.var.labels = c("$j = \\text{Total}$", "$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$"),
          #se = list(se.1, se.2, se.3, se.4, se.5, se.6),
          covariate.labels = c("Log Shortfall"),
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE,
          label = "",
          style = "qje",
          column.sep.width = "3pt",
         # float.env = "sidewaystable",
          align = TRUE)
```

### 2020-04-02 Data/Clean.R
Simplify selection of variables

@@ -324,7 +324,7 @@ bank.data <- bank.data %>% ungroup() %>% rowwise() %>%
    filter(not.collect == 0) %>%
    # Remove auxiliar variables
    select(bvdid, name, Country, year, RWA.r, RW.s, RWA.s,  
-          contains(c("log.", ".ratio", "_gr", ".PD", "_", "q.")),
+          matches("log.|.ratio|_gr|.PD|_|q."),
           ROE, ROA, subordinated, Zscore, leverage) 
 
 ##============================================================================##
@@ -427,8 +427,7 @@ bank.year.decomposition <- bank.year.decomposition %>%
          other.gain = SA.gain + IRB.gain + mix.gain + rollout.gain,
          PD.gain    = int.gain + ext.gain,
          total.gain = RW-mean.RW)%>%
-  select(bvdid, name, Country, year,
-         IRB.gain, mix.gain, rollout.gain, int.gain, ext.gain, SA.gain, other.gain, PD.gain, total.gain)%>%
+  select(bvdid, name, Country, year, contains(".gain"))%>%
   ungroup() %>% group_by(bvdid) %>%
   mutate_at(vars(matches(".gain")), .funs = list(gr= ~ . - lag(., order_by = year))) 
 
@@ -546,9 +545,7 @@ cross.section.decomposition <- cross.section.decomposition %>%
          other.gain = SA.gain + IRB.gain + mix.gain + rollout.gain,
          PD.gain    = int.gain + ext.gain,
          total.gain = RW-mean.RW)%>%
-  select(bvdid, name, Country,
-         IRB.gain, mix.gain, rollout.gain, int.gain, ext.gain, SA.gain, other.gain, PD.gain, total.gain,
-         IRB.gain.alt, mix.gain.alt, rollout.gain.alt)%>%
+  select(bvdid, name, Country, contains(".gain"))%>%
   ungroup()
 
 # Save clean data frame
@@ -641,8 +638,7 @@ for(i in c("Wholesale", "Retail", "Equity")) {
          ext.gain     = (q.SA-t0.q.SA)*(RW.s),
          PD.gain      = int.gain + ext.gain,
          total.gain = RW-t0.RW)%>% ungroup() %>%
-    select(j, RW, t0.RW, IRB.gain, mix.gain, rollout.gain, int.gain, 
-           ext.gain, SA.gain, other.gain, PD.gain, total.gain)%>%
[…]

### 2020-04-02 Analysis/Draft/Draft_bookdown.Rmd
Add code that allow control of column width in stargazer command

@@
 source("../../Auxiliar/longtable.R") ## longtable for stargazer
+source("../../Auxiliar/adj_stargazer.R") ## adjust the first column width for stargazer
 
@@
 ```{r summary_control, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-bank.data.win %>% select(
+table <- capture.output({bank.data.win %>% ungroup()%>% select(
   log.gdp_gr,
@@
           digits = 3,
-          float = FALSE)
+          float = FALSE)})
+
+table <- gsub("lccccc",
+              "p{0.4\\textwidth}ccccc", table, fixed = T)
+
+cat(table)
 ```

### 2020-04-14 Data/Clean.R
Remove some redundancies in Clean.R + add exchange rate conversion of BankScope dataset

@@ -157,82 +157,70 @@ bank.data <- SA.data %>%
             EAD.SA = sum(EAD, na.rm = TRUE))%>%
   ungroup()%>%
   full_join(sample, by=c("bvdid", "name", "Country","year"))%>%
-  mutate(RW.SA = 100*RWA.SA/EAD.SA)
- 
-  
-## Aggregated IRB exposures at bank-portfolio-year level 
-for(i in c("Wholesale","Retail","Equity")) {
-bank.data <- IRB.data %>% 
-  group_by(name,bvdid,Country,year,Portfolio_1) %>%
-  filter(Portfolio_1 %in% c(i)) %>%
-  summarise(!!(paste("gini", "PD" ,        i, sep=".")) := Gini(PD,w.EAD, na.rm = TRUE),
-            !!(paste("sd"  , "PD" ,        i, sep=".")) := sqrt(wtd.var(PD,w.EAD, na.rm = TRUE))*100,
-            !!(paste("mean", "PD" ,        i, sep=".")) := wtd.mean(PD,w.EAD, na.rm = TRUE)*100,
-            !!(paste(        "RWA",        i, sep=".")) := sum(RWA, na.rm = TRUE),
-            !!(paste(        "RWA", "hat", i, sep=".")) := sum(RWA.hat, na.rm = TRUE),
-            !!(paste(        "RWA", "til", i, sep=".")) := sum(RWA.til, na.rm = TRUE),
-            !!(paste(        "RWA", "bar", i, sep=".")) := sum(RWA.bar, na.rm = TRUE),
-            !!(paste(        "EAD",        i, sep=".")) := sum(EAD, na.rm = TRUE))%>%
-  ungroup()%>%
-  full_join(sample, by=c("bvdid","name","Country","year"))%>%
-  mutate(!!(paste("RW" , "hat", i, sep=".")) := 100*get(paste("RWA","hat", i , sep="."))/get(paste("EAD", i , sep=".")),
-         !!(paste("RWA", "s"  , i, sep=".")) := get(paste("RWA","hat", i, sep="."))-get(paste("RWA", i , sep=".")),
-         !!(paste("RWA", "r"  , i, sep=".")) := get(paste("RWA","hat", i, sep="."))/get(paste("RWA", i , sep="."))-1,
-         !!(paste("RWA", "c"  , i, sep=".")) := get(paste("RWA","bar", i, sep="."))/get(paste("RWA","til", i , sep="."))-1,
[…]

### 2020-04-14 Data/Import.R
Add exchange rate conversiont of Pillar 3 dataset

@@ -13,6 +13,8 @@ library(tidyverse)    ## Data manipulation, pipe operator
 library(readxl)       ## Command to open xlsx files                                                     
 library(data.table)   ## Command to bind lists
 library(stringr)      ## Command to clean variables names in Orbis
+library(zoo)          ## Command to replace missing values by other group's mean value 
+
 rm(list = ls()) 
 
 ##============================================================================##
@@ -39,11 +41,13 @@ EBA.indicator <- read_xlsx("Data/Raw/Pillar3/Auxiliar/Auxiliar.xlsx",sheet = 3) 
 country.code <- read.csv("Data/Raw/WDI/WDICountry.csv") %>%
   rename_all(tolower)%>%
   select(country.code = ï..country.code,
-         Country = x2.alpha.code)
+         Country = x2.alpha.code,
+         currency.unit)%>%
+  mutate(currency.unit = ifelse(country.code=="EMU", "Euro", as.character(currency.unit)))
 
 WDI <- read.csv("Data/Raw/WDI/WDIData.csv") %>%
   filter(Indicator.Code %in% c("NY.GDP.DEFL.ZS","NY.GDP.PCAP.KD","NY.GDP.PCAP.KN", "PA.NUS.FCRF", "PX.REX.REER",
-                               "FS.AST.DOMS.GD.ZS","FS.AST.PRVT.GD.ZS","FD.AST.PRVT.GD.ZS")) %>%
+                               "FS.AST.DOMS.GD.ZS","FS.AST.PRVT.GD.ZS","FD.AST.PRVT.GD.ZS", "FP.CPI.TOTL")) %>%
   rename_all(funs(str_replace_all(.,"([\\w])([0-9]+)$", "\\1\\.\\2"))) %>%
   reshape(varying   = c(grep("X.", names(.))),
                       direction = 'long', 
@@ -55,6 +59,7 @@ WDI <- read.csv("Data/Raw/WDI/WDIData.csv") %>%
   rename_all(tolower) %>%
   inner_join(country.code)%>%
   select(deflator        = x.ny.gdp.defl.zs,
+         cpi             = x.fp.cpi.totl,
          gdppc.us        = x.ny.gdp.pcap.kd,
          gdppc           = x.ny.gdp.pcap.kn,
          er              = x.pa.nus.fcrf,
@@ -63,7 +68,17 @@ WDI <- read.csv("Data/Raw/WDI/WDIData.csv") %>%
          priv_credit_gdp = x.fs.ast.prvt.gd.zs,
[…]

### 2020-04-14 Analysis/Draft/Draft_bookdown.Rmd
Remove RW decomposition of main analysis and improve results section

@@
 ---
-title: "Credit Risk Distribution and Capital Savings"
-author:
-- affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan
-    2, 5037 AB Tilburg, NL, The Netherlands
-  name: Lucas Avezum
+title: "The Countercyclical Component of Capital Regulation"
 date: '\emph{\today}'
@@
   word_document: default
+author:
+- name: "Lucas Avezum"
+  affiliation: "CentER and EBC, Tilburg University"
 biblio-style: apsr
@@
 csl: ../../Auxiliar/cambridge-journal-of-economics.csl
-anonymous: no
+#anonymous: no
 fontsize: 11pt
 geometry: margin= 1in
-header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float}
+header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float, siunitx, tabularx}
 keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
@@
   larger average component increases capital charges, a larger shape component generates
-  capital savings. I document that these capital savings are important sources of
-  risk-weight variability across banks. I further investigate the relationship of
-  capital savings with the business cycle and find it to be countercyclical. The results
-  suggest that a recalibration of the current IRB framework can significantly reduce
-  the procyclicality of capital requirements.
+  capital savings.
+I investigate the relationship of these capital savings with the business cycle and find it to   be countercyclical.
+I further document that the shape are important sources of across banks.
+The results suggest   that a recalibration of the current IRB framework can have sizeble effects on the procyclicality of capital       regulation.
 spacing: double
@@
 library(ggpubr)               ## More plots
+library(splitstackshape)     ## Command to expand data
+
 #library(textablr)
@@
 
-# Institutional Background
+# Institutional Background and Capital Savings Measures
 
@@
 \begin{equation}
-\label{measure1}
+\label{eq:measure1}
 \text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}}.
@@
 \begin{equation}
-\label{measure2}
+\label{eq:measure2}
 \text{RWA}^{\$}=\text{RWA}^c-\text{RWA}.
@@
 
+Because this is a measure in US Dollars, it is affected by the size of the IRB portfolio.
+Therefore, differently from the main analysis, *Capital savings* are mechanically larger for larger banks and they increase as the portfolio rolls out to the IRB approach.
+
 An alternative way to measure *Capital savings* is in monetary units as follows:
@@
 \begin{equation}
-\label{measure3}
+\label{eq:measure3}
 \text{RW}^s=\frac{\text{RWA}^{\$}}{\text{RWA}}.
@@
 
-## RW Decomposition
-
-In this section I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
-I divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
-Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
-I further breakdown RW_{IRB} into its portfolio components, $\text{RW}_{IRB}=\sum_j q_{j}\text{RW}_{j}$, in which the $j$ subscript stands for the different portfolios under the IRB approach: Wholesale, retail and equity.
-Equation \ref{eq5} shows the final decomposition:
-
-\begin{equation}
-\begin{aligned}
-\label{eq5}
-\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j} \text{RW}_{j}^{c} \Delta q_{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
-&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin},
-\end{aligned}
-\end{equation}
-in which  $\Delta$ stands for changes relative to a benchmark.
-For the cross-section decomposition, the benchmark is the average bank.
-For the time decomposition, the benchmark is either the year when the IRB approach was first introduced at the bank level or 2008.
-The superscript $B$ stands for benchmark.
-For instance, $q_SA^B$ is the share of the portfolio under the standardized approach for the benchmark.
-
-I decompose the variation of RW relative to the benchmark in six components, two of them relating to *Capital savings*.
-The first, $(1-q^{B}_{SA})\Delta \text{RW}^{s}$, measures the intensity of *Capital savings*.
-That is, I fix the share of the portfolio under the IRB approach ($q_{SA}^B$) and variation in the component comes only from changes of $RW^s$ relative to the benchmark.
-The second component, $\text{RW}^{s}\Delta q^{B}_{SA}$, measures the extensive margin of *Capital savings*, i.e., variation in this component comes only from changes of the share under the IRB approach relative to the benchmark.
-Note that the magnitude of the second component is moderated by how intense *Capital savings* are in absolute terms.
-
-The remaining components are extensively discussed by @cannata2012inside.
-The first on the right-hand side of equation \ref{eq5} is the intensity of RW under the standardized approach ($q^{B}_{SA} \Delta \text{RW}_{SA}$).
-The contribution of this component to $\Delta \text{RW}$ comes from variaton of RW in portfolios under the standardized approach while fixing the share of this portfolio for all banks.
-The second term measures the intensity of RW under the IRB approach excluding the *Capital savings*, or differently, the intensity of the average PD ($(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j$).
-The contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of RW in the three portfolios under the IRB approach relative to the respective benchmark RW while holding constant the share of each IRB portfolio and the share of total IRB.
-The third component measure the contribution of the IRB portfolio mix.
-Because some types of portfolios have on average lower RW, the composition within IRB across types of portfolios matter to how large overall RW is.
-For instance retail portfolios tend to have lower RW relative to wholesale portfolios because the former usually are colletarized loans, such as mortgages, while the latter is mostly composed of corporate loans.
-Hence, the contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of the three portfolios shares within the IRB portfolio relative to the respective benchmarks while holding constant the respective intensity of average PD ($RW_j^c$) and the share of total IRB.
-Lastly, $(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}$ is the roll-out effect, i.e., the contribution to $\Delta \text{RW}$ coming from moving the portfolio from the standardized approach to the IRB approach.
-
-Figure \ref{fig:crossdecomposition1} plots the intensive and extensive margin of each bank in my sample using the average bank as benchmark.
-The scatter plot shows that the correlation between the two *Capital savings* components is not statistically significant.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
-If Goldman Sachs is included the correlation becomes statistically significant.} The figure suggests that banks benefit from the two components separately, some benefting more of the extensive margin, such as Northern Trust, and others more of the intensive margin, such as BayernLB.
-
-Figure \ref{fig:crossdecomposition2} shows that the net effect of the intensive and extensive margin is negatively correlated to the net effect of the other components.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
-If Goldman Sachs is included the correlation becomes even more negative.} This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
-
-As Figure \ref{fig:rsqrtdecomp} shows *Capital savings* components account for 27% of the RW variation across banks.
-Each component accounts for around 13.5% of the RW variation and are second and third in rank, losing only to the average PD contribution.
-I derive the $\text{R}^2$ contribution of each RW component using the Shapley-Owen decomposition method.
-This method consider the contribution of each variable to total $\text{R}^2$ as the sum of $\text{R}^2$ from regressions using all possible variable combinations that include the variable of interest and weighting each regression by the number of variable permutations.
-In sum, figure \ref{fig:rsqrtdecomp} offers strong evindence of the relative importance of *Capital savings* as source of variation to RW across banks.
-
-Figure \ref{fig:timedecomposition} shows the evolution of RW components relative to the time of IRB introduction (left-hand side plot), and relative to 2008 (right-hand side plot).
-The negative values for the extensive margin implies that banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach.
-However, during the same period, the intensity of *Capital savings* has, on average, decreased in time.
-That is, for the same average PD, capital requirements have increased from 2008 to 2018 and since the year IRB approach was introduced.
-
-The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}.
-Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.
-[^1] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
-This distribution compression leads to a reduction in *Capital savings*.
-This explanation is in line with the hypothesis that *Capital savings* are countercyclical.
-Efforts to increase capital ratio by risk reduction are increasingly more difficult.
-
-[^1]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].
-
 # Capital Savings and the Business Cycle
@@
 
-## A simple theoretical framework
+## A simple theoretical framework {#theory}
 
@@
 
-In Table \ref{tab1}, I regress the change in the logarithm of *Capital savings* on the change in the logatithm of real GDP per capita.
-In all regressions the measure of *Capital savings* follows equation \ref{eq4}.
-The difference between regressions in the table is the portfolio used to calculate the *Capital savings*.
-The real GDP per capita used  comes from the country where the bank has its headquarter.
-All regressions include bank and year fixed effects and standard errors clustered at the country level.
-
-Column 1 shows that total *Capital savings* are countercyclical as the growth rate of tota l*Capital savings* is negatively associated to the GDP growth rate.
+Throughout this section, I report results from estimating the following specification:
+
+\begin{equation*}
+\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*}
+
+The dependent variable is the yearly change in a variable associated to portfolio $j$ for bank $i$ at year $t$.
+The change in the logatithm of real GDP per capita used comes from the country where the bank has its headquarter.
+All regressions include bank and year fixed effects and robust standard errors adjusted for clustering at the country level are reported in parentheses.
+
+Column 1 Table \ref{tab:main} shows results using equation \ref{eq:measure1}, my prefered measure of *Capital savings*.
+shows that total *Capital savings* are countercyclical.
 According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average.
@@
 
-The remaining columns of Table \ref{tab1} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios.
+The remaining columns of Table \ref{tab:main} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios.
 The estimated coefficient in Column 2 is negative suggesting that *Capital savings* generated by retail portfolios are also countercyclical.
@@
 On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant.
-According to the regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation.
+According to regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation.
 Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios.
@@
 
-A concern to the previous analysis is the possibility of that I could be overestimating the average impact of GDP growth rate on *Capital savings*.
-First, Because the measure used in Table \ref{tab1} is dimentionless, i.e., it is a measure of intensity of *Capital savings*, if the portfolio of small banks, in terms of asset, are more impacted during the business cycle than the portfolio of large banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank.
-To account for differences in *Capital savings* size, in Table \ref{tab2} the measure of *Capital savings* follows equation \ref{eq3}, which is a measure of *Capital savings* in monetary units.
-
-The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.
-
-Second, because both measures in Tables \ref{tab1} and \ref{tab2} are relative to the IRB portfolio, if the portfolio of banks with lower share of their portfolio under the IRB approach are more impacted during the business cycle than the portfolio of high IRB share banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank.
-The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.
-
-Lastly,  In Table \ref{tab3} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution.
-Again, the results shown in Table \ref{tab3} suppport the evidence from the last three tables: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
-
-## Capital Savings and Portfolio Realocation
-
-Ideally we want to fix the porfolio at some point in time and compare its and new loans PD distribution.
-Unfortunately, I cannot track individual assets of a portfolio.
-In this section I compare moments of the PD distribution between good and bad times without disantangling the direct effect of the business cycle on the distribution from the realocation effect.
-
-The results from Tables \ref{tab1} to XX raises the questions of why it appears that only corporate portfolios are impacted during the business cycle. 
-I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases.
+In this section, I test the robustness of the main results to alternative measures of *Capital savings*.
+In Panel A of Table \ref{tab:alt}, the dependent variable measures *Capital savings* in US dollars following equation \ref{eq:measure2}.
+As before, the estimated coefficients suggest that the countercyclical effect is mostly driven by wholesale portfolios.
+Differently from Table \ref{tab:main}, the coefficient of GDP growth rate estimated on the bank portfolio is now negative and statistically significant.
+Because coefficients in Panel A of Table \ref{tab:alt} are  impacted by the size of the portfolio, while in \ref{tab:main} they are not, the results of regressions 6 in both tables suggest that part of the *Capital savings* countercyclicality comes from an increase in exposure to other banks during recessions.
+
+In panel B of Table \ref{tab:alt} the measure of *Capital savings* in US dollars is rescaled by the amount of EAD in the respective portfolio following equation \ref{eq:measure3}.
+This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity, risk-weights.
+Although the coefficient on the wholesale portfolio is no longer statiscally significant, total *Capital savings* and those coming from corporate portfolios are still statistically different from zero at the 1%  confidence level.
+The attenuation of the effect on wholesale portfolio appears to come from large, albeit not significant, procyclical *Capital savings* in sovereign portfolios.
+Moreover, the small and statistically insignificant coefficient in regression 6 reinforces the claim that countercyclical *Capital savings* in the bank portfolio are due to the extensive and not the intensive margin.
+
+Lastly, in Panel C of Table \ref{tab:alt} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution.
+Again, the results shown in Panel C suppport the evidence from Table \ref{tab:alt}: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
+
+## Capital Savings and Portfolio Reallocation
+
+The results from Tables \ref{tab:main} and \ref{tab:alt} raises the question of why it appears that only *Capital savings* from corporate portfolios are impacted during the business cycle.
+In this section I argue that portfolio reallocation during the business cycle explains part of *Capital savings* cyclicality.
+Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans.
+Unfortunately, I can only observe assets aggregated by PD bands which does not allow me to track individual assets.
+As an alternative approach, I analyze the relationship of PD distribution moments with GDP to understand asset allocation during the business cycle.
+
+According to the theoretical framework presented in section 4.1, an increase in average PD reduces the amount of *Capital savings* if other moments are kept constant.
+This relationship is illustrated in panel A of Figure \ref{fig:pdmapcases} and empirically tested in Table \ref{tab:moments_contribution}.
+In all columns of this table average PD is negatively associated with *Capital savings*.
+Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios.
+In sum, the estimated coefficients of the average PD and the PD variance provide strong empirical support to the theoretical framework presented in section 4.1.
+At the same time, however, I cannot refute the hypothesis that portfolio reallocation does not impact *Capital savings* because the estimated effect of these two moments on *Capital savings* are in line with a scenario without any portfolio reallocation.
+
+Extending the analysis to the skewness and kurtosis of the PD distribution proves to be very helpful. Consider a
+
+Panel A in table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle.
+As expected, in most of the regressions, changes in average PD are negatively associated with GDP growth.
+This association is the known driver of the procyclical character of the current model-based capital regulation and its interpretation is fairly mechanical.
+During recessions average credit quality deteriorates, or in other words, average PD increases.
+The only exception to the rule is the sovereign portfolio.
+Column 5 of panel A shows that the coefficient of GDP growth on changes on average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero.
+One possible reason for this result is that, during economic downturns, banks may search for safe assets mostly in the form of safe haven government bonds.
+According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to counterbalance the deteriorations of the prior portfolio.
+It also helps explain the insignificant results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.
+
+As explained, if no other moment is affected during the business cycle we should observe procyclical *Capital savings*. However, it was shown in
+
+Common  
+I further analyse the relationship of the PD distribution moments with GDP and find that during economic expansions: average, variance and kurtosis decrease while the skewness increases.
 This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
@@
 ```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
-bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
-  select(name, Country, RWA.r)%>%
+bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
+  select(name, Country, RWA.r.Total)%>%
   group_by(name, Country) %>%
@@
   #       Country = fct_reorder(Country, ordering)) %>%
- ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
-                      y = RWA.r)) +
+ ggplot(mapping = aes(x = reorder(name, RWA.r.Total, FUN = mean, na.rm=TRUE),
+                      y = RWA.r.Total-1)) +
 geom_point(alpha = 0.5,size = 2) +
@@
 ```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
-bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
-  select(Country,name, year, RWA.r)%>%
+bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid) %>% 
+  select(Country,name, year, RWA.r.Total)%>%
   group_by(Country,name) %>%
  summarise_if(is.numeric,mean,na.rm= TRUE)%>%
- ggplot(mapping = aes(x = reorder(Country, RWA.r, FUN = mean, na.rm=TRUE),
-                          y = RWA.r)) +
+ ggplot(mapping = aes(x = reorder(Country, RWA.r.Total, FUN = mean, na.rm=TRUE),
+                          y = RWA.r.Total-1)) +
 geom_point(alpha = 0.5,size = 2) +
@@
 plot2 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year,  mean.PD) %>%
+  select(bvdid, year,  mean.PD.Total) %>%
   mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=mean.PD)) +
+  ggplot(mapping = aes(x = year,y=mean.PD.Total)) +
   geom_point(alpha = 0.3) +
@@
 plot3 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, var.PD) %>%
+  select(bvdid, year, var.PD.Total) %>%
   mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
+  ggplot(mapping = aes(x = year,y=sqrt(var.PD.Total))) +
   geom_point(alpha = 0.3) +
@@
   ungroup()%>%
-  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
   mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
@@
 ```
+
+```{r PD_distribution_data, eval=FALSE, echo=FALSE}
+
+IRB.data <- as.data.frame(pillar3.data) %>%
+  filter(Method2 %in% c("IRB"),
+         PD  < 0.99, EAD > 0) %>%
+  mutate(LGD        = ifelse(is.na(LGD), 0.45, LGD),
+         w.PD       = PD*EAD*LGD,
+         w.EAD      = LGD*EAD)
+
+PD.distribution <- data.frame()
+
+for(i in c("Total","Wholesale","Retail","Corporate","Sovereign","Banks")){
+PD.distribution <- IRB.data %>%
+  select(bvdid, name, year, w.EAD, PD, Portfolio_1, Portfolio_2) %>%
+  {if(i %in% c("Total")){
+     .
+    }else if(i %in% c("Wholesale","Retail")){
+     filter(., Portfolio_1 %in% c(i))
+    }else{
+     filter(., Portfolio_2 %in% c(i))}} %>%
+  mutate(w.EAD = round(w.EAD/100)) %>%
+  na.omit(cols=c("w.EAD")) %>%
+  expandRows("w.EAD") %>%
+  rbind(PD.distribution)}
+
+```
+
+```{r}
+density_cap = "\\label{fig:pdmapcases}The figure shows the two possible outcomes of a negative credit risk shock on a portfolio.
+The left-hand side plot shows the case where capital savings are procyclical.
+The right-hand side plot shows the case where capital savings are countercyclical.
+"
+```
+
+```{r, density, eval=FALSE, fig.height = 3, fig.width = 6,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= denisty_cap}
+
+ggplot(PD.distribution) + geom_density(aes(x=PD))
+
+view(select(pillar3.data, name, year, PD, Portfolio_1, Portfolio_2))
+```
+
 # Tables {-}
@@
 # Dependent variables
-dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
+dep.vars <- c(".Total", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
 dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
@@
 # Model formulas
-formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-   felm(x, data = filter(bank.data.win, RWA.s>0))
+   felm(x, data = bank.data.win)
 })
@@
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} for portfolio $j$.
-In all regressions the measure of \\textit{Capital savings} follows equation \\ref{measure1}.
+          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$.
+In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}.
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
@@
           type             = "latex",
-          font.size        = "small",
+          font.size        = "footnotesize",
           notes.label      = "",
@@
           notes.append     = FALSE,
-          label            = "tab1",
+          label            = "tab:main",
           style            = "qje",
@@
 
-\newpage
-
 \begin{table}[!htbp] \centering  
 \caption{Capital Savings and the Business Cycle: Robustness to Alternative Measures}  
-\label{tab2:reg}
+\label{tab:alt}
 \footnotesize
 \begin{tabular}{@{\extracolsep{1pt}}lccccc}
-\multicolumn{6}{c}{\parbox[t]{16cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}.
+\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}.
 \end{equation*} \newline
-The dependent variable is the change in the logarithm of \textit{Capital savings} for portfolio $j$.
-In Panel A the measure of \textit{Capital savings} follows equation \ref{eq5}.
-In Panel B the measure of \textit{Capital savings} follows equation \ref{eq6}.
+The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$.
+The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:measure3} in panel B.
 In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution.
@@
 \hline \\
-\end{tabular}
+\end{tabular}\\
 \begin{tabular}{@{\extracolsep{1pt}}cccccc}
@@
 reg.list <- lapply(formulas, function(x) {
- felm(x, data = filter(bank.data.win, RWA.s>0))
+   felm(x, data = bank.data.win)
 })
@@
 ```{r table_cyclicality_alt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se              = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          #label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+```
+
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+ \\
+\multicolumn{6}{c}{Panel B: $\text{Y}=\text{Log RW}^{s}$ } \\
+\end{tabular}
+```{r reg_cyclicality_rw, echo=FALSE}
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("RW.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+   felm(x, data = bank.data.win)
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_rw, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se              = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          #label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
+
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\\
+\multicolumn{6}{c}{Panel C: $\text{Y}=\text{Log Gini}$ } \\
+\end{tabular}
+
+```{r reg_cyclicality_gini, echo=FALSE}
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+   felm(x, data = bank.data.win)
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se              = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          #label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
+\end{table}
+
+\newpage
+
+```{r reg_cyclicality_components, echo=FALSE}
+
+dep.vars2 <- c("log.RWA.SA_gr","RW.SA_gr ", "q.SA_gr", "q.Wholesale_gr", "q.Retail_gr")
+dep.text2 <- c("$\\Delta\\text{Log RWA}_{i,t}^{SA}$", "$\\Delta\\text{Log RW}_{i,t}^{SA}$",
+               "$\\Delta\\text{q}_{i,t}^{SA}$", "$\\Delta\\text{q}_{i,t}^{Wholesale}$", "$\\Delta\\text{q}_{i,t}^{Retail}$")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+reg.list <- lapply(formulas, function(x) {
+ felm(x, data = filter(bank.data.win, year>2007))
+})
+
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_cyclicality_components, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
 stargazer(reg.list,
+          title = "Other RWA Components and the Business Cycle",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{13.5cm}{The table examines the relationship between the business cycle and other components of RWA.
+The dependent variable is the change in the logarithm of risk-weighted assets for portfolios under the standardize approach in column 1, the change in risk-weights for portfolios under the standardize approach in column 2, the change in the ratio of exposure under the standardize approach to total exposure in column 3, the change in the wholesale share of the IRB portfolio in column 4, and the change in the retail share of the IRB portfolio in column 5.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels   = dep.text2,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "small",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab:other",
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE)
+
+```
+
+\newpage
+
+```{r reg_moments, echo=FALSE}
+
+ind.text <- c("$\\Delta \\text{Mean(PD}_{i,t}$)", "$\\Delta \\text{Var(PD}_{i,t}$)", "$\\Delta \\text{Skew(PD}_{i,t}$)", "$\\Delta \\text{Kurt(PD}_{i,t}$)")
+
+reg.list <- list()
+# Run the models
+for(i in dep.vars){
+formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ mean + var + skew + kurt  | year + bvdid | 0 | Country "))
+ data = rename(bank.data.win,
+               mean    = paste0("mean.PD"  , i, "_gr"),
+               var     = paste0("var.PD"   , i, "_gr"),
+               skew    = paste0("skew.PD"  , i, "_gr"),
+               kurt    = paste0("kurt.PD"  , i, "_gr"))%>%
+   felm(data = ., formula = formula.text ) -> reg.list[[i]]
+  }
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          title = "PD moments and Capital Savings",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$.
+In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab:moments_contribution",
+          style            = "qje",
+          column.sep.width = "1pt",
+          align            = TRUE)
+
+```
+
+\newpage
+
+\begin{table}[!htbp] \centering  
+\caption{PD moments and the Business Cycle}  
+\label{tab:moments_cyclicality}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}.
+\end{equation*} \newline
+The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$.
+The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:measure3} in panel B.
+In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution.
+$\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
+[-1.8ex] \\
+\hline \\
+\end{tabular}\\
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\multicolumn{6}{c}{Panel A: $\text{Y}=\Delta \text{Mean(PD}_{i,t}\text{)}$} \\
+\end{tabular}
+
+```{r reg_mean, echo=FALSE}
+
+# Model formulas
+formula.text <- paste0(paste0("mean.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = bank.data.win)
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_mean, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
           table.layout     = "=c=d#-t-as-",
@@
           column.sep.width = "2pt",
-          float            = FALSE,
-          align            = TRUE)
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
 
@@
 \\
-\multicolumn{6}{c}{Panel B: $\text{Y}=\text{Log RW}^{s}$ } \\
+\multicolumn{6}{c}{Panel B: $\text{Y}=\Delta \text{Var(PD}_{i,t}\text{)}$}
 \end{tabular}
-```{r reg_cyclicality_rw, echo=FALSE}
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("RW.s", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+
+```{r reg_var, echo=FALSE}
+
+# Model formulas
+formula.text <- paste0(paste0("var.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
 
+# Run the models
 reg.list <- lapply(formulas, function(x) {
- felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
+  felm(x, data = bank.data.win)
+})
+
+# Robust standard errors
 se.list  <-  lapply(reg.list, function(x){
@@
 
-```{r table_cyclicality_rw, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
+```{r table_var, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
           table.layout     = "=c=d#-t-as-",
@@
 \\
-\multicolumn{6}{c}{Panel C: $\text{Y}=\text{Log Gini}$ } \\
+\multicolumn{6}{c}{Panel C: $\text{Y}=\Delta \text{Skew(PD}_{i,t}\text{)}$}
 \end{tabular}
 
-```{r reg_cyclicality_gini, echo=FALSE}
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0("gini.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+```{r reg_skew, echo=FALSE}
+
+# Model formulas
+formula.text <- paste0(paste0("skew.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
 
+# Run the models
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
+  felm(x, data = bank.data.win)
+})
+
+# Robust standard errors
 se.list  <-  lapply(reg.list, function(x){
@@
 
-```{r table_cyclicality_gini, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          title = "Alternative measure: PD distribution Gini Coefficient",
+```{r table_skew, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
           table.layout     = "=c=d#-t-as-",
@@
           notes.append     = FALSE,
-          label            = "tab3",
+          label            = "tab2",
           style            = "qje",
           column.sep.width = "2pt",
-          float            = FALSE, 
-          align            = TRUE)
-
-```
-
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
 \end{table}
 
-\newpage
-
-```{r reg_cyclicality_components, echo=FALSE}
-dep.vars2 <- c("log.RWA.SA_gr","RW.SA_gr ", "q.SA_gr", "q.Wholesale_gr", "q.Retail_gr")
-dep.text2 <- c("$\\Delta\\text{Log RWA}_{i,t}^{SA}$", "$\\Delta\\text{Log RW}_{i,t}^{SA}$",
-               "$\\Delta\\text{q}_{i,t}^{SA}$", "$\\Delta\\text{q}_{i,t}^{Wholesale}$", "$\\Delta\\text{q}_{i,t}^{Retail}$")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+\begin{table}[!htbp] \centering
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\multicolumn{6}{c}{\parbox[c]{15.8cm}{\centering \tablename\ \thetable{} -- continued from previous page}}\\
+\hline \\
+\end{tabular}\\
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\multicolumn{6}{c}{Panel D: $\text{Y}=\Delta \text{Kurt(PD}_{i,t}\text{)}$} 
+\end{tabular}
+
+```{r reg_kurt, echo=FALSE}
+
+# Model formulas
+formula.text <- paste0(paste0("kurt.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
 
+# Run the models
 reg.list <- lapply(formulas, function(x) {
- felm(x, data = filter(bank.data.win, RWA.s>0))
-})
-
+  felm(x, data = bank.data.win)
+})
+
+# Robust standard errors
 se.list  <-  lapply(reg.list, function(x){
@@
 
-```{r table_cyclicality_components, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          title = "Other RWA Components and the Business Cycle",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{14cm}{The table examines the relationship between the business cycle and other components of RWA.
-The dependent variable is the change in the logarithm of risk-weighted assets for portfolios under the standardize approach in column 1, the change in the logarithm of risk-weights for portfolios under the standardize approach in column 2, the change in the logarithm of the ratio of exposure under the standardize approach to total exposure in column 3, the change in the logarithm of the share of the IRB portfolio as wholesale exposure in column 4, and the change in the logarithm of the share of the IRB portfolio as retail exposure in column 5.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-All regressions include bank and year fixed effects.
-Robust standard errors adjusted for clustering at the country level are reported in parentheses.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          dep.var.labels   = dep.text2,
+```{r table_kurt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
           covariate.labels = ind.text,
@@
           column.sep.width = "2pt",
-          align            = TRUE)
-
-```
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
+
+\end{table}
 
@@
 
-```{r reg_moments, echo=FALSE}
-
+```{r reg_contribution, eval=TRUE, echo=FALSE}
 # Variables in the model 
-dep.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr", "log.RWA.r_gr")
-aux.1 <- c("$\\Delta \\text{Mean(PD}_{i,t}$)", "$\\Delta \\text{Variance(PD}_{i,t}$)", "$\\Delta \\text{Skewness(PD}_{i,t}$)", "$\\Delta \\text{Kurtosis(PD}_{i,t}$)")
-aux.2 <- c( "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
-dep.text <- c(aux.1, aux.2)
-ind.vars <- c("mean.PD_gr", "var.PD_gr", "skew.PD_gr", "kurt.PD_gr")
-ind.vars  <- paste(ind.vars, collapse=" + ")
-ind.vars <- c(rep("log.gdp_gr", 4), ind.vars)
-ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$", aux.1)
-
-# Fixed effects
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
+# Dependent variables
+dep.vars2 <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat.Total_gr", "log.RWA.r.Total_gr")
+dep.text2 <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
 
 # Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
+  felm(x, data = filter(bank.data.win, !is.na(log.CAR_gr)))
 })
@@
 
-```
-
-```{r table_moments, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+# Effect 
+effect.list  <-  lapply(reg.list, function(x){
+  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
+})
+effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
+effect.list <-  sprintf("%.1f \\%%",effect.list)
+
+fe.list <- list(c("Effect", effect.list),
+                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
+```
+
+```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
 stargazer(reg.list,
-          title = "Alternative measure: PD distribution Gini Coefficient",
+          title = "Capital Savings Effect on Capital Ratio Procyclicality",
           table.layout = "n=c=d#-t-as-",
@@
 
-          dep.var.labels   = dep.text,
+          dep.var.labels   = dep.text2,
           covariate.labels = ind.text,
@@
           notes.append     = FALSE,
-          label            = "tab3",
+          label            = "tab:contribution",
           style            = "qje",
@@
 
-```{r reg_contribution, eval=TRUE, echo=FALSE}
+```{r reg_extension, echo=FALSE, include=FALSE}
+
 # Variables in the model 
-# Dependent variables
-dep.vars <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat_gr", "log.RWA.r_gr")
-dep.text <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
-# Independent variables
-ind.vars <- paste0("log.gdp_gr")
-ind.text <- c("$\\Delta\\text{Log GDP}_{i,t-1}$")
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0 & !is.na(log.CAR_gr)))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-# Effect 
-effect.list  <-  lapply(reg.list, function(x){
-  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
-})
-effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
-effect.list <-  sprintf("%.1f \\%%",effect.list)
-
-fe.list <- list(c("Effect", effect.list),
-                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
-```
-
-```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          title = "Alternative measure: PD distribution Gini Coefficient",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{r} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-All regressions include bank and year fixed effects.
-Robust standard errors adjusted for clustering at the country level are reported in parentheses.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          #se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "small",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "tab3",
-          style            = "qje",
-          column.sep.width = "2pt",
-          #float.env        = "sidewaystable",
-          align            = TRUE)
-
-```
-
-\newpage
-
-```{r reg_extension, echo=FALSE, include=FALSE}
-
-# Variables in the model 
-# Dependent variables
-dep.vars <- c("log.RWA.r","log.RWA.r.Retail", "log.RWA.r.Wholesale", "log.RWA.r.Corporate", "log.RWA.r.Sovereign", "log.RWA.r.Banks")
-dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
 # Independent variables
@@
            "bank_credit_gdp",
-           #"pillarI_stringency",
-           "cap_reg",
+           "pillarI_stringency",
+           #"cap_reg",
            "Sup_Power",
@@
            "log.asset",
-           "Zscore",
            "ROE",
@@
               "$\\text{Credit-to-GDP}_{i,t}$",
-              #"$\\text{Pillar I stringency}_{i,t}$", 
-              "$\\text{Capital stringency}_{i,t}$", 
+              "$\\text{Pillar I stringency}_{i,t}$", 
+              #"$\\text{Capital stringency}_{i,t}$", 
               "$\\text{Official Supervisory power}_{i,t}$",
@@
               "$\\text{Log Total Assets}_{i,t}$",
-              "$\\text{Z-score}_{i,t}$",
               # "$\\text{Loans / Total Assets}_{i,t-1}$", 
@@
 # Model formulas
-formula.text <- paste0(paste0(dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
+formula.text <- paste0(paste0("log.RWA.r", dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, RWA.s>0))
+  felm(x, data = bank.data.win)
 })
@@
 stargazer(reg.list,
-          title = "capital Savings and Bank Characteristics",
+          title = "Correlates of Capital Savings",
           table.layout = "n=c=d#-t-as-",
@@
           notes.append     = FALSE,
-          label            = "tab6",
+          label            = "tab:extension",
           style            = "qje",
-          column.sep.width = "-1pt",
+          column.sep.width = "-3pt",
           #float.env        = "sidewaystable",
@@
 \caption{Summary Statistics}  
-\label{tab1:summary}
+\label{tab:summary}
 \footnotesize
 \begin{tabular}{@{\extracolsep{1pt}}lccccc}
-\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
+\multicolumn{6}{c}{\parbox[t]{13.5cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
 \end{tabular}
 
-```{r summary_components, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-bank.data.win %>% select(
-  capital.ratio_gr,
-  log.CAR_gr,
-  log.K_gr, 
-  log.RWA.IRB_gr,
-  log.RWA.r_gr,
+```{r summary1,  fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+table <-capture.output({bank.data.win %>%
+        filter(year>2007) %>% 
+  select(
+  log.RWA.r.Total_gr,
   log.RWA.r.Retail_gr,
@@
   log.RWA.r.Banks_gr,
-  log.RWA.hat_gr, 
-  log.RWA.hat.Retail_gr,
-  log.RWA.hat.Wholesale_gr, 
-  log.RWA.hat.Corporate_gr, 
-  log.RWA.hat.Sovereign_gr, 
-  log.RWA.hat.Banks_gr) %>%
+  log.RWA.r.Total, 
+  log.RWA.r.Retail,
+  log.RWA.r.Wholesale, 
+  log.RWA.r.Corporate, 
+  log.RWA.r.Sovereign, 
+  log.RWA.r.Banks,
+  log.RWA.s.Total_gr,
+  log.RWA.s.Retail_gr,
+  log.RWA.s.Wholesale_gr, 
+  log.RWA.s.Corporate_gr, 
+  log.RWA.s.Sovereign_gr, 
+  log.RWA.s.Banks_gr,
+  RW.s.Total_gr,
+  RW.s.Retail_gr,
+  RW.s.Wholesale_gr, 
+  RW.s.Corporate_gr, 
+  RW.s.Sovereign_gr, 
+  RW.s.Banks_gr,
+  gini.PD.Total_gr,
+  gini.PD.Retail_gr,
+  gini.PD.Wholesale_gr, 
+  gini.PD.Corporate_gr, 
+  gini.PD.Sovereign_gr, 
+  gini.PD.Banks_gr,
+  log.CAR_gr,
+  log.K_gr, 
+  log.RWA.IRB_gr,
+  log.RWA.hat.Total_gr,
+  log.RWA.SA_gr,
+  RW.SA_gr, 
+  q.SA_gr, 
+  q.Wholesale_gr, 
+  q.Retail_gr) %>%
   as.data.frame() %>% 
@@
           covariate.labels = 
-            c("$\\Delta \\text{Log Capital Ratio}$",
-              "$\\Delta \\text{Log Adjusted Capital Ratio}$",
+            c("\\hspace{0.0cm}$\\Delta \\text{Log RWA}^s$ total",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^s$ retail",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^s$ wholesale",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$ corporate",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$ sovereign", 
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$ banks",
+              "\\hspace{0.0cm}$ \\text{Log RWA}^s$ total",
+              "\\hspace{0.5cm}$ \\text{Log RWA}^s$ retail",
+              "\\hspace{0.5cm}$ \\text{Log RWA}^s$ wholesale",
+              "\\hspace{1.0cm}$ \\text{Log RWA}^s$ corporate",
+              "\\hspace{1.0cm}$ \\text{Log RWA}^s$ sovereign", 
+              "\\hspace{1.0cm}$ \\text{Log RWA}^s$ banks",
+              "\\hspace{0.0cm}$\\Delta \\text{Log RWA}^{\\$}$ total",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ retail",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ wholesale",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ corporate",
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ sovereign", 
+              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ banks",
+              "\\hspace{0.0cm}$\\Delta \\text{RW}^s$ total",
+              "\\hspace{0.5cm}$\\Delta \\text{RW}^s$ retail",
+              "\\hspace{0.5cm}$\\Delta \\text{RW}^s$ wholesale",
+              "\\hspace{1.0cm}$\\Delta \\text{RW}^s$ corporate",
+              "\\hspace{1.0cm}$\\Delta \\text{RW}^s$ sovereign", 
+              "\\hspace{1.0cm}$\\Delta \\text{RW}^s$ banks",
+              "\\hspace{0.0cm}$\\Delta$ Log Gini total",
+              "\\hspace{0.5cm}$\\Delta$ Log Gini retail",
+              "\\hspace{0.5cm}$\\Delta$ Log Gini wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log Gini corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log Gini sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log Gini banks",
+              "$\\Delta \\text{Log capital ratio}$",
               "\\hspace{0.5cm}$\\Delta \\text{Log capital}$",
-              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}$",
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\text{IRB}}$",
               "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^c$",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^c$ Retail",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^c$ Wholesale",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^c$ Corporate",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^c$ Sovereign", 
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^c$ Banks",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^s$",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Retail",
-              "\\hspace{1.5cm}$\\Delta \\text{Log RWA}^s$ Wholesale",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Corporate",
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Sovereign", 
-              "\\hspace{2.0cm}$\\Delta \\text{Log RWA}^s$ Banks"),
+              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\text{SA}}$",
+              "\\hspace{0.5cm}$\\Delta \\text{RW}^{\\text{SA}}$",
+              "$\\Delta q^{\\text{SA}}$",
+              "$\\Delta q^{\\text{IRB, Wholesale}}$",
+              "$\\Delta q^{\\text{IRB, Retail}}$"),
           notes.label = "",
@@
           digits = 3,
-          float = FALSE)
+          align = TRUE,
+          float = FALSE)})
+
+table <- c(table[1:8],"Panel A: Main dependent variables \\\\",table[-(1:8)])
+
+table <- c(table[1:21],"Panel B: Alternative measures \\\\",table[-(1:21)])
+table <- c(table[1:21]," \\\\[-1.8ex]",table[-(1:21)])
+
+table <- c(table[1:41],"Panel C: Other dependet variables \\\\",table[-(1:41)])
+table <- c(table[1:41],"\\\\[-1.8ex]",table[-(1:41)])
+
+table <- gsub("}lD{.}{.}{-3}",  "}p{0.35\\textwidth}D{.}{.}{4.1}", table, fixed = T)
+
+#table <- gsub("{-3}", "{3.4}", table, fixed = T)
+
+cat(table)
+
 ```
@@
 \begin{table}[!htbp] \centering
-\caption{Summary Statistics}  
-\label{tab1:summary}
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\multicolumn{6}{c}{\parbox[c]{13cm}{\centering \tablename\ \thetable{} -- continued from previous page}}\\
+\end{tabular}\\
 \footnotesize
-\begin{tabular}{@{\extracolsep{1pt}}lccccc}
-\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
-\end{tabular}
-
-```{r summary_other_measures, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-bank.data.win %>% select(
-  RW.s_gr,
-  RW.s.Retail_gr,
-  RW.s.Wholesale_gr, 
-  RW.s.Corporate_gr, 
-  RW.s.Sovereign_gr, 
-  RW.s.Banks_gr,
-  log.RWA.s_gr,
-  log.RWA.s.Retail_gr,
-  log.RWA.s.Wholesale_gr, 
-  log.RWA.s.Corporate_gr, 
-  log.RWA.s.Sovereign_gr, 
-  log.RWA.s.Banks_gr,
-  gini.PD_gr,
-  gini.PD.Retail_gr,
-  gini.PD.Wholesale_gr, 
-  gini.PD.Corporate_gr, 
-  gini.PD.Sovereign_gr, 
-  gini.PD.Banks_gr,
-  log.RWA.r, 
-  log.RWA.r.Retail,
-  log.RWA.r.Wholesale, 
-  log.RWA.r.Corporate, 
-  log.RWA.r.Sovereign, 
-  log.RWA.r.Banks_gr) %>%
-  as.data.frame() %>% 
-stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
-          label = "tab:summary",
-          omit.summary.stat = c("p25","p75"),
-          covariate.labels = 
-            c("\\hspace{0.0cm}$\\Delta \\text{Log RW}^s$",
-              "\\hspace{0.5cm}$\\Delta \\text{Log RW}^s$ Retail",
-              "\\hspace{0.5cm}$\\Delta \\text{Log RW}^s$ Wholesale",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RW}^s$ Corporate",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RW}^s$ Sovereign", 
-              "\\hspace{1.0cm}$\\Delta \\text{Log RW}^s$ Banks",
-              "\\hspace{0.0cm}$\\Delta \\text{Log RWA}^{\\$}$",
-              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ Retail",
-              "\\hspace{0.5cm}$\\Delta \\text{Log RWA}^{\\$}$ Wholesale",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ Corporate",
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ Sovereign", 
-              "\\hspace{1.0cm}$\\Delta \\text{Log RWA}^{\\$}$ Banks",
-              "\\hspace{0.0cm}$\\Delta$ Log Gini",
-              "\\hspace{0.5cm}$\\Delta$ Log Gini Retail",
-              "\\hspace{0.5cm}$\\Delta$ Log Gini Wholesale",
-              "\\hspace{1.0cm}$\\Delta$ Log Gini Corporate",
-              "\\hspace{1.0cm}$\\Delta$ Log Gini Sovereign", 
-              "\\hspace{1.0cm}$\\Delta$ Log Gini Banks",
-              "\\hspace{0.0cm}$ \\text{Log RWA}^r$",
-              "\\hspace{0.5cm}$ \\text{Log RWA}^r$ Retail",
-              "\\hspace{0.5cm}$ \\text{Log RWA}^r$ Wholesale",
-              "\\hspace{1.0cm}$ \\text{Log RWA}^r$ Corporate",
-              "\\hspace{1.0cm}$ \\text{Log RWA}^r$ Sovereign", 
-              "\\hspace{1.0cm}$ \\text{Log RWA}^r$ Banks"),
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          summary = TRUE,
-          digits = 3,
-          float = FALSE)
-```
-
-\end{table}
-
-\begin{table}[!htbp] \centering  
-\caption{Summary Statistics}  
-\label{tab1:summary}
-\footnotesize
-\begin{tabular}{@{\extracolsep{1pt}}lccccc}
-\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
-\end{tabular}
-
-```{r summary_portfolio, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-bank.data.win %>% select(
-  mean.PD_gr, 
+
+```{r summary3, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
+table <- capture.output({bank.data.win %>%
+                         filter(year>2007) %>% 
+  select(
+  mean.PD.Total_gr, 
   mean.PD.Retail_gr,
@@
   mean.PD.Banks_gr,
-  var.PD_gr, 
-  skew.PD_gr,
-  kurt.PD_gr) %>%
-  as.data.frame() %>% 
-stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
-          label = "tab:summary",
-          omit.summary.stat = c("p25","p75"),
-          covariate.labels = 
-            c("$\\Delta$ Log PD Mean",
-              "\\hspace{0.5cm}$\\Delta$ Log PD Mean Retail",
-              "\\hspace{0.5cm}$\\Delta$ Log PD Mean Wholesale",
-              "\\hspace{1.0cm}$\\Delta$ Log PD Mean Corporate",
-              "\\hspace{1.0cm}$\\Delta$ Log PD Mean Sovereign", 
-              "\\hspace{1.0cm}$\\Delta$ Log PD Mean Banks", 
-              "$\\Delta$ Log PD Variance",
-              "$\\Delta$ Log PD Skewness", 
-              "$\\Delta$ Log PD Kurtosis"),
-          notes.label = "",
-          notes.align = "c", 
-          notes.append = FALSE,
-          digits = 3,
-          summary = TRUE,
-          float = FALSE)
-```
-
-\end{table}
-
-\begin{table}[!htbp] \centering  
-\caption{Summary Statistics}  
-\label{tab1:summary}
-\footnotesize
-\begin{tabular}{@{\extracolsep{1pt}}lccccc}
-\multicolumn{6}{c}{\parbox[t]{12cm}{The table shows summary statistics for banks that have introduced the IRB approach between 2007 and 2018. All variables are winsorized at 5\%}} \\
-\end{tabular}
-
-```{r summary_control, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-table <- capture.output({bank.data.win %>% ungroup()%>% select(
+  var.PD.Total_gr, 
+  var.PD.Retail_gr,
+  var.PD.Wholesale_gr, 
+  var.PD.Corporate_gr, 
+  var.PD.Sovereign_gr, 
+  var.PD.Banks_gr,
+  skew.PD.Total_gr,
+  skew.PD.Retail_gr,
+  skew.PD.Wholesale_gr, 
+  skew.PD.Corporate_gr, 
+  skew.PD.Sovereign_gr, 
+  skew.PD.Banks_gr,
+  kurt.PD.Total_gr,
+  kurt.PD.Retail_gr,
+  kurt.PD.Wholesale_gr, 
+  kurt.PD.Corporate_gr, 
+  kurt.PD.Sovereign_gr, 
+  kurt.PD.Banks_gr,
   log.gdp_gr,
@@
           covariate.labels = 
-            c("$\\Delta \\text{Log GDP}$",
+            c("$\\Delta$ Log PD mean total",
+              "\\hspace{0.5cm}$\\Delta$ Log PD mean retail",
+              "\\hspace{0.5cm}$\\Delta$ Log PD mean wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log PD mean corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log PD mean sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD mean banks", 
+              "$\\Delta$ Log PD variance total",
+              "\\hspace{0.5cm}$\\Delta$ Log PD variance retail",
+              "\\hspace{0.5cm}$\\Delta$ Log PD variance wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log PD variance corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log PD variance sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD variance banks", 
+              "$\\Delta$ Log PD skewness total", 
+              "\\hspace{0.5cm}$\\Delta$ Log PD skewness retail",
+              "\\hspace{0.5cm}$\\Delta$ Log PD skewness wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log PD skewness corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log PD skewness sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD skewness banks", 
+              "$\\Delta$ Log PD kurtosis total",
+              "\\hspace{0.5cm}$\\Delta$ Log PD kurtosis retail",
+              "\\hspace{0.5cm}$\\Delta$ Log PD kurtosis wholesale",
+              "\\hspace{1.0cm}$\\Delta$ Log PD kurtosis corporate",
+              "\\hspace{1.0cm}$\\Delta$ Log PD kurtosis sovereign", 
+              "\\hspace{1.0cm}$\\Delta$ Log PD kurtosis banks",
+              "$\\Delta \\text{Log GDP}$",
               "Log GDP",
@@
           notes.append = FALSE,
+          digits = 3,
           summary = TRUE,
-          digits = 3,
+          align = TRUE,
           float = FALSE)})
 
-table <- gsub("lccccc",
-              "p{0.4\\textwidth}ccccc", table, fixed = T)
+table <- c(table[1:8],"Panel D: PD distribution moments \\\\",table[-(1:8)])
+
+table <- c(table[1:33],"Panel E: Independent variables \\\\",table[-(1:33)])
+table <- c(table[1:33]," \\\\[-1.8ex]",table[-(1:33)])
+
+table <- gsub("}lD{.}{.}{-3}",  "}p{0.35\\textwidth}D{.}{.}{4.1}", table, fixed = T)
+
+#table <- gsub("{-3}", "{3.4}", table, fixed = T)
 
@@
 \end{table}
+
+## RW Decomposition {-}
+
+In this section I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
+I divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
+Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
+I further breakdown RW_{IRB} into its portfolio components, $\text{RW}_{IRB}=\sum_j q_{j}\text{RW}_{j}$, in which the $j$ subscript stands for the different portfolios under the IRB approach: Wholesale, retail and equity.
+Equation \ref{eq5} shows the final decomposition:
+
+\begin{equation}
+\begin{aligned}
+\label{eq5}
+\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j} \text{RW}_{j}^{c} \Delta q_{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
+&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin},
+\end{aligned}
+\end{equation}
+in which  $\Delta$ stands for changes relative to a benchmark.
+For the cross-section decomposition, the benchmark is the average bank.
+For the time decomposition, the benchmark is either the year when the IRB approach was first introduced at the bank level or 2008.
+The superscript $B$ stands for benchmark.
+For instance, $q_SA^B$ is the share of the portfolio under the standardized approach for the benchmark.
+
+I decompose the variation of RW relative to the benchmark in six components, two of them relating to *Capital savings*.
+The first, $(1-q^{B}_{SA})\Delta \text{RW}^{s}$, measures the intensity of *Capital savings*.
+That is, I fix the share of the portfolio under the IRB approach ($q_{SA}^B$) and variation in the component comes only from changes of $RW^s$ relative to the benchmark.
+The second component, $\text{RW}^{s}\Delta q^{B}_{SA}$, measures the extensive margin of *Capital savings*, i.e., variation in this component comes only from changes of the share under the IRB approach relative to the benchmark.
+Note that the magnitude of the second component is moderated by how intense *Capital savings* are in absolute terms.
+
+The remaining components are extensively discussed by @cannata2012inside.
+The first on the right-hand side of equation \ref{eq5} is the intensity of RW under the standardized approach ($q^{B}_{SA} \Delta \text{RW}_{SA}$).
+The contribution of this component to $\Delta \text{RW}$ comes from variaton of RW in portfolios under the standardized approach while fixing the share of this portfolio for all banks.
+The second term measures the intensity of RW under the IRB approach excluding the *Capital savings*, or differently, the intensity of the average PD ($(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j$).
+The contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of RW in the three portfolios under the IRB approach relative to the respective benchmark RW while holding constant the share of each IRB portfolio and the share of total IRB.
+The third component measure the contribution of the IRB portfolio mix.
+Because some types of portfolios have on average lower RW, the composition within IRB across types of portfolios matter to how large overall RW is.
+For instance retail portfolios tend to have lower RW relative to wholesale portfolios because the former usually are colletarized loans, such as mortgages, while the latter is mostly composed of corporate loans.
+Hence, the contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of the three portfolios shares within the IRB portfolio relative to the respective benchmarks while holding constant the respective intensity of average PD ($RW_j^c$) and the share of total IRB.
+Lastly, $(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}$ is the roll-out effect, i.e., the contribution to $\Delta \text{RW}$ coming from moving the portfolio from the standardized approach to the IRB approach.
+
+Figure \ref{fig:crossdecomposition1} plots the intensive and extensive margin of each bank in my sample using the average bank as benchmark.
+The scatter plot shows that the correlation between the two *Capital savings* components is not statistically significant.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
+If Goldman Sachs is included the correlation becomes statistically significant.} The figure suggests that banks benefit from the two components separately, some benefting more of the extensive margin, such as Northern Trust, and others more of the intensive margin, such as BayernLB.
+
+Figure \ref{fig:crossdecomposition2} shows that the net effect of the intensive and extensive margin is negatively correlated to the net effect of the other components.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
+If Goldman Sachs is included the correlation becomes even more negative.} This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
+
+As Figure \ref{fig:rsqrtdecomp} shows *Capital savings* components account for 27% of the RW variation across banks.
+Each component accounts for around 13.5% of the RW variation and are second and third in rank, losing only to the average PD contribution.
+I derive the $\text{R}^2$ contribution of each RW component using the Shapley-Owen decomposition method.
+This method consider the contribution of each variable to total $\text{R}^2$ as the sum of $\text{R}^2$ from regressions using all possible variable combinations that include the variable of interest and weighting each regression by the number of variable permutations.
+In sum, figure \ref{fig:rsqrtdecomp} offers strong evindence of the relative importance of *Capital savings* as source of variation to RW across banks.
+
+Figure \ref{fig:timedecomposition} shows the evolution of RW components relative to the time of IRB introduction (left-hand side plot), and relative to 2008 (right-hand side plot).
+The negative values for the extensive margin implies that banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach.
+However, during the same period, the intensity of *Capital savings* has, on average, decreased in time.
+That is, for the same average PD, capital requirements have increased from 2008 to 2018 and since the year IRB approach was introduced.
+
+The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}.
+Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.
+[^1] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
+This distribution compression leads to a reduction in *Capital savings*.
+This explanation is in line with the hypothesis that *Capital savings* are countercyclical.
+Efforts to increase capital ratio by risk reduction are increasingly more difficult.
+
+[^1]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].

### 2020-04-14 Analysis/Draft/Draft_bookdown.Rmd
Start portfolio reallocation section

@@
 source("../../Auxiliar/longtable.R") ## longtable for stargazer
-source("../../Auxiliar/adj_stargazer.R") ## adjust the first column width for stargazer
+#source("../../Auxiliar/adj_stargazer.R") ## adjust the first column width for stargazer
 
@@
 
-The results from Tables \ref{tab:main} and \ref{tab:alt} raises the question of why it appears that only *Capital savings* from corporate portfolios are impacted during the business cycle.
-In this section I argue that portfolio reallocation during the business cycle explains part of *Capital savings* cyclicality.
-Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans.
-Unfortunately, I can only observe assets aggregated by PD bands which does not allow me to track individual assets.
-As an alternative approach, I analyze the relationship of PD distribution moments with GDP to understand asset allocation during the business cycle.
+In this section I argue that portfolio reallocation during the business cycle explains part of the *Capital savings* cyclicality.
+As method, I use the observed movements on moments of the PD distribution during the business cycle that are not in line with the expected movements from a static portfolio as evidence in favor of a portfolio reallocation effect on *Capital savings*.[^1]
+
+[^1]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans.
+Unfortunately, in my dataset, I only observe assets aggregated by PD bands without knowledge if amount changes in a band are due to new loans or changes in credit quality of assets already held.
 
@@
 In all columns of this table average PD is negatively associated with *Capital savings*.
-Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios.
-In sum, the estimated coefficients of the average PD and the PD variance provide strong empirical support to the theoretical framework presented in section 4.1.
-At the same time, however, I cannot refute the hypothesis that portfolio reallocation does not impact *Capital savings* because the estimated effect of these two moments on *Capital savings* are in line with a scenario without any portfolio reallocation.
-
-Extending the analysis to the skewness and kurtosis of the PD distribution proves to be very helpful. Consider a
 
@@
 The only exception to the rule is the sovereign portfolio.
-Column 5 of panel A shows that the coefficient of GDP growth on changes on average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero.
+Column 5 of panel A shows that the coefficient of GDP growth rate on changes in the average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero.
 One possible reason for this result is that, during economic downturns, banks may search for safe assets mostly in the form of safe haven government bonds.
-According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to counterbalance the deteriorations of the prior portfolio.
-It also helps explain the insignificant results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.
-
-As explained, if no other moment is affected during the business cycle we should observe procyclical *Capital savings*. However, it was shown in
-
-Common  
-I further analyse the relationship of the PD distribution moments with GDP and find that during economic expansions: average, variance and kurtosis decrease while the skewness increases.
-This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.
+According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to compensate any deterioration of the prior held portfolio.
+Together with column 5 of Table \ref{tab:moments_contribution}, it also helps explain the positive (and insignificant) results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.
+
+Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios.
+And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, Panel B in table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios.
+In sum, the estimated coefficients of the average PD and the PD variance in Table \ref{tab:moments_contribution} and panels A and B in Table \ref{tab:moments_cyclicality} provide strong empirical support to the theoretical framework presented in section 4.1.
+
+At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted exactly because the estimated effect of these two moments on *Capital savings* are strongly in line with a model without any portfolio reallocation.
+Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness and kurtosis.
+It is trivial that the equivalent scenario of panel A of figure \ref{fig:pdmapcases} wouldn't affect either of these moments.
+On the other hand, in the second scenario the PD distribution should become more skewed to the right.
+Even if we think that some assets will become safer during bad times, it is reasonable to expect a larger effect on the right tail than on the left tail of the PD distribution.
+Consequently, considering a static portfolio, we expect the skewness of the PD distribution to be negatively associated with the GDP growth rate.
+
+Panel C in table \ref{tab:moments_cyclicality} provides evidence that the observed relationship of PD distribution skewness and the business cycle can only be explained with a within portfolion reallocation.
 
@@
 Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.
-[^1] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
+[^2] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
 This distribution compression leads to a reduction in *Capital savings*.
@@
 
-[^1]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].
+[^2]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].

### 2020-04-15 Analysis/Draft/Draft_bookdown.Rmd
Add text on skewness to draft

@@
 Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios.
-And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, Panel B in table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios.
+And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, panel B in Table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios.
 In sum, the estimated coefficients of the average PD and the PD variance in Table \ref{tab:moments_contribution} and panels A and B in Table \ref{tab:moments_cyclicality} provide strong empirical support to the theoretical framework presented in section 4.1.
 
-At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted exactly because the estimated effect of these two moments on *Capital savings* are strongly in line with a model without any portfolio reallocation.
+At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted, exactly because the estimated effect of these two moments on *Capital savings* are strongly aligned with the theoretical framework which portfolio reallocation is not present.
 Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness and kurtosis.
-It is trivial that the equivalent scenario of panel A of figure \ref{fig:pdmapcases} wouldn't affect either of these moments.
+It is trivial that the equivalent case with a continuous PD distribution to the scenario in panel A of figure \ref{fig:pdmapcases} wouldn't affect either of these moments.
 On the other hand, in the second scenario the PD distribution should become more skewed to the right.
 Even if we think that some assets will become safer during bad times, it is reasonable to expect a larger effect on the right tail than on the left tail of the PD distribution.
-Consequently, considering a static portfolio, we expect the skewness of the PD distribution to be negatively associated with the GDP growth rate.
-
-Panel C in table \ref{tab:moments_cyclicality} provides evidence that the observed relationship of PD distribution skewness and the business cycle can only be explained with a within portfolion reallocation.
+Consequently, without portfolio reallocation, we should expect the skewness of the PD distribution to be negatively associated with the GDP growth rate.
+
+Panel C in table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle cannot be explained without portfolion reallocation.
+All the statistical significant coefficients are positive, contrary to the expectation if banks wouldn't reallocate their portfolio.
+Considering an economic downturn, the positive coefficients suggest that banks fight the deterioration of their portfolio credit quality by reducing risk taking in new loans.
+The combination of these two movements, higher PD of existing exposure and lower PD of new assets, results in a less skewed PD distribution during recessions.
+
+Table \ref{tab:moments_contribution} shows that lower skewness translates to higher *Capital savings*.
+The effect is statistical significant for all except sovereign portfolios.
+According to column 1, an increase of 0.182 percentage points in PD distribution skewness growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.009 percentage points on average, representing 20% of total *Capital savings* standard deviation.
+The effect is even stronger on corporate portfolios accounting for 40% of their *Capital savings* standard deviation.
 
@@
 for(i in dep.vars){
-formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ mean + var + skew + kurt  | year + bvdid | 0 | Country "))
+formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ mean + var + skew   | year + bvdid | 0 | Country "))
  data = rename(bank.data.win,
@@
                var     = paste0("var.PD"   , i, "_gr"),
-               skew    = paste0("skew.PD"  , i, "_gr"),
-               kurt    = paste0("kurt.PD"  , i, "_gr"))%>%
+               skew    = paste0("skew.PD"  , i, "_gr"))%>%
    felm(data = ., formula = formula.text ) -> reg.list[[i]]

### 2020-04-18 Analysis/Draft/Draft_bookdown.Rmd
Finish moments section first draft

@@
 In all columns of this table average PD is negatively associated with *Capital savings*.
+Note that, in Table \ref{tab:moments_contribution}, not only the explained but also explanatory variables change across regressions by portfolio $j$.
 
@@
 
-Panel C in table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle cannot be explained without portfolion reallocation.
+Panel C in Table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle cannot be explained without portfolion reallocation.
 All the statistical significant coefficients are positive, contrary to the expectation if banks wouldn't reallocate their portfolio.
@@
 
-Table \ref{tab:moments_contribution} shows that lower skewness translates to higher *Capital savings*.
+Table \ref{tab:moments_contribution} shows that lower skewness translates into higher *Capital savings*.
 The effect is statistical significant for all except sovereign portfolios.
@@
 The effect is even stronger on corporate portfolios accounting for 40% of their *Capital savings* standard deviation.
+The evidence of Panel C in Table \ref{tab:moments_cyclicality} and Table \ref{tab:moments_contribution} suggest that as banks flight to safety in economic recessions, the increase in *Capital savings* due to the associated less skewed PD distribution helps banks to maintain their capital ratios.
+This countercyclical effect driven by the skewness of the PD distribution is on top of any gain from a lower average PD.
 

### 2020-05-10 Analysis/Draft/Draft_bookdown.Rmd
Several changes to the draft text. Version before I started the RW-lending project

@@
 I investigate the relationship of these capital savings with the business cycle and find it to   be countercyclical.
-I further document that the shape are important sources of across banks.
-The results suggest   that a recalibration of the current IRB framework can have sizeble effects on the procyclicality of capital       regulation.
+The results suggest that a recalibration of the current IRB framework can reduce considerably the procyclicality of capital regulation.
 spacing: double
@@
 library(splitstackshape)      ## Command to expand data
-
+library(latex2exp)            ## Use latex code in plots
+library(lemon)                ## Plot brackets
 #library(textablr)
@@
 
-# Introduction
-
-The regulation on bank capital requirements have evolved substantially since the first international standard, the 1988 Basel Accord.
+\section{Introduction} \label{sec:intro}
+
+The regulation on bank capital requirements has evolved substantially since the first international standard, the 1988 Basel Accord.
 At the heart of these changes is the link of capital charges to asset risk [@basel2004international; @BCBS2017basel; @EBA2019policy].
@@
 First, linking capital charges to asset risks exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
-As the overall credit quality detiorates during economic downturns, capital requirements tighten, forcing banks to cut down lending.
+As the overall credit quality deteriorates during economic downturns, capital requirements tighten, forcing banks to cut down lending.
 Second, the inherent flexibility of the IRB approach may allow for differences in capital requirements that do not reflect portfolio risk but rather modelling choices [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits;@berg2017analysis; @ferribank].
@@
 I also find that *Capital savings* measures are negatively correlated to other RW components.
-This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
+This correlation implies that banks with higher average portfolio riskiness benefit the most from *Capital savings*.
 Lastly, although banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach (known as the roll-out effect), the intensity of *Capital savings* has, on average, decreased in time.
@@
 
-I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic expasions average PD and PD variance decrease while the PD kurtosis and PD skewness increase.
-This evidence implies that distributions become more symmetric and with less extreme values during recessions, suggesting a flight to safety behaviour by banks.
+I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic expansions average PD and PD variance decrease while the PD kurtosis and PD skewness increase.
+This evidence implies that distributions become more symmetric and with less extreme values during recessions, suggesting a flight to safety behavior by banks.
 As the recession reduces their existing portfolio credit quality (evidenced by the negative coefficient of average PD) banks respond by reducing risk-taking.
@@
 Finally, section 5 concludes.
-
-# Institutional Background and Capital Savings Measures
-
-## Capital Requirements under the IRB approach
-
-In June 2004, the Basel Committee issued a revised framework on international convergence of capital measurements and capital standards [@basel2004international], which serves as the basis for national rulemaking and implementation processes.
-Since then, financial institutions may choose between two approaches to calculate the capital requirement for credit risk; the standardized approach (essentially a slightly modified version of the first accord) and the IRB approach.
+\section{Institutional Background and Hypotheses} \label{sec:background}
+
+\subsection{Capital Requirements under the IRB approach} \label{sec:cap_req}
+
+In June 2004, the Basel Committee issued a revised framework on international convergence of capital measurements and capital standards [@basel2004international], known as Basel II, which serves as the basis for national rulemaking and implementation processes.
+At the heart of the revision was the link of capital charges to asset risk.
+Since the first accord from 1988 (Basel I), minimum capital requirements are calculated as a percentage of risk-weighted assets.
+Hence, the approach to define risk-weights is of major importance.
+In Basel I, risk-weights were flat within portfolio categories.
+For instance, all sovereign exposures were weighted by 0% and all corporate exposures by 100%.
+However, regulating capital via a simple capital to asset ratio can incentivize banks to take on excessive risk as the cost in terms of capital requirements is the same within a portfolio category [@koehn1980regulation; @kim1988risk; @gordy2010risk].
+Therefore, since Basel II, financial institutions may choose between two approaches to calculate capital requirements for credit risk; the standardized approach (essentially a slightly modified version of the first accord) and the IRB approach.
 
 In the IRB approach, capital requirements are derived from risk-weight formulas that take as input banks' own estimates of credit risk parameters.
+The rationale for allowing internal models is the better alignment of risk estimates to actual portfolio risk [@barakova2014banks] and the promotion of better risk management practices [@cucinelli2018credit].
 The basic regulatory formula derives from a model characterized by the property that the risk-weight of each asset depends only on the estimated credit risk parameters for this asset and not on the composition of the portfolio, i.e.
@@
 This feature leads to a bottom-up approach where capital requirements are determined on the asset level and total requirement is simply the sum of asset individual requirement.
-
-Equation \ref{eq1} is the basic risk-weight formula:
+Equation \ref{eq:RW_function} is the risk-weight formula at its most simple form:
 
 \begin{equation}
-\label{eq1}
-\text{RW}= 12.5\frac{\sum_{i}\text{EAD}_i\text{LGD}_i N\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\text{PD}_i)+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)}{\sum_{i}\text{EAD}_i} ,
+\label{eq:RW_function}
+\text{RW}(\text{LGD}_i, \text{PD}_i)= 12.5\text{LGD}_i\bigg[ N\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\text{PD}_i)+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)-\text{PD}_i\bigg],
 \end{equation}
-
 where $i$ is an asset in the portfolio and $N$ is the standard normal cumulative distribution.
-Equation \ref{eq1} calculates the average RW for the porftolio.
-The total amount of RWA for this portfolio is $\text{RW}\times \sum_i\text{EAD}_i$ and total capital requirements $\text{RWA}\times 8\%$.
-
-Basel accords segment credit portfolio into several categories, each having different additional features (maturity adjustments, risk-dependent correlation coefficient, among others) while keeping the basic structure of equation \ref{eq1}.
-The credit risk parameters estimated by banks, which are used as input on equation \ref{eq1}, are: Exposure at default (EAD), loss given default (LGD), and probability of default (PD).
-
-## Probability of Default Distribution and Capital Savings
-
-Figure \ref{fig:pdmap} plots the mapping from PD to RW according to equation \ref{eq1}.
-For my analysis, the crucial characteristic is that this mapping is concave.
-The concavity of the PD-RW mapping together with the bottom-up method to aggregate capital requirements imply that actual capital requirements are lower than if they were calculated using the average PD to the entire portfolio.
+Basel accords segment credit portfolios into several categories (for instance, corporate, sovereign or retail), each having different additional features (maturity adjustments, risk-dependent correlation coefficient, among others) while keeping the basic structure of equation \ref{eq:RW_function}.
+Because of the portfolio invariant feature of the model, average risk-weight $\text{RW}^f$ is the exposure weighted average of all individual assets' risk-weights, as shown in equation \ref{eq:RW_factual}:
+
+\begin{equation}
+\label{eq:RW_factual}
+\text{RW}^f= \frac{\sum_{i}\text{EAD}_i \text{RW}(\text{LGD}_i, \text{PD}_i)}{\sum_{i}\text{EAD}_i}= \sum_{i}\text{q}_i \text{RW}(\text{LGD}_i, \text{PD}_i).
+\end{equation}
+where $q_i= EAD_i/\sum_jEAD_j$.
+Thus, to calculated their capital requirements banks estimate three credit risk parameters: the exposure at default (EAD), the loss given default (LGD), and the probability of default (PD).[^1] The total amount of RWA for this portfolio is $\text{RW}^f\times \sum_i\text{EAD}_i$ and total capital requirements $\text{RWA}\times 8\%$.
+
+[^1]: For wholesale exposures (corporate, sovereign and banks) the effective maturity of the asset is another parameter and for SMEs annual sales turnover is used to adjust the correlation coefficient.
+
+\subsection{Probability of Default Distribution and Capital Savings} \label{sec:PD_capital_savings}
+
+Figure \ref{fig:pdmap} plots the mapping from PD to RW according to equation \ref{eq:RW_function}.
+For this paper, the important characteristic of this mapping is that it is a concave function.
+The concavity of the PD-RW mapping together with the bottom-up method to aggregate capital requirements entail that actual capital requirements are lower than if they were calculated using the average PD to the entire portfolio.
 For clarity, consider two portfolios.
-The first contains a low PD asset and a high PD asset in equal proportions.
-The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD.
-Because of the concavity of the IRB formula, by Jensen’s inequality, the capital requirement for the second portfolio is greater than the first.
-More generally, total capital requirements depend not only on banks' portfolio average PD but also on the shape of the PD distribution.
+The first contains a low PD asset and a high PD asset in proportions $q$ and $1-q$.
+The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD weighted by their respective shares.
+Because of the concavity of the IRB formula and Jensen’s inequality, the capital requirement for the second portfolio is greater than the first.
 Figure \ref{fig:pdmap} illustrates the comparison.
-$\text{RW}^a$ and $\text{RW}^c$ are the average RW for the first portfolio and the second portfolio, respectively.
-The difference between  $\text{RW}^c$ and $\text{RW}^a$ is the monetary amount of *Capital savings* per unit of EAD due to the concavity of the IRB formula.
-
-I propose a method that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
-The method is a generalization of the example in Figure \ref{fig:pdmap}: for each bank-year combination, compare the actual capital requirement to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio’s average PD.
-
-I define the counterfactual RW as:
+$\text{RW}^f$ and $\text{RW}^c$ are the average RW for the first portfolio and the second portfolio, respectively.
+More generally, total capital requirements do not depend only on the portfolio average PD but also on the shape of the PD distribution.
+
+I propose a method that decomposes the contribution of the PD distribution to capital requirements into the average component and the shape component.
+The method is a generalization of the example in Figure \ref{fig:pdmap}: for each bank-year combination, I compare the actual capital requirement ($\text{RW}^f$) to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio’s average PD.
+The counterfactual RW is the average component of the PD distribution to capital requirements:
 
 \begin{equation}
-\label{eq2}
-\text{RW}^c= 12.5\frac{\sum_{i}\text{EAD}_i\text{LGD}_iN\bigg(\sqrt{\frac{1}{1-R}}N^{-1}(\overline{\text{PD}})+\sqrt{\frac{R}{1-R}}N^{-1}(0.999)\bigg)}{\sum_i \text{EAD}_i},
+\label{eq:RW_counterfactual}
+\text{RW}^c= \sum_{i}\text{q}_i \text{RW}(\text{LGD}_i,\overline{\text{PD}}),
 \end{equation}
-in which the PD parameter for all assets is replaced by the portfolio weighted average:
+where the PD parameter for all assets is replaced by the portfolio weighted average. The latter is defined as the following:
 
@@
 
-The main measure of *Capital savings* used in this paper is the ratio of equation \ref{eq2} to equation \ref{eq1} as follows:
+The difference between $\text{RW}^c$ and $\text{RW}^f$ measures any contribution to capital requirements from the PD distribution that is no related to the location of this distribution, i.e., it is a measure of the shape component.
+It can also be interpreted as the amount of *Capital savings* per unit of EAD due to the concavity of the IRB formula and the shape of the PD distribution.
+Equation \ref{eq:RW_savings} shows one of the possibilities to measure *Capital savings*:
+
+\begin{equation}
+\label{eq:RW_savings}
+\text{RW}^{s}=\text{RW}^c-\text{RW}^f.
+\end{equation}
+
+Note that I defined *Capital savings* such that positive values translate to lower actual capital requirements and consequently $\text{RW}^s \geq 0$.
+
+\subsection{Hypotheses on Capital Savings during the Business Cycle} \label{sec:hypotheses}
+
+Overall, capital requirements under the IRB approach are procyclical.
+During economic downturns credit risk increases which raises capital requirements forcing banks to either raise equity or cut down lending.
+Theory and empirical evidence suggest that banks opt for the latter as raising equity is also more expensive during recessions [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
+
+Between the two capital requirement components, the average component is a source of the procyclicality as long the relationship between the business cycle and PD is monotonic.
+Consider again the example shown in Figure \ref{fig:pdmap} with a portfolio containing a safe asset with PD equal to $PD_L$ and a risky asset with PD equal to $PD_H$ in proportions $q$ and $1-q$.
+In Appendix XX, I show that for $\text{PD} \in (0,0.25)$, $\text{RW}'>0$, $\text{RW}''<0$ and $\text{RW}'''>0$.
+I also show in Appendix XX that, on average, XX% of assets' PD lies within this range.
+For simplicity, I assume a constant $\text{LGD}_i=\text{LGD}$.
+So, the average component for this portfolio is:
+
+\begin{equation}
+\label{eq:RW_counter_example}
+\text{RW}^{c}(q,\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H).
+\end{equation}
+
+Because $\text{RW}'>0$ and as long PD estimates are affected monotonic and negatively by GDP growth, the average component of capital requirements increases during economic downturns.
+
+The direction of *Capital savings* during the business cycle is not trivial even when assuming a monotonic relationship between PD estimates and the business cycle.
+To explain the ambiguity, I consider three cases.
+But first, note that using the example in Figure \ref{fig:pdmap}, *Capital savings* can be written as the following:
+
+\begin{equation}
+\label{eq:RW_savings_example}
+\text{RW}^{s}(q,\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}(\text{PD}_L)+(1-q)\text{RW}(\text{PD}_H).
+\end{equation}
+
+We can find the effect of changes in the PD distribution to *Capital savings* by taking the total derivative of \ref{eq:RW_savings}:
+
+\begin{equation}
+\label{eq:RW_savings_diff}
+\begin{aligned}
+d\text{RW}^{s}(q,\text{PD}_L, \text{PD}_H)
+&=\bigg[q\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}'(\text{PD}_L)\bigg]d\text{PD}_L\\
+&+\bigg[(1-q)\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-(1-q)\text{RW}'(\text{PD}_H)\bigg]d\text{PD}_H\\
+&+\bigg[\text{RW}(\text{PD}_H)-\text{RW}(\text{PD}_L)-(\text{PD}_H-\text{PD}_L)\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]dq\\
+\end{aligned}
+\end{equation}
+
+Next, consider the case that all assets are affected equally during the business cycles in term of PD, i.e.,  $d\text{PD}_H=d\text{PD}_L=d\text{PD}$.
+Also, consider for now no portfolio reallocation, i.e., $dq=0$.
+In this scenario equation \ref{eq:RW_savings_diff} becomes:
+
+\begin{equation}
+\label{eq:first_case}
+\begin{aligned}
+d\text{RW}^{s}
+&=\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\bigg(q\text{RW}'(\text{PD}_L)+(1-q)\text{RW}'(\text{PD}_H)\bigg)\bigg]d\text{PD}.\\
+\end{aligned}
+\end{equation}
+
+Because $\text{RW}'$ is a convex function and Jensen's inequality, if $d\text{PD}_H=d\text{PD}_H=d\text{PD}$ then $d\text{RW}^{s}/d\text{PD}<0$.
+In this case, *Capital savings* are procyclical, decreasing when PD estimates increase, i.e., during economic downturns.
+Figure \ref{fig:pdmapcasesym} shows graphically this first case.
+
+Next, consider the case riskier assets are more affected during the business cycles in term of PD, i.e.,  $d\text{PD}_H>d\text{PD}_L$.For simplicity consider the extreme case of $d\text{PD}_H=d\text{PD}$ and $d\text{PD}_L=0$.
+In this scenario equation \ref{eq:RW_savings_diff} becomes:
+
+\begin{equation}
+\label{eq:second_case}
+\begin{aligned}
+d\text{RW}^{s}
+&=(1-q)\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_H)\bigg]d\text{PD}\\
+\end{aligned}
+\end{equation}
+
+Because $\text{RW}''<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H$, if $d\text{PD}_H=d\text{PD}$ and $d\text{PD}$ then $d\text{RW}^{s}/d\text{PD}>0$.
+In this case, *Capital savings* are countercyclical, increasing when PD estimates increase.
+In other words, the marginal capital requirement decreases during economic downturns.
+Figure \ref{fig:pdmapcaseassym} shows graphically this second case.
+
+Finally, consider the case where $d\text{PD}_H=d\text{PD}$, $d\text{RW}^{s}=0$ and $dq=0$. Using equation \ref{eq:RW_savings_diff}, we can find $d\text{PD}_L$ satisfying these conditions:
+
+\begin{equation}
+\label{eq:third_case}
+\begin{aligned}
+d\text{PD}_{L}
+&=d\text{PD}\frac{(1-q)}{q}\frac{\bigg[\text{RW}'(\text{PD}_H)-\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]}{\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_L)\bigg]}\\
+\end{aligned}
+\end{equation}
+
+Because $\text{RW}'<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H>\text{PD}_L$, both denominator and numerator of equation \ref{eq:third_case} are negative and consequently, $d\text{PD}_L>0$.
+From the first case we also know that $d\text{PD}_L<d\text{PD}$.
+Hence, the difference between the impact of the business cycle on riskier and safer assets have to be sufficiently large such that *Capital savings* are countercyclical.
+
+Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the shape of the PD distribution.
+
+\section{Data and Capital Savings Measures} \label{sec:data}
+
+The main measure of *Capital savings* used in this paper is the ratio of equation \ref{eq:RW_counterfactual} to equation \ref{eq:RW_factual} as follows:
 
@@
 $\text{RW}^r$ is the rate at which a bank saves capital.
-For instance, a value of 0.5 means this portfolio would require 50% more capital in the abscense of *Capital savings* to keep the same capital ratio.
-A value of zero means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
+For instance, a value of 1.5 means this portfolio would require 50% more capital in the absence of *Capital savings* to keep the same capital ratio.
+A value of one means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
 Any PD distribution with positive variance results in positive *Capital savings*.
@@
 
-An alternative way to measure *Capital savings* is in monetary units as follows:
-
-\begin{equation}
-\label{eq:measure3}
-\text{RW}^s=\frac{\text{RWA}^{\$}}{\text{RWA}}.
-\end{equation}
-
-# Data
+In the analysis I also consider the measure used in section \ref{sec:hypotheses} ($\text{RW}^s$), which is equivalent to dividing equation \ref{eq:measure2} by total EAD and therefore it is another measure of the intensity of *Capital savings*.
+Lastly, I calculate the Gini coefficient of the PD distribution as a robustness measure to my approach.
+According to the analysis shown in section \ref{sec:hypotheses} PD distributions with larger Gini coefficients should also generate larger *Capital savings*.
 
@@
 
-Figure \ref{fig:savingsbank} shows that the *Capital savings ratio* calculated with the compiled dataset.
+Figure \ref{fig:savingsbank} shows that the *Capital savings* calculated with the compiled dataset using the main measure \ref{eq:measure1} (minus one to improve readability).
 Each point is a bank-year observations and the bars indicates the average value across time for each bank.
-There is substancial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
+There is substantial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
 Variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
-*Capital savings ratio* also varies between and within countries.
+*Capital savings* also varies between and within countries.
 Figure \ref{fig:savingscountry}  aggregates the dataset by country.
@@
 
-# Capital Savings and the Business Cycle
+Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
+Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks].
+Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
+
+The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
+First, I control for bank size using the log of total assets (in millions of US dollars).
+The expectation regarding the impact of bank size is ambiguous.
+On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
+On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
+I also include the regulatory capital ratio as control.
+Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
+However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
+Thus, the expected effect of capital ratio on capital savings is also ambiguous.
+Next, I add banks' profitability using return on asset (ROA) and return on equity (ROE) with both measured before taxes.
+Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
+Finally, I study the relationship of *Capital savings* to risk-taking.
+I add the ratio of non-performing loans to total loans to test this proposition.
+
+\section{Capital Savings and the Business Cycle} \label{sec:cycle}
 
@@
 Next I show that *Capital savings* are countercyclical, i.e., given an average portfolio PD, banks need more capital in good times than in bad times to reach the same capital ratio.
-Finally, I provide evidence that this relationship is not purely mechanical but depends on how banks realocate their assets through the business cycle.
-
-## A simple theoretical framework {#theory}
-
-Overall, capital requirements are procyclical: during economic downturns credit risk in the economy increases which raises capital requirements, under the IRB approach forcing banks to cut down lending.
-However, as credit conditions changes (and banks react to it) the amount of *Capital savings* also changes.
-The expected direction of *Capital savings* cyclicality is ambiguous.
-On the one hand, if the entire portfolio is equally affected, capital savings are procyclical.
-On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
-The ambiguity arises because *Capital savings* can only increase, when credit conditions deteriorates, if the credit distribution widens.
-In the first situation, there is only a shift of the location parameters of the distribution and, consequently, a reduction of *Capital savings*.
-
-I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
-Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
-However, as credit conditions deteriorates (and banks react to these new credit conditions) the PD distribution and, consequently, the amount of *Capital savings* also change.
-A priori the direction of the economic cycle effect on *Capital savings* is ambiguous.
-At one extreme, if the entire portfolio is equally affected, then there is only a shift of the location parameter of the PD distribution.
-Because the PD-RW formula elasticity is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
-On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
-The ambiguity arises because *Capital savings* can only increase during a recession if the shape of the PD distribution is affect during the cycle.
-
-I use a simple example to illustrate this ambiguity.
-Consider a bank with the asset side of its balance sheet composed of \$100 of a safer asset and \$100 of a riskier asset and three states of the economy.
-Without loss of generality, assume $\text{LGD}=1$ and $\text{R}=0.15$.
-In the initial state, safer and riskier assets have a 0.1% and 0.2% probability of default, respectively.
-According to equation \ref{eq1}, the actual RWA of this bank is \$63.9.
-The weighted average of this portfolio in terms of $\text{PD}$ is 0.15% which yields a counterfactual RWA of \$64.7 according to equation \ref{eq2}.
-Hence, in the initial state *Capital savings* equal \$0.8.
-
-Consider now a credit risk shock.
-In state A, both safer and riskier assets are affected, having their PDs raised to 0.2% and 0.3%, respectively.
-In state B, only the riskier asset is affected, having its PD equal to 0.4%.
-Note that the average PD in both states are equal at 0.25%.
-However, relative to the initial state, the capital savings measure for state A decreases to \$0.45 while it increases to \$4.35 in state B.
-The mechanism of this example is simple: relative to the initial state, in state A only the location parameter of the PD distribution changes (standard deviation is equal to 0.05% in both states) while in state B also other moments change (standard deviation increases to 0.15%).
-Note that in both state A and B total capital requirements increase as average credit risk is higher relative to the initial state in both states.
-This net procyclical effect is in line with the results of @behn2016procyclical.
-However, due to the increased portfolio dispersion, capital savings are countercyclical in state B.
-Conversely, because the PD-RW formula elasticity is lower for higher PD estimates, in stata A *Capital savings* are procyclical.
-Therefore, capital savings are procyclical in state A.
-Table \ref{tab1} summarizes the example and Figure \ref{fig:pdmapcases}
-
-Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the PD distribution.
-
-## Empirical Results
+Finally, I provide evidence that this relationship is not purely mechanical but depends on how banks reallocate their assets through the business cycle.
 
@@
 The dependent variable is the yearly change in a variable associated to portfolio $j$ for bank $i$ at year $t$.
-The change in the logatithm of real GDP per capita used comes from the country where the bank has its headquarter.
+The change in the logarithm of real GDP per capita used comes from the country where the bank has its headquarter.
 All regressions include bank and year fixed effects and robust standard errors adjusted for clustering at the country level are reported in parentheses.
 
-Column 1 Table \ref{tab:main} shows results using equation \ref{eq:measure1}, my prefered measure of *Capital savings*.
+Column 1 Table \ref{tab:main} shows results using equation \ref{eq:measure1}, my preferred measure of *Capital savings*.
 shows that total *Capital savings* are countercyclical.
 According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average.
-The effect is economically sizeable representing 14% of total *Capital savings* standard deviation in the sample.
+The effect is economically sizable representing 14% of total *Capital savings* standard deviation in the sample.
 
@@
 
-## Robustness to Alternative Measures
+To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
+The capital ratio decomposition is shown in equation \ref{eq5}.
+The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
+To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
+Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab4} shows that *Capital savings* are the only countercyclical component of capital ratios.
+The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:decomposition}
+\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
+&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^r_{i,t} 
+\end{aligned}
+\end{equation}
+
+\subsection{Robustness to Alternative Measures and Samples} \label{sec:robustness}
 
@@
 This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity, risk-weights.
-Although the coefficient on the wholesale portfolio is no longer statiscally significant, total *Capital savings* and those coming from corporate portfolios are still statistically different from zero at the 1%  confidence level.
+Although the coefficient on the wholesale portfolio is no longer statistically significant, total *Capital savings* and those coming from corporate portfolios are still statistically different from zero at the 1%  confidence level.
 The attenuation of the effect on wholesale portfolio appears to come from large, albeit not significant, procyclical *Capital savings* in sovereign portfolios.
@@
 Lastly, in Panel C of Table \ref{tab:alt} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution.
-Again, the results shown in Panel C suppport the evidence from Table \ref{tab:alt}: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
-
-## Capital Savings and Portfolio Reallocation
+Again, the results shown in Panel C support the evidence from Table \ref{tab:alt}: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
+
+\section{Capital Savings and Portfolio Reallocation} \label{sec:reallocation}
 
 In this section I argue that portfolio reallocation during the business cycle explains part of the *Capital savings* cyclicality.
-As method, I use the observed movements on moments of the PD distribution during the business cycle that are not in line with the expected movements from a static portfolio as evidence in favor of a portfolio reallocation effect on *Capital savings*.[^1]
-
-[^1]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans.
+As method, I use the observed movements on moments of the PD distribution during the business cycle that are not in line with the expected movements from a static portfolio as evidence in favor of a portfolio reallocation effect on *Capital savings*.[^2]
+
+[^2]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PD estimates evolution to the PD estimates of new loans.
 Unfortunately, in my dataset, I only observe assets aggregated by PD bands without knowledge if amount changes in a band are due to new loans or changes in credit quality of assets already held.
@@
 
-Panel C in Table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle cannot be explained without portfolion reallocation.
+Panel C in Table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle cannot be explained without portfolio reallocation.
 All the statistical significant coefficients are positive, contrary to the expectation if banks wouldn't reallocate their portfolio.
@@
 
-## Economic Impact of Capital Savings Cyclicality
-
-To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
-The capital ratio decomposition is shown in equation \ref{eq5}.
-The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
-To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
-Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab4} shows that *Capital savings* are the only countercyclical component of capital ratios.
-The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
-
-\begin{equation}
-\begin{aligned}
-\label{eq5}
-\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
-&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^r_{i,t} 
-\end{aligned}
-\end{equation}
-
-# Determinants of Capital Savings
+\subsection{Selection and Reallocation Across Portfolios} \label{sec:cross_reallocation}
+
+\section{Additional Results} \label{sec:results}
 
@@
 
-# Conclusion and Discussion
+\section{Conclusion and Discussion} \label{sec:conclusion}
 
@@
 
-# References {-}
+\section*{References}
 
@@
 ```{r, results="hide"}
-pd_map_cap <-  "\\label{fig:pdmap}The figure plots the mapping from probabilities of default to capital requirementes per unit of exposure according to equation \\ref{eq1}.
-$EAD$ and $LGD$ are set equal to one."
+pd_map_cap <-  "\\label{fig:pdmap}The figure plots the mapping from probabilities of default to capital requirementes per unit of exposure according to equation \\ref{eq:RW_function}.
+$LGD$ is set equal to one and the correlation coefficient to 0.15."
 
@@
   geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=1)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
   geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=5)+
-  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=5)+
-  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
-               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.5)+
-
-  labs(x = "PD", 
-         y = "Risk-weight") +
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
+
   theme_bw()+
-  theme(axis.ticks = element_blank(),
+  theme(#axis.ticks = element_blank(),
         axis.line = element_line(),
-        axis.title.x = element_blank(),
+        axis.title.x = element_text( hjust = 1),
+        axis.title.y = element_text( hjust = 1, angle = 0),
         panel.grid = element_blank())+
   expand_limits(x = 0, y = 0)+
-  scale_x_continuous(breaks=c(PD.l,PD.a,PD.h,PD.i),labels=c("Low PD","Mean PD","High PD","PD"),expand = c(0, 0))+
-  scale_y_continuous("",breaks=c(K.s,K.a,K.i),labels=c(expression(paste(RW^{a})),expression(paste(RW^{c})),"RW"),expand = c(0, 0)) +
-  annotate("text", x = (PD.l+PD.a)/3 , y = (K.s+K.a)/2, label = "Capital \n Savings", size = 3)
+  scale_x_continuous(name= "PD", 
+                     breaks=c(PD.l,PD.a,PD.h),
+                     labels= unname(TeX(c("$PD_{L}", "$\\bar{PD}", "$PD_{H}"))),
+                     expand = c(0, 0))+
+  scale_y_continuous(name= "RW",
+                     breaks=c(K.s,K.a),
+                     labels=unname(TeX(c("$RW^{f}", "$RW^{c}" ))),
+                     expand = c(0, 0)) 
+```
+
+```{r parameters curve 2, echo=FALSE, include=FALSE}
+set.seed(123)
+LGD <- 1
+M   <- 2.5
+grid  <- seq(0.0001, 1, by=0.00001) 
+data <- mapping_retail(grid,1,0.15,0.999)
+PD.l <- 0.01
+PD.h <- 0.13
+PD.a <- (PD.l+PD.h)/2
+PD.h2 <- 0.19
+PD.h3 <- 0.16
+PD.l3 <- 0.04
+PD.a2 <- (PD.l+PD.h2)/2
+PD.l.a <- (PD.l+PD.l3)/2
+PD.h.a <- (PD.h+PD.h3)/2
+PD.h.a2 <- (PD.h+PD.h2)/2
+K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
+K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
+K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
+K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
+K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
+K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
+K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
+K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
+K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
+K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
+aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
+                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
+                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
+```
+
+```{r}
+cyclicality_cap_symm = "\\label{fig:pdmapcasesymm}The Figure shows the case where capital savings are procyclical.
+The dotted lines correspond to the potfolio before the credit risk shock (dPD) and the dashed lines to the portfolio after the the shock.
+The segmentes in the plot show the size of the capital savings before the shock."
+```
+
+```{r, cyclicalitysymm, fig.height = 3, fig.width = 4,results="hide", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap_symm}
+
+ ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K))+ 
+  geom_line(alpha = 1) + 
+  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5)+
+  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+  geom_segment(aes(x=PD.l, xend=PD.l, yend=K.a, y=K.s),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a+0.081, yend=K.s+0.081),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+  geom_segment(aes(x=PD.l, xend=PD.l, yend=K.a+0.081, y=K.s+0.081),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+
+  theme_bw()+
+  theme(#axis.ticks = element_blank(),
+        axis.line = element_line(),
+        axis.title.x = element_text( hjust = 1),
+        axis.title.y = element_text( hjust = 1, angle = 0),
+        panel.grid = element_blank())+
+  expand_limits(x = 0, y = 0)+
+ scale_x_continuous(name = "PD", breaks=c(PD.l.a,PD.h.a),labels=c("dPD","dPD"), expand = c(0, 0))+
+  scale_y_continuous(name = "RW", breaks = NULL, expand = c(0, 0)) +
+coord_flex_cart(bottom = brackets_horizontal(length = unit(0.075, "npc")))
+```
+```{r}
+cyclicality_cap_asym = "\\label{fig:pdmapcaseassym}The Figure shows the case where capital savings are countercyclical.
+The dotted lines correspond to the potfolio before the credit risk shock (dPD) and the dashed lines to the portfolio after the the shock.
+The segmentes in the plot show the size of the capital savings before the shock.
+"
+```
+
+```{r, cyclicalityasym, fig.height = 3, fig.width = 4,results="hide", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap_asym}
+
+ggplot(data = subset(data,subset = PD <0.2),
+            mapping = aes(x = PD, y=K))+ 
+  geom_line(alpha = 1) + 
+  geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
+  geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=5)+
+  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5)+
+  geom_segment(aes(x=PD.l, xend=PD.l, y=K.a, yend=K.s),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+  geom_segment(aes(x=PD.l, xend=PD.l, yend=K.a, y=K.s),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+  geom_segment(aes(x=PD.l+0.01, xend=PD.l+0.01, y=K.a+0.025, yend=K.s+0.025),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+  geom_segment(aes(x=PD.l+0.01, xend=PD.l+0.01, yend=K.a+0.025, y=K.s+0.025),
+               arrow = arrow(length = unit(0.15, "cm"),type = "closed"),linejoin='mitre',size = 0.1)+
+ theme_bw()+
+  theme(#axis.ticks = element_blank(),
+        axis.line = element_line(),
+        axis.title.x = element_text( hjust = 1),
+        axis.title.y = element_text( hjust = 1, angle = 0),
+        panel.grid = element_blank())+
+  expand_limits(x = 0, y = 0)+
+ scale_x_continuous(name = "PD",breaks=c(PD.h.a2),labels=c("dPD"), expand = c(0, 0))+
+  scale_y_continuous(name = "RW",breaks=NULL,expand = c(0, 0)) +
+  coord_flex_cart(bottom = brackets_horizontal(length = unit(0.15, "npc")))
 
@@
 savings_cap = "\\label{fig:savingsbank}The figure plots the log of banks capital savings.
-This measure is calculated using equation \\ref{eq3}.
+This measure is calculated using equation \\ref{eq:measure1}.
 Each point in the plot is a bank-year observation and bars are the averages values for each bank."
@@
     coord_flip() +
-  labs(y =  "RWA savings ratio") +
+  labs(y =  unname(TeX(c("$RWA^{s}-1" )))) +
   scale_fill_grey()+
@@
 savings_cap_country = "\\label{fig:savingscountry}The figure plots the banks capital savings ratio.
-This measure is calculated using equation \\ref{eq4}.
+This measure is calculated using equation \\ref{eq:measure1}.
 For each bank I used the average across time such that each point in the plot is a bank observation and bars are the averages values for each country."
@@
     labs(x=NULL) +
-  labs(y =  "RWA savings ratio") +
+  labs(y =  unname(TeX(c("$RWA^{s}-1" )))) +
   coord_flip() +
@@
 
-```
-
-```{r}
-cross.dec_cap1 = "\\label{fig:crossdecomposition1} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
-The figure and calculations include all banks in the sample.
-The first plot shows the total RW for each bank relative to the average total RW.
-Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
-Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
-```
-
-```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap1, include=TRUE}
- plot1 <- cross.section.decomposition %>% 
-  mutate(total= -PD.gain)%>%
-  select(name, Country,total, int.gain, ext.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition, levels = c("int.gain","ext.gain"),
-                                       labels = c("Intensive margin", "Extensive margin")))%>%
-ggplot(mapping = aes(x = reorder(name, total),
-                       y = value)) +
-geom_bar(stat = 'identity')+
-  #geom_line(aes(y = -total,group=1))  +
-  labs(x=NULL, y= NULL) +
-    coord_flip() +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  facet_wrap(~condition)
-
-plot2 <- cross.section.decomposition %>% 
-  select(name, Country, int.gain, ext.gain) %>%
-  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
-  ggplot(mapping = aes(x = int.gain,
-                       y = ext.gain)) +
-    geom_point()+
-    geom_smooth(method = "lm", color = "black")  +
-  labs(x=NULL, y= NULL) +
-   theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  scale_fill_grey()+
-  stat_cor()
-
-  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
-            ncol = 1, nrow = 2, align = "v", axis = "lr")
-```
-
-```{r}
-cross.dec_cap2 = "\\label{fig:crossdecomposition2} The figure shows the result of Equation \\ref{eq5} using the average bank as benchmark and the average across time for each bank.
-The figure and calculations include all banks in the sample.
-The first plot shows the total RW for each bank relative to the average total RW.
-Intensive margin is the contribution to total RW of increasing the capital savings while holding the portfolio composition fixed.
-Extensive margin is the contribution to total RW of increasing the share of the portfolio under the IRB approach while holding the intensive margin fixed."
-```
-
-```{r, fig.height = 7.9, fig.width = 6, results="hide",fig.cap= cross.dec_cap2, include=TRUE}
-plot1 <- cross.section.decomposition %>% 
-  mutate(total= -PD.gain)%>%
-  select(name, Country,total, other.gain, PD.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition, levels = c("PD.gain","other.gain"),
-                                       labels = c("Capital savings", "Other components")))%>%
-ggplot(mapping = aes(x = reorder(name, total),
-                       y = value)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL, y= NULL) +
-    coord_flip() +theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  facet_wrap(~condition)
-
-plot2 <- cross.section.decomposition %>%
-  select(name, Country, other.gain, PD.gain) %>%
-  mutate_if(is.numeric,Winsorize,probs = c(0.02, 1) ,na.rm = TRUE)%>% 
-  ggplot(mapping = aes(x = PD.gain,
-                       y = other.gain)) +
-    geom_point()+
-    geom_smooth(method = "lm", color = "black")  +
-  labs(x=NULL, y= NULL) +
-   theme_bw() +
-  theme(panel.grid.minor = element_blank())+
-  scale_fill_grey()+
-  stat_cor(label.y= -0.1)
-
-  plot_grid(plot1, plot2, rel_heights = c(3.5,1) ,hjust = 0.5,
-            ncol = 1, nrow = 2, align = "v", axis = "lr")
-```
-
-```{r}
-rsqrt_cap = "\\label{fig:rsqrtdecomp} The figure shows the $R^2$ share for each of the RW components derived using equation Equation \\ref{eq5}." 
-```
-
-```{r,rsqrt, fig.height = 4, fig.width = 5,out.width = "75%", results="hide", fig.cap= rsqrt_cap, include=TRUE,fig.align = 'center'}
-
-unrestrict <- cross.section.decomposition %>% 
-  mutate(epsilon = rnorm(nrow(.),0,0.0000001),
-         total.gain = total.gain+epsilon)%>% 
-  lm(data = ., total.gain ~ IRB.gain + mix.gain + rollout.gain + int.gain + ext.gain + SA.gain)%>%
-  relaimpo::calc.relimp(type = c("lmg"), rela = FALSE )
-
-  data.frame(share  = unrestrict@lmg) %>%
-   add_rownames(var = "component")%>%
-  mutate(component = factor(component,
-                      levels = c("IRB.gain", "ext.gain", "int.gain", "SA.gain", "mix.gain", "rollout.gain"),
-                      labels = c("PD avg.", "Ext. comp.", "Int. comp.", "SA", "Mix", "Rollout"))) %>%
-ggplot(mapping = aes(x = reorder(component, share),
-                       y = share, fill = component)) +
-geom_bar(stat = 'identity')+
-  labs(x=NULL, y = expression(paste("Share of ", R^{2}))) +
-    theme_bw()+
-  theme(panel.grid.minor = element_blank(),
-        strip.background = element_blank(),
-       # strip.text = element_blank()
-       legend.position = "none",
-        axis.text.x = element_text(hjust = 1))+
-  scale_fill_manual(values = c("#C3BCB2","#353535",  "#353535", "#C3BCB2","#C3BCB2", "#C3BCB2"))+
-scale_y_continuous(labels = scales::percent_format(accuracy = 1,
-                                 decimal.mark = '.')) +
-    coord_flip()
-
-```
-
-```{r}
-time.dec_cap = "\\label{fig:timedecomposition}The figure shows the evolution of average RW components in time.
-Plot a shows the average contribution of each RW component relative to the time of IRB introduction (t=0).
-The figure and calculations include only banks with observations for at least 9 consecutive years.
-Plot b shows the average contribution of each component relative to 2008.
-The figure and calculations include all banks in the sample.
-The black line in both plots is the average change in RW relative to the initial period."
-```
-
-```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
-
-plot1 <-time.decomposition %>% 
-  select(time, RW, t0.RW,  int.gain, ext.gain, other.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain")),
-         RW = RW - t0.RW,
-         time = as.factor(time)) %>%
-  ggplot(mapping = aes(x = time,
-                       y    = value, 
-                       fill = condition)) +
-geom_bar(stat = 'identity')+
-geom_line(aes(group=1, y = RW))  +
-  theme_bw()+
-  labs( y = 'Change from t=0') +
-  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
- theme(legend.position = c(0.3, 0.2), 
-        legend.title = element_blank(), 
-        legend.background=element_blank())+
-  scale_fill_grey(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")
-
-year.decomposition <- year.decomposition %>%
-  mutate(RW_2 = RW - t0.RW,
-         year = as.factor(year)) 
-plot2 <- year.decomposition %>% 
-  select(year,RW, int.gain, ext.gain, other.gain) %>%
-  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
-  mutate(condition = factor(condition,levels=c("int.gain","ext.gain","other.gain"))) %>%
-  ggplot(mapping = aes(x = year)) +
-geom_bar(aes(y = value, 
-          fill = condition),
-          stat = 'identity')+
-geom_line(data = year.decomposition, 
-          aes(group=1,y = RW_2))  +
-  theme_bw()+
-  labs(y = 'Change from 2008', x=NULL) +
-  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
-theme(legend.position='none')+
-  scale_fill_grey()
-grid.arrange(plot1, plot2, ncol=2)
-
-```
-
-```{r}
-time.var_cap = "\\label{fig:timevariables}The figure shows the time series of banks' capital ratio, portfolio average PD of their portfolio, and PD standard deviation of their portfolio.
-The black line is average value across all banks in the sample and the shaded area shows the first and third quartile."
-```
-
-```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= time.var_cap}
-
-mean_cl_quantile <- function(x, q=c(0.25, 0.75), na.rm=TRUE){
-  dat <- data.frame(y=mean(x, na.rm = TRUE),
-                    ymin = quantile(x, probs = q[1],na.rm = na.rm),
-                    ymax = quantile(x, probs = q[2],na.rm = na.rm))
-  return(dat)
-}
-scaleFUN <- function(x) sprintf("%.1f", x)
-
-plot1 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, capital.ratio) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=capital.ratio)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black") +
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'Capital Ratio (%)', x=NULL) + scale_y_continuous(labels=scaleFUN)
-
-plot2 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year,  mean.PD.Total) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=mean.PD.Total)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'Average PD (%)', x=NULL)
-
-plot3 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, var.PD.Total) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=sqrt(var.PD.Total))) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'SD PD (%)', x=NULL)
-
-plot4 <-bank.data %>% filter(year>2007)%>%
-  select(bvdid, year, pillarI_stringency) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year,y=pillarI_stringency)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(stat = 'summary',
-            fun.data = mean_cl_quantile,
-            color = "black")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'GDP growth (%)', x=NULL)
-
-plot_grid(plot1, plot2, plot3, plot4, hjust = 0.5,
-            ncol = 2, nrow = 2, align = "v", axis = "lr")
-```
-```{r test, fig.height = 5, fig.width = 7, results="hide",}
-bank.data %>% filter(year>2007, bvdid %in% pillar3.data$bvdid)%>%
-  select(bvdid, year, log.gdp_gr, Country, pillarI_stringency,cap_reg) %>%
-  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
-  ggplot(mapping = aes(x = year)) +
-  geom_line(mapping = aes(y=pillarI_stringency))+
-  geom_line(mapping = aes(y=cap_reg))+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  labs(y = 'GDP growth (%)', x=NULL) +
-  facet_wrap(~Country)
-```
-
-<!-- \begin{table}[!htbp] \centering  -->
-<!--   \caption{Cyclicality of Capital Savings}  -->
-<!--   \label{tab1}  -->
-<!-- \small  -->
-<!-- \begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3}}  -->
-<!-- \hline  -->
-<!-- \hline \\[-1.8ex]  -->
-<!-- \\[-1.8ex] & \multicolumn{3}{c}{States}  & \multicolumn{2}{c}{$\Delta$ Initial}  \\ \cline{2-4}\cline{5-6} -->
-<!-- \\[-1.8ex] & \multicolumn{1}{c}{Initial} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} \\  -->
-<!-- \hline \\[-1.8ex]  -->
-<!--  $\text{RWA}$  & 63.88 & 94.21 & 90.29 & 30.32 & 26.42 \\  -->
-<!--   $\text{RWA}^c$ & 64.65 & 94.65 & 94.65 & 29.96 & 29.96  \\  -->
-<!--   $\text{RWA}^s$ & 0.81 & 0.45 & 4.35 & -0.36 & 3.54  \\  -->
-<!--   \hline \\[-1.8ex] -->
-<!--  Avg. PD & 0.15\% & 0.25\% & 0.25\% &  &   \\  -->
-<!--   S.d. PD & 0.05\% & 0.05\% & 0.23\% &  &   \\  -->
-<!-- \hline \\[-1.8ex]  -->
-<!-- \end{tabular}  -->
-<!-- \end{table}  -->
-
-```{r parameters curve 2, echo=FALSE, include=FALSE}
-set.seed(123)
-LGD <- 1
-M   <- 2.5
-grid  <- seq(0.0001, 1, by=0.00001) 
-data <- mapping_retail(grid,1,0.15,0.999)
-PD.l <- 0.01
-PD.h <- 0.13
-PD.a <- (PD.l+PD.h)/2
-PD.h2 <- 0.19
-PD.h3 <- 0.16
-PD.l3 <- 0.04
-PD.a2 <- (PD.l+PD.h2)/2
-K.l  <- as.numeric(mapping_retail(PD.l,1,0.15,0.999)["K"])
-K.h  <- as.numeric(mapping_retail(PD.h,1,0.15,0.999)["K"])
-K.a  <- as.numeric(mapping_retail(PD.a,1,0.15,0.999)["K"])
-K.l3 <- as.numeric(mapping_retail(PD.l3,1,0.15,0.999)["K"])
-K.h2 <- as.numeric(mapping_retail(PD.h2,1,0.15,0.999)["K"])
-K.h3 <- as.numeric(mapping_retail(PD.h3,1,0.15,0.999)["K"])
-K.a2  <- as.numeric(mapping_retail(PD.a2,1,0.15,0.999)["K"])
-K.s  <- ((K.h-K.l)/(PD.h-PD.l))*PD.a+K.l-((K.h-K.l)/(PD.h-PD.l))*PD.l
-K.s2 <- ((K.h2-K.l)/(PD.h2-PD.l))*PD.a2+K.l-((K.h2-K.l)/(PD.h2-PD.l))*PD.l
-K.s3 <- ((K.h3-K.l3)/(PD.h3-PD.l3))*PD.a2+K.l3-((K.h3-K.l3)/(PD.h3-PD.l3))*PD.l3
-aux.df <- data.frame(PD.l = PD.l, PD.h = PD.h, PD.a = PD.a, PD.l3 = PD.l3, PD.a2,
-                     K.l = K.l, K.h = K.h, K.a = K.a, K.s = K.s, K.l3 = K.l3, K.h2 = K.h2, 
-                     K.h3 = K.h3, K.s3 = K.s3, K.s2 = K.s2)
-```
-
-```{r}
-cyclicality_cap = "\\label{fig:pdmapcases}The figure shows the two possible outcomes of a negative credit risk shock on a portfolio.
-The left-hand side plot shows the case where capital savings are procyclical.
-The right-hand side plot shows the case where capital savings are countercyclical.
-"
-```
-
-```{r, cyclicality, fig.height = 3, fig.width = 6,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= cyclicality_cap}
-
-p1 <- ggplot(data = subset(data,subset = PD <0.2),
-            mapping = aes(x = PD, y=K))+ 
-  geom_line(alpha = 1) + 
-  #geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
-  #geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
-  #geom_segment(aes(x = PD.l3, y = 0, xend = PD.l3, yend = K.l3), data = aux.df, linetype=5, colour = "#008EC6")+
-  #geom_segment(aes(x = PD.h3, y = 0, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5, colour = "#008EC6")+
-  geom_segment(aes(x = PD.l3, y = K.l3, xend = PD.h3, yend = K.h3), data = aux.df, linetype=5, colour = "#008EC6")+
-  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5, colour = "#008EC6")+
-  geom_segment(aes(x = PD.a2, y = K.s3, xend = 0, yend = K.s3), data = aux.df, linetype=5, colour = "#008EC6")+
-  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5, colour = "#008EC6")+
-  theme_bw()+
-  theme(axis.ticks = element_blank(),
-        axis.line = element_line(),
-        axis.title.x = element_blank(),
-        axis.title.y = element_blank(),
-        #axis.text.x = element_blank(),
-        panel.grid = element_blank())+
-  expand_limits(x = 0, y = 0)+
- scale_x_continuous(breaks=c(PD.h2),labels=c("PD"), expand = c(0, 0))+
-  scale_y_continuous(breaks=c(K.h2),labels=c("RW"),expand = c(0, 0))
-
-p2 <- ggplot(data = subset(data,subset = PD <0.2),
-            mapping = aes(x = PD, y=K))+ 
-  geom_line(alpha = 1) + 
-  #geom_segment(aes(x = PD.l, y = 0, xend = PD.l, yend = K.l), data = aux.df, linetype=3)+
-  #geom_segment(aes(x = PD.h, y = 0, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h, yend = K.h), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.a, xend = PD.a, yend = 0), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.s, xend = 0, yend = K.s), data = aux.df, linetype=3)+
-  geom_segment(aes(x = PD.a, y = K.a, xend = 0, yend = K.a), data = aux.df, linetype=3)+
-  #geom_segment(aes(x = PD.h2, y = 0, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5, colour = "#CC9933")+
-  geom_segment(aes(x = PD.l, y = K.l, xend = PD.h2, yend = K.h2), data = aux.df, linetype=5, colour = "#CC9933")+
-  geom_segment(aes(x = PD.a2, y = K.a2, xend = PD.a2, yend = 0), data = aux.df, linetype=5, colour = "#CC9933")+
-  geom_segment(aes(x = PD.a2, y = K.s2, xend = 0, yend = K.s2), data = aux.df, linetype=5, colour = "#CC9933")+
-  geom_segment(aes(x = PD.a2, y = K.a2, xend = 0, yend = K.a2), data = aux.df, linetype=5, colour = "#CC9933")+
- theme_bw()+
-  theme(axis.ticks = element_blank(),
-        axis.line = element_line(),
-        axis.title.x = element_blank(),
-        axis.title.y = element_blank(),
-        #axis.text.x = element_blank(),
-        panel.grid = element_blank())+
-  expand_limits(x = 0, y = 0)+
- scale_x_continuous(breaks=c(PD.h2),labels=c("PD"), expand = c(0, 0))+
-  scale_y_continuous(breaks=c(K.h2),labels=c("RW"),expand = c(0, 0))
-
-grid.arrange(p1, p2, ncol=2)
 ```
@@
   ungroup()%>%
+  mutate(year = as.character(year))%>%
   mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
-  group_by(bvdid)
+  group_by(bvdid)%>%
+  mutate(year = as.numeric(year))%>%
+  mutate_at(vars(matches("log.RWA.r")), .funs = list(n = ~sum(!is.na(.))))
 
@@
 
+```{r reg_contribution, eval=TRUE, echo=FALSE}
+# Variables in the model 
+# Dependent variables
+dep.vars2 <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat.Total_gr", "log.RWA.r.Total_gr")
+dep.text2 <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
+
+# Model formulas
+formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = filter(bank.data.win, !is.na(log.CAR_gr)))
+})
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+# Effect 
+effect.list  <-  lapply(reg.list, function(x){
+  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
+})
+effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
+effect.list <-  sprintf("%.1f \\%%",effect.list)
+
+fe.list.car <- list(c("Effect", effect.list),
+                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
+```
+
+```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          title = "Capital Savings Effect on Capital Ratio Procyclicality",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{r} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline  Each column uses one component of the capital ratio according to equation \\ref{eq:decomposition} as dependent variable.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+
+          dep.var.labels   = dep.text2,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list.car,
+          #keep            = c("log.gdp_gr"),
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab:contribution",
+          style            = "qje",
+          column.sep.width = "2pt",
+          #float.env        = "sidewaystable",
+          align            = TRUE)
+
+```
+
 \begin{table}[!htbp] \centering  
@@
 The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$.
-The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:measure3} in panel B.
+The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:RW_savings} in panel B.
 In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution.
@@
 
-ind.text <- c("$\\Delta \\text{Mean(PD}_{i,t}$)", "$\\Delta \\text{Var(PD}_{i,t}$)", "$\\Delta \\text{Skew(PD}_{i,t}$)", "$\\Delta \\text{Kurt(PD}_{i,t}$)")
+ind.text.moments <- c("$\\Delta \\text{Mean(PD}_{i,t}$)", "$\\Delta \\text{Var(PD}_{i,t}$)", "$\\Delta \\text{Skew(PD}_{i,t}$)", "$\\Delta \\text{Kurt(PD}_{i,t}$)")
 
@@
           dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
+          covariate.labels = ind.text.moments,
           #se               = se.list,
@@
 \end{equation*} \newline
-The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$.
-The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:measure3} in panel B.
-In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution.
+The dependent variable is the change in a PD distribution moment.
+In panel A the dependent variable is the mean PD, in panel B the variance of PD and in panel C the PD skewness.
 $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
@@
 \end{table}
-
-\begin{table}[!htbp] \centering
-\begin{tabular}{@{\extracolsep{1pt}}cccccc}
-\multicolumn{6}{c}{\parbox[c]{15.8cm}{\centering \tablename\ \thetable{} -- continued from previous page}}\\
-\hline \\
-\end{tabular}\\
-\footnotesize
-\begin{tabular}{@{\extracolsep{1pt}}cccccc}
-\multicolumn{6}{c}{Panel D: $\text{Y}=\Delta \text{Kurt(PD}_{i,t}\text{)}$} 
-\end{tabular}
-
-```{r reg_kurt, echo=FALSE}
-
-# Model formulas
-formula.text <- paste0(paste0("kurt.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_kurt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-table <- capture.output(
-{stargazer(reg.list,
-          table.layout     = "=c=d#-t-as-",
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          #se              = se.list,
-          add.lines        = fe.list,
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "tab2",
-          style            = "qje",
-          column.sep.width = "2pt",
-          align            = TRUE,
-          float            = FALSE)})
-
-table <- gsub("{-3}",
-              "{5.5}"
-              , table, fixed = T)
-
-cat(table)
-
-```
-
-\end{table}
-
-\newpage
-
-```{r reg_contribution, eval=TRUE, echo=FALSE}
-# Variables in the model 
-# Dependent variables
-dep.vars2 <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat.Total_gr", "log.RWA.r.Total_gr")
-dep.text2 <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
-
-# Model formulas
-formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, !is.na(log.CAR_gr)))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-# Effect 
-effect.list  <-  lapply(reg.list, function(x){
-  100*abs(x$coefficients)/abs(reg.list[[1]]$coefficients)
-})
-effect.list[5] <- 100*reg.list[[5]]$coefficients/reg.list[[1]]$coefficients
-effect.list <-  sprintf("%.1f \\%%",effect.list)
-
-fe.list <- list(c("Effect", effect.list),
-                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
-```
-
-```{r table_contribution, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          title = "Capital Savings Effect on Capital Ratio Procyclicality",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{r} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-All regressions include bank and year fixed effects.
-Robust standard errors adjusted for clustering at the country level are reported in parentheses.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text2,
-          covariate.labels = ind.text,
-          #se               = se.list,
-          add.lines        = fe.list,
-          #keep            = c("log.gdp_gr"),
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "tab:contribution",
-          style            = "qje",
-          column.sep.width = "2pt",
-          #float.env        = "sidewaystable",
-          align            = TRUE)
-
-```
 
@@
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{r}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{s}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
 \\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of bank control variables.
+$X_{i,t-1}$ is a vector of independent variables.
 All regressions include bank and year fixed effects.
@@
 \newpage
-
-# Appendix {-}
+\section*{Appendix A} \label{sec:appendixa}
 
@@
 \renewcommand{\thetable}{A\arabic{table}}
+
+\newpage
+\section*{Appendix B} \label{sec:appendixb}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{B\arabic{table}}
 
@@
   skew.PD.Banks_gr,
-  kurt.PD.Total_gr,
-  kurt.PD.Retail_gr,
-  kurt.PD.Wholesale_gr, 
-  kurt.PD.Corporate_gr, 
-  kurt.PD.Sovereign_gr, 
-  kurt.PD.Banks_gr,
   log.gdp_gr,
@@
               "\\hspace{1.0cm}$\\Delta$ Log PD skewness banks", 
-              "$\\Delta$ Log PD kurtosis total",
-              "\\hspace{0.5cm}$\\Delta$ Log PD kurtosis retail",
-              "\\hspace{0.5cm}$\\Delta$ Log PD kurtosis wholesale",
-              "\\hspace{1.0cm}$\\Delta$ Log PD kurtosis corporate",
-              "\\hspace{1.0cm}$\\Delta$ Log PD kurtosis sovereign", 
-              "\\hspace{1.0cm}$\\Delta$ Log PD kurtosis banks",
               "$\\Delta \\text{Log GDP}$",
@@
 
-## RW Decomposition {-}
-
-In this section I extend the decomposition method proposed by @cannata2012inside to study the relative contribution of *Capital savings* to total RW in time and across banks.
-I divide equation \ref{eq3} by EAD, which results in $\text{RW}_{IRB} =  \text{RW}^c +\text{RW}^s$.
-Total RW is the sum of IRB and SA portfolios RWs weighted by their respective shares, i.e., $\text{RW}=(1-q_{SA})\text{RW}_{IRB} + q_{SA}\text{RW}_{SA}$, where $q_{SA}$ is the share of the credit portfolio under the standardised approach.
-I further breakdown RW_{IRB} into its portfolio components, $\text{RW}_{IRB}=\sum_j q_{j}\text{RW}_{j}$, in which the $j$ subscript stands for the different portfolios under the IRB approach: Wholesale, retail and equity.
-Equation \ref{eq5} shows the final decomposition:
-
-\begin{equation}
-\begin{aligned}
-\label{eq5}
-\Delta \text{RW} & = q^{B}_{SA} \Delta \text{RW}_{SA}+(1-q^{B}_{SA})\sum_{j} q^{B}_{j}\Delta \text{RW}^{c}_j  + (1-q^{B}_{SA})\sum_{j} \text{RW}_{j}^{c} \Delta q_{j}  + (\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}\\
-&- \underbrace{(1-q^{B}_{SA})\Delta \text{RW}^{s}}_\text{Intensive margin}+ \underbrace{\text{RW}^{s}\Delta q_{SA}}_\text{Extensive margin},
-\end{aligned}
-\end{equation}
-in which  $\Delta$ stands for changes relative to a benchmark.
-For the cross-section decomposition, the benchmark is the average bank.
-For the time decomposition, the benchmark is either the year when the IRB approach was first introduced at the bank level or 2008.
-The superscript $B$ stands for benchmark.
-For instance, $q_SA^B$ is the share of the portfolio under the standardized approach for the benchmark.
-
-I decompose the variation of RW relative to the benchmark in six components, two of them relating to *Capital savings*.
-The first, $(1-q^{B}_{SA})\Delta \text{RW}^{s}$, measures the intensity of *Capital savings*.
-That is, I fix the share of the portfolio under the IRB approach ($q_{SA}^B$) and variation in the component comes only from changes of $RW^s$ relative to the benchmark.
-The second component, $\text{RW}^{s}\Delta q^{B}_{SA}$, measures the extensive margin of *Capital savings*, i.e., variation in this component comes only from changes of the share under the IRB approach relative to the benchmark.
-Note that the magnitude of the second component is moderated by how intense *Capital savings* are in absolute terms.
-
-The remaining components are extensively discussed by @cannata2012inside.
-The first on the right-hand side of equation \ref{eq5} is the intensity of RW under the standardized approach ($q^{B}_{SA} \Delta \text{RW}_{SA}$).
-The contribution of this component to $\Delta \text{RW}$ comes from variaton of RW in portfolios under the standardized approach while fixing the share of this portfolio for all banks.
-The second term measures the intensity of RW under the IRB approach excluding the *Capital savings*, or differently, the intensity of the average PD ($(1-q^{B}_{SA})\sum_{j=1}^{P} q^{B}_{j}\Delta \text{RW}^{c}_j$).
-The contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of RW in the three portfolios under the IRB approach relative to the respective benchmark RW while holding constant the share of each IRB portfolio and the share of total IRB.
-The third component measure the contribution of the IRB portfolio mix.
-Because some types of portfolios have on average lower RW, the composition within IRB across types of portfolios matter to how large overall RW is.
-For instance retail portfolios tend to have lower RW relative to wholesale portfolios because the former usually are colletarized loans, such as mortgages, while the latter is mostly composed of corporate loans.
-Hence, the contribution of this component to $\Delta \text{RW}$ comes from the sum of the variaton of the three portfolios shares within the IRB portfolio relative to the respective benchmarks while holding constant the respective intensity of average PD ($RW_j^c$) and the share of total IRB.
-Lastly, $(\text{RW}_{SA}-\text{RW}^{c})\Delta q_{SA}$ is the roll-out effect, i.e., the contribution to $\Delta \text{RW}$ coming from moving the portfolio from the standardized approach to the IRB approach.
-
-Figure \ref{fig:crossdecomposition1} plots the intensive and extensive margin of each bank in my sample using the average bank as benchmark.
-The scatter plot shows that the correlation between the two *Capital savings* components is not statistically significant.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
-If Goldman Sachs is included the correlation becomes statistically significant.} The figure suggests that banks benefit from the two components separately, some benefting more of the extensive margin, such as Northern Trust, and others more of the intensive margin, such as BayernLB.
-
-Figure \ref{fig:crossdecomposition2} shows that the net effect of the intensive and extensive margin is negatively correlated to the net effect of the other components.\footnote{Goldman Sachs is considered an outlier and excluded of the scatter plot in figure \ref{fig:crossdecomposition1}.
-If Goldman Sachs is included the correlation becomes even more negative.} This correlation implies that banks with higher average portfolio riskness benefit the most from *Capital savings*.
-
-As Figure \ref{fig:rsqrtdecomp} shows *Capital savings* components account for 27% of the RW variation across banks.
-Each component accounts for around 13.5% of the RW variation and are second and third in rank, losing only to the average PD contribution.
-I derive the $\text{R}^2$ contribution of each RW component using the Shapley-Owen decomposition method.
-This method consider the contribution of each variable to total $\text{R}^2$ as the sum of $\text{R}^2$ from regressions using all possible variable combinations that include the variable of interest and weighting each regression by the number of variable permutations.
-In sum, figure \ref{fig:rsqrtdecomp} offers strong evindence of the relative importance of *Capital savings* as source of variation to RW across banks.
-
-Figure \ref{fig:timedecomposition} shows the evolution of RW components relative to the time of IRB introduction (left-hand side plot), and relative to 2008 (right-hand side plot).
-The negative values for the extensive margin implies that banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach.
-However, during the same period, the intensity of *Capital savings* has, on average, decreased in time.
-That is, for the same average PD, capital requirements have increased from 2008 to 2018 and since the year IRB approach was introduced.
-
-The follow-up to the 2008 financial crisis offer an plausible explanation for the dynamics of *Capital savings* in time, which is illustrated in Figure \label{fig:timevariables}.
-Increased supervisory pressure to increase capital ratios (top plot), especially after the Basell III, led banks to deleverage and reduce risk-taking.
-[^2] The shift for safer assets not only reduced the average PD on banks' portfolios (middle plot) but also compressed their PD distribution (bottom plot).
-This distribution compression leads to a reduction in *Capital savings*.
-This explanation is in line with the hypothesis that *Capital savings* are countercyclical.
-Efforts to increase capital ratio by risk reduction are increasingly more difficult.
-
-[^2]: Although supervisors favored raising additional capital as means to achieve the required capital ratios [@EBA2011recommendation], banks respond to capital requirements mostly by reducing RWA [@gropp2018banks].
+```{r reg_constant, echo=FALSE}
+
+reg.list <- list()
+# Run the models
+for(i in dep.vars){
+formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ log.gdp_gr | year + bvdid | 0 | Country "))
+ data = filter(bank.data.win,
+               get(paste0("log.RWA.r"  , i, "_gr_n"))>9)%>%
+   felm(data = ., formula = formula.text ) -> reg.list[[i]]
+  }
+
+# Robust standard errors
+se.list  <-  lapply(reg.list, function(x){
+  summary(x, robust=TRUE)$coefficients[,2]
+})
+
+```
+
+```{r table_constant, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+stargazer(reg.list,
+          title = "Capital Savings and the Business Cycle",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$.
+In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          #se               = se.list,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab:constant",
+          style            = "qje",
+          column.sep.width = "1pt",
+          align            = TRUE)
+```

### 2020-05-10 Analysis/Draft/Draft_bookdown.Rmd
Several changes to the draft text. Version before I started the RW-lending project

@@
 As is common in the literature, I calculate *Capital savings* in terms of risk-weighted assets (RWA), which are simply capital requirements times 12.5, or risk-weights (RW), which are RWA divided by assets.
-The sample consists of 52 large banking groups that have adopted the IRB approach.
-The dataset covers 13 countries and goes from 2007 to 2018.
+The sample consists of 59 large banking groups that have adopted the IRB approach.
+The dataset covers 15 countries and goes from 2007 to 2018.
 Data at such granularity is not available in the common balance sheet datasets, and although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
 
-With this novel dataset, I start by documenting the relative importance of *Capital savings* to banks' overall RW.
-By extending the decomposition method proposed by @cannata2012inside, I find that *Capital savings* accounts for 27% of the RW variation across banks.
-I also find that *Capital savings* measures are negatively correlated to other RW components.
-This correlation implies that banks with higher average portfolio riskiness benefit the most from *Capital savings*.
-Lastly, although banks benefited from higher *Capital savings* as they moved their portfolio from the standardized to the IRB approach (known as the roll-out effect), the intensity of *Capital savings* has, on average, decreased in time.
-That is, for the same average PD, capital requirements have increased from 2008 to 2018.
-
-I then extend the analysis to investigate the relationship between *Capital savings* and the business cycle.
+With this novel dataset, I investigate the relationship between *Capital savings* and the business cycle.
 Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
@@
 I find *Capital savings* to be countercyclical.
-I estimate that a 1.7% GDP growth rate (one standard deviation) decreases *Capital savings* by 0.83 percentage points on average.
-In other words, banks have to increase capital by 0.83 percentage points to keep their capital ratio constant.
-The effect is economically significant as it represents a quarter of the capital ratio standard deviation in the sample.
-In a capital ratio decomposition exercise, I also find that the the cyclicality of *Capital savings* reduces the procyclicality of capital ratios by 11.6%.
+I estimate that a 1.6% GDP growth rate (one standard deviation) decreases *Capital savings* by 0.7 percentage points on average.
+In other words, banks have to increase capital by 0.7 percentage points to keep their capital ratio constant.
+The effect is also economically significant.
+In a capital ratio decomposition exercise, I also find that the the cyclicality of *Capital savings* reduces the procyclicality of capital ratios by 23.7%.
 The results are robust to different measures of the shape component, such as, *Capital savings* in terms of ratios, *Capital savings* in monetary units, and using Gini coefficients.
 
-I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic expansions average PD and PD variance decrease while the PD kurtosis and PD skewness increase.
+I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic expansions average PD and PD variance decrease while the PD skewness increase.
 This evidence implies that distributions become more symmetric and with less extreme values during recessions, suggesting a flight to safety behavior by banks.
@@
 The remainder of this paper is organized as follows.
-Section 2 describes the capital requirement formula, the *Capital savings* measure, and the hypothesis concerning this measure.
-Section 3 describes the collected data and analyses the relative importance of *Capital savings* in explaining differences in risk-weights across banks and time.
-Section 4 presents evidence on the countercyclical feature of *Capital savings*.
-Finally, section 5 concludes.
+Section \ref{sec:background describes the capital requirement formula under the IRB approach, why it creates *Capital savings*, and the hypothesis concerning the cyclicality of *Capital savings*.
+Section \ref{sec:data} describes the collected data and the measures of *Capital savings*.
+Section \ref{sec:cycle} presents evidence on the countercyclical feature of *Capital savings*.
+Section \ref{sec:reallocation} tests if *Capital savings* are affected by portfolio reallocation during the business cycle.
+Section \ref{sec:additional_results} presentes additional results.
+Finally, section \ref{sec:conclusion} concludes.
+
 \section{Institutional Background and Hypotheses} \label{sec:background}
@@
 Consider again the example shown in Figure \ref{fig:pdmap} with a portfolio containing a safe asset with PD equal to $PD_L$ and a risky asset with PD equal to $PD_H$ in proportions $q$ and $1-q$.
-In Appendix XX, I show that for $\text{PD} \in (0,0.25)$, $\text{RW}'>0$, $\text{RW}''<0$ and $\text{RW}'''>0$.
-I also show in Appendix XX that, on average, XX% of assets' PD lies within this range.
+In Appendix \ref{sec:appendixa}, I show that for $\text{PD} \in (0,0.25)$, $\text{RW}'>0$, $\text{RW}''<0$ and $\text{RW}'''>0$.
+I also show in \ref{sec:appendixa} that, on average, XX% of assets' PD lies within this range.
 For simplicity, I assume a constant $\text{LGD}_i=\text{LGD}$.
@@
 
-Because $\text{RW}'$ is a convex function and Jensen's inequality, if $d\text{PD}_H=d\text{PD}_H=d\text{PD}$ then $d\text{RW}^{s}/d\text{PD}<0$.
+Because $\text{RW}'$ is a convex function and Jensen's inequality, if $\text{dPD}_H=\text{dPD}_H=\text{dPD}$ then $\text{dRW}^{s}/\text{dPD}<0$.
 In this case, *Capital savings* are procyclical, decreasing when PD estimates increase, i.e., during economic downturns.
-Figure \ref{fig:pdmapcasesym} shows graphically this first case.
-
-Next, consider the case riskier assets are more affected during the business cycles in term of PD, i.e.,  $d\text{PD}_H>d\text{PD}_L$.For simplicity consider the extreme case of $d\text{PD}_H=d\text{PD}$ and $d\text{PD}_L=0$.
+Figure \ref{fig:pdmapcasesymm} shows graphically this first case.
+
+Next, consider the case riskier assets are more affected during the business cycles in term of PD, i.e.,  $\text{dPD}_H>\text{dPD}_L$.For simplicity consider the extreme case of $\text{dPD}_H=\text{dPD}$ and $\text{dPD}_L=0$.
 In this scenario equation \ref{eq:RW_savings_diff} becomes:
@@
 \begin{aligned}
-d\text{RW}^{s}
+\text{dRW}^{s}
 &=(1-q)\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_H)\bigg]d\text{PD}\\
@@
 
-Because $\text{RW}''<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H$, if $d\text{PD}_H=d\text{PD}$ and $d\text{PD}$ then $d\text{RW}^{s}/d\text{PD}>0$.
+Because $\text{RW}''<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H$, if $\text{dPD}_H=\text{dPD}$ and $\text{dPD}$ then $\text{dRW}^{s}/\text{dPD}>0$.
 In this case, *Capital savings* are countercyclical, increasing when PD estimates increase.
@@
 
-Finally, consider the case where $d\text{PD}_H=d\text{PD}$, $d\text{RW}^{s}=0$ and $dq=0$. Using equation \ref{eq:RW_savings_diff}, we can find $d\text{PD}_L$ satisfying these conditions:
+Finally, consider the case where $\text{dPD}_H=\text{dPD}$, $\text{dRW}^{s}=0$ and $\text{d}q=0$. Using equation \ref{eq:RW_savings_diff}, we can find $\text{dPD}_L$ satisfying these conditions:
 
@@
 \begin{aligned}
-d\text{PD}_{L}
-&=d\text{PD}\frac{(1-q)}{q}\frac{\bigg[\text{RW}'(\text{PD}_H)-\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]}{\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_L)\bigg]}\\
+\text{dPD}_{L}
+&=\text{dPD}\frac{(1-q)}{q}\frac{\bigg[\text{RW}'(\text{PD}_H)-\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]}{\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_L)\bigg]}\\
 \end{aligned}
@@
 
-Because $\text{RW}'<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H>\text{PD}_L$, both denominator and numerator of equation \ref{eq:third_case} are negative and consequently, $d\text{PD}_L>0$.
-From the first case we also know that $d\text{PD}_L<d\text{PD}$.
+Because $\text{RW}'<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H>\text{PD}_L$, both denominator and numerator of equation \ref{eq:third_case} are negative and consequently, $\text{dPD}_L>0$.
+From the first case we also know that $\text{dPD}_L<\text{dPD}$.
 Hence, the difference between the impact of the business cycle on riskier and safer assets have to be sufficiently large such that *Capital savings* are countercyclical.
@@
 To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
-The capital ratio decomposition is shown in equation \ref{eq5}.
+The capital ratio decomposition is shown in equation \ref{eq:decomposition}.
 The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
 To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
-Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab4} shows that *Capital savings* are the only countercyclical component of capital ratios.
+Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab:contribution} shows that *Capital savings* are the only countercyclical component of capital ratios.
 The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
@@
 
-In panel B of Table \ref{tab:alt} the measure of *Capital savings* in US dollars is rescaled by the amount of EAD in the respective portfolio following equation \ref{eq:measure3}.
+In panel B of Table \ref{tab:alt} the measure of *Capital savings* in US dollars is rescaled by the amount of EAD in the respective portfolio following equation \ref{eq:RW_savings}.
 This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity, risk-weights.
@@
 According to the theoretical framework presented in section 4.1, an increase in average PD reduces the amount of *Capital savings* if other moments are kept constant.
-This relationship is illustrated in panel A of Figure \ref{fig:pdmapcases} and empirically tested in Table \ref{tab:moments_contribution}.
+This relationship is illustrated in Figure \ref{fig:pdmapcasesymm} and empirically tested in Table \ref{tab:moments_contribution}.
 In all columns of this table average PD is negatively associated with *Capital savings*.
@@
 Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness and kurtosis.
-It is trivial that the equivalent case with a continuous PD distribution to the scenario in panel A of figure \ref{fig:pdmapcases} wouldn't affect either of these moments.
+It is trivial that the equivalent case with a continuous PD distribution to the scenario in Figure \ref{fig:pdmapcasesymm} wouldn't affect either of these moments.
 On the other hand, in the second scenario the PD distribution should become more skewed to the right.
@@
 
-\section{Additional Results} \label{sec:results}
+\section{Additional Results} \label{sec:additional_results}
 
@@
 *Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are economically relevant.
-They explain 27% of the RW variation across banks.
-*Capital savings* are also countercyclical and reduce the procyclicality of RWA by 11.6%.
+*Capital savings* are also countercyclical and reduce the procyclicality of RWA by 23.7%.
 Hence, the concavity of IRB formulas could be explored to reduce procyclicality of capital regulation.
@@
 
+p_w <- c(0.01, 0.99)        
+bank.data.win2 <- bank.data %>% 
+  ungroup()%>%
+  mutate(year = as.character(year))%>%
+  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
+  group_by(bvdid)%>%
+  mutate(year = as.numeric(year))%>%
+  mutate_at(vars(matches("log.RWA.r")), .funs = list(n = ~sum(!is.na(.))))
+
 ```
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win, !is.na(log.CAR_gr)))
+  felm(x, data = filter(bank.data.win2, !is.na(log.CAR_gr)))
 })
@@
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{r} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{s} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
 \\end{equation*} \\newline  Each column uses one component of the capital ratio according to equation \\ref{eq:decomposition} as dependent variable.
@@
 \begin{tabular}{@{\extracolsep{1pt}}cccccc}
-\multicolumn{6}{c}{Panel A: $\text{Y}=\Delta \text{Mean(PD}_{i,t}\text{)}$} \\
+\multicolumn{6}{c}{Panel A: $\text{Y}=\Delta \text{Log mean(PD}_{i,t}\text{)}$} \\
 \end{tabular}
@@
 \\
-\multicolumn{6}{c}{Panel B: $\text{Y}=\Delta \text{Var(PD}_{i,t}\text{)}$}
+\multicolumn{6}{c}{Panel B: $\text{Y}=\Delta \text{Log var(PD}_{i,t}\text{)}$}
 \end{tabular}
@@
 \\
-\multicolumn{6}{c}{Panel C: $\text{Y}=\Delta \text{Skew(PD}_{i,t}\text{)}$}
+\multicolumn{6}{c}{Panel C: $\text{Y}=\Delta \text{Log skew(PD}_{i,t}\text{)}$}
 \end{tabular}
@@
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{s}_{i,t} = \\gamma X_{i,t-1}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{s}_{i,t} = \\gamma X_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
 \\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.

### 2020-05-10 Analysis/Draft/Draft_bookdown.Rmd

@@
 Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab:contribution} shows that *Capital savings* are the only countercyclical component of capital ratios.
-The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
+The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 23.7\%.
 
@@
 *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
           dep.var.labels   = dep.text2,

### 2020-07-17 Data/Clean.R
Version of clean file before first revision of draft.

@@ -176,8 +176,8 @@ for(i in c("Wholesale", "Retail", "Equity", "Corporate", "Sovereign", "Banks")) 
     full_join(sample, by=c("bvdid", "year"))%>%
     mutate(!!(paste("RW" ,"hat", i, sep=".")) := get(paste("RWA","hat", i , sep="."))/get(paste("EAD", i , sep=".")),
            !!(paste("RWA","s"  , i, sep=".")) := get(paste("RWA","hat", i, sep="."))-get(paste("RWA", i , sep=".")),
-           !!(paste("RWA","r"  , i, sep=".")) := get(paste("RWA","hat", i, sep="."))/get(paste("RWA", i , sep="."))-1,
-           !!(paste("RWA","c"  , i, sep=".")) := get(paste("RWA","bar", i, sep="."))/get(paste("RWA","til" ,i , sep="."))-1,
+           !!(paste("RWA","r"  , i, sep=".")) := get(paste("RWA","hat", i, sep="."))/get(paste("RWA", i , sep=".")),
+           !!(paste("RWA","c"  , i, sep=".")) := get(paste("RWA","bar", i, sep="."))/get(paste("RWA","til" ,i , sep=".")),
            !!(paste("RW" ,"s"  , i, sep=".")) := ifelse(get(paste("EAD", i, sep=".")) == 0, 0,
                                                  get(paste("RWA","s", i , sep="."))/get(paste("EAD", i , sep=".")))) %>% 
     select(bvdid, year, contains(i)) %>%
@@ -225,7 +225,8 @@ bank.data <- bank.data%>%
   # mutate(survey = ifelse(year<2001, 1, 
   #                        ifelse(between(year, 2001, 2004), 2, 
   #                               ifelse(between(year, 2005, 2007), 3, 4))))%>%
-  left_join(BRSS, by = c("Country", "year"))
+  left_join(select(BRSS, Country, cap_reg, Sup_Power, year, pillarI_stringency), by = c("Country", "year"))%>%
+  left_join(distinct(select(BRSS, Country, country.name), country.name, .keep_all = TRUE ), by = c("Country"))
 
 #------------------------------------------------------------------------------#
 # Basic cleaning and variable calculation                                      #
@@ -300,7 +301,7 @@ bank.data <- bank.data %>% ungroup() %>% rowwise() %>%
           Zscore     = (r.mean.ROA + r.mean.CAR)/r.sd.ROA)%>%
    # Growth rates
    ungroup() %>% group_by(bvdid) %>%
[…]

### 2020-07-17 Analysis/Draft/Draft_bookdown.Rmd
Draft version of first revision by supervisors

--- 2020-05-10 Analysis/Draft/Draft_bookdown.Rmd
@@
     fig_caption: yes
+    #citation_package: natbib
     latex_engine: pdflatex
@@
 citecolor: black
-csl: ../../Auxiliar/cambridge-journal-of-economics.csl
+csl: ../../Auxiliar/Citation_styles/american-marketing-association.csl
 #anonymous: no
@@
 geometry: margin= 1in
-header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float, siunitx, tabularx}
+header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float, siunitx, tabularx, appendix} 
 keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
@@
 spacing: double
+indent: true
 subtitle: ''
@@
 library(cowplot)              ## Create grid for plots and tables
+library(multipanelfigure)     ## Create grid for figures
 library(DescTools)            ## Command to winsorize variables
@@
 library(lemon)                ## Plot brackets
+library(mosaic)               ## Package to calculate derivatives
 #library(textablr)
@@
 The regulation on bank capital requirements has evolved substantially since the first international standard, the 1988 Basel Accord.
-At the heart of these changes is the link of capital charges to asset risk [@basel2004international; @BCBS2017basel; @EBA2019policy].
+At the heart of these changes is the link of capital charges to asset risk [@BCBS2006international; @BCBS2017basel; @EBA2019policy].
 Risk sensitivity of capital requirements was present in the first accord and it was greatly enhanced in the Basel II framework by giving banks the option to calculate capital requirements using risk estimates obtained from their own models, known as the internal ratings-based (IRB) approach.
@@
 However, both risk sensitivity and the use of internal models may have adverse consequences.
-First, linking capital charges to asset risks exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
-As the overall credit quality deteriorates during economic downturns, capital requirements tighten, forcing banks to cut down lending.
-Second, the inherent flexibility of the IRB approach may allow for differences in capital requirements that do not reflect portfolio risk but rather modelling choices [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits;@berg2017analysis; @ferribank].
+First, linking capital charges to asset risk exacerbate the procyclicality of lending [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
+As the overall credit quality deteriorates during economic downturns, capital requirements tighten, forcing banks to reduce lending.
+Second, the inherent flexibility of the IRB approach may allow for differences in capital requirements across banks that do not reflect portfolio risk but rather modelling choices [@le2012revisiting; @mariathasan2014manipulation; @behn2016limits;@berg2017analysis; @ferribank].
 
 This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements.
-Under the IRB approach, banks supply their own risk estimates, such as probability of default (PD), to formulas set by the regulation which determine the amount of capital banks have to hold.
+Under the IRB approach, banks supply their own risk estimates, such as the probability of default (PD), to formulas set by the regulation which determine the amount of capital banks have to hold.
 Two characteristics of the framework are important for this analysis.
 First, the mapping from PD estimates to capital requirements is a concave function.
-Second capital requirements are calculated for each asset separately and then added up to reach total capital charges.
-Together, these characteristics imply that actual capital requirements are lower than if they were calculated using the average PD to the entire portfolio.
+Second, capital requirements are calculated for each asset separately and then added up to obtain the total capital charge.
+Because of these characteristics, actual capital requirements for a portfolio are lower than implied by the average PD of this portfolio.
 For clarity, consider two portfolios.
@@
 The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD.
-Because of the concavity of the IRB formula, capital requirements for the second portfolio are greater than the first.
+Because of the concavity of the IRB formula, capital requirements for the second portfolio are greater than capital requirements for the first.
 More generally, total capital requirements depend not only on banks' portfolio average PD but also on the shape of the PD distribution.
@@
 I propose a measure that decomposes the contribution of the PD distribution to capital requirements into an average component and a shape component.
-The method is straightforward: for each bank-year combination, compare actual capital requirements to the counterfactual case where requirements for the entire portfolio are calculated using the portfolio’s average PD.
-I label the shape component *Capital savings* because it reduces total capital requirements.
-
-To calculate measures of *Capital savings*, information on the distribution of banks' regulatory risk parameters is needed.
-I hand-collected information on the distribution of credit risk parameters from banks' Pillar-3 reports.
-As is common in the literature, I calculate *Capital savings* in terms of risk-weighted assets (RWA), which are simply capital requirements times 12.5, or risk-weights (RW), which are RWA divided by assets.
+The average component is a counterfactual scenario such that the capital requirements for the entire portfolio are calculated using the portfolio’s average PD.
+The shape component is obtained as the difference between the average component and the actual amount of capital requirements.
+I label the shape component *Capital savings* because an increase of the shape component reduces total capital requirements.
+
+To calculate *Capital savings*, I hand-collected information on the distribution of credit risk parameters from banks' Pillar-3 reports.
 The sample consists of 59 large banking groups that have adopted the IRB approach.
 The dataset covers 15 countries and goes from 2007 to 2018.
-Data at such granularity is not available in the common balance sheet datasets, and although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
 
 With this novel dataset, I investigate the relationship between *Capital savings* and the business cycle.
-Prior evidence suggests that capital requirements are procyclical: during economic downturns banks tend to deleverage instead of raising capital.
-However, as credit conditions deteriorate (and banks react to these new credit conditions) the PD distribution and, consequently, the amount of *Capital savings*, also change.
-A priori, the direction of the business cycle effect on *Capital savings* is ambiguous.
-At one extreme, if the entire portfolio is equally affected, then there is only a shift of the location parameter of the PD distribution.
-Because the PD-RW formula elasticity is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
-On the other hand, if riskier assets are more sensitive to credit shocks, then capital savings are countercyclical.
-The ambiguity arises because *Capital savings* can only increase during a recession if the shape of the PD distribution is affected during the cycle.
+Between the two capital requirement components, the average component is a source of the procyclicality as long the relationship between the business cycle and PD estimates is monotonic.
+Conversely, the direction of *Capital savings* during the business cycle is not trivial even when assuming a monotonic relationship between PD estimates and the business cycle.
+The ambiguity arises because *Capital savings* can only increase during a recession, and therefore be countercyclical, if the shape of the PD distribution is affected during the business cycle.
+If the PD estimates of the the assets in a portfolio are equally affected during the business cycle, then there is only a shift of the location parameter of the PD distribution.
+Because the PD elasticity of capital requirements is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
+On the other hand, if riskier assets are sufficiently more sensitive to credit shocks than safer assets, then *Capital savings* are countercyclical.
 
 I find *Capital savings* to be countercyclical.
-I estimate that a 1.6% GDP growth rate (one standard deviation) decreases *Capital savings* by 0.7 percentage points on average.
-In other words, banks have to increase capital by 0.7 percentage points to keep their capital ratio constant.
+I estimate that a 1.7% GDP growth rate (one standard deviation) decreases the growth rate of *Capital savings* by 0.8 percentage points on average.
+In other words, given an average PD, banks have to increase capital by 0.8 percentage points to keep their capital ratio constant.
 The effect is also economically significant.
-In a capital ratio decomposition exercise, I also find that the the cyclicality of *Capital savings* reduces the procyclicality of capital ratios by 23.7%.
-The results are robust to different measures of the shape component, such as, *Capital savings* in terms of ratios, *Capital savings* in monetary units, and using Gini coefficients.
-
-I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic expansions average PD and PD variance decrease while the PD skewness increase.
-This evidence implies that distributions become more symmetric and with less extreme values during recessions, suggesting a flight to safety behavior by banks.
-As the recession reduces their existing portfolio credit quality (evidenced by the negative coefficient of average PD) banks respond by reducing risk-taking.
+The variation of *Capital savings* during the business cycle reduces the procyclicality of capital ratio by 23.7% on average.
+The results also suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios.
+The results are robust to different measures of the shape component, to the inclusion of several control variables and the alternative sample selections.
+I also show that selection into the IRB approach does not explain the cyclicality of *Capital savings*.
+
+I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic recessions average and variance of PD distributions increase while the their skewness decrease.
+Although the effect of the average PD is stronger, which characterizes capital requirements as procyclical, both the higher variance and the lower skewness of the PD distribution mitigate the increase of capital requirements during recessions.
+Moreover, I argue that the change of skewness during the business cycle is consistent with a portfolio reallocation effect on *Capital savings*.
 
 This papers contributes to two strands of the banking literature.
-First, several studies document significant RW variability across otherwise similar banks.
+First, several studies document significant capital requirements variability across otherwise similar banks.
 Some take this evidence as indicative of either excessive subjectivity in the current capital requirement rules or risk measurement manipulation by banks to decrease capital charges [see, @mariathasan2014manipulation; @ferribank].
-Others are more cautious and emphasize that part of RW variability is desirable since it reflects differences in risk among banks [@cannata2012inside; @BCBS2016regulatory;@berg2017analysis; @EBA2019result].
-I contribute to the literature by showing that *Capital savings* are an important source of the RW variation across banks.
+Others are more cautious and emphasize that part of the capital requirements variability is desirable since it reflects differences in risk among banks [@cannata2012inside; @BCBS2016regulatory;@berg2017analysis; @EBA2019result].
+I contribute to the literature by showing that *Capital savings* are an important source of capital requirements variation across banks.
 
@@
 The remainder of this paper is organized as follows.
-Section \ref{sec:background describes the capital requirement formula under the IRB approach, why it creates *Capital savings*, and the hypothesis concerning the cyclicality of *Capital savings*.
+Section \ref{sec:background} describes the capital requirement formula under the IRB approach, why it creates *Capital savings*, and the hypothesis concerning the cyclicality of *Capital savings*.
 Section \ref{sec:data} describes the collected data and the measures of *Capital savings*.
@@
 Section \ref{sec:reallocation} tests if *Capital savings* are affected by portfolio reallocation during the business cycle.
-Section \ref{sec:additional_results} presentes additional results.
+Section \ref{sec:additional_results} presents additional results.
 Finally, section \ref{sec:conclusion} concludes.
@@
 
-In June 2004, the Basel Committee issued a revised framework on international convergence of capital measurements and capital standards [@basel2004international], known as Basel II, which serves as the basis for national rulemaking and implementation processes.
-At the heart of the revision was the link of capital charges to asset risk.
-Since the first accord from 1988 (Basel I), minimum capital requirements are calculated as a percentage of risk-weighted assets.
-Hence, the approach to define risk-weights is of major importance.
-In Basel I, risk-weights were flat within portfolio categories.
+In June 2004, the Basel Committee issued a revised framework on international convergence of capital measurements and capital standards [@BCBS2006international], known as Basel II, which serves as the basis for national rulemaking and implementation processes.
+The link of capital charges to asset risk was at the heart of the revision.
+Since the first accord from 1988 (Basel I), minimum capital requirements are calculated as a percentage of risk-weighted assets (RWAs).
+Hence, choosing the approach that defines risk-weights is crucial.
+In Basel I, RWs were constant within portfolio categories.
 For instance, all sovereign exposures were weighted by 0% and all corporate exposures by 100%.
@@
 
-In the IRB approach, capital requirements are derived from risk-weight formulas that take as input banks' own estimates of credit risk parameters.
-The rationale for allowing internal models is the better alignment of risk estimates to actual portfolio risk [@barakova2014banks] and the promotion of better risk management practices [@cucinelli2018credit].
-The basic regulatory formula derives from a model characterized by the property that the risk-weight of each asset depends only on the estimated credit risk parameters for this asset and not on the composition of the portfolio, i.e.
+Under the IRB approach, banks are required to estimate their credit risk parameters using internal risk models.
+The risk parameters are the probability of default (PD), the loss given default (LGD), and the exposure at default (EAD).
+The main rationale for allowing internal models is the better alignment of regulatory risk estimates to actual portfolio risk [@barakova2014banks] as banks are expected to have better knowledge of and more resources to monitor their portfolio than supervisors have.
+Because the internal models have to be approved by the supervisors before implemented for regulatory purposes, regulators also expected the IRB approach to promote stronger risk management practices in the banking industry [@BCBS2006international; @cucinelli2018credit].
+
+The estimated risk parameters are then used in regulatory formulas for calculating risk-weights.
+The regulatory formulas comes from a model characterized by the property that the risk-weight of each asset depends only on the estimated credit risk parameters for this asset and not on the composition of the portfolio, i.e.
 the model is portfolio invariant.
@@
 \end{equation}
-where $i$ is an asset in the portfolio and $N$ is the standard normal cumulative distribution.
-Basel accords segment credit portfolios into several categories (for instance, corporate, sovereign or retail), each having different additional features (maturity adjustments, risk-dependent correlation coefficient, among others) while keeping the basic structure of equation \ref{eq:RW_function}.
-Because of the portfolio invariant feature of the model, average risk-weight $\text{RW}^f$ is the exposure weighted average of all individual assets' risk-weights, as shown in equation \ref{eq:RW_factual}:
+where the subscript $i$ indexes an asset in the portfolio, $N$ is the standard normal cumulative distribution, and $R$ is a correlation coefficient.
+Basel accords segment credit portfolios into several categories each having different additional features while keeping the basic structure of equation \ref{eq:RW_function}.
+For instance, the correlation coefficient is a function of PD for wholesale exposures while it is constant at 15% for real estate exposures.[^1] Although the regulatory formulas are common for all banks the scope that banks are allowed to use their internal models varies.
+Under the foundation IRB (F-IRB) approach, banks can apply or be approved to use only internal PD estimates and are required to use values defined by the regulator for the other parameters.
+Only banks approved to the advanced IRB (A-IRB) approach can use internal estimates for all the risk parameters.
+
+Because of the portfolio invariant feature of the model, average risk-weight, $\text{RW}^f$, is the exposure weighted average of all individual assets' risk-weights, as shown in equation \ref{eq:RW_factual}:[^2]
 
@@
 \label{eq:RW_factual}
-\text{RW}^f= \frac{\sum_{i}\text{EAD}_i \text{RW}(\text{LGD}_i, \text{PD}_i)}{\sum_{i}\text{EAD}_i}= \sum_{i}\text{q}_i \text{RW}(\text{LGD}_i, \text{PD}_i).
+\text{RW}^f= \sum_{i}q_i \text{RW}(\text{LGD}_i, \text{PD}_i).
 \end{equation}
-where $q_i= EAD_i/\sum_jEAD_j$.
-Thus, to calculated their capital requirements banks estimate three credit risk parameters: the exposure at default (EAD), the loss given default (LGD), and the probability of default (PD).[^1] The total amount of RWA for this portfolio is $\text{RW}^f\times \sum_i\text{EAD}_i$ and total capital requirements $\text{RWA}\times 8\%$.
-
-[^1]: For wholesale exposures (corporate, sovereign and banks) the effective maturity of the asset is another parameter and for SMEs annual sales turnover is used to adjust the correlation coefficient.
+where $q_i= \text{EAD}_i/\sum_j\text{EAD}_j$. The total amount of RWAs for this portfolio is $\text{RW}^f\times \sum_i\text{EAD}_i$ and total capital requirements $\text{RWA}\times 8\%$.
+
+[^1]: Some other differences are maturity adjustments for wholesale exposures and the use of annual sales turnover to adjust the correlation coefficient for SMEs.
+
+[^2]: The superscript $f$ (for factual) is used to differentiate the average value from the RW function.
 
@@
 
-Figure \ref{fig:pdmap} plots the mapping from PD to RW according to equation \ref{eq:RW_function}.
-For this paper, the important characteristic of this mapping is that it is a concave function.
-The concavity of the PD-RW mapping together with the bottom-up method to aggregate capital requirements entail that actual capital requirements are lower than if they were calculated using the average PD to the entire portfolio.
+Figure \ref{fig:pdmap} plots the mapping from PD to risk-weights according to equation \ref{eq:RW_function}.
+For the purpose of this paper, the important characteristic of this mapping is that it is a concave function.
+The concavity of the regulatory formula together with the bottom-up method to aggregate capital requirements entail that total capital requirements for a portfolio do not depend only on the portfolio average PD but also on the shape of the PD distribution.
 For clarity, consider two portfolios.
 The first contains a low PD asset and a high PD asset in proportions $q$ and $1-q$.
-The second has the same total amount as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD weighted by their respective shares.
+The second has the same amount of exposure as the first but contains only a mean PD asset, where mean PD is the mean of low PD and high PD weighted by their respective shares.
 Because of the concavity of the IRB formula and Jensen’s inequality, the capital requirement for the second portfolio is greater than the first.
 Figure \ref{fig:pdmap} illustrates the comparison.
-$\text{RW}^f$ and $\text{RW}^c$ are the average RW for the first portfolio and the second portfolio, respectively.
-More generally, total capital requirements do not depend only on the portfolio average PD but also on the shape of the PD distribution.
-
+$\text{RW}^f$ and $\text{RW}^c$ are the average risk-weight for the first portfolio and the second portfolio, respectively.
+More generally, any PD distribution with positive variance is charged with capital requirements lower than implied by the average of the distribution.
 I propose a method that decomposes the contribution of the PD distribution to capital requirements into the average component and the shape component.
-The method is a generalization of the example in Figure \ref{fig:pdmap}: for each bank-year combination, I compare the actual capital requirement ($\text{RW}^f$) to the counterfactual case where the requirement for the entire portfolio is calculated using the portfolio’s average PD.
-The counterfactual RW is the average component of the PD distribution to capital requirements:
+The method is a generalization of the example in Figure \ref{fig:pdmap}.
+The average component is the amount of capital requirements if the portfolio’s average PD is used for the entire portfolio.
+Similar to $\text{RW}^c$ in Figure \ref{fig:pdmap}, the average component is defined as the following:
 
@@
 \label{eq:RW_counterfactual}
-\text{RW}^c= \sum_{i}\text{q}_i \text{RW}(\text{LGD}_i,\overline{\text{PD}}),
+\text{RW}^c= \sum_{i}q_i \text{RW}(\text{LGD}_i,\overline{\text{PD}}),
 \end{equation}
@@
 
-The difference between $\text{RW}^c$ and $\text{RW}^f$ measures any contribution to capital requirements from the PD distribution that is no related to the location of this distribution, i.e., it is a measure of the shape component.
+The difference between $\text{RW}^c$ and $\text{RW}^f$ measures any contribution to capital requirements from the PD distribution that is not related to the location of this distribution, i.e., it is a measure of the shape component.
 It can also be interpreted as the amount of *Capital savings* per unit of EAD due to the concavity of the IRB formula and the shape of the PD distribution.
-Equation \ref{eq:RW_savings} shows one of the possibilities to measure *Capital savings*:
+I define this intensity measure of *Capital savings* as the following:
 
@@
 
-Note that I defined *Capital savings* such that positive values translate to lower actual capital requirements and consequently $\text{RW}^s \geq 0$.
+The total amount of *Capital savings* is simply $\text{RW}^{s}$ multiplied by total EAD in the portfolio as follows:
+
+\begin{equation}
+\label{eq:measure2}
+\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}^f=\text{RW}^{s}\times \sum_i\text{EAD}_i.
+\end{equation}
+
+Because this is a measure in monetary units, $\text{RWA}^{\$}$ is mechanically larger for larger banks and they increase as the portfolio rolls out to the IRB approach.
+
+An alternative way to measure the intensity of *Capital savings* is the ratio of equation \ref{eq:RW_counterfactual} to equation \ref{eq:RW_factual} as follows:
+
+\begin{equation}
+\label{eq:measure1}
+\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}^f}.
+\end{equation}
+
+$\text{RWA}^s$ is the rate at which a bank saves capital.
+For instance, a value of 1.5 means this portfolio would require 50% more capital in the absence of *Capital savings* to keep the same capital ratio.
+A value of one means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
+Any PD distribution with positive variance results in positive *Capital savings*.
+
+$\text{RWA}^s$ is also my preferred measure of *Capital savings* because it is the most straightforward measure to evaluate the impact of *Capital savings* on the capital ratio during the business cycle.
+In the empirical analysis, I study the behavior of variables during the cycle by relating their growth rates to GDP growth rate.
+Hence, I regress the change in the logarithm of my variables of interest on GDP growth rate.
+Taking my empirical approach into account, $\text{RWA}^s$ becomes the natural measure of *Capital savings* as the following decomposition of the log of capital ratio shows:
+
+\begin{equation}
+\begin{aligned}
+\label{eq:decomposition}
+\text{Log}\bigg(\frac{\text{Capital}}{\text{RWA}^f}\bigg) &= \text{Log Capital} -  \text{Log RWA}^f \\
+&= \text{Log Capital} -  \text{Log}\bigg(\frac{\text{RWA}^c}{\text{RWA}^s}\bigg) \\
+&= \text{Log Capital} -  \text{Log RWA}^c + \text{Log RWA}^s
+\end{aligned}
+\end{equation}
+
+If I regress all the change of these four variables - $\frac{\text{Capital}}{\text{RWA}^f}$, $\text{Capital}$, $\text{Log RWA}^c$, and $\text{Log RWA}^s$ - on GDP growth rate I obtain four estimates $\hat\beta_{\text{Cap.
+ratio}}$, $\hat\beta_{\text{Capital}}$, $\hat\beta_{\text{RWA}^c}$, and $\hat\beta_{\text{RWA}^s}$, respectively.
+Using the predicted values implied by these estimated in Equation \ref{eq:decomposition} we get the following  equality: $\hat\beta_{\text{Cap.
+ratio}} = \hat\beta_{\text{Capital}} -\hat\beta_{\text{RWA}^c} + \hat\beta_{\text{RWA}^s}$, which can be used to obtain the relative importance of changes of each of these three variables to changes of the capital ratio during the business cycle.
+
+Note that I defined *Capital savings* such that positive values translate to lower actual capital requirements and consequently $\text{RW}^s \geq 0$, $\text{RWA}^{\$} \geq 0$, and $\text{RWA}^s \geq 1$.
+
+Lastly, for robustness of the empirical analysis, I also calculate the Gini coefficient of the PD distribution as an alternative measure to my approach.
 
@@
 During economic downturns credit risk increases which raises capital requirements forcing banks to either raise equity or cut down lending.
-Theory and empirical evidence suggest that banks opt for the latter as raising equity is also more expensive during recessions [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
-
-Between the two capital requirement components, the average component is a source of the procyclicality as long the relationship between the business cycle and PD is monotonic.
-Consider again the example shown in Figure \ref{fig:pdmap} with a portfolio containing a safe asset with PD equal to $PD_L$ and a risky asset with PD equal to $PD_H$ in proportions $q$ and $1-q$.
-In Appendix \ref{sec:appendixa}, I show that for $\text{PD} \in (0,0.25)$, $\text{RW}'>0$, $\text{RW}''<0$ and $\text{RW}'''>0$.
-I also show in \ref{sec:appendixa} that, on average, XX% of assets' PD lies within this range.
-For simplicity, I assume a constant $\text{LGD}_i=\text{LGD}$.
-So, the average component for this portfolio is:
+Theory and empirical evidence suggest that banks opt for the latter as raising equity is costly, even more so during recessions [@danielsson2001academic; @kashyap2004cyclical; @repullo2012procyclical; @behn2016procyclical].
+
+Between the two capital requirement components, the average component is a source of the procyclicality as long the relationship between the business cycle and PD is negative and monotonic.
+Consider again the example shown in Figure \ref{fig:pdmap} with a portfolio containing a share $q$ of a safe asset with PD equal to $\text{PD}_L$ and a share $1-q$ of a risky asset with PD equal to $\text{PD}_H$ such that $\text{PD}_H>\text{PD}_L$.
+For practical reasons I restrict the analysis to the  $\text{PD} \in [0,0.25)$.
+Figure \ref{fig:densities} in Appendix \ref{sec:appendixa} shows that the percentage of PD within the $[0,0.25)$ interval is between 99.35% and 99.98%.
+Figure \ref{fig:derivatives} in Appendix \ref{sec:appendixa} shows that for $\text{PD} \in [0,0.25)$, $\frac{\partial\text{RW}}{\partial\text{PD}}>0$, $\frac{\partial^2\text{RW}}{\partial\text{PD}^2}<0$, and $\frac{\partial^3\text{RW}}{\partial\text{PD}^3}>0$.
+I assume, without loss of generality, a constant $\text{LGD}_i=\text{LGD}$ and constant $q$.
+Using the weighted average of this portfolio in equation \ref{eq:RW_counterfactual}, the average component for this portfolio is equal to the following:
 
@@
 \label{eq:RW_counter_example}
-\text{RW}^{c}(q,\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H).
+\text{RW}^{c}(\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H).
 \end{equation}
 
-Because $\text{RW}'>0$ and as long PD estimates are affected monotonic and negatively by GDP growth, the average component of capital requirements increases during economic downturns.
-
-The direction of *Capital savings* during the business cycle is not trivial even when assuming a monotonic relationship between PD estimates and the business cycle.
-To explain the ambiguity, I consider three cases.
-But first, note that using the example in Figure \ref{fig:pdmap}, *Capital savings* can be written as the following:
+Because $\frac{\partial\text{RW}}{\partial\text{PD}}>0$ and as long PD estimates are affected monotonic and negatively by GDP growth, the average component of capital requirements increases during economic downturns.
+
+Conversely, the direction of *Capital savings* during the business cycle is not trivial even when assuming a monotonic relationship between PD estimates and the business cycle.
+The ambiguity arises because *Capital savings* can only increase during a recession, and therefore be countercyclical, if the shape of the PD distribution is affected during the business cycle.
+If the PD estimates are equally affected during the business cycle, then there is only a shift of the location parameter of the PD distribution.
+Because the PD elasticity of capital requirements is lower for higher PD estimates, in this scenario *Capital savings* are procyclical.
+On the other hand, if riskier assets are sufficiently more sensitive to credit shocks than safer assets, then *Capital savings* are countercyclical.
+To illustrate the ambiguity, I consider three cases.
+
+First, note that using the example in Figure \ref{fig:pdmap}, *Capital savings* can be written as the following:
 
@@
 \label{eq:RW_savings_example}
-\text{RW}^{s}(q,\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}(\text{PD}_L)+(1-q)\text{RW}(\text{PD}_H).
+\text{RW}^{s}(\text{PD}_L, \text{PD}_H)=\text{RW}(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}(\text{PD}_L)+(1-q)\text{RW}(\text{PD}_H).
 \end{equation}
 
-We can find the effect of changes in the PD distribution to *Capital savings* by taking the total derivative of \ref{eq:RW_savings}:
+We can find the effect of changes in the PD distribution to *Capital savings* by taking the total derivative of \ref{eq:RW_savings_example}:
 
@@
 \begin{aligned}
-d\text{RW}^{s}(q,\text{PD}_L, \text{PD}_H)
+d\text{RW}^{s}(\text{PD}_L, \text{PD}_H)
 &=\bigg[q\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-q\text{RW}'(\text{PD}_L)\bigg]d\text{PD}_L\\
 &+\bigg[(1-q)\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-(1-q)\text{RW}'(\text{PD}_H)\bigg]d\text{PD}_H\\
-&+\bigg[\text{RW}(\text{PD}_H)-\text{RW}(\text{PD}_L)-(\text{PD}_H-\text{PD}_L)\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]dq\\
 \end{aligned}
@@
 Next, consider the case that all assets are affected equally during the business cycles in term of PD, i.e.,  $d\text{PD}_H=d\text{PD}_L=d\text{PD}$.
-Also, consider for now no portfolio reallocation, i.e., $dq=0$.
 In this scenario equation \ref{eq:RW_savings_diff} becomes:
@@
 
-Because $\text{RW}'$ is a convex function and Jensen's inequality, if $\text{dPD}_H=\text{dPD}_H=\text{dPD}$ then $\text{dRW}^{s}/\text{dPD}<0$.
-In this case, *Capital savings* are procyclical, decreasing when PD estimates increase, i.e., during economic downturns.
+Because $\frac{\partial^2\text{RW}}{\partial\text{PD}^2}$ is a convex function and Jensen's inequality, if $d\text{PD}_H=d\text{PD}_H=d\text{PD}$ then $d\text{RW}^{s}/d\text{PD}<0$.
+In this case, *Capital savings* are procyclical, decreasing during economic downturns.
 Figure \ref{fig:pdmapcasesymm} shows graphically this first case.
 
-Next, consider the case riskier assets are more affected during the business cycles in term of PD, i.e.,  $\text{dPD}_H>\text{dPD}_L$.For simplicity consider the extreme case of $\text{dPD}_H=\text{dPD}$ and $\text{dPD}_L=0$.
-In this scenario equation \ref{eq:RW_savings_diff} becomes:
+Next, consider the case that riskier assets are more affected during the business cycles in term of PD, i.e.,  $\text{dPD}_H>\text{dPD}_L$.
+For simplicity, consider the extreme case of $\text{dPD}_H=\text{dPD}$ and $\text{dPD}_L=0$.
+In this scenario, equation \ref{eq:RW_savings_diff} becomes:
 
@@
 
-Because $\text{RW}''<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H$, if $\text{dPD}_H=\text{dPD}$ and $\text{dPD}$ then $\text{dRW}^{s}/\text{dPD}>0$.
+Because $\frac{\partial^2\text{RW}}{\partial\text{PD}^2}<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H$ $\text{dRW}^{s}/\text{dPD}>0$.
 In this case, *Capital savings* are countercyclical, increasing when PD estimates increase.
 In other words, the marginal capital requirement decreases during economic downturns.
-Figure \ref{fig:pdmapcaseassym} shows graphically this second case.
-
-Finally, consider the case where $\text{dPD}_H=\text{dPD}$, $\text{dRW}^{s}=0$ and $\text{d}q=0$. Using equation \ref{eq:RW_savings_diff}, we can find $\text{dPD}_L$ satisfying these conditions:
+Figure \ref{fig:pdmapcaseassym} shows the second case graphically.
+
+Finally, consider the case where $\text{dPD}_H=\text{dPD}$ and $\text{dRW}^{s}=0$. Using equation \ref{eq:RW_savings_diff}, we can find $\text{dPD}_L$ satisfying these conditions:
 
@@
 \begin{aligned}
-\text{dPD}_{L}
-&=\text{dPD}\frac{(1-q)}{q}\frac{\bigg[\text{RW}'(\text{PD}_H)-\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]}{\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_L)\bigg]}\\
+d\text{PD}_{L}
+&=d\text{PD}\frac{(1-q)}{q}\frac{\bigg[\text{RW}'(\text{PD}_H)-\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)\bigg]}{\bigg[\text{RW}'(q\text{PD}_L+(1-q)\text{PD}_H)-\text{RW}'(\text{PD}_L)\bigg]}\\
 \end{aligned}
@@
 
-Because $\text{RW}'<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H>\text{PD}_L$, both denominator and numerator of equation \ref{eq:third_case} are negative and consequently, $\text{dPD}_L>0$.
-From the first case we also know that $\text{dPD}_L<\text{dPD}$.
+Because $\frac{\partial\text{RW}}{\partial\text{PD}}<0$ and $\text{PD}_H>q\text{PD}_L+(1-q)\text{PD}_H>\text{PD}_L$, both denominator and numerator of equation \ref{eq:third_case} are negative and consequently, $\text{dPD}_L>0$.
+We also know from the first case that $\text{dPD}_L<\text{dPD}$, otherwise $\text{dRW}^{s}$ would be negative.
 Hence, the difference between the impact of the business cycle on riskier and safer assets have to be sufficiently large such that *Capital savings* are countercyclical.
-
-Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the economic cycle affects the shape of the PD distribution.
+If that is not the case, then the procyclical effect on *Capital savings* shown in equation \ref{eq:first_case} dominates.
+
+Whether *Capital savings* are pro- or countercyclical is an empirical question, as the direction of the cyclicality depends on how the shape of the PD distribution changes during the business cycle.
 
 \section{Data and Capital Savings Measures} \label{sec:data}
-
-The main measure of *Capital savings* used in this paper is the ratio of equation \ref{eq:RW_counterfactual} to equation \ref{eq:RW_factual} as follows:
-
-\begin{equation}
-\label{eq:measure1}
-\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}}.
-\end{equation}
-
-$\text{RW}^r$ is the rate at which a bank saves capital.
-For instance, a value of 1.5 means this portfolio would require 50% more capital in the absence of *Capital savings* to keep the same capital ratio.
-A value of one means that there are no *Capital savings* and implies a portfolio with a degenerate PD distribution.
-Any PD distribution with positive variance results in positive *Capital savings*.
-
-An alternative way to measure *Capital savings* is in monetary units as follows:
-
-\begin{equation}
-\label{eq:measure2}
-\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}.
-\end{equation}
-
-Because this is a measure in US Dollars, it is affected by the size of the IRB portfolio.
-Therefore, differently from the main analysis, *Capital savings* are mechanically larger for larger banks and they increase as the portfolio rolls out to the IRB approach.
-
-In the analysis I also consider the measure used in section \ref{sec:hypotheses} ($\text{RW}^s$), which is equivalent to dividing equation \ref{eq:measure2} by total EAD and therefore it is another measure of the intensity of *Capital savings*.
-Lastly, I calculate the Gini coefficient of the PD distribution as a robustness measure to my approach.
-According to the analysis shown in section \ref{sec:hypotheses} PD distributions with larger Gini coefficients should also generate larger *Capital savings*.
 
@@
 Although very detailed, datasets from credit registers are restricted to one country, and may still lack the relevant information.
-Therefore, to overcome data limitations I collected information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports.
-The sample consists of 50 large banking groups, from 12 countries, that have adopted the IRB approach.
+Therefore, to overcome data limitations, I collect information on the distribution of credit risk parameters (PD, EAD and LGD) from banks' individual Pillar-3 reports.
+The sample consists of 59 large banking groups, from 15 countries, that have adopted the IRB approach.
 For each bank, I collected consolidated information from the year the IRB approach was approved until 2018.
-The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate and sovereign for wholesale, and real estate and qualifying revolving credit for retail).
+The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories (for instance, corporate, sovereign, and banks for wholesale, and real estate, qualifying revolving credit, and others for retail).
 This level of portfolio breakdown is important because regulatory formulas vary across categories, and consequently, more precise knowledge of portfolio composition reduces measurements errors.
 
-Figure \ref{fig:savingsbank} shows that the *Capital savings* calculated with the compiled dataset using the main measure \ref{eq:measure1} (minus one to improve readability).
+Figure \ref{fig:savingsbank} shows *Capital savings* calculated according to equation \ref{eq:measure1} i.e., the ratio of the average component of RWAs to actual RWAs.
 Each point is a bank-year observations and the bars indicates the average value across time for each bank.
-There is substantial variation across banks with Goldman Sachs saving the highest at an average rate of 120% and Mediobanca the lowest at an average rate of 18.4%.
-Variation across time is considerable with, for instance, *Capital savings ratio* for Deutsche Bank going from 108.7% in 2010 to 61.4% in 2014.
-*Capital savings* also varies between and within countries.
+There is substantial variation across banks with Goldman Sachs saving the highest at an average rate of 220% and Mediobanca the lowest at an average rate of 118%.
+Variation across time is considerable with, for instance, *Capital savings* for Deutsche Bank going from 209% in 2010 to 161% in 2014.
+*Capital savings* also vary between and within countries.
 Figure \ref{fig:savingscountry}  aggregates the dataset by country.
-In the figure, each point is a bank observation and the bars are the average of these points.
+In the figure, each point is the average *Capital savings* for a bank and the bars are the average of across banks within a country.
 Variation between countries reaches its maximum at 50% (Denmark versus Norway) and within at 75% (between BNY Mellon and Goldman Sachs in the United States).
 
-Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
-Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks].
-Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
-
-The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
-First, I control for bank size using the log of total assets (in millions of US dollars).
-The expectation regarding the impact of bank size is ambiguous.
-On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
-On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
-I also include the regulatory capital ratio as control.
-Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
-However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
-Thus, the expected effect of capital ratio on capital savings is also ambiguous.
-Next, I add banks' profitability using return on asset (ROA) and return on equity (ROE) with both measured before taxes.
-Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
-Finally, I study the relationship of *Capital savings* to risk-taking.
-I add the ratio of non-performing loans to total loans to test this proposition.
-
+Figure \ref{fig:savingsportfoliobank} shows *Capital savings* for wholesale and retail portfolios using the ranking of Figure \ref{fig:savingsbank}.
+The Figure suggests that *Capital savings* from wholesale portfolios are positively correlated to total *Capital savings*.
+This correlation seems not to be true for retail portfolios.
+
+Lastly, Figure \ref{fig:averagebank} plots, in terms of risk-weights, *Capital savings*, the average component, and total risk-weights, following equations \ref{eq:RW_savings}, \ref{eq:RW_counterfactual}, and \ref{eq:RW_factual}, respectively.
+Banks are ranked by total risk-weight.
+From equation \ref{eq:RW_savings}, the third column is obtained by subtracting the first column from the second, i.e., total risk-weights are equal to the average component minus *Capital savings*.
+As expected, the average component is larger than the *Capital savings* generated by the shape component.
+Still, because of *Capital savings*, the variation and level of total risk-weights are substantially lower than implied by the average component.
+Total risk-weights are 17 percentage points, on average, lower than the average component and their standard deviation is 6 percentage points lower.
 \section{Capital Savings and the Business Cycle} \label{sec:cycle}
@@
 In this section I investigate the relationship between *Capital savings* and the business cycle.
-I start by explaining the theoretical ambiguous relationship between *Capital savings* and the business cycle.
-Next I show that *Capital savings* are countercyclical, i.e., given an average portfolio PD, banks need more capital in good times than in bad times to reach the same capital ratio.
-Finally, I provide evidence that this relationship is not purely mechanical but depends on how banks reallocate their assets through the business cycle.
-
-Throughout this section, I report results from estimating the following specification:
+I show that *Capital savings* are countercyclical, i.e., for the same average PD, capital requirements are high when the economic is booming and low during recessions.
+
+Throughout this section, I report the results from estimating the following specification:
 
@@
 The dependent variable is the yearly change in a variable associated to portfolio $j$ for bank $i$ at year $t$.
-The change in the logarithm of real GDP per capita used comes from the country where the bank has its headquarter.
-All regressions include bank and year fixed effects and robust standard errors adjusted for clustering at the country level are reported in parentheses.
-
-Column 1 Table \ref{tab:main} shows results using equation \ref{eq:measure1}, my preferred measure of *Capital savings*.
-shows that total *Capital savings* are countercyclical.
-According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average.
-The effect is economically sizable representing 14% of total *Capital savings* standard deviation in the sample.
+The change in the logarithm of real GDP per capita comes from the country where the bank has its headquarter.
+All regressions include bank and year fixed effects and standard errors are clustered at the country-level.
+
+Table \ref{tab:main} shows the results using equation \ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure of *Capital savings*.
+According to regression 1, an increase of 1.7 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.8 percentage points on average.
+The effect is economically sizable representing 17% of total *Capital savings* standard deviation in the sample.
 
 The remaining columns of Table \ref{tab:main} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios.
-The estimated coefficient in Column 2 is negative suggesting that *Capital savings* generated by retail portfolios are also countercyclical.
+The negative estimated coefficient in Column 2 suggests that *Capital savings* generated by retail portfolios are also countercyclical.
 However, this coefficient is not statistically significant.
 On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant.
-According to regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation.
+According to regression 3, one standard deviation increase in GDP growth rate decreases the growth rate of *Capital savings* from wholesale portfolios by 0.9 percentage points on average.
+The effect is 19% of the dependent variable's standard deviation.
 Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios.
-Only in regression 4 the coefficient of GDP growth rate is negative and statistically significant.
-
-To shed light on the importance of *Capital savings*, I regress all the components of capital ratio on GDP growth rate.
-The capital ratio decomposition is shown in equation \ref{eq:decomposition}.
-The contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
-To guarantee that the sum of all effects add up to 100\%, I use in this analysis the non-winsorized data and restrict the sample to banks which I observe their capital ratio.
-Although the coefficient for *Capital savings* are no longer statistically significant, Table \ref{tab:contribution} shows that *Capital savings* are the only countercyclical component of capital ratios.
-The economic magnitude is also meaningful: *Capital savings* reduce the procyclicality of capital ratios by 11.6\%.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:decomposition}
-\Delta \text{Log Cap. Ratio}_{i,t} &= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}_{i,t} \\
-&= \Delta \text{Log Capital}_{i,t} - \Delta \text{Log RWA}^c_{i,t} + \Delta \text{Log RWA}^r_{i,t} 
-\end{aligned}
-\end{equation}
+Only in regression 4 the coefficient of GDP growth rate is both negative and statistically significant.
+
+Next, I evaluate the impact of *Capital savings* on the capital ratio during the business cycle.
+I regress the yearly change of each capital ratio component on GDP growth rate.
+As explained in Section \ref{sec:PD_capital_savings}, the effect of GDP growth rate on each component sum up to the effect on the capital ratio.
+Consequently, the contribution of each component to the cyclicality of the capital ratio is the ratio of their estimated coefficient to the coefficient for the capital ratio regression.
+
+Table \ref{tab:contribution} shows that the growth rate of *Capital savings* is the only countercyclical component of capital ratio growth rate.
+As expected, both the amount of capital and RWAs are procyclical which translate to capital ratios also being procyclical.
+The results also confirm the expectation that the procyclical effect of the average component dominates over the countercyclical *Capital savings*.
+Nevertheless, the contribution of *Capital savings* is economically relevant, reducing the procyclicality of capital ratios by 23.7\% on average.[^3]
+
+[^3]: The coefficient on total *Capital savings* differs slightly from Table \ref{tab:main} because in Table \ref{tab:contribution} the sample is restrict to banks with observable capital ratio and is winsorized at 1% instead of 5%.
+The process of winsorizing is also the reason why the contributions do not add-up to 100%.
 
@@
 In this section, I test the robustness of the main results to alternative measures of *Capital savings*.
-In Panel A of Table \ref{tab:alt}, the dependent variable measures *Capital savings* in US dollars following equation \ref{eq:measure2}.
+In Panel A of Table \ref{tab:alt}, the dependent variable measures the amount *Capital savings* in US dollars following equation \ref{eq:measure2}.
 As before, the estimated coefficients suggest that the countercyclical effect is mostly driven by wholesale portfolios.
@@
 In panel B of Table \ref{tab:alt} the measure of *Capital savings* in US dollars is rescaled by the amount of EAD in the respective portfolio following equation \ref{eq:RW_savings}.
-This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity, risk-weights.
+This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity — i.e., risk-weights.
 Although the coefficient on the wholesale portfolio is no longer statistically significant, total *Capital savings* and those coming from corporate portfolios are still statistically different from zero at the 1% confidence level.
-The attenuation of the effect on wholesale portfolio appears to come from large, albeit not significant, procyclical *Capital savings* in sovereign portfolios.
-Moreover, the small and statistically insignificant coefficient in regression 6 reinforces the claim that countercyclical *Capital savings* in the bank portfolio are due to the extensive and not the intensive margin.
-
-Lastly, in Panel C of Table \ref{tab:alt} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution.
-Again, the results shown in Panel C support the evidence from Table \ref{tab:alt}: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
+The attenuation of the effect on wholesale portfolio appears to come from large, albeit not statistically significant, procyclical *Capital savings* in sovereign portfolios.
+Moreover, the small and statistically insignificant coefficient in regression 6 reinforces the claim that countercyclical *Capital savings* in the bank portfolio are mostly due to the extensive and not the intensive margin.
+
+In Panel C of Table \ref{tab:alt} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution.
+Again, the results in Panel C support the evidence from Table \ref{tab:main} i.e., *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.
+
+Lastly, Appendix \ref{sec:appendixc} shows that the results are robust to the addition of bank- and country-level controls (Table \ref{tab:extension}), to including only banks with at least 10 observations (Panel A of Table \ref{tab:robustness}), and to winsorizing the sample to 1% level instead of 5% (Panel B of Table \ref{tab:robustness}).
 
@@
 In this section I argue that portfolio reallocation during the business cycle explains part of the *Capital savings* cyclicality.
-As method, I use the observed movements on moments of the PD distribution during the business cycle that are not in line with the expected movements from a static portfolio as evidence in favor of a portfolio reallocation effect on *Capital savings*.[^2]
-
-[^2]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PD estimates evolution to the PD estimates of new loans.
+As method, I study the changes of the PD distribution moments during the business cycle and check if these changes are consistent with a portfolio reallocation effect on *Capital savings*.[^4]
+
+[^4]: Ideally, for this purpose, I would track individual assets during the cycle and compare their PD estimates evolution to the PD estimates of new loans.
 Unfortunately, in my dataset, I only observe assets aggregated by PD bands without knowledge if amount changes in a band are due to new loans or changes in credit quality of assets already held.
 
-According to the theoretical framework presented in section 4.1, an increase in average PD reduces the amount of *Capital savings* if other moments are kept constant.
+According to the theoretical framework presented in section \ref{sec:hypotheses}, an increase in average PD reduces the amount of *Capital savings* holding other moments constant.
 This relationship is illustrated in Figure \ref{fig:pdmapcasesymm} and empirically tested in Table \ref{tab:moments_contribution}.
@@
 
-Panel A in table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle.
+Panel A in Table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle.
 As expected, in most of the regressions, changes in average PD are negatively associated with GDP growth.
-This association is the known driver of the procyclical character of the current model-based capital regulation and its interpretation is fairly mechanical.
-During recessions average credit quality deteriorates, or in other words, average PD increases.
-The only exception to the rule is the sovereign portfolio.
-Column 5 of panel A shows that the coefficient of GDP growth rate on changes in the average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero.
-One possible reason for this result is that, during economic downturns, banks may search for safe assets mostly in the form of safe haven government bonds.
-According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to compensate any deterioration of the prior held portfolio.
-Together with column 5 of Table \ref{tab:moments_contribution}, it also helps explain the positive (and insignificant) results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.
+This association is the driver of the procyclical character of the current model-based capital regulation.
+The only exception is the sovereign portfolio.
+Column 5 of Panel A shows that the coefficient of GDP growth rate on changes in the average PD of sovereign portfolios is positive, although not statistically different from zero.
+A possible explanation for the positive and insignificant result is that during economic downturns banks may search for safe assets mostly in the form of safe haven government bonds.
+According to column 5 of Panel A, this reallocation of portfolio appears to be sufficient to compensate any deterioration of the prior held portfolio.
+Together with column 5 in Table \ref{tab:moments_contribution}, Column 5 of Panel A in Table \ref{tab:moments_cyclicality} also helps explain the positive and insignificant results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.
 
 Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios.
-And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, panel B in Table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios.
-In sum, the estimated coefficients of the average PD and the PD variance in Table \ref{tab:moments_contribution} and panels A and B in Table \ref{tab:moments_cyclicality} provide strong empirical support to the theoretical framework presented in section 4.1.
-
-At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted, exactly because the estimated effect of these two moments on *Capital savings* are strongly aligned with the theoretical framework which portfolio reallocation is not present.
-Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness and kurtosis.
-It is trivial that the equivalent case with a continuous PD distribution to the scenario in Figure \ref{fig:pdmapcasesymm} wouldn't affect either of these moments.
+And in line with the evidence from Tables \ref{tab:main} and \ref{tab:alt}, Panel B in Table \ref{tab:moments_cyclicality} shows that the effect of the business cycle on the PD variance is concentrated on corporate portfolios.
+In sum, the estimated coefficients of the average PD and the PD variance in Table \ref{tab:moments_contribution} and Panels A and B in Table \ref{tab:moments_cyclicality} provide strong empirical support to the theoretical framework presented in section \ref{sec:hypotheses}.
+
+At the same time, however, using only the evidence provided by the average and variance of the PD distribution, the hypothesis that portfolio reallocation does not impact *Capital savings* cannot be refuted exactly because the estimated effect of these two moments on *Capital savings* are strongly aligned with the theoretical framework, which does not allow for portfolio reallocation.
+Consider now a negative shock to a continuous PD distribution, such that we can measure its skewness.
+It is trivial that the equivalent case with a continuous PD distribution to the scenario in Figure \ref{fig:pdmapcasesymm} would not affect the skewness of the distribution.
+Still in this case, only a change in the location parameter occurs.
 On the other hand, in the second scenario the PD distribution should become more skewed to the right.
@@
 
-Panel C in Table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle cannot be explained without portfolio reallocation.
-All the statistical significant coefficients are positive, contrary to the expectation if banks wouldn't reallocate their portfolio.
-Considering an economic downturn, the positive coefficients suggest that banks fight the deterioration of their portfolio credit quality by reducing risk taking in new loans.
-The combination of these two movements, higher PD of existing exposure and lower PD of new assets, results in a less skewed PD distribution during recessions.
+Panel C in Table \ref{tab:moments_cyclicality} provides evidence that the observed relationship between PD distribution skewness and the business cycle is consistent with portfolio reallocation.
+All the statistical significant coefficients are positive, contrary to the expectation if banks would not reallocate their portfolio.
+The positive coefficients suggest that banks respond to the deterioration of their portfolios' credit quality by reducing risk taking in new loans.
+The combination of these two movements, higher PD of existing exposures and lower PD of new assets, results in a less skewed PD distribution during recessions.
 
@@
 The effect is statistical significant for all except sovereign portfolios.
-According to column 1, an increase of 0.182 percentage points in PD distribution skewness growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.009 percentage points on average, representing 20% of total *Capital savings* standard deviation.
-The effect is even stronger on corporate portfolios accounting for 40% of their *Capital savings* standard deviation.
-The evidence of Panel C in Table \ref{tab:moments_cyclicality} and Table \ref{tab:moments_contribution} suggest that as banks flight to safety in economic recessions, the increase in *Capital savings* due to the associated less skewed PD distribution helps banks to maintain their capital ratios.
-This countercyclical effect driven by the skewness of the PD distribution is on top of any gain from a lower average PD.
+According to column 1, an increase of 18.2 percentage points in PD distribution skewness growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.9 percentage points on average, representing 20% of total *Capital savings* standard deviation.
+The effect is even stronger on corporate portfolios accounting for 41% of their *Capital savings* standard deviation.
+
+The evidence of Panel C in Table \ref{tab:moments_cyclicality} and Table \ref{tab:moments_contribution} suggest that part of the increase in *Capital savings* in economic recessions is consistent with a portfolio reallocation effect.
+This countercyclical effect, driven by the skewness of the PD distribution, is on top of any gain from a lower average PD also a result of the reduced risk taking.
 
@@
 
-\section{Additional Results} \label{sec:additional_results}
-
-Besides the business cycle, I also test if bank-specific characteristics are associated to *Capital savings*.
-Several works relate bank-specific characteristics to either total RWA or average RW [@vallascas2013risk; @mariathasan2014manipulation; @ferribank; @gropp2018banks ].
-In this section, I use these prior evidence to develop my expectations on how each of these variables may affect capital savings.
-
-Table \ref{tab:summary} shows summary statistics for the all the dependent variables, the main variable of interest (GDP growth rate), and all the other variables used in the analysis.
-
-The first set of variables captures banks’ opportunities and incentives to affect RWA by means of capital arbitrage.
-First, I control for bank size using the log of total assets (in millions of US dollars).
-The expectation regarding the impact of bank size is ambiguous.
-On the one hand, larger banks might be able to adopt more sophisticated internal risk models and better exploit modelling choices to increase capital savings.
-On the other hand, larger banks might be subject to closer regulatory scrutiny and, as a result, may find it more difficult to manipulate their risk estimates.
-I also include the regulatory capital ratio as control.
-Better capitalized banks could be subject to less regulatory scrutiny [@calem1999impact] and, consequently, in a better position to engage in regulatory arbitrage.
-However, the incentives to manipulate risk parameters should be lower for banks with high capital ratios as they are at a greater distance to the minimum requirement.
-Thus, the expected effect of capital ratio on capital savings is also ambiguous.
-Next, I add bank profitability using net income over total assets.
-Profitable banks face fewer incentives to understate the riskiness of their assets and to engage in regulatory arbitrage more generally.
-
-The second set of variables captures differences in the regulatory treatment of bank activities on capital savings.
-For instance, I control for for differences in banks’ business models by including the ratios of customer loans to total assets and net interest income to operational revenue.
-Under Basel accords, the risk weights formulas are concave for lending activities, such as customer loans.
-As a result, I expect banks with higher shares of these variables to have more *Capital savings*.
-
-Finally, I study the relationship of *Capital savings* to risk-taking.
-The RW decomposition analysis suggests that riskier banks benefit more from *Capital savings*.
-I add the ratio of non-performing loans to total loans to test this proposition.
-I also include bank funding using the ratio of total deposits to total assets.
-Because deposits are a more stable and cheap form of funding, banks with a larger deposit base are, therefore, more likely to take risks.
-The effect of deposits funding on *Capital savings* depends on how risk-taking is manifested by the deposit funding.
-If banks increase their entire portfolio riskiness, i.e., if risk-taking is manifested simply as a shift in the PD distribution location parameter, then deposits to total assets are negatively correlated to *Capital savings*.
-Conversely, *Capital savings* increase if banks take risk by having fatter upper-tails in the PD distribution.
+In this section, I show that selection into IRB is not driving my results.
+Although banks cannot opt out of IRB approach once it has been approved, they can try to (re)allocate assets to portfolios under the less risk-sensitive standardized approach.
+Consequently, the countercyclical effect of *Capital savings* can potentially be confounded with this across approach reallocation.
+
+Table \ref{tab:other} provides little support for cross approach reallocation during the business cycle.
+The growth rate of the total amount (column 1), the average risk-weight (column 2), and the share under the standardized approach (column 3) are all positively associated with GDP growth rate but none of the coefficients are statistically significant.
+The positive coefficient in column 3 is also not in line with the expectation that banks would move assets to the standardized approach during economic downturns.
+The positive coefficient in column 1 is most likely driven by the balance sheet expansion during economics booms while the positive coefficient in column 2 is reflecting portfolio reallocation within the standardized approach during the business cycle, for instance, from sovereign to corporate lending.
+
+Lastly, I test if banks reallocate across portfolios under the IRB approach during the business cycle.
+According to columns 4 and 5 in Table \ref{tab:other}, it seems that banks increase retail exposure at the expense of wholesale exposure during periods of economic expansion.
+However, I find this reallocation to be statistically insignificant.
+
+Overall, Table \ref{tab:other} supports ruling out selection concerns to the result that *Capital savings* are a countercyclical component of capital requirements.
 
@@
 This paper contributes to understanding the cyclicality and variability of capital requirements by identifying the effects of an unexplored feature of the IRB framework on capital requirements.
-*Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are economically relevant.
-*Capital savings* are also countercyclical and reduce the procyclicality of RWA by 23.7%.
-Hence, the concavity of IRB formulas could be explored to reduce procyclicality of capital regulation.
+*Capital savings* due to the shape of PD distribution and the concavity of the IRB formulas are countercyclical and economically relevant.
+*Capital savings* reduce the procyclicality of the capital ratio by 23.7%.
+Overall, this paper provides evidence that the concavity of IRB formulas could be explored to reduce the procyclicality of current capital regulation.
 
 \section*{References} 
-
+\noindent
 <div id="refs"></div>
@@
 ```{r, results="hide"}
-pd_map_cap <-  "\\label{fig:pdmap}The figure plots the mapping from probabilities of default to capital requirementes per unit of exposure according to equation \\ref{eq:RW_function}.
-$LGD$ is set equal to one and the correlation coefficient to 0.15."
+pd_map_cap <-  "\\label{fig:pdmap}The figure plots the mapping from PD to risk-weights according to equation \\ref{eq:RW_function}. $LGD$ is set equal to one and the correlation coefficient to 0.15."
 
@@
 ```{r}
-cyclicality_cap_symm = "\\label{fig:pdmapcasesymm}The Figure shows the case where capital savings are procyclical.
-The dotted lines correspond to the potfolio before the credit risk shock (dPD) and the dashed lines to the portfolio after the the shock.
-The segmentes in the plot show the size of the capital savings before the shock."
+cyclicality_cap_symm = "\\label{fig:pdmapcasesymm}The figure shows how changes in PD can lead to procyclical \\textit{Capital savings}.
+The dotted lines correspond to the portfolio before the credit risk shock and the dashed lines to the portfolio after the the shock.
+The arrow segments in the plot show the size of the capital savings before the shock."
 ```
@@
 ```{r}
-cyclicality_cap_asym = "\\label{fig:pdmapcaseassym}The Figure shows the case where capital savings are countercyclical.
-The dotted lines correspond to the potfolio before the credit risk shock (dPD) and the dashed lines to the portfolio after the the shock.
-The segmentes in the plot show the size of the capital savings before the shock.
+cyclicality_cap_asym = "\\label{fig:pdmapcaseassym}The figure shows how changes in PD can lead to countercyclical \\textit{Capital savings}.
+The dotted lines correspond to the portfolio before the credit risk shock and the dashed lines to the portfolio after the the shock.
+The arrow segments in the plot show the size of the capital savings before the shock.
 "
@@
 ```{r}
-savings_cap = "\\label{fig:savingsbank}The figure plots the log of banks capital savings.
-This measure is calculated using equation \\ref{eq:measure1}.
-Each point in the plot is a bank-year observation and bars are the averages values for each bank."
-```
-
-```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
+savings_cap = "\\label{fig:savingsbank}The figure plots \\textit{Capital savings} using equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure.
+Each point in the plot is a bank-year observation and bars are the averages across time for each bank."
+```
+
+```{r, include = TRUE, fig.height = 7.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
 bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
@@
  ggplot(mapping = aes(x = reorder(name, RWA.r.Total, FUN = mean, na.rm=TRUE),
-                      y = RWA.r.Total-1)) +
+                      y = RWA.r.Total)) +
+geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+    labs(x=NULL) +
+    coord_flip(ylim=c(1, 2.5)) +
+  labs(y =  unname(TeX(c("$RWA^{s}$" )))) +
+  scale_fill_grey()+
+    theme_bw()
+
+```
+
+```{r}
+savings_cap_country = "\\label{fig:savingscountry}The figure plots \\textit{Capital savings} using equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure.
+Each point is the across time average \\textit{Capital savings} for a bank and the bars are the average across banks within a country."
+```
+
+```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
+bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid) %>% 
+  select(country.name,name, year, RWA.r.Total)%>%
+  group_by(country.name,name) %>%
+ summarise_if(is.numeric,mean,na.rm= TRUE)%>%
+ ggplot(mapping = aes(x = reorder(country.name, RWA.r.Total, FUN = mean, na.rm=TRUE),
+                          y = RWA.r.Total)) +
+geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+    labs(x=NULL) +
+  labs(y =  unname(TeX(c("$RWA^{s}" )))) +
+  coord_flip(ylim=c(1, 2.5)) +
+  theme_bw()
+
+```
+```{r}
+savings_portfolio_cap = "\\label{fig:savingsportfoliobank}The figure plots \\textit{Capital savings} using equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs, as measure.
+The left-hand side plots shows \\textit{Capital savings} for wholesale portfolios and the right-hand side for retail portfolios.
+Each point in the plot is a bank-year observation and bars are the averages across time for each bank."
+```
+
+```{r, include = TRUE, fig.height = 7.5, fig.width = 6,results="hide",fig.cap= savings_portfolio_cap, fig.align = 'center'}
+bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
+  select(name, Country, RWA.r.Total, RWA.r.Retail, RWA.r.Wholesale)%>%
+  pivot_longer(cols = c("RWA.r.Retail","RWA.r.Wholesale"), names_to = "Portfolio", values_to = "RWA.r") %>%
+  mutate(Portfolio = factor(Portfolio, 
+                            levels = c("RWA.r.Wholesale","RWA.r.Retail"), 
+                            labels = c("Wholesale","Retail")))%>%
+ ggplot(mapping = aes(x = reorder(name, RWA.r.Total, FUN = mean, na.rm=TRUE),
+                      y = RWA.r)) +
+geom_point(alpha = 0.5,size = 2) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+    labs(x=NULL) +
+    coord_flip(ylim=c(1, 3.5)) +
+  labs(y =  unname(TeX(c("$RWA^{s}$" )))) +
+  scale_fill_grey()+
+    theme_bw() +
+  facet_wrap(~Portfolio)
+
+```
+
+```{r}
+average_cap = "\\label{fig:averagebank}The figure plots, in terms of risk-weights, \\textit{Capital savings} ($\\text{RW}^s$), the average component ($\\text{RW}^c$), and total risk-weights ($\\text{RW}^f$), following equations \\ref{eq:RW_savings}, \\ref{eq:RW_counterfactual}, and \\ref{eq:RW_factual}, respectively.
+Each point in the plot is a bank-year observation and bars are the averages across time for each bank."
+```
+
+```{r, include = TRUE, fig.height = 7.5, fig.width = 6,results="hide",fig.cap= average_cap, fig.align = 'center'}
+RW_summary <-  bank.data %>%
+  mutate(RW.ranking = RW.IRB) %>%
+  select(name, Country, RW.ranking, RW.s.Total, RW.hat.Total, RW.IRB)%>%
+  pivot_longer(cols = c("RW.hat.Total","RW.s.Total","RW.IRB"), names_to = "Portfolio", values_to = "RW") %>%
+  mutate(Portfolio = factor(Portfolio, 
+                            levels = c("RW.s.Total", "RW.hat.Total", "RW.IRB"), 
+                            labels = c("j = s","j = c", "j = f")))%>%
+  group_by(Portfolio) %>%
+  dplyr::summarize(mean = round(mean(RW, na.rm = TRUE),2),
+            sd   = round(sd(RW, na.rm = TRUE),2)) %>%
+  mutate(lab = paste("Mean = ", mean, "\n St. Dev. = ", sd, sep = ""))
+
+bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
+  mutate(RW.ranking = RW.IRB) %>%
+  select(name, Country, RW.ranking, RW.s.Total, RW.hat.Total, RW.IRB)%>%
+  pivot_longer(cols = c("RW.hat.Total","RW.s.Total","RW.IRB"), names_to = "Portfolio", values_to = "RW") %>%
+  mutate(Portfolio = factor(Portfolio, 
+                            levels = c("RW.s.Total", "RW.hat.Total", "RW.IRB"), 
+                            labels = c("j = s","j = c", "j = f")))%>%
+ ggplot(mapping = aes(x = reorder(name, RW.ranking, FUN = mean, na.rm=TRUE),
+                      y = RW)) +
 geom_point(alpha = 0.5,size = 2) +
@@
     coord_flip() +
-  labs(y =  unname(TeX(c("$RWA^{s}-1" )))) +
+  labs(y =  unname(TeX(c("$RW^{j}$" )))) +
   scale_fill_grey()+
-    theme_bw()
-
-```
-
-```{r}
-savings_cap_country = "\\label{fig:savingscountry}The figure plots the banks capital savings ratio.
-This measure is calculated using equation \\ref{eq:measure1}.
-For each bank I used the average across time such that each point in the plot is a bank observation and bars are the averages values for each country."
-```
-
-```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
-bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid) %>% 
-  select(Country,name, year, RWA.r.Total)%>%
-  group_by(Country,name) %>%
- summarise_if(is.numeric,mean,na.rm= TRUE)%>%
- ggplot(mapping = aes(x = reorder(Country, RWA.r.Total, FUN = mean, na.rm=TRUE),
-                          y = RWA.r.Total-1)) +
-geom_point(alpha = 0.5,size = 2) +
-  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
-    labs(x=NULL) +
-  labs(y =  unname(TeX(c("$RWA^{s}-1" )))) +
-  coord_flip() +
-  theme_bw()
+    theme_bw() +
+  facet_wrap(~Portfolio)+
+  geom_text(data = RW_summary, aes(label = lab), x = 3, y = 0.85, size=3.5)
 
@@
 p_w <- c(0.05, 0.95)        
-bank.data.win <- bank.data %>% 
+bank.data.win5 <- bank.data %>% 
   ungroup()%>%
@@
 p_w <- c(0.01, 0.99)        
-bank.data.win2 <- bank.data %>% 
+bank.data.win1 <- bank.data %>% 
   ungroup()%>%
@@
 
-```{r PD_distribution_data, eval=FALSE, echo=FALSE}
-
-IRB.data <- as.data.frame(pillar3.data) %>%
-  filter(Method2 %in% c("IRB"),
-         PD  < 0.99, EAD > 0) %>%
-  mutate(LGD        = ifelse(is.na(LGD), 0.45, LGD),
-         w.PD       = PD*EAD*LGD,
-         w.EAD      = LGD*EAD)
-
-PD.distribution <- data.frame()
-
-for(i in c("Total","Wholesale","Retail","Corporate","Sovereign","Banks")){
-PD.distribution <- IRB.data %>%
-  select(bvdid, name, year, w.EAD, PD, Portfolio_1, Portfolio_2) %>%
-  {if(i %in% c("Total")){
-     .
-    }else if(i %in% c("Wholesale","Retail")){
-     filter(., Portfolio_1 %in% c(i))
-    }else{
-     filter(., Portfolio_2 %in% c(i))}} %>%
-  mutate(w.EAD = round(w.EAD/100)) %>%
-  na.omit(cols=c("w.EAD")) %>%
-  expandRows("w.EAD") %>%
-  rbind(PD.distribution)}
-
-```
-
-```{r}
-density_cap = "\\label{fig:pdmapcases}The figure shows the two possible outcomes of a negative credit risk shock on a portfolio.
-The left-hand side plot shows the case where capital savings are procyclical.
-The right-hand side plot shows the case where capital savings are countercyclical.
-"
-```
-
-```{r, density, eval=FALSE, fig.height = 3, fig.width = 6,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= denisty_cap}
-
-ggplot(PD.distribution) + geom_density(aes(x=PD))
-
-view(select(pillar3.data, name, year, PD, Portfolio_1, Portfolio_2))
-```
+\newpage
 
@@
 
-```{r reg_cyclicality_main, echo=FALSE}
-
-# Variables in the model 
+```{r tables_aux, echo=FALSE}
+
 # Dependent variables
@@
 dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
+
 # Independent variables
@@
 
+```
+
+```{r reg_cyclicality_main, echo=FALSE}
+
 # Model formulas
@@
 reg.list <- lapply(formulas, function(x) {
-   felm(x, data = bank.data.win)
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
+   felm(x, data = bank.data.win5)
 })
@@
 \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$.
-In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}.
+In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs.
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
@@
           covariate.labels = ind.text,
-          #se               = se.list,
           add.lines        = fe.list,
@@
 # Dependent variables
-dep.vars2 <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat.Total_gr", "log.RWA.r.Total_gr")
-dep.text2 <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
+dep.vars.temp <- c("log.CAR_gr", "log.K_gr", "log.RWA.IRB_gr", "log.RWA.hat.Total_gr", "log.RWA.r.Total_gr")
+dep.text.temp <- c("$\\Delta\\text{Log capital ratio}_{i,t}$","$\\Delta\\text{Log capital}_{i,t}$", "$\\Delta\\text{Log RWA}_{i,t}^{f}$", "$\\Delta\\text{Log RWA}_{i,t}^{c}$", "$\\Delta\\text{Log RWA}_{i,t}^{s}$")
 
 # Model formulas
-formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formula.text <- paste0(paste0(dep.vars.temp)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = filter(bank.data.win2, !is.na(log.CAR_gr)))
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
+  felm(x, data = filter(bank.data.win1, !is.na(log.CAR_gr)))
 })
@@
 
-fe.list.car <- list(c("Effect", effect.list),
-                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
+fe.list.car <- list(c("Contribution", effect.list),
+                c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars.temp))),
+                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars.temp))))
 ```
@@
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,j,t}^{s} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  Each column uses one component of the capital ratio according to equation \\ref{eq:decomposition} as dependent variable.
+          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Y}_{i,t}^{s} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+\\end{equation*} \\newline Where $\\Delta\\text{Y}_{i,t}$ is one component of the capital ratio according to equation \\ref{eq:decomposition}.
+The dependent variables are the change in the logarithm of the capital ratio in column 1, of total capital in column 2, of total RWAs in column 3, of the average component in column 4, and of \\textit{Capital savings} in column 5.
+The average component is the amount of capital requirements if the portfolio’s average PD is used for the entire portfolio.
+\\textit{Capital savings} is the ratio of the average component of RWAs to actual RWAs (equation \\ref{eq:measure1}).
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+The table shows the contribution of each component to the cyclicality of capital ratios, which is the respective coefficient divided by the coefficient in column 1.
 All regressions include bank and year fixed effects.
@@
 *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text2,
+          dep.var.labels   = dep.text.temp,
           covariate.labels = ind.text,
-          #se               = se.list,
           add.lines        = fe.list.car,
-          #keep            = c("log.gdp_gr"),
           keep.stat        = c("n","rsq"),
@@
           column.sep.width = "2pt",
-          #float.env        = "sidewaystable",
           align            = TRUE)
@@
 The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$.
-The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:RW_savings} in panel B.
+In Panel A, \textit{Capital savings} are measure in monetary units as RWAs, following equation \ref{eq:measure2}.
+In Panel B, \textit{Capital savings} are measure in terms of risk-weights, following equation equation \ref{eq:RW_savings}.
 In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution.
@@
 \begin{tabular}{@{\extracolsep{1pt}}cccccc}
-\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}$} \\
+\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}_{i,t}$} \\
 \end{tabular}
@@
 reg.list <- lapply(formulas, function(x) {
-   felm(x, data = bank.data.win)
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
+   felm(x, data = bank.data.win5)
 })
@@
           covariate.labels = ind.text,
-          #se              = se.list,
           add.lines        = fe.list,
@@
           notes.append     = FALSE,
-          #label            = "tab2",
           style            = "qje",
@@
  \\
-\multicolumn{6}{c}{Panel B: $\text{Y}=\text{Log RW}^{s}$ } \\
+\multicolumn{6}{c}{Panel B: $\text{Y}=\text{Log RW}^{s}_{i,t}$ } \\
 \end{tabular}
@@
 \\
-\multicolumn{6}{c}{Panel C: $\text{Y}=\text{Log Gini}$ } \\
+\multicolumn{6}{c}{Panel C: $\text{Y}=\text{Log Gini}_{i,t}$ } \\
 \end{tabular}
@@
 
-```{r reg_cyclicality_components, echo=FALSE}
-
-dep.vars2 <- c("log.RWA.SA_gr","RW.SA_gr ", "q.SA_gr", "q.Wholesale_gr", "q.Retail_gr")
-dep.text2 <- c("$\\Delta\\text{Log RWA}_{i,t}^{SA}$", "$\\Delta\\text{Log RW}_{i,t}^{SA}$",
-               "$\\Delta\\text{q}_{i,t}^{SA}$", "$\\Delta\\text{q}_{i,t}^{Wholesale}$", "$\\Delta\\text{q}_{i,t}^{Retail}$")
-
-# Creating formulas and running the models
-formula.text <- paste0(paste0(dep.vars2)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
-formulas <- lapply(formula.text, formula)
-
-reg.list <- lapply(formulas, function(x) {
- felm(x, data = filter(bank.data.win, year>2007))
-})
-
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_cyclicality_components, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-stargazer(reg.list,
-          title = "Other RWA Components and the Business Cycle",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{13.5cm}{The table examines the relationship between the business cycle and other components of RWA.
-The dependent variable is the change in the logarithm of risk-weighted assets for portfolios under the standardize approach in column 1, the change in risk-weights for portfolios under the standardize approach in column 2, the change in the ratio of exposure under the standardize approach to total exposure in column 3, the change in the wholesale share of the IRB portfolio in column 4, and the change in the retail share of the IRB portfolio in column 5.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-All regressions include bank and year fixed effects.
-Robust standard errors adjusted for clustering at the country level are reported in parentheses.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-          dep.var.labels   = dep.text2,
-          covariate.labels = ind.text,
-          #se               = se.list,
-          add.lines        = fe.list,
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "small",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "tab:other",
-          style            = "qje",
-          column.sep.width = "2pt",
-          align            = TRUE)
-
-```
-
-\newpage
-
 ```{r reg_moments, echo=FALSE}
 
-ind.text.moments <- c("$\\Delta \\text{Mean(PD}_{i,t}$)", "$\\Delta \\text{Var(PD}_{i,t}$)", "$\\Delta \\text{Skew(PD}_{i,t}$)", "$\\Delta \\text{Kurt(PD}_{i,t}$)")
+ind.text.temp <- c("$\\Delta \\text{Log mean(PD}_{i,t}^j$)", "$\\Delta \\text{Log var(PD}_{i,t}^j$)", "$\\Delta \\text{Log skew(PD}_{i,t}^j$)", "$\\Delta \\text{Kurt(PD}_{i,t}^j$)")
 
@@
 for(i in dep.vars){
-formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ mean + var + skew   | year + bvdid | 0 | Country "))
- data = rename(bank.data.win,
+formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ mean + var + skew | year + bvdid | 0 | Country"))
+ data = rename(bank.data.win5,
                mean    = paste0("mean.PD"  , i, "_gr"),
@@
 
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
 ```
@@
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log mean(PD}_{i,t}^j\\text{)}+\\gamma\\Delta \\text{Log var(PD}_{i,t}^j\\text{)}+\\kappa\\Delta \\text{Log skew(PD}_{i,t}^j\\text{)}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
 \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$.
-In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs.
+$\\Delta\\text{Log mean(PD}_{i,t}^j$) is the change in the logarithm of the average of the portfolio $j$'s PD distribution.
+$\\Delta\\text{Log var(PD}_{i,t}^j$) is the change in the logarithm of the variance of the portfolio $j$'s PD distribution.
+$\\Delta\\text{Log skew(PD}_{i,t}^j$) is the change in the logarithm of the skewness of the portfolio $j$'s PD distribution.
 All regressions include bank and year fixed effects.
@@
           dep.var.labels   = dep.text,
-          covariate.labels = ind.text.moments,
-          #se               = se.list,
+          covariate.labels = ind.text.temp,
           add.lines        = fe.list,
@@
 \end{equation*} \newline
-The dependent variable is the change in a PD distribution moment.
-In panel A the dependent variable is the mean PD, in panel B the variance of PD and in panel C the PD skewness.
+The dependent variable is the change in the logarithm of the mean in Panel A, of the variance in Panel B, and of the skewness in Panel C of the portfolio $j$'s PD distribution.
 $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
@@
 \begin{tabular}{@{\extracolsep{1pt}}cccccc}
-\multicolumn{6}{c}{Panel A: $\text{Y}=\Delta \text{Log mean(PD}_{i,t}\text{)}$} \\
+\multicolumn{6}{c}{Panel A: $\text{Y}=\Delta \text{Log mean(PD}_{i,t}^j\text{)}$} \\
 \end{tabular}
@@
 # Model formulas
-formula.text <- paste0(paste0("mean.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formula.text <- paste0(paste0("mean.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
+  felm(x, data = bank.data.win5)
 })
 
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
+```
+
+```{r table_mean, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          label            = "tab2",
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
+
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\\
+\multicolumn{6}{c}{Panel B: $\text{Y}=\Delta \text{Log var(PD}_{i,t}^j\text{)}$}
+\end{tabular}
+
+```{r reg_var, echo=FALSE}
+
+# Model formulas
+formula.text <- paste0(paste0("var.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
+formulas <- lapply(formula.text, formula)
+
+# Run the models
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = bank.data.win5)
 })
@@
 
-```{r table_mean, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+```{r table_var, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
@@
 \\
-\multicolumn{6}{c}{Panel B: $\text{Y}=\Delta \text{Log var(PD}_{i,t}\text{)}$}
+\multicolumn{6}{c}{Panel C: $\text{Y}=\Delta \text{Log skew(PD}_{i,t}^j\text{)}$}
 \end{tabular}
 
-```{r reg_var, echo=FALSE}
+```{r reg_skew, echo=FALSE}
 
 # Model formulas
-formula.text <- paste0(paste0("var.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+formula.text <- paste0(paste0("skew.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
 formulas <- lapply(formula.text, formula)
@@
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
+  felm(x, data = bank.data.win5)
 })
 
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_var, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+```
+
+```{r table_skew, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 
@@
 ```
-
-\begin{tabular}{@{\extracolsep{1pt}}cccccc}
-\\
-\multicolumn{6}{c}{Panel C: $\text{Y}=\Delta \text{Log skew(PD}_{i,t}\text{)}$}
-\end{tabular}
-
-```{r reg_skew, echo=FALSE}
-
-# Model formulas
-formula.text <- paste0(paste0("skew.PD", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
+\end{table}
+
+\newpage
+
+```{r reg_cyclicality_components, echo=FALSE}
+
+dep.vars.temp <- c("log.RWA.SA_gr","RW.SA_gr ", "q.SA_gr", "q.Wholesale_gr", "q.Retail_gr")
+dep.text.temp <- c("$\\Delta\\text{Log RWA}_{i,t}^{SA}$", "$\\Delta\\text{Log RW}_{i,t}^{SA}$",
+               "$\\Delta\\text{q}_{i,t}^{SA}$", "$\\Delta\\text{q}_{i,t}^{Wholesale}$", "$\\Delta\\text{q}_{i,t}^{Retail}$")
+
+# Creating formulas and running the models
+formula.text <- paste0(paste0(dep.vars.temp)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
 formulas <- lapply(formula.text, formula)
 
-# Run the models
 reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
+ felm(x, data = filter(bank.data.win5, year>2007))
 })
 
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_skew, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
-
-table <- capture.output(
-{stargazer(reg.list,
-          table.layout     = "=c=d#-t-as-",
-          dep.var.labels   = dep.text,
+```
+
+```{r table_cyclicality_components, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+stargazer(reg.list,
+          title = "Other RWA Components and the Business Cycle",
+          table.layout = "n=c=d#-t-as-",
+          notes = "\\parbox[t]{13.5cm}{The table examines the relationship between the business cycle and other components of RWA.
+The dependent variable is the change in the logarithm of risk-weighted assets for portfolios under the standardize approach in column 1, the change in risk-weights for portfolios under the standardize approach in column 2, the change in the ratio of exposure under the standardize approach to total exposure in column 3, the change in the wholesale share of the IRB portfolio in column 4, and the change in the retail share of the IRB portfolio in column 5.
+$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
+          dep.var.labels   = dep.text.temp,
           covariate.labels = ind.text,
-          #se              = se.list,
           add.lines        = fe.list,
@@
           type             = "latex",
-          font.size        = "footnotesize",
+          font.size        = "small",
           notes.label      = "",
@@
           notes.append     = FALSE,
-          label            = "tab2",
+          label            = "tab:other",
           style            = "qje",
           column.sep.width = "2pt",
-          align            = TRUE,
-          float            = FALSE)})
-
-table <- gsub("{-3}",
-              "{5.5}"
-              , table, fixed = T)
-
-cat(table)
-
-```
-\end{table}
+          align            = TRUE)
+
+```
 
@@
 
-```{r reg_extension, echo=FALSE, include=FALSE}
-
-# Variables in the model 
-# Independent variables
-aux.1 <- c("log.gdp",
-           "bank_credit_gdp",
-           "pillarI_stringency",
-           #"cap_reg",
-           "Sup_Power",
-           "NPL.ratio",
-           "log.asset",
-           "ROE",
-           "ROA")
-ind.vars  <- paste(aux.1, collapse=" + ")
-ind.text <- c("$\\text{Log GDP}_{i,t}$",
-              "$\\text{Credit-to-GDP}_{i,t}$",
-              "$\\text{Pillar I stringency}_{i,t}$", 
-              #"$\\text{Capital stringency}_{i,t}$", 
-              "$\\text{Official Supervisory power}_{i,t}$",
-               "$\\text{NPL Ratio}_{i,t}$",
-              "$\\text{Log Total Assets}_{i,t}$",
-              # "$\\text{Loans / Total Assets}_{i,t-1}$", 
-              #"$\\text{Net Interest Income / Oper. Rev.}_{i,t-1}$",
-             # "$\\text{Net Income / Total Assets}_{i,t-1}$", 
-              "$\\text{ROE}_{i,t}$",
-              "$\\text{ROA}_{i,t}$")
-
-fe.list <- list(c("Bank FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))),
-                c("Year FE",    rep("\\multicolumn{1}{c}{Yes}", length(dep.vars))))
-
-# Model formulas
-formula.text <- paste0(paste0("log.RWA.r", dep.vars)," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country")
-formulas <- lapply(formula.text, formula)
-
-# Run the models
-reg.list <- lapply(formulas, function(x) {
-  felm(x, data = bank.data.win)
-})
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
-})
-
-```
-
-```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
-
-stargazer(reg.list,
-          title = "Correlates of Capital Savings",
-          table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\text{Log RWA}^{s}_{i,t} = \\gamma X_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
-\\end{equation*} \\newline  The first line of the table show the dependent variable in each regression.
-$\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
-$X_{i,t-1}$ is a vector of independent variables.
-All regressions include bank and year fixed effects.
-Robust standard errors adjusted for clustering at the country level are reported in parentheses.
-*, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
-
-          dep.var.labels   = dep.text,
-          covariate.labels = ind.text,
-          se               = se.list,
-          add.lines        = fe.list,
-          keep.stat        = c("n","rsq"),
-          header           = FALSE, 
-          type             = "latex",
-          font.size        = "footnotesize",
-          notes.label      = "",
-          notes.align      = "c", 
-          notes.append     = FALSE,
-          label            = "tab:extension",
-          style            = "qje",
-          column.sep.width = "-3pt",
-          #float.env        = "sidewaystable",
-          align            = TRUE)
+\appendix
+
+\section{} \label{sec:appendixa}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+```{r}
+derivatives = "\\label{fig:derivatives} The figure plots risk-weights as a function of probability of default (top-left) and its first, second, and third derivative with respect to probability of default (top-right, bottom-left, and bottom-right, respectively)."
+```
+```{r, fig.cap= derivatives, fig.align="center"}
+p1 <- ggdraw() + draw_image("../../Auxiliar/Pictures/function.jpg", scale = 0.9)
+p2 <- ggdraw() + draw_image("../../Auxiliar/Pictures/first_derivative.jpg", scale = 0.9)
+p3 <- ggdraw() + draw_image("../../Auxiliar/Pictures/second_derivative.jpg", scale = 0.9)
+p4 <- ggdraw() + draw_image("../../Auxiliar/Pictures/third_derivative.jpg", scale = 0.9)
+
+plot_grid(p1,p2,p3,p4, ncol = 2)  
+```
+```{r}
+densities = "\\label{fig:densities} The figure plots probability of default histograms by portfolio for all banks and years.
+The last bin contains all observations with PD higher or equal to 25%.
+Defaulted exposures are excluded."
+```
+
+```{r, fig.height = 6, fig.width = 6, results="hide",fig.cap= densities, include=TRUE}
+
+PD_summary <-  pillar3.data %>%
+  select(name, PD, EAD, Portfolio_2, Portfolio_1) %>%
+  filter(EAD>0, PD<1, Portfolio_2 %in% c("Banks", "Corporate", "Sovereign") | Portfolio_1 %in% c("Retail"))%>%
+  na.omit(cols=c("EAD", "Portfolio")) %>%
+  mutate(EAD = round(EAD/10))%>%
+  expandRows("EAD")%>%
+  mutate(Portfolio = ifelse(Portfolio_1 == "Retail", "Retail", as.character(Portfolio_2)),
+         low25PD = ifelse(PD<0.25,1,0),
+         low5PD = ifelse(PD<0.05,1,0)) %>%
+           group_by(Portfolio) %>%
+  dplyr::summarize(mean = round(100*mean(PD),2),
+            sd   = round(100*sd(PD),2),
+            p25 = round(100*sum(low25PD)/n(),2),
+            p5 = round(100*sum(low5PD)/n(),2)) %>%
+  mutate(lab = paste("Mean = ", mean, "% \n St. Dev. = ", sd, "% \n P(PD<25%) = ",p25,"% \n P(PD<5%) = ",p5, "%", sep = ""))
+
+ pillar3.data %>%
+ select(name, PD, EAD, Portfolio_2, Portfolio_1) %>%
+  filter(EAD>0, PD<1, Portfolio_2 %in% c("Banks", "Corporate", "Sovereign") | Portfolio_1 %in% c("Retail"))%>%
+  mutate(Portfolio = ifelse(Portfolio_1 == "Retail", "Retail", as.character(Portfolio_2)),
+         EAD = round(EAD/10),
+         PD = ifelse(PD>0.25, 0.25, PD)) %>%
+  na.omit(cols=c("EAD", "Portfolio")) %>%
+  expandRows("EAD")%>%
+  ggplot(mapping = aes(x = PD)) +
+    geom_histogram(aes(y=..density..), bins = 50)+
+  labs(x=NULL, y= NULL) +
+   theme_bw() +
+  theme(panel.grid.minor = element_blank())+
+  scale_fill_grey() +
+  facet_wrap(~ Portfolio) +
+  geom_text(data = PD_summary, aes(label = lab), x = 0.15, y = 150)
+
 ```
@@
 \newpage
-\section*{Appendix A} \label{sec:appendixa}
-
-\setcounter{table}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-
-\newpage
-\section*{Appendix B} \label{sec:appendixb}
+\section{}  \label{sec:appendixb}
 
@@
 ```{r summary1,  fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-table <-capture.output({bank.data.win %>%
+table <-capture.output({bank.data.win5 %>%
         filter(year>2007) %>% 
@@
 
-table <- c(table[1:41],"Panel C: Other dependet variables \\\\",table[-(1:41)])
+table <- c(table[1:41],"Panel C: Other dependent variables \\\\",table[-(1:41)])
 table <- c(table[1:41],"\\\\[-1.8ex]",table[-(1:41)])
@@
 ```{r summary3, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
-table <- capture.output({bank.data.win %>%
+table <- capture.output({bank.data.win5 %>%
                          filter(year>2007) %>% 
@@
   skew.PD.Banks_gr,
+  log.asset_gr,
+  NPL.ratio_gr,
+  log.asset,
+  NPL.ratio,
+  deposit.ratio,
+  NII.ratio,
+  Zscore,
+  ROE,
+  ROA,
   log.gdp_gr,
+  credit_gdp_gr,
   log.gdp,
   credit_gdp,
-  pillarI_stringency,
   cap_reg,
-  Sup_Power,
-  log.asset,
-  NPL.ratio,
-  Zscore,
-  ROE,
-  ROA) %>%
+  Sup_Power) %>%
   as.data.frame() %>% 
@@
               "\\hspace{1.0cm}$\\Delta$ Log PD skewness banks", 
+              "$\\Delta \\text{Log total assets}$",
+              "$\\Delta \\text{NPL / loans}$",
+              "Log total assets",
+              "NPL / loans",
+              "Deposit / total assets",
+              "NII / oper. rev.",
+              "Z-score",
+              "ROE",
+              "ROA",
                "$\\Delta \\text{Log GDP}$",
+              "$\\Delta \\text{Credit / GDP}$",
               "Log GDP",
               "Credit / GDP",
-              "Capital Stringency Index",
-              "Pillar I Stringency",
-              "Official Supervisory Power",
-              "Log Total Assets",
-              "Non-performing Loans / Loans",
-              "Z-score",
-              "ROE",
-              "ROA"),
+              "Capital stringency",
+              "Official supervisory power"),
           notes.label = "",
@@
 
-table <- c(table[1:33],"Panel E: Independent variables \\\\",table[-(1:33)])
-table <- c(table[1:33]," \\\\[-1.8ex]",table[-(1:33)])
+table <- c(table[1:27],"Panel E: Other bank-level variables \\\\",table[-(1:27)])
+table <- c(table[1:27]," \\\\[-1.8ex]",table[-(1:27)])
+
+table <- c(table[1:39],"Panel F: Country-level variables \\\\",table[-(1:39)])
+table <- c(table[1:39]," \\\\[-1.8ex]",table[-(1:39)])
 
@@
 
-#table <- gsub("{-3}", "{3.4}", table, fixed = T)
-
 cat(table)
@@
 
-```{r reg_constant, echo=FALSE}
-
-reg.list <- list()
+\newpage
+\section{}  \label{sec:appendixc}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{C\arabic{table}}
+
+```{r reg_extension, echo=FALSE, include=FALSE}
+
+# Variables in the model 
+# Independent variables
+aux.1 <- c("log.gdp_gr",
+           "bank_credit_gdp_gr",
+           "log.gdp",
+           "cap_reg",
+           "Sup_Power",
+           "log.asset_gr",
+           "NPL.ratio_gr",
+           "log.asset",
+           "NPL.ratio",
+           "deposit.ratio",
+           "NII.ratio")
+ind.vars.temp  <- paste(aux.1, collapse=" + ")
+ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$",
+              "$\\Delta\\text{Credit / GDP}_{i,t}$",
+              "$\\text{Log GDP}_{i,t}$",
+              "$\\text{Capital stringency}_{i,t}$", 
+              "$\\text{Official supervisory power}_{i,t}$",
+              "$\\Delta\\text{Log total assets}_{i,t}$",
+              "$\\Delta\\text{NPL / loans}_{i,t}$",
+              "$\\text{Log total assets}_{i,t}$",
+              "$\\text{NPL / loans}_{i,t}$",
+              "$\\text{Deposit / total assets}_{i,t}$",
+              "$\\text{NII / oper. rev.}_{i,t}$")
+
+# Model formulas
+formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars.temp)," | year + bvdid | 0 | Country")
+formulas <- lapply(formula.text, formula)
+
 # Run the models
-for(i in dep.vars){
-formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ log.gdp_gr | year + bvdid | 0 | Country "))
- data = filter(bank.data.win,
-               get(paste0("log.RWA.r"  , i, "_gr_n"))>9)%>%
-   felm(data = ., formula = formula.text ) -> reg.list[[i]]
-  }
-
-# Robust standard errors
-se.list  <-  lapply(reg.list, function(x){
-  summary(x, robust=TRUE)$coefficients[,2]
+reg.list <- lapply(formulas, function(x) {
+  felm(x, data = bank.data.win5)
 })
@@
 
-```{r table_constant, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+```{r table_extension, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE, include=TRUE, eval=TRUE}
+
 stargazer(reg.list,
-          title = "Capital Savings and the Business Cycle",
+          title = "Capital Savings and the Business Cycle: Control Variables",
           table.layout = "n=c=d#-t-as-",
-          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
+          notes = "\\parbox[t]{17cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}^{s,j}_{i,t} =\\beta \\Delta\\text{Log GDP}_{i,t} + \\gamma \\Delta  X_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}.
 \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$.
-In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}.
+In all regressions the measure of \textit{Capital savings} follows equation \ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs.
 $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+$X_{i,t}$ is a vector of independent variables.
 All regressions include bank and year fixed effects.
@@
           notes.append     = FALSE,
-          label            = "tab:constant",
+          label            = "tab:extension",
           style            = "qje",
-          column.sep.width = "1pt",
+          column.sep.width = "-3pt",
           align            = TRUE)
 ```
+
+\begin{table}[!htbp] \centering  
+\caption{Capital Savings and the Business Cycle: Robustness to Alternative Samples}  
+\label{tab:robustness}
+\footnotesize
+\begin{tabular}{@{\extracolsep{1pt}}lccccc}
+\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Log RWA}_{i,t}^{s,j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}.
+\end{equation*} \newline The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$.
+In all regressions the measure of \textit{Capital savings} follows equation \ref{eq:measure1}, the ratio of the average component of RWAs to actual RWAs.
+$\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate.
+All regressions include bank and year fixed effects.
+Robust standard errors adjusted for clustering at the country level are reported in parentheses.
+*, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
+[-1.8ex] \\
+\hline \\
+\end{tabular}\\
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\multicolumn{6}{c}{Panel A: Balance panel} \\
+\end{tabular}
+
+```{r reg_constant, echo=FALSE}
+
+reg.list <- list()
+# Run the models
+for(i in dep.vars){
+formula.text <- formula(paste0("log.RWA.r", i, "_gr ~ log.gdp_gr | year + bvdid | 0 | Country"))
+ data = filter(bank.data.win5,
+               get(paste0("log.RWA.r"  , i, "_gr_n"))>9)%>%
+   felm(data = ., formula = formula.text ) -> reg.list[[i]]
+  }
+
+```
+
+```{r table_constant, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
+
+\begin{tabular}{@{\extracolsep{1pt}}cccccc}
+\\
+\multicolumn{6}{c}{Panel B: Sample winsorized at 1\%}
+\end{tabular}
+
+```{r reg_winsorized, echo=FALSE}
+
+# Run the models
+formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr ~ log.gdp_gr | year + bvdid | 0 | Country"))
+formulas <- lapply(formula.text, formula)
+reg.list <- lapply(formulas, function(x) {
+   felm(x, data = bank.data.win1)
+})
+
+```
+
+```{r table_winsorized, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
+
+table <- capture.output(
+{stargazer(reg.list,
+          table.layout     = "=c=d#-t-as-",
+          dep.var.labels   = dep.text,
+          covariate.labels = ind.text,
+          add.lines        = fe.list,
+          keep.stat        = c("n","rsq"),
+          header           = FALSE, 
+          type             = "latex",
+          font.size        = "footnotesize",
+          notes.label      = "",
+          notes.align      = "c", 
+          notes.append     = FALSE,
+          style            = "qje",
+          column.sep.width = "2pt",
+          align            = TRUE,
+          float            = FALSE)})
+
+table <- gsub("{-3}",
+              "{5.5}"
+              , table, fixed = T)
+
+cat(table)
+
+```
+
+\end{table}

### 2021-04-23 Analysis/Draft/Draft_bookdown.Rmd
version dissertation pre defense

@@
 thanks: Preliminary draft.
-JEL: G21, G28, G11
+JEL: G11, G21, G28 
 urlcolor: black
@@
 library(lemon)                ## Plot brackets
-library(mosaic)               ## Package to calculate derivatives
+#library(mosaic)               ## Package to calculate derivatives
 #library(textablr)
@@
 The results also suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios.
-The results are robust to different measures of the shape component, to the inclusion of several control variables and the alternative sample selections.
+The results are robust to different measures of the shape component, to the inclusion of several control variables and alternative sample selections.
 I also show that selection into the IRB approach does not explain the cyclicality of *Capital savings*.
 
-I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic recessions average and variance of PD distributions increase while the their skewness decrease.
+I further analyse the relationship of the PD distribution moments with GDP growth and find that during economic recessions the average and the variance of PD distributions increase while the their skewness decrease.
 Although the effect of the average PD is stronger, which characterizes capital requirements as procyclical, both the higher variance and the lower skewness of the PD distribution mitigate the increase of capital requirements during recessions.
@@
 
-```{r, fig.height = 3, fig.width = 4,results="hide", fig.cap = pd_map_cap, fig.align = 'center', echo=FALSE, message=FALSE}
+```{r mapping, fig.height = 3, fig.width = 4,results="hide", fig.cap = pd_map_cap, fig.align = 'center', echo=FALSE, message=FALSE}
 p <- ggplot(data = subset(data,subset = PD <0.2),
@@
 
-```{r, include = TRUE, fig.height = 7.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
+```{r savings_by_bank, include = TRUE, fig.height = 5.5, fig.width = 4,results="hide",fig.cap= savings_cap, fig.align = 'center'}
 bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
@@
                       y = RWA.r.Total)) +
-geom_point(alpha = 0.5,size = 2) +
-  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+geom_point(alpha = 0.5,size = 1) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3, aes(width=0.6)) +
     labs(x=NULL) +
@@
   scale_fill_grey()+
-    theme_bw()
+    theme_bw()+
+  theme(text = element_text(size=8))
 
@@
 
-```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
+```{r savings_by_country, include=TRUE, fig.height = 3, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
 bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid) %>% 
@@
 
-```{r, include = TRUE, fig.height = 7.5, fig.width = 6,results="hide",fig.cap= savings_portfolio_cap, fig.align = 'center'}
+```{r savings_by_portfolio, include = TRUE, fig.height = 5.5, fig.width = 5,results="hide",fig.cap= savings_portfolio_cap, fig.align = 'center'}
+
+RW_summary <-  bank.data %>%ungroup()%>%
+  select(name, Country, RWA.r.Total, RWA.r.Retail, RWA.r.Wholesale)%>%
+  dplyr::summarise(across(c("RWA.r.Retail","RWA.r.Wholesale"), 
+                          ~round(cor(RWA.r.Total,., use="complete.obs"),2),))%>%
+  pivot_longer(cols = everything(), names_to = "Portfolio", values_to = "Corr")%>%
+   mutate(Portfolio = factor(Portfolio, 
+                            levels = c("RWA.r.Wholesale","RWA.r.Retail"), 
+                            labels = c("Wholesale","Retail")))%>%
+  mutate(lab = paste("Correlation = ", Corr, sep = ""))
+
 bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
@@
                       y = RWA.r)) +
-geom_point(alpha = 0.5,size = 2) +
-  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+geom_point(alpha = 0.5,size = 1) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3, aes(width=0.6)) +
     labs(x=NULL) +
@@
     theme_bw() +
+  theme(text = element_text(size=8))+
   facet_wrap(~Portfolio)
+#+
+ # geom_text(data = RW_summary, aes(label = lab), x = 4, y = 2.8, size=2.5)
 
@@
 
-```{r, include = TRUE, fig.height = 7.5, fig.width = 6,results="hide",fig.cap= average_cap, fig.align = 'center'}
+```{r decomposition, include = TRUE, fig.height = 5.5, fig.width = 5,results="hide",fig.cap= average_cap, fig.align = 'center'}
 RW_summary <-  bank.data %>%
@@
                       y = RW)) +
-geom_point(alpha = 0.5,size = 2) +
-  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
+geom_point(alpha = 0.5,size = 1) +
+  stat_summary(fun.y="mean", geom="bar",alpha = 0.3, aes(width=0.6)) +
     labs(x=NULL) +
@@
     theme_bw() +
+  theme(text = element_text(size=8))+
   facet_wrap(~Portfolio)+
-  geom_text(data = RW_summary, aes(label = lab), x = 3, y = 0.85, size=3.5)
+  geom_text(data = RW_summary, aes(label = lab), x = 3, y = 0.85, size=2.5)
 
@@
 ```
-```{r, fig.cap= derivatives, fig.align="center"}
+```{r derivatives, fig.cap= derivatives, fig.align="center", fig.width = 5}
 p1 <- ggdraw() + draw_image("../../Auxiliar/Pictures/function.jpg", scale = 0.9)
@@
 
-```{r, fig.height = 6, fig.width = 6, results="hide",fig.cap= densities, include=TRUE}
+```{r densities, fig.height = 5, fig.width = 5, results="hide",fig.cap= densities, include=TRUE}
 
@@
   facet_wrap(~ Portfolio) +
-  geom_text(data = PD_summary, aes(label = lab), x = 0.15, y = 150)
+  geom_text(data = PD_summary, aes(label = lab), x = 0.15, y = 150, size=2.5)
 
